### 2019-08-05 paper/main.tex
add current paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{color}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\newcommand{\bxput}[3]{put$_{BX}$ \ {#1} \ {#2} \ {#3}}
\newcommand{\bxget}[2]{get$_{BX}$ \ {#1} \ {#2}}
\newcommand{\revput}[3]{put$_{REV}$ \ {#1} \ {#2} \ {#3}}
\newcommand{\revget}[2]{get$_{REV}$ \ {#1} \ {#2}}
\newcommand{\sif}[3]{if \ {#1} \ then \ {#2} \ else \ {#3}}
\newcommand{\product}[2]{({#1} `cons' {#2})}
\newcommand{\tuple}[2]{(#1 :: #2)}
\newcommand{\rearrS}[2]{rearrS \ #1 \ #2}
\newcommand{\rearrV}[2]{rearrV \ #1 \ #2}
\newcommand{\pg}[3]{pg \ #1 \ (#2, #3)}
\newcommand{\casebx}[6]{case [#1][#2] $\to$ #3, [#4][#5] $\to$ #6}
\newcommand{\xpgone}[7]{kpg [\![#1]\!]}
\newcommand{\xpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
\newcommand{\xpgtwo}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
\newcommand{\xpgtwod}[5]{cpg [\![#1]\!]}

\begin{document}
%
\title{Bx to Rev (tentative) \thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{Introduction}

% Story ->
% reversible is needed not to lose any information -> pg is introduced.
% pg is not good for lassoc -> cpg (xpg2) is introduced.
% cpg can be improved -> kpg(xpg) is introduced.
% kpg is slower than pg for rassoc. combined version -> xpg.

% Evaluation result: Comparison with OCaml implementation of miniBiGUL

\begin{itemize}
\item Importance of BX, BX is a solution of view update problem in database.
\item Explanation of put-based BX: BiGUL.
\item Current status of BiGUL: Fastest BX language in the world
\item A problem: Efficiency of compose evaluation. If we do not save the intermediate states, the number of get is quadratic. This is not good.
\item To solve this problem we use an idea from reversible computation: not to lose any information.
\item Based on this idea we introduce pg.
\item We extend pg with several ideas and produce faster implementation.
\end{itemize}

\subsection{Contribution}

\begin{itemize}
\item Improvement of evaluation efficiency
\item Reveal a part of relationship between BX and Reversible.
  \begin{itemize}
  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
  \item A pair of put and get is self-inverse (involution).
  \item If we can obtain good parameter (dummy) for a BiGUL program,
    put(s, dummy) can be a complement function.
  \end {itemize}
\item Introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
\end{itemize}

\subsection{Construction of this paper}

\begin{itemize}
\item 
\end{itemize}

\section{Bidirectional Programming Language: BiGUL}

\begin{itemize}
\item our target language is miniBiGUL (BiGUL minus adaptive case) : very well-behaved
\item get = s $\to$ v
\item put = s * v $\to$ s
\item explanation about get and put.
\item We assume putget and getput hold (explanation about putget and getput)
\item explanation about put-based BX language: BiGUL
\item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
\end{itemize}

\subsection{Semantics of miniBiGUL}

%\begin{figure}
\bxput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{s}{fail}\\
\bxput{Replace}{s}{v} = v\\
\bxput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} = \product{\bxput{$bx_1$}{$s_1$}{$v_1$}}{\bxput{$bx_2$}{$s_2$}{$v_2$}}\\
\bxput{(\rearrS{$f$}{bx})}{s}{v} = $f^{-1}$(\bxput{bx}{($f$ s)}{v})\\
\bxput{(\rearrV{$g$}{bx})}{s}{v} = \bxput{bx}{s}{($g$ v)}\\
\bxput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}
      {s}{v} = \\
      \hspace{10mm} if {cond$_{sv1}$ s v}\\
      \hspace{10mm} then let s' = \bxput{bx$_1$}{s}{v}\\
      \hspace{10mm} else let s' = \bxput{bx$_2$}{s}{v}\\
      \hspace{10mm} fi cond$_{s1}$ s'; return s'\\
\bxput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} = \bxput{bx$_1$}{s}{(\bxput{bx$_2$}{(\bxget{bx$_1$}{s})}{v})}\\
%\end{figure}

\noindent 
%\begin{figure}
\bxget{(Skip $h$)}{s} = {$h$ s}\\
\bxget{Replace}{s} = s\\
\bxget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} =
\product{\bxget{$bx_1$}{$s_1$}}{\bxget{$bx_2$}{$s_2$}}\\
\bxget{(\rearrS{$f$}{bx})}{s} = \bxget{bx}{($f$ s)}\\
\bxget{(\rearrV{$g$}{bx})}{s} = $g^{-1}$ (\bxget{bx}{s})\\
\bxget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
      \hspace{10mm} if {cond$_{s1}$ s}\\
      \hspace{10mm} then let v' = \bxget{bx$_1$}{s}\\
      \hspace{10mm} else let v' = \bxget{bx$_2$}{s}\\
      \hspace{10mm} fi {cond$_{sv1}$ s v'}; return v'\\
\bxget{(bx$_1$ $\circ$ bx$_2$)}{s} = \bxget{bx$_2$}{(\bxget{bx$_1$}{s})}\\
%\end{figure}

\begin{itemize}
\item To implement this in OCaml, we need some way to produce inverse function for $f$ and $g$.
\item Avoiding this, our implementation requires users to write inverse function f$^{-1}$ and g$^{-1}$.
\end{itemize}

\section{Skip:Reversible langauge: Rwhile}

% Maybe we should skip

\begin{itemize}
\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
\item Using reversible language, we can obtain inverse programs easily.
\item Explanation of Rwhile: we need fi-condition.
\item syntax, how to work.
\end{itemize}

\section{Skip:First approach}

% We should skip this

\begin{itemize}
\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
\item put$_{REV}$ = s * v $\to$ s * g
\item get$_{REV}$ = s $\to$ v * g
\item introduce (put$_{REV}^{-1}$: s * g $\to$ s * v) and (get$_{REV}^{-1}$:v * g $\to$ s)
\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
\end{itemize}

\noindent
\revput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{(s, v)}{fail}

\noindent
\revput{Replace}{s}{v} = (v, s)

\noindent
\revput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} =\\
\hspace{5mm}let (s1, g1) = \revput{$bx_1$}{$s_1$}{$v_1$} in\\
\hspace{5mm}let (s2, g2) = \revput{$bx_2$}{$s_2$}{$v_2$} in\\
\hspace{5mm}((s1, s2), (g1, g2))

\noindent
\revput{(\rearrS{$f$}{bx})}{s}{v} =\\
\hspace{5mm}let (s1, g1) = \revput{bx}{($f$ s)}{v} in\\
\hspace{5mm}($f^{-1}$ s1, g1)

\noindent
\revput{(\rearrV{$g$}{bx})}{s}{v} =\\
\hspace{5mm} let (s1, g1) = \revput{bx}{s}{($g$ v)} in\\
\hspace{5mm} (s1, g1)

\noindent
\revput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v} =\\
\hspace{5mm} if {cond$_{sv1}$ s v}\\
\hspace{5mm} then let (s', g) = \revput{bx$_1$}{s}{v} in\\
\hspace{5mm} else let (s', g) = \revput{bx$_2$}{s}{v} in\\
\hspace{5mm} fi cond$_{s1}$ s'; return (s', g)

\noindent
\revput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
\hspace{5mm} let (s1, g1) = \revget{bx$_1$}{s} in\\
\hspace{5mm} let (s2, g2) = \revput{bx$_2$}{s1}{v} in\\
\hspace{5mm} let (s3, g3) = \revput{bx$_1$}{s}{s2} in (* need to copy s *)\\
\hspace{5mm} (s3, (g1, g2, g3))

\vspace{5mm}
\noindent
\revget{(Skip $h$)}{s} = {($h$ s, s)}

\noindent
\revget{Replace}{s} = (s, \textcolor{red}{Nil})

\noindent
\revget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} = \\
\hspace{5mm} let (v1, g1) = \revget{$bx_1$}{$s_1$} in\\
\hspace{5mm} let (v2, g2) = \revget{$bx_2$}{$s_2$} in\\
\hspace{5mm} ((v1, v2), (g1, g2))

\noindent
%% from here
\revget{(\rearrS{$f$}{bx})}{s}{v} =\\
\hspace{5mm}let (v1, g1) = \revget{bx}{($f$ s)} in\\
\hspace{5mm}(v1, g1)

\noindent
\revget{(\rearrV{$g$}{bx})}{s}{v} =\\
\hspace{5mm} let (v1, g1) = \revget{bx}{s} in\\
\hspace{5mm} ($g^{-1}$ v1, g1)

\noindent
\revget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
\hspace{5mm} if {cond$_{s1}$ s}\\
\hspace{5mm} then let (v', g) = \revget{bx$_1$}{s} in\\
\hspace{5mm} else let (v', g) = \revget{bx$_2$}{s} in\\
\hspace{5mm} fi cond$_{sv1}$ s v'; return (v', g)

\noindent
\revget{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
\hspace{5mm} let (v1, g1) = \revget{bx$_1$}{s} in\\
\hspace{5mm} let (v2, g2) = \revget{bx$_1$}{s} in\\
\hspace{5mm} (v2, (g1, g2))

\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

\section{``pg'' approach}

\subsection{Self-inverse function: pg}

\begin{itemize}
\item Information loss causes re-evaluation of get.
\item To prevent this, we introduced the idea of reversible computation.
\item In reversible computation we need to keep all information: we can not lose any information.
\item In BiGul we can do this with a pair of put and get.
\item pg bx (s, v) = (put bx (s, v), get bx (s))
\item pg: s * v $\to$ s * v
\item No garbage
\item This function is self-inverse. (involution, (f (f input)) = input)
\item \textcolor{red}{Constraint: In case expression, put and get must always choose the same branch. This constraint is not satisfied in general: (e.g, SNOC, the different length of two inputs.) To remove this constraint we need to use garbage.}
\end{itemize}

Proof.\\
pg bx (pg bx (s, v)) = (put bx (put bx (s, v), get bx (s)), get bx (put bx (s, v)))\\
= (put bx (put bx (s, v), get bx (s)), v)\\
= (put bx (s, get bx (s)), v)\\
= (s, v)\\

\subsection{Skip:Construction of pg}

\pg{Skip h}{s}{v} \\
= (\sif{$h$ s = v}{s}{fail}, {$h$ s})\\
= \sif{$h$ s = v}{(s, $h$ s)}{fail}\\
= \sif{$h$ s = v}{(s, v)}{fail}\\

\noindent
\pg{Replace}{s}{v}\\
= (v, s)\\

\noindent
\pg{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}}\\
= (\product{\bxput{$bx_1$}{$s_1$}{$v_1$}}{\bxput{$bx_2$}{$s_2$}{$v_2$}}, \product{\bxget{$bx_1$}{$s_1$}}{\bxget{$bx_2$}{$s_2$}})\\
= let (s1, v1) = \pg{$bx_1$}{$s_1$}{$v_1$} in\\
  let (s2, v2) = \pg{$bx_2$}{$s_2$}{$v_2$} in\\
  (\product{s1}{s2}, \product{v1}{v2})\\

  \noindent
\pg{(\rearrS{$f$}{bx})}{s}{v}\\
= ($f^{-1}$(\bxput{bx}{($f$ s)}{v}), \bxget{bx}{($f$ s)})\\
= let (s, v) = \pg{bx}{($f$ s)}{v} in\\
($f^{-1}$ s, v)\\

\noindent
\pg{(\rearrV{$g$}{bx})}{s}{v}\\
= (\bxput{bx}{s}{($g$ v)}, $g^{-1}$ (\bxget{bx}{s}))\\
= let (s, v) = \pg{bx}{s}{($g$ v)} in\\
  (s, $g^{-1}$ v)\\

\noindent
\pg{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v}\\
= (if {cond$_{sv1}$ s v}\\
\hspace{4mm} then let s' = \bxput{bx$_1$}{s}{v}\\
\hspace{4mm} else let s' = \bxput{bx$_2$}{s}{v}\\
\hspace{4mm} fi cond$_{v1}$ s'; return s',\\
\hspace{4mm} if {cond$_{s1}$ s}\\
\hspace{4mm} then let v' = \bxget{bx$_1$}{s}\\
\hspace{4mm} else let v' = \bxget{bx$_2$}{s}\\
\hspace{4mm} fi cond$_{sv1}$ s v'; return v')\\
= if {cond$_{sv1}$ s v} \&\& {cond$_{s1}$ s}\\
\hspace{3mm} then let s' = \pg{bx$_1$}{s}{v}\\
\hspace{3mm} else let s' = \pg{bx$_2$}{s}{v}\\
\hspace{3mm} fi cond$_{v1}$ s' \&\& cond$_{sv1}$ s v'; return s'\\
\textcolor{red}{A restriction is introduced here. In general we need to consider additional two cases: {cond$_{sv1}$ s v} \&\& not {cond$_{s1}$ s}, not {cond$_{sv1}$ s v} \&\& {cond$_{s1}$ s}. In these two cases, the evaluated bxs are different in put and get. We need to keep garbage. }

\noindent
\pg{(bx$_1$ $\circ$ bx$_2$)}{s}{v}\\
= (\bxput{bx$_1$}{s}{(\bxput{bx$_2$}{(\bxget{bx$_1$}{s})}{v})}, \bxget{bx$_2$}{(\bxget{bx$_1$}{s})})\\
= let a = \bxget{bx$_1$}{s} in\\
\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
\hspace{3mm} let (d, e) = \pg{bx$_1$}{s}{b} in\\
\hspace{3mm} (d, c)\\
= let (a, f) = \pg{bx$_1$}{s}{dummy} in\\
\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
\hspace{3mm} let (d, e) = \pg{bx$_1$}{s}{b} in\\
\hspace{3mm} (d, c)\\
= let (a, f) = \pg{bx$_1$}{s}{dummy} in\\
\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
\hspace{3mm} let (d, e) = \pg{bx$_1$}{\textcolor{red}{f}}{b} in\\
\hspace{3mm} (d, c)\\
\textcolor{red}{Last transformation uses putput.}

\subsection{How to Evaluate Bidirectional Programs using pg}

\begin{itemize}
\item Because we combine put and get, we need to supply view for evaluation of get.
\item introduce dummy?
\item automatic analysis for domain \& range?
\item 
\item partial evaluation of put(s(D), dummy(S)).
\item \textcolor{red}{Not good? Basically put (BX) is often a recursion on the source. If source is static, partial evaluation will work well. On the other hand, view is small data. So partial evaluation might not work well.}
\item \textcolor{red}{lazy evaluation is also a possibility?}
\end{itemize}

\subsection{cpg}

\begin{itemize}
\item If the BX programs are left assoc of composition, pg will be exponential. Because it calls pg bx$_1$ twice.
\item To solve this problem, we introduce accumulation of changes.
\end{itemize}

\noindent
$
\begin{array}{lll}
\xpgtwod{Skip \ h}{ks}{kv}{s'}{v'} & = &(ks, kv, s', v')\\
\\

\xpgtwod{Replace}{ks}{kv}{s'}{v'} & = & (kv, ks, v', s')\\
\\

\xpgtwod{bx1 \ prod \ bx2}{ks}{kv}{s'}{v'} & = &
(ks1, kv1, s1', v1') \Leftarrow \xpgtwo{bx1}{fst \circ ks}{fst \circ kv}{fst \ s'}{fst \ v'};\\
& & (ks2, kv2, s2', v2') \Leftarrow \xpgtwo{bx2}{snd \circ ks}{snd \circ kv}{snd \ s'}{snd \ v'};\\
& & (con \ ks1 \ ks2, con \ kv1 \ kv2, \\
& & (Con \ s1' \ s2'), (Con \ v1' \ v2'))\\
\\

\xpgtwod{RearrS \ f1 \ f2 \ bx}{ks}{kv}{s'}{v'} & = &
(ks, kv, s', v') \Leftarrow \xpgtwo{bx}{f1 \circ ks}{kv}{f1 \ s'}{v'};\\
& & (f2 \circ ks, kv, s', v')\\
\\

\xpgtwod{RearrV \ g1 \ g2 \ bx}{ks}{kv}{s'}{v'} & = &
(ks, kv, s', v') \Leftarrow \xpgtwo{bx}{ks}{g1 \circ kv}{s'}{g1 \ v'};\\
& & (ks, g2 \circ kv, ks', g2 \ v')\\
\\

\xpgtwod{Case \ c \ bx1 \ bx2}{ks}{kv}{s'}{v'}
& & if \ c \ s' \ v' \ then \ \xpgtwo{bx1}{ks}{kv}{s'}{v'}\\
& & else \ \xpgtwo{bx2}{ks}{kv}{s'}{v'}\\
\\

\xpgtwod{bx1 \circ bx2}{ks}{kv}{s'}{v'}
& = & (ks1, kv1, \underline{s1'}, v1') \Leftarrow \xpgtwo{bx1}{ks}{id}{s'}{const$\_$dummy \ v'};\\
& & (ks2, kv2, s2', v2') \Leftarrow \xpgtwo{bx2}{kv1}{kv}{v1'}{v'};\\
& & (ks1 \circ ks2, kv2,  ks1 \ s2', v2')\\
\end{array}
$

\begin{itemize}
\item This is faster than pg for left assoc BX programs.
\item (Slower for right assoc programs)
\item []
\item $s1'$ in composition is redundant.  
\item Next, we will extend this to avoid this redundant computation.
\end{itemize}

\subsection{kpg}

$fst = \lambda (Con (x1, x2)). x1$\\
\noindent 
$snd = \lambda (Con (x1, x2)). x2$\\
$con = \lambda ks1. \lambda ks2. \lambda x. Con(ks1 \ x, ks2 \ x)$

\noindent 
$\xpg{bx}{ks}{kv}{ks'}{kv'}{s'}{v'}$\\

$ks, kv$ (from cpg. For keeping modification information = used for compose)

$ks', kv'$ (from kpg. For evaluating actual values)

$s', v'$ (from kpg. For keeping evaluated values)

\vspace{5mm}

\noindent
$
\begin{array}{lll}
\xpgone{Skip \ h}{ks}{kv}{ks'}{kv'}{s'}{v'} & = & s' \Leftarrow ks' \ s';\\
& & v' \Leftarrow kv' \ v';\\
& & (ks, kv, id, id, s', v')\\
\\

\xpgone{Replace}{ks}{kv}{ks'}{kv'}{s'}{v'} & = & (kv, ks, kv', ks', v', s')\\
\\

\xpgone{bx1 \ prod \ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
\textcolor{red}{let \ s' = ks' \ s' \ in \ let \ v' = kv' \ v' \ in}\\
& & \textcolor{red}{let \ ks' = id \ in \ let \ kv' = id \ in}\\
& & (ks1, kv1, ks1', kv1', s1', v1') \Leftarrow \xpg{bx1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s'}{v'};\\
& & (ks2, kv2, ks2', kv2', s2', v2') \Leftarrow \xpg{bx2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s'}{v'};\\
& & (con \ ks1 \ ks2, con \ kv1 \ kv2, \\
& & con \ (ks1' \circ fst) \ (ks2' \circ snd), con \ (kv1' \circ fst) \ (kv2' \circ snd),\\
& & (Con \ s1' \ s2'), (Con \ v1' \ v2'))\\
\\

\xpgone{RearrS \ f1 \ f2 \ bx}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
(ks, kv, ks', kv', s', v') \Leftarrow \xpg{bx}{f1 \circ ks}{kv}{f1 \circ ks'}{kv'}{s'}{v'};\\
& & (f2 \circ ks, kv, f2 \circ ks', kv', s', v')\\
\\

\xpgone{RearrV \ g1 \ g2 \ bx}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
(ks, kv, ks', kv', s', v') \Leftarrow \xpg{bx}{ks}{g1 \circ kv}{ks'}{g1 \circ kv'}{s'}{v'};\\
& & (ks, g2 \circ kv, ks', g2 \circ kv', s', v')\\
\\

\xpgone{Case \ c \ bx1 \ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'}
& = & s' \Leftarrow ks' s';\\
& & v' \Leftarrow kv' v';\\
& & if \ c \ s' \ v' \ then \ \xpg{bx1}{ks}{kv}{id}{id}{s'}{v'}\\
& & else \ \xpg{bx2}{ks}{kv}{id}{id}{s'}{v'}\\
\\

\xpgone{bx1 \circ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'}
& = & (ks1, kv1, \underline{ks1'}, kv1', \underline{s1'}, v1') \Leftarrow \xpg{bx1}{ks}{id}{ks'}{id}{s'}{const$\_$dummy \ v'};\\
& & (ks2, kv2, ks2', kv2', s2', v2') \Leftarrow \xpg{bx2}{kv1}{kv}{kv1'}{kv'}{v1'}{v'};\\
& & (ks1 \circ ks2, kv2, ks1 \circ ks2', kv2', s2', v2')\\
\end{array}
$

\begin{itemize}
\item In this definition of kpg, s' and v' will have actual value only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
\item In composition $ks1'$ and $s1'$ are still redundant. But, they are not evaluated.
\item In prod, Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s' and v'.
\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
\end{itemize}

\section{xpg}

\begin{itemize}
\item pg is fastest for right assoc programs and kpg is fastest for left assoc programs.
\item To combine their strength, we can use the following definition.
\item If there is a composition, we use xpg for evaluation of $bx_1$ and pg for evaluation of $bx_2$.
\item If $bx_1$ is simple program (= Right assoc), this is almost the same with pg.
\item If $bx_1$ is complicated program (= Left assoc), this is almost the same with kpg.
\end{itemize}

$
\begin{array}{lll}
  xpg [\![E]\!]_{(s, v)} & = & same \ with \ pg \ if \ E \ is \ not \ compose\\
  xpg [\![bx_1 \circ bx_2]\!]_{(s, v)} & = &
  let \ (ks, kv, ks', kv', s', v') = kpg \ bx_1 \ (\lambda \_. s) \ id \ id \ id \ in\\
  & & let \ (s', v') = pg \ bx_2 \ (kv' \ v') \ v \ in\\
  & & (ks \ s', v')\\
\end{array}
$

\section{Experimental result}

\begin{itemize}
\item Evaluation time for miniBiGul, pg, cpg, kpg, xpg.
\item Memory allocation of miniBiGul, pg, cpg, kpg, xpg.
\item []
\item xpg looks the best one (maybe)
\item 
\end{itemize}

\subsection{Theoretical Computational Complexity}

\section{Conclusion and Future Work}

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%

### 2019-08-13 paper/main.tex
Update main.tex

@@
 \usepackage{color}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
 % Used for displaying a sample figure. If possible, figure files should
@@
 
-\newcommand{\bxput}[3]{put$_{BX}$ \ {#1} \ {#2} \ {#3}}
-\newcommand{\bxget}[2]{get$_{BX}$ \ {#1} \ {#2}}
-\newcommand{\revput}[3]{put$_{REV}$ \ {#1} \ {#2} \ {#3}}
-\newcommand{\revget}[2]{get$_{REV}$ \ {#1} \ {#2}}
-\newcommand{\sif}[3]{if \ {#1} \ then \ {#2} \ else \ {#3}}
-\newcommand{\product}[2]{({#1} `cons' {#2})}
+\setlength{\parindent}{0pt}
+\renewcommand{\indent}{\hspace*{0pt}}
+
+\newcommand{\tab}{\hspace*{5mm}}
+\newcommand{\qtab}{\hspace*{5mm} \ \quad}
+
+\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
+\newcommand{\product}[2]{#1 \ prod \ #2}
 \newcommand{\tuple}[2]{(#1 :: #2)}
@@
 \newcommand{\rearrV}[2]{rearrV \ #1 \ #2}
-\newcommand{\pg}[3]{pg \ #1 \ (#2, #3)}
-\newcommand{\casebx}[6]{case [#1][#2] $\to$ #3, [#4][#5] $\to$ #6}
-\newcommand{\xpgone}[7]{kpg [\![#1]\!]}
-\newcommand{\xpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\xpgtwo}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\xpgtwod}[5]{cpg [\![#1]\!]}
+\newcommand{\casebx}[6]{case [#1][#2] \to #3, [#4][#5] \to #6}
+
+\newcommand{\bxput}[3]{put \, [\![#1]\!] \ #2 \ #3}
+\newcommand{\bxputinline}[1]{put \, [\![#1]\!]}
+\newcommand{\bxget}[2]{get \, [\![#1]\!] \ #2}
+\newcommand{\bxgetinline}[1]{get \, [\![#1]\!]}
+
+\newcommand{\revput}[3]{put$_{REV}$ \ {#1} \ {#2} \ {#3}}
+\newcommand{\revget}[2]{get$_{REV}$ \ {#1} \ {#2}}
+
+\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
+\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
+
+\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
+\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
+
+\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
+\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
+
+\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
+\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
 
@@
 
-%\begin{figure}
-\bxput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{s}{fail}\\
-\bxput{Replace}{s}{v} = v\\
-\bxput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} = \product{\bxput{$bx_1$}{$s_1$}{$v_1$}}{\bxput{$bx_2$}{$s_2$}{$v_2$}}\\
-\bxput{(\rearrS{$f$}{bx})}{s}{v} = $f^{-1}$(\bxput{bx}{($f$ s)}{v})\\
-\bxput{(\rearrV{$g$}{bx})}{s}{v} = \bxput{bx}{s}{($g$ v)}\\
-\bxput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}
-      {s}{v} = \\
-      \hspace{10mm} if {cond$_{sv1}$ s v}\\
-      \hspace{10mm} then let s' = \bxput{bx$_1$}{s}{v}\\
-      \hspace{10mm} else let s' = \bxput{bx$_2$}{s}{v}\\
-      \hspace{10mm} fi cond$_{s1}$ s'; return s'\\
-\bxput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} = \bxput{bx$_1$}{s}{(\bxput{bx$_2$}{(\bxget{bx$_1$}{s})}{v})}\\
-%\end{figure}
+$\bxput{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
 
-\noindent 
-%\begin{figure}
-\bxget{(Skip $h$)}{s} = {$h$ s}\\
-\bxget{Replace}{s} = s\\
-\bxget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} =
-\product{\bxget{$bx_1$}{$s_1$}}{\bxget{$bx_2$}{$s_2$}}\\
-\bxget{(\rearrS{$f$}{bx})}{s} = \bxget{bx}{($f$ s)}\\
-\bxget{(\rearrV{$g$}{bx})}{s} = $g^{-1}$ (\bxget{bx}{s})\\
-\bxget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
-      \hspace{10mm} if {cond$_{s1}$ s}\\
-      \hspace{10mm} then let v' = \bxget{bx$_1$}{s}\\
-      \hspace{10mm} else let v' = \bxget{bx$_2$}{s}\\
-      \hspace{10mm} fi {cond$_{sv1}$ s v'}; return v'\\
-\bxget{(bx$_1$ $\circ$ bx$_2$)}{s} = \bxget{bx$_2$}{(\bxget{bx$_1$}{s})}\\
-%\end{figure}
+$\bxput{Replace}{s}{v} = v$
+
+$\bxput{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
+    \tab \product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}$
+
+$\bxput{\rearrS{f}{bx}}{s}{v} = f^{-1}(\bxput{bx}{(f \ s)}{v})$
+
+$\bxput{\rearrV{g}{bx}}{s}{v} = \bxput{bx}{s}{(g \ v)}$
+
+$\bxput{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} = \\
+    \tab \text{if } {cond_{sv1} \ s \ v} \\
+    \tab \text{then let } s' = \bxput{bx_1}{s}{v} \\
+    \tab \text{else let } s' = \bxput{bx_2}{s}{v} \\
+    \tab \text{fi } cond_{s1} \ s'; \text{ return } s'$
+
+$\bxput{(bx_1 \circ bx_2)}{s}{v} = \bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}$
+
+\vspace{5mm}
+
+$\bxget{Skip \ h}{s} = h \ s$
+
+$\bxget{Replace}{s} = s$
+
+$\bxget{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
+    \tab \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}$
+
+$\bxget{\rearrS{f}{bx}}{s} = \bxget{bx}{(f \ s)}$
+
+$\bxget{\rearrV{g}{bx}}{s} = g^{-1} (\bxget{bx}{s})$
+
+$\bxget{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s} =\\
+    \tab \text{if } {cond_{s1} \ s}\\
+    \tab \text{then let } v' = \bxget{bx_1}{s}\\
+    \tab \text{else let } v' = \bxget{bx_2}{s}\\
+    \tab \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v'$
+
+$\bxget{bx_1 \circ bx_2}{s} = \bxget{bx_2}{(\bxget{bx_1}{s})}$
 
@@
 
-\noindent
-\revput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{(s, v)}{fail}
+% \noindent
+% \revput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{(s, v)}{fail}
 
-\noindent
-\revput{Replace}{s}{v} = (v, s)
+% \noindent
+% \revput{Replace}{s}{v} = (v, s)
 
-\noindent
-\revput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} =\\
-\hspace{5mm}let (s1, g1) = \revput{$bx_1$}{$s_1$}{$v_1$} in\\
-\hspace{5mm}let (s2, g2) = \revput{$bx_2$}{$s_2$}{$v_2$} in\\
-\hspace{5mm}((s1, s2), (g1, g2))
+% \noindent
+% \revput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} =\\
+% \hspace{5mm}let (s1, g1) = \revput{$bx_1$}{$s_1$}{$v_1$} in\\
+% \hspace{5mm}let (s2, g2) = \revput{$bx_2$}{$s_2$}{$v_2$} in\\
+% \hspace{5mm}((s1, s2), (g1, g2))
 
-\noindent
-\revput{(\rearrS{$f$}{bx})}{s}{v} =\\
-\hspace{5mm}let (s1, g1) = \revput{bx}{($f$ s)}{v} in\\
-\hspace{5mm}($f^{-1}$ s1, g1)
+% \noindent
+% \revput{(\rearrS{$f$}{bx})}{s}{v} =\\
+% \hspace{5mm}let (s1, g1) = \revput{bx}{($f$ s)}{v} in\\
+% \hspace{5mm}($f^{-1}$ s1, g1)
 
-\noindent
-\revput{(\rearrV{$g$}{bx})}{s}{v} =\\
-\hspace{5mm} let (s1, g1) = \revput{bx}{s}{($g$ v)} in\\
-\hspace{5mm} (s1, g1)
+% \noindent
+% \revput{(\rearrV{$g$}{bx})}{s}{v} =\\
+% \hspace{5mm} let (s1, g1) = \revput{bx}{s}{($g$ v)} in\\
+% \hspace{5mm} (s1, g1)
 
-\noindent
-\revput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v} =\\
-\hspace{5mm} if {cond$_{sv1}$ s v}\\
-\hspace{5mm} then let (s', g) = \revput{bx$_1$}{s}{v} in\\
-\hspace{5mm} else let (s', g) = \revput{bx$_2$}{s}{v} in\\
-\hspace{5mm} fi cond$_{s1}$ s'; return (s', g)
+% \noindent
+% \revput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v} =\\
+% \hspace{5mm} if {cond$_{sv1}$ s v}\\
+% \hspace{5mm} then let (s', g) = \revput{bx$_1$}{s}{v} in\\
+% \hspace{5mm} else let (s', g) = \revput{bx$_2$}{s}{v} in\\
+% \hspace{5mm} fi cond$_{s1}$ s'; return (s', g)
 
-\noindent
-\revput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
-\hspace{5mm} let (s1, g1) = \revget{bx$_1$}{s} in\\
-\hspace{5mm} let (s2, g2) = \revput{bx$_2$}{s1}{v} in\\
-\hspace{5mm} let (s3, g3) = \revput{bx$_1$}{s}{s2} in (* need to copy s *)\\
-\hspace{5mm} (s3, (g1, g2, g3))
+% \noindent
+% \revput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
+% \hspace{5mm} let (s1, g1) = \revget{bx$_1$}{s} in\\
+% \hspace{5mm} let (s2, g2) = \revput{bx$_2$}{s1}{v} in\\
+% \hspace{5mm} let (s3, g3) = \revput{bx$_1$}{s}{s2} in (* need to copy s *)\\
+% \hspace{5mm} (s3, (g1, g2, g3))
 
-\vspace{5mm}
-\noindent
-\revget{(Skip $h$)}{s} = {($h$ s, s)}
+% \vspace{5mm}
+% \noindent
+% \revget{(Skip $h$)}{s} = {($h$ s, s)}
 
-\noindent
-\revget{Replace}{s} = (s, \textcolor{red}{Nil})
+% \noindent
+% \revget{Replace}{s} = (s, \textcolor{red}{Nil})
 
-\noindent
-\revget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} = \\
-\hspace{5mm} let (v1, g1) = \revget{$bx_1$}{$s_1$} in\\
-\hspace{5mm} let (v2, g2) = \revget{$bx_2$}{$s_2$} in\\
-\hspace{5mm} ((v1, v2), (g1, g2))
+% \noindent
+% \revget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} = \\
+% \hspace{5mm} let (v1, g1) = \revget{$bx_1$}{$s_1$} in\\
+% \hspace{5mm} let (v2, g2) = \revget{$bx_2$}{$s_2$} in\\
+% \hspace{5mm} ((v1, v2), (g1, g2))
 
-\noindent
-%% from here
-\revget{(\rearrS{$f$}{bx})}{s}{v} =\\
-\hspace{5mm}let (v1, g1) = \revget{bx}{($f$ s)} in\\
-\hspace{5mm}(v1, g1)
+% \noindent
+% %% from here
+% \revget{(\rearrS{$f$}{bx})}{s}{v} =\\
+% \hspace{5mm}let (v1, g1) = \revget{bx}{($f$ s)} in\\
+% \hspace{5mm}(v1, g1)
 
-\noindent
-\revget{(\rearrV{$g$}{bx})}{s}{v} =\\
-\hspace{5mm} let (v1, g1) = \revget{bx}{s} in\\
-\hspace{5mm} ($g^{-1}$ v1, g1)
+% \noindent
+% \revget{(\rearrV{$g$}{bx})}{s}{v} =\\
+% \hspace{5mm} let (v1, g1) = \revget{bx}{s} in\\
+% \hspace{5mm} ($g^{-1}$ v1, g1)
 
-\noindent
-\revget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
-\hspace{5mm} if {cond$_{s1}$ s}\\
-\hspace{5mm} then let (v', g) = \revget{bx$_1$}{s} in\\
-\hspace{5mm} else let (v', g) = \revget{bx$_2$}{s} in\\
-\hspace{5mm} fi cond$_{sv1}$ s v'; return (v', g)
+% \noindent
+% \revget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
+% \hspace{5mm} if {cond$_{s1}$ s}\\
+% \hspace{5mm} then let (v', g) = \revget{bx$_1$}{s} in\\
+% \hspace{5mm} else let (v', g) = \revget{bx$_2$}{s} in\\
+% \hspace{5mm} fi cond$_{sv1}$ s v'; return (v', g)
 
-\noindent
-\revget{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
-\hspace{5mm} let (v1, g1) = \revget{bx$_1$}{s} in\\
-\hspace{5mm} let (v2, g2) = \revget{bx$_1$}{s} in\\
-\hspace{5mm} (v2, (g1, g2))
+% \noindent
+% \revget{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
+% \hspace{5mm} let (v1, g1) = \revget{bx$_1$}{s} in\\
+% \hspace{5mm} let (v2, g2) = \revget{bx$_1$}{s} in\\
+% \hspace{5mm} (v2, (g1, g2))
 
@@
 
-\section{``pg'' approach}
-
+\section{pg}
 \subsection{Self-inverse function: pg}
@@
 
-Proof.\\
-pg bx (pg bx (s, v)) = (put bx (put bx (s, v), get bx (s)), get bx (put bx (s, v)))\\
-= (put bx (put bx (s, v), get bx (s)), v)\\
-= (put bx (s, get bx (s)), v)\\
-= (s, v)\\
+For the approaches so far, when evaluating a composition of BX programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
+This is a kind of information loss.
+One question is whether it is possible to calculate the above programs without losing information.
+And the answer is yes.
+It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
+In miniBiGUL, we can do that by tupling $put$ and $get$.
+The such pair is accepted as input of a function named $pg$ to produce a new one that contains the actual result of a BX program.
+
+\textbf{Definition:} \\
+$\pg{bx}{s}{v} = (\bxput{bx}{s}{v}, \bxget{bx}{s})$ \\
+
+$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
+
+Proof:\\
+$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
+    \tab = \pg{bx}{(\bxput{bx}{s}{v})}{(\bxget{bx}{s})} \quad [pg \text{ definition}] \\
+    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}) \ (\bxget{bx}{s}), \bxget{bx}{(\bxput{bx}{s}{v})})  \quad [pg \text{ definition}] \\
+    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}, \bxget{bx}{s}), v) \quad [\text{PUTGET}] \\
+    \tab = (put \ [\![bx]\!] \ (s, \bxget{bx}{s}), v) \quad [\text{PUTPUT}] \\
+    \tab = (s, v) \quad [\text{GETPUT}]$
 
@@
 
-\pg{Skip h}{s}{v} \\
-= (\sif{$h$ s = v}{s}{fail}, {$h$ s})\\
-= \sif{$h$ s = v}{(s, $h$ s)}{fail}\\
-= \sif{$h$ s = v}{(s, v)}{fail}\\
+$\pg{Skip \ h}{s}{v} \\
+    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
+    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
+    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
 
-\noindent
-\pg{Replace}{s}{v}\\
-= (v, s)\\
+There is a trick in the construction of $\pginline{Skip \ h}$.
+The first equality is simply based on the definitions of $pg$, $\bxputinline{Skip \ h}$ and $\bxgetinline{Skip \ h}$.
+The second one tuples two results of $put$ and $get$ in the body of the if-expression.
+And the last one is quite obivious.
+What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\bxgetinline{Skip \ h}$.\\
 
-\noindent
-\pg{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}}\\
-= (\product{\bxput{$bx_1$}{$s_1$}{$v_1$}}{\bxput{$bx_2$}{$s_2$}{$v_2$}}, \product{\bxget{$bx_1$}{$s_1$}}{\bxget{$bx_2$}{$s_2$}})\\
-= let (s1, v1) = \pg{$bx_1$}{$s_1$}{$v_1$} in\\
-  let (s2, v2) = \pg{$bx_2$}{$s_2$}{$v_2$} in\\
-  (\product{s1}{s2}, \product{v1}{v2})\\
+$\pg{Replace}{s}{v}\\
+    \tab = (v, s)$
 
-  \noindent
-\pg{(\rearrS{$f$}{bx})}{s}{v}\\
-= ($f^{-1}$(\bxput{bx}{($f$ s)}{v}), \bxget{bx}{($f$ s)})\\
-= let (s, v) = \pg{bx}{($f$ s)}{v} in\\
-($f^{-1}$ s, v)\\
+$\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
+    \tab = (\product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}, \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}) \\
+    \tab = \text{let } (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \text{ in} \\
+        \qtab \text{let } (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \text{ in} \\
+        \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$
 
-\noindent
-\pg{(\rearrV{$g$}{bx})}{s}{v}\\
-= (\bxput{bx}{s}{($g$ v)}, $g^{-1}$ (\bxget{bx}{s}))\\
-= let (s, v) = \pg{bx}{s}{($g$ v)} in\\
-  (s, $g^{-1}$ v)\\
+$\pg{\rearrS{f}{bx}}{s}{v} \\
+    \tab = (f^{-1}(\bxput{bx}{(f \ s)}{v}), \bxget{bx}{(f \ s)}) \\
+    \tab = \text{let } (s, v) = \pg{bx}{(f \ s)}{v} \text{ in} \\
+        \qtab (f^{-1} s, v)$
 
-\noindent
-\pg{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v}\\
-= (if {cond$_{sv1}$ s v}\\
-\hspace{4mm} then let s' = \bxput{bx$_1$}{s}{v}\\
-\hspace{4mm} else let s' = \bxput{bx$_2$}{s}{v}\\
-\hspace{4mm} fi cond$_{v1}$ s'; return s',\\
-\hspace{4mm} if {cond$_{s1}$ s}\\
-\hspace{4mm} then let v' = \bxget{bx$_1$}{s}\\
-\hspace{4mm} else let v' = \bxget{bx$_2$}{s}\\
-\hspace{4mm} fi cond$_{sv1}$ s v'; return v')\\
-= if {cond$_{sv1}$ s v} \&\& {cond$_{s1}$ s}\\
-\hspace{3mm} then let s' = \pg{bx$_1$}{s}{v}\\
-\hspace{3mm} else let s' = \pg{bx$_2$}{s}{v}\\
-\hspace{3mm} fi cond$_{v1}$ s' \&\& cond$_{sv1}$ s v'; return s'\\
-\textcolor{red}{A restriction is introduced here.
-In general we need to consider additional two cases: {cond$_{sv1}$ s v} \&\& not {cond$_{s1}$ s}, not {cond$_{sv1}$ s v} \&\& {cond$_{s1}$ s}.
-In these two cases, the evaluated bxs are different in put and get.
-We need to keep garbage.
-}
+$\pg{\rearrV{g}{bx}}{s}{v} \\
+    \tab = (\bxput{bx}{s}{(g v)}, g^{-1} (\bxget{bx}{s})) \\
+    \tab = \text{let } (s, v) = \pg{bx}{s}{(g v)} \text{ in} \\
+        \qtab (s, g^{-1} v)$\\
 
-\noindent
-\pg{(bx$_1$ $\circ$ bx$_2$)}{s}{v}\\
-= (\bxput{bx$_1$}{s}{(\bxput{bx$_2$}{(\bxget{bx$_1$}{s})}{v})}, \bxget{bx$_2$}{(\bxget{bx$_1$}{s})})\\
-= let a = \bxget{bx$_1$}{s} in\\
-\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
-\hspace{3mm} let (d, e) = \pg{bx$_1$}{s}{b} in\\
-\hspace{3mm} (d, c)\\
-= let (a, f) = \pg{bx$_1$}{s}{dummy} in\\
-\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
-\hspace{3mm} let (d, e) = \pg{bx$_1$}{s}{b} in\\
-\hspace{3mm} (d, c)\\
-= let (a, f) = \pg{bx$_1$}{s}{dummy} in\\
-\hspace{3mm} let (b, c) = \pg{bx$_2$}{a}{v} in\\
-\hspace{3mm} let (d, e) = \pg{bx$_1$}{\textcolor{red}{f}}{b} in\\
-\hspace{3mm} (d, c)\\
-\textcolor{red}{Last transformation uses putput.}
+Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrS{f}{bx}]\!]$ and $[\![\rearrV{g}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
+
+$\pg{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} \\
+    \tab = (\text{if } {cond_{sv1} \ s \ v} \\
+    \qtab \text{then let } s' = \bxput{bx_1}{s}{v} \\
+    \qtab \text{else let } s' = \bxput{bx_2}{s}{v} \\
+    \qtab \text{fi } cond_{s1} \ s'; \text{ return } s', \\
+    \qtab \quad \text{if } {cond_{s1} \ s}\\
+    \qtab \quad \text{then let } v' = \bxget{bx_1}{s}\\
+    \qtab \quad \text{else let } v' = \bxget{bx_2}{s}\\
+    \qtab \quad \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v') \\
+    \tab = \text{if } {cond_{sv1} \ s \ v} \ and \ {cond_{s1} \ s} \\
+        \qtab \text{then let } (s', v') = \pg{bx_1}{s}{v}\\
+        \qtab \text{else let } (s', v') = \pg{bx_2}{s}{v}\\
+        \qtab \text{fi } cond_{v1} \ s' \ and \ cond_{sv1} \ s \ v'; \text{ return } (s',v')$\\
+
+A restriction for $\pginline{case}$ needs to be introduced here.
+We know that there is one entering condition and one exit condition when evaluating $\bxputinline{case}$ as well as $\bxgetinline{case}$.
+If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
+This means that two entering conditions of $\bxputinline{case}$ and $\bxgetinline{case}$ are not always simultaneously satified.
+The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv1} \ s \ v) \ and \ (not (cond_{s1} \ s)))$ and $((not (cond_{sv1} \ s \ v) \ and \ (cond_{s1} \ s))$, which are restricted in this paper.
+This does not happen for the others, which the construction of $\pginline{case}$ uses above.
+\\
+
+$\pg{bx_1 \circ bx_2}{s}{v} \\
+    \hspace*{5mm} = (\bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}, \bxget{bx_2}{(\bxget{bx_1}{s})}) \\
+    \hspace*{5mm} = \text{let } a = \bxget{bx_1}{s} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
+        \hspace*{5mm} \ \quad (d, c) \\
+    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
+        \hspace*{5mm} \ \quad (d, c) \\
+    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
+        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \text{ in} \\
+        \hspace*{5mm} \ \quad (d, c)$\\
+
+The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
+The first two equalities comes from mentioned definitions and some basic transformations.
+The third one rewrites $(\text{let } a = \bxget{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$.
+It is possible if we consider $(\bxget{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid.
+The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\bxput{bx_1}{s}{dummy}$.
+This transformation is feasible due to the PUTPUT law.
+Then, $e$ in the result pair $(d,e)$ of the new let-expression equals $dummy$.
+So we can realize that there is no loss information when computing a composition.
 
@@
 
-\noindent
-$
-\begin{array}{lll}
-\xpgtwod{Skip \ h}{ks}{kv}{s'}{v'} & = &(ks, kv, s', v')\\
-\\
+$fst = \lambda (Con (x1, x2)). x1$\\
+$snd = \lambda (Con (x1, x2)). x2$\\
+$con = \lambda ks1. \lambda ks2. \lambda x. Con(ks1 \ x, ks2 \ x)$\\
 
-\xpgtwod{Replace}{ks}{kv}{s'}{v'} & = & (kv, ks, v', s')\\
-\\
+$\cpg{bx}{ks}{kv}{s}{v}$ \\
+    \tab $ks, kv$ evaluate actual values\\
+    \tab $s, v$ keep evaluated values\\
+\vspace{3mm}
 
-\xpgtwod{bx1 \ prod \ bx2}{ks}{kv}{s'}{v'} & = &
-(ks1, kv1, s1', v1') \Leftarrow \xpgtwo{bx1}{fst \circ ks}{fst \circ kv}{fst \ s'}{fst \ v'};\\
-& & (ks2, kv2, s2', v2') \Leftarrow \xpgtwo{bx2}{snd \circ ks}{snd \circ kv}{snd \ s'}{snd \ v'};\\
-& & (con \ ks1 \ ks2, con \ kv1 \ kv2, \\
-& & (Con \ s1' \ s2'), (Con \ v1' \ v2'))\\
-\\
+$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
 
-\xpgtwod{RearrS \ f1 \ f2 \ bx}{ks}{kv}{s'}{v'} & = &
-(ks, kv, s', v') \Leftarrow \xpgtwo{bx}{f1 \circ ks}{kv}{f1 \ s'}{v'};\\
-& & (f2 \circ ks, kv, s', v')\\
-\\
+$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
 
-\xpgtwod{RearrV \ g1 \ g2 \ bx}{ks}{kv}{s'}{v'} & = &
-(ks, kv, s', v') \Leftarrow \xpgtwo{bx}{ks}{g1 \circ kv}{s'}{g1 \ v'};\\
-& & (ks, g2 \circ kv, ks', g2 \ v')\\
-\\
+$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
+    \tab (ks1, kv1, s1, v1) \Leftarrow \cpg{bx1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
+    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
+    \qtab ((con \ ks1 \ ks2, con \ kv1 \ kv2, Con \ s1 \ s2, Con \ v1 \ v2)$\\
 
-\xpgtwod{Case \ c \ bx1 \ bx2}{ks}{kv}{s'}{v'}
-& & if \ c \ s' \ v' \ then \ \xpgtwo{bx1}{ks}{kv}{s'}{v'}\\
-& & else \ \xpgtwo{bx2}{ks}{kv}{s'}{v'}\\
-\\
+$\cpg{RearrS \ f1 \ f2 \ bx}{ks}{kv}{s}{v} =\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f1 \circ ks}{kv}{f1 \ s}{v};\\
+    \qtab (f2 \circ ks, kv, s, v)$\\
 
-\xpgtwod{bx1 \circ bx2}{ks}{kv}{s'}{v'}
-& = & (ks1, kv1, \underline{s1'}, v1') \Leftarrow \xpgtwo{bx1}{ks}{id}{s'}{const$\_$dummy \ v'};\\
-& & (ks2, kv2, s2', v2') \Leftarrow \xpgtwo{bx2}{kv1}{kv}{v1'}{v'};\\
-& & (ks1 \circ ks2, kv2,  ks1 \ s2', v2')\\
-\end{array}
-$
+$\cpg{RearrV \ g1 \ g2 \ bx}{ks}{kv}{s}{v} =\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g1 \circ kv}{s}{g1 \ v};\\
+    \qtab (ks, g2 \circ kv, ks, g2 \ v)$\\
+
+$\cpg{Case \ c \ bx1 \ bx2}{ks}{kv}{s}{v} =\\
+    \tab \text{if} \ c \ s \ v \ \text{then}\\
+        \qtab \cpg{bx1}{ks}{kv}{s}{v}\\
+    \tab \text{else} \\ 
+        \qtab \cpg{bx2}{ks}{kv}{s}{v}$\\
+
+$\cpg{bx1 \circ bx2}{ks}{kv}{s}{v} =\\
+    \tab (ks1, kv1, \underline{s1}, v1) \Leftarrow \cpg{bx1}{ks}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{kv1}{kv}{v1}{v};\\
+        \qtab (ks1 \circ ks2, kv2,  ks1 \ s2, v2)$\\
 
@@
 \item (Slower for right assoc programs)
-\item []
 \item $s1'$ in composition is redundant.  
@@
 
-$fst = \lambda (Con (x1, x2)). x1$\\
-\noindent 
-$snd = \lambda (Con (x1, x2)). x2$\\
-$con = \lambda ks1. \lambda ks2. \lambda x. Con(ks1 \ x, ks2 \ x)$
+$\kpg{bx}{ks'}{kv'}{ks}{kv}{s}{v}$ \\
+    \tab $ks', kv'$ keep modification information\\
+    \tab $ks, kv$ evaluate actual values\\
+    \tab $s, v$ keep evaluated values\\
 
-\noindent 
-$\xpg{bx}{ks}{kv}{ks'}{kv'}{s'}{v'}$\\
+$\kpg{Skip \ h}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \tab s \Leftarrow ks \ s;\\
+    \tab v \Leftarrow kv \ v;\\
+    \tab \text{if} \ h \ s = v \ \text{then} \ (ks', kv', id, id, s, v) \ \text{else fail}$\\
 
-$ks, kv$ (from cpg. For keeping modification information = used for compose)
+$\kpg{Replace}{ks'}{kv'}{ks}{kv}{s}{v} = (kv', ks', kv, ks, v, s)$\\
 
-$ks', kv'$ (from kpg. For evaluating actual values)
+$\kpg{bx1 \ prod \ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \textcolor{red}{s \Leftarrow ks \ s; v \Leftarrow kv \ v; ks \Leftarrow id; kv \Leftarrow id;}\\
+    (ks1', kv1', ks1, kv1, s1, v1) \Leftarrow \kpg{bx1}{fst \circ ks'}{fst \circ kv'}{fst \circ ks}{fst \circ kv}{s}{v};\\
+    (ks2', kv2', ks2, kv2, s2, v2) \Leftarrow \kpg{bx2}{snd \circ ks'}{snd \circ kv'}{snd \circ ks}{snd \circ kv}{s}{v};\\
+    \tab ( con \ ks1' \ ks2', con \ kv1' \ kv2', \\
+    \qtab con \ (ks1 \circ fst) \ (ks2 \circ snd), con \ (kv1 \circ fst) \ (kv2 \circ snd), \\
+    \qtab (Con \ s1 \ s2), (Con \ v1 \ v2))$\\
 
-$s', v'$ (from kpg. For keeping evaluated values)
+$\kpg{RearrS \ f1 \ f2 \ bx}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \tab (ks', kv', ks, kv, s, v) \Leftarrow \kpg{bx}{f1 \circ ks'}{kv'}{f1 \circ ks}{kv}{s}{v};\\
+    \qtab (f2 \circ ks', kv', f2 \circ ks, kv, s, v)$\\
 
-\vspace{5mm}
+$\kpg{RearrV \ g1 \ g2 \ bx}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \tab (ks', kv', ks, kv, s, v) \Leftarrow \kpg{bx}{ks'}{g1 \circ kv'}{ks}{g1 \circ kv}{s}{v};\\
+    \qtab (ks', g2 \circ kv', ks, g2 \circ kv, s, v)$\\
 
-\noindent
-$
-\begin{array}{lll}
-\xpgone{Skip \ h}{ks}{kv}{ks'}{kv'}{s'}{v'} & = & s' \Leftarrow ks' \ s';\\
-& & v' \Leftarrow kv' \ v';\\
-& & (ks, kv, id, id, s', v')\\
-\\
+$\kpg{Case \ c \ bx1 \ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \tab s \Leftarrow ks \ s;\\
+    \tab v \Leftarrow kv \ v;\\
+    \tab \text{if} \ c \ s \ v \ \text{then}\\
+        \qtab \kpg{bx1}{ks'}{kv'}{id}{id}{s}{v}\\
+    \tab \text{else}\\\
+        \qtab \kpg{bx2}{ks'}{kv'}{id}{id}{s}{v}$\\
 
-\xpgone{Replace}{ks}{kv}{ks'}{kv'}{s'}{v'} & = & (kv, ks, kv', ks', v', s')\\
-\\
-
-\xpgone{bx1 \ prod \ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
-\textcolor{red}{let \ s' = ks' \ s' \ in \ let \ v' = kv' \ v' \ in}\\
-& & \textcolor{red}{let \ ks' = id \ in \ let \ kv' = id \ in}\\
-& & (ks1, kv1, ks1', kv1', s1', v1') \Leftarrow \xpg{bx1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s'}{v'};\\
-& & (ks2, kv2, ks2', kv2', s2', v2') \Leftarrow \xpg{bx2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s'}{v'};\\
-& & (con \ ks1 \ ks2, con \ kv1 \ kv2, \\
-& & con \ (ks1' \circ fst) \ (ks2' \circ snd), con \ (kv1' \circ fst) \ (kv2' \circ snd),\\
-& & (Con \ s1' \ s2'), (Con \ v1' \ v2'))\\
-\\
-
-\xpgone{RearrS \ f1 \ f2 \ bx}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
-(ks, kv, ks', kv', s', v') \Leftarrow \xpg{bx}{f1 \circ ks}{kv}{f1 \circ ks'}{kv'}{s'}{v'};\\
-& & (f2 \circ ks, kv, f2 \circ ks', kv', s', v')\\
-\\
-
-\xpgone{RearrV \ g1 \ g2 \ bx}{ks}{kv}{ks'}{kv'}{s'}{v'} & = &
-(ks, kv, ks', kv', s', v') \Leftarrow \xpg{bx}{ks}{g1 \circ kv}{ks'}{g1 \circ kv'}{s'}{v'};\\
-& & (ks, g2 \circ kv, ks', g2 \circ kv', s', v')\\
-\\
-
-\xpgone{Case \ c \ bx1 \ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'}
-& = & s' \Leftarrow ks' s';\\
-& & v' \Leftarrow kv' v';\\
-& & if \ c \ s' \ v' \ then \ \xpg{bx1}{ks}{kv}{id}{id}{s'}{v'}\\
-& & else \ \xpg{bx2}{ks}{kv}{id}{id}{s'}{v'}\\
-\\
-
-\xpgone{bx1 \circ bx2}{ks}{kv}{ks'}{kv'}{s'}{v'}
-& = & (ks1, kv1, \underline{ks1'}, kv1', \underline{s1'}, v1') \Leftarrow \xpg{bx1}{ks}{id}{ks'}{id}{s'}{const$\_$dummy \ v'};\\
-& & (ks2, kv2, ks2', kv2', s2', v2') \Leftarrow \xpg{bx2}{kv1}{kv}{kv1'}{kv'}{v1'}{v'};\\
-& & (ks1 \circ ks2, kv2, ks1 \circ ks2', kv2', s2', v2')\\
-\end{array}
-$
+$\kpg{bx1 \circ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
+    \tab (ks1', kv1', \underline{ks1}, kv1, \underline{s1}, v1) \Leftarrow \kpg{bx1}{ks'}{id}{ks}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks2', kv2', ks2, kv2, s2, v2) \Leftarrow \kpg{bx2}{kv1'}{kv'}{kv1}{kv}{v1}{v};\\
+    \qtab (ks1' \circ ks2', kv2', ks1' \circ ks2, kv2, s2, v2)$
 
@@
 
-$
-\begin{array}{lll}
-  xpg [\![E]\!]_{(s, v)} & = & same \ with \ pg \ if \ E \ is \ not \ compose\\
-  xpg [\![bx_1 \circ bx_2]\!]_{(s, v)} & = &
-  let \ (ks, kv, ks', kv', s', v') = kpg \ bx_1 \ (\lambda \_. s) \ id \ id \ id \ in\\
-  & & let \ (s', v') = pg \ bx_2 \ (kv' \ v') \ v \ in\\
-  & & (ks \ s', v')\\
-\end{array}
-$
+$\xpg{E}{s}{v} = \text{same with } pg \text{ if } E \text{ is not compose}$
+
+$\xpg{bx_1 \circ bx_2}{s}{v} =\\
+    \tab (ks1',kv1', ks1, kv1, s1, v1) \Leftarrow \kpg{bx_1}{id}{id}{\lambda \_.s}{id}{s}{v};\\
+    \tab (s2, v2) \Leftarrow \xpg{bx_2}{kv1 \ v1}{v};\\
+    \qtab (ks1' \ s2, v2)$
 

### 2019-08-13 paper/main.tex
Update main.tex

@@
 
-\newcommand{\bxput}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\bxputinline}[1]{put \, [\![#1]\!]}
-\newcommand{\bxget}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\bxgetinline}[1]{get \, [\![#1]\!]}
+\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
+\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
+\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
+\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
 
-\newcommand{\revput}[3]{put$_{REV}$ \ {#1} \ {#2} \ {#3}}
-\newcommand{\revget}[2]{get$_{REV}$ \ {#1} \ {#2}}
+\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
+\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
 
@@
 % reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg2) is introduced.
+% pg is not good for lassoc -> cpg (xpg_2) is introduced.
 % cpg can be improved -> kpg(xpg) is introduced.
@@
 
-$\bxput{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
+$\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
 
-$\bxput{Replace}{s}{v} = v$
+$\putbx{Replace}{s}{v} = v$
 
-$\bxput{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
-    \tab \product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}$
+$\putbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
+    \tab \product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}$
 
-$\bxput{\rearrS{f}{bx}}{s}{v} = f^{-1}(\bxput{bx}{(f \ s)}{v})$
+$\putbx{\rearrS{f}{bx}}{s}{v} = f^{-1}(\putbx{bx}{(f \ s)}{v})$
 
-$\bxput{\rearrV{g}{bx}}{s}{v} = \bxput{bx}{s}{(g \ v)}$
+$\putbx{\rearrV{g}{bx}}{s}{v} = \putbx{bx}{s}{(g \ v)}$
 
-$\bxput{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} = \\
-    \tab \text{if } {cond_{sv1} \ s \ v} \\
-    \tab \text{then let } s' = \bxput{bx_1}{s}{v} \\
-    \tab \text{else let } s' = \bxput{bx_2}{s}{v} \\
-    \tab \text{fi } cond_{s1} \ s'; \text{ return } s'$
+$\putbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} = \\
+    \tab \text{if } {cond_{sv_1} \ s \ v} \\
+    \tab \text{then let } s' = \putbx{bx_1}{s}{v} \\
+    \tab \text{else let } s' = \putbx{bx_2}{s}{v} \\
+    \tab \text{fi } cond_{s_1} \ s'; \ \text{return} \ s'$
 
-$\bxput{(bx_1 \circ bx_2)}{s}{v} = \bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}$
+$\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
 
@@
 
-$\bxget{Skip \ h}{s} = h \ s$
+$\getbx{Skip \ h}{s} = h \ s$
 
-$\bxget{Replace}{s} = s$
+$\getbx{Replace}{s} = s$
 
-$\bxget{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
-    \tab \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}$
+$\getbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
+    \tab \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}$
 
-$\bxget{\rearrS{f}{bx}}{s} = \bxget{bx}{(f \ s)}$
+$\getbx{\rearrS{f}{bx}}{s} = \getbx{bx}{(f \ s)}$
 
-$\bxget{\rearrV{g}{bx}}{s} = g^{-1} (\bxget{bx}{s})$
+$\getbx{\rearrV{g}{bx}}{s} = g^{-1} (\getbx{bx}{s})$
 
-$\bxget{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s} =\\
-    \tab \text{if } {cond_{s1} \ s}\\
-    \tab \text{then let } v' = \bxget{bx_1}{s}\\
-    \tab \text{else let } v' = \bxget{bx_2}{s}\\
-    \tab \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v'$
+$\getbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
+    \tab \text{if} \ {cond_{s_1} \ s}\\
+    \tab \text{then let} \ v' = \getbx{bx_1}{s}\\
+    \tab \text{else let} \ v' = \getbx{bx_2}{s}\\
+    \tab \text{fi} \ {cond_{sv_1} \ s \ v'}; \ \text{return} \ v'$
 
-$\bxget{bx_1 \circ bx_2}{s} = \bxget{bx_2}{(\bxget{bx_1}{s})}$
+$\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
 
@@
 
-% \noindent
-% \revput{(Skip $h$)}{s}{v} = \sif{$h$ s = v}{(s, v)}{fail}
+$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
 
-% \noindent
-% \revput{Replace}{s}{v} = (v, s)
+$\putrev{Replace}{s}{v} = (v, s)$
 
-% \noindent
-% \revput{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}}{\tuple{$v_1$}{$v_2$}} =\\
-% \hspace{5mm}let (s1, g1) = \revput{$bx_1$}{$s_1$}{$v_1$} in\\
-% \hspace{5mm}let (s2, g2) = \revput{$bx_2$}{$s_2$}{$v_2$} in\\
-% \hspace{5mm}((s1, s2), (g1, g2))
+$\putrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
+    \tab \text{let} \ (s_1, g_1) = \putrev{bx_1}{s_1}{v_1} \ \text{in}\\
+    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_2}{v_2} \ \text{in}\\
+        \qtab ((s_1, s_2), (g_1, g_2))$
 
-% \noindent
-% \revput{(\rearrS{$f$}{bx})}{s}{v} =\\
-% \hspace{5mm}let (s1, g1) = \revput{bx}{($f$ s)}{v} in\\
-% \hspace{5mm}($f^{-1}$ s1, g1)
+$\putrev{\rearrS{f}{bx}}{s}{v} =\\
+    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{(f \ s)}{v} \ \text{in}\\
+        \qtab (f^{-1} s_1, g_1)$
 
-% \noindent
-% \revput{(\rearrV{$g$}{bx})}{s}{v} =\\
-% \hspace{5mm} let (s1, g1) = \revput{bx}{s}{($g$ v)} in\\
-% \hspace{5mm} (s1, g1)
+$\putrev{\rearrV{g}{bx}}{s}{v} =\\
+    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{s}{(g \ v)} \ \text{in}\\
+        \qtab (s_1, g_1)$
 
-% \noindent
-% \revput{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s}{v} =\\
-% \hspace{5mm} if {cond$_{sv1}$ s v}\\
-% \hspace{5mm} then let (s', g) = \revput{bx$_1$}{s}{v} in\\
-% \hspace{5mm} else let (s', g) = \revput{bx$_2$}{s}{v} in\\
-% \hspace{5mm} fi cond$_{s1}$ s'; return (s', g)
+$\putrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} =\\
+    \tab \text{if} \ cond_{sv_1} \ s \ v\\
+    \tab \text{then let} \ (s', g) = \putrev{bx_1}{s}{v} \ \text{in}\\
+    \tab \text{else let} \ (s', g) = \putrev{bx_2}{s}{v} \ \text{in}\\
+    \tab \text{fi} \ cond_{s_1} \ s'; \ \text{return} \ (s', g)$
 
-% \noindent
-% \revput{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
-% \hspace{5mm} let (s1, g1) = \revget{bx$_1$}{s} in\\
-% \hspace{5mm} let (s2, g2) = \revput{bx$_2$}{s1}{v} in\\
-% \hspace{5mm} let (s3, g3) = \revput{bx$_1$}{s}{s2} in (* need to copy s *)\\
-% \hspace{5mm} (s3, (g1, g2, g3))
+$\putrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab \text{let} \ (s_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
+    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_1}{v} \ \text{in}\\
+    \tab \text{let} \ (s_3, g_3) = \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
+        \qtab (s_3, (g_1, g_2, g_3))$
 
-% \vspace{5mm}
-% \noindent
-% \revget{(Skip $h$)}{s} = {($h$ s, s)}
+\vspace{5mm}
 
-% \noindent
-% \revget{Replace}{s} = (s, \textcolor{red}{Nil})
+$\getrev{Skip \ h}{s} = {(h \ s, s)}$
 
-% \noindent
-% \revget{\product{$bx_1$}{$bx_2$}}{\tuple{$s_1$}{$s_2$}} = \\
-% \hspace{5mm} let (v1, g1) = \revget{$bx_1$}{$s_1$} in\\
-% \hspace{5mm} let (v2, g2) = \revget{$bx_2$}{$s_2$} in\\
-% \hspace{5mm} ((v1, v2), (g1, g2))
+$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
 
-% \noindent
-% %% from here
-% \revget{(\rearrS{$f$}{bx})}{s}{v} =\\
-% \hspace{5mm}let (v1, g1) = \revget{bx}{($f$ s)} in\\
-% \hspace{5mm}(v1, g1)
+$\getrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
+    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s_1} \ \text{in}\\
+    \tab \text{let} \ (v_2, g_2) = \getrev{bx_2}{s_2} \ \text{in}\\
+        \qtab ((v_1, v_2), (g_1, g_2))$
 
-% \noindent
-% \revget{(\rearrV{$g$}{bx})}{s}{v} =\\
-% \hspace{5mm} let (v1, g1) = \revget{bx}{s} in\\
-% \hspace{5mm} ($g^{-1}$ v1, g1)
+$\getrev{\rearrS{f}{bx}}{s}{v} =\\
+    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{(f \ s)} \ \text{in}\\
+        \qtab (v_1, g_1)$
 
-% \noindent
-% \revget{(\casebx{cond$_{sv1}$}{cond$_{s1}$}{bx$_1$}{cond$_{sv2}$}{cond$_{s2}$}{bx$_2$})}{s} =\\
-% \hspace{5mm} if {cond$_{s1}$ s}\\
-% \hspace{5mm} then let (v', g) = \revget{bx$_1$}{s} in\\
-% \hspace{5mm} else let (v', g) = \revget{bx$_2$}{s} in\\
-% \hspace{5mm} fi cond$_{sv1}$ s v'; return (v', g)
+$\getrev{\rearrV{g}{bx}}{s}{v} =\\
+    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{s} \ \text{in}\\
+        \qtab (g^{-1} v_1, g_1)$
 
-% \noindent
-% \revget{(bx$_1$ $\circ$ bx$_2$)}{s}{v} =\\
-% \hspace{5mm} let (v1, g1) = \revget{bx$_1$}{s} in\\
-% \hspace{5mm} let (v2, g2) = \revget{bx$_1$}{s} in\\
-% \hspace{5mm} (v2, (g1, g2))
+$\getrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
+    \tab \text{if} \ cond_{s_1} \ s\\
+    \tab \text{then let} \ (v', g) = \getrev{bx_1}{s} \ \text{in}\\
+    \tab \text{else let} \ (v', g) = \getrev{bx_2}{s} \ \text{in}\\
+    \tab \text{fi} \ cond_{sv_1} \ s \ v'; \ \text{return} \ (v', g)$
+
+$\getrev{(bx_1 \circ bx_2)}{s}{v} =\\
+    \tab let (v_1, g_1) = \getrev{bx_1}{s} in\\
+    \tab let (v_2, g_2) = \getrev{bx_1}{s} in\\
+        \qtab (v_2, (g_1, g_2))$
 
@@
 \textbf{Definition:} \\
-$\pg{bx}{s}{v} = (\bxput{bx}{s}{v}, \bxget{bx}{s})$ \\
+$\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$ \\
 
@@
 $pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\bxput{bx}{s}{v})}{(\bxget{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}) \ (\bxget{bx}{s}), \bxget{bx}{(\bxput{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\bxput{bx}{s}{v}, \bxget{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \bxget{bx}{s}), v) \quad [\text{PUTPUT}] \\
+    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
+    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
+    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
+    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
     \tab = (s, v) \quad [\text{GETPUT}]$
@@
 There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\bxputinline{Skip \ h}$ and $\bxgetinline{Skip \ h}$.
+The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
 The second one tuples two results of $put$ and $get$ in the body of the if-expression.
 And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\bxgetinline{Skip \ h}$.\\
+What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
 
@@
 $\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
-    \tab = (\product{(\bxput{bx_1}{s_1}{v_1})}{(\bxput{bx_2}{s_2}{v_2})}, \product{(\bxget{bx_1}{s_1})}{(\bxget{bx_2}{s_2})}) \\
-    \tab = \text{let } (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \text{ in} \\
-        \qtab \text{let } (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \text{ in} \\
+    \tab = (\product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}, \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}) \\
+    \tab = \text{let} \ (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \ \text{in} \\
+        \qtab \text{let} \ (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \ \text{in} \\
         \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$
@@
 $\pg{\rearrS{f}{bx}}{s}{v} \\
-    \tab = (f^{-1}(\bxput{bx}{(f \ s)}{v}), \bxget{bx}{(f \ s)}) \\
-    \tab = \text{let } (s, v) = \pg{bx}{(f \ s)}{v} \text{ in} \\
+    \tab = (f^{-1}(\putbx{bx}{(f \ s)}{v}), \getbx{bx}{(f \ s)}) \\
+    \tab = \text{let} \ (s, v) = \pg{bx}{f \ s}{v} \ \text{in} \\
         \qtab (f^{-1} s, v)$
@@
 $\pg{\rearrV{g}{bx}}{s}{v} \\
-    \tab = (\bxput{bx}{s}{(g v)}, g^{-1} (\bxget{bx}{s})) \\
-    \tab = \text{let } (s, v) = \pg{bx}{s}{(g v)} \text{ in} \\
+    \tab = (\putbx{bx}{s}{(g v)}, g^{-1} (\getbx{bx}{s})) \\
+    \tab = \text{let} \ (s, v) = \pg{bx}{s}{g \ v} \ \text{in} \\
         \qtab (s, g^{-1} v)$\\
@@
 
-$\pg{\casebx{cond_{sv1}}{cond_{s1}}{bx_1}{cond_{sv2}}{cond_{s2}}{bx_2}}{s}{v} \\
-    \tab = (\text{if } {cond_{sv1} \ s \ v} \\
-    \qtab \text{then let } s' = \bxput{bx_1}{s}{v} \\
-    \qtab \text{else let } s' = \bxput{bx_2}{s}{v} \\
-    \qtab \text{fi } cond_{s1} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if } {cond_{s1} \ s}\\
-    \qtab \quad \text{then let } v' = \bxget{bx_1}{s}\\
-    \qtab \quad \text{else let } v' = \bxget{bx_2}{s}\\
-    \qtab \quad \text{fi } {cond_{sv1} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if } {cond_{sv1} \ s \ v} \ and \ {cond_{s1} \ s} \\
-        \qtab \text{then let } (s', v') = \pg{bx_1}{s}{v}\\
-        \qtab \text{else let } (s', v') = \pg{bx_2}{s}{v}\\
-        \qtab \text{fi } cond_{v1} \ s' \ and \ cond_{sv1} \ s \ v'; \text{ return } (s',v')$\\
+$\pg{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} \\
+    \tab = (\text{if} \ {cond_{sv_1} \ s \ v} \\
+    \qtab \text{then let} \ s' = \putbx{bx_1}{s}{v} \\
+    \qtab \text{else let} \ s' = \putbx{bx_2}{s}{v} \\
+    \qtab \text{fi} \ cond_{s_1} \ s'; \text{ return } s', \\
+    \qtab \quad \text{if} \ {cond_{s_1} \ s}\\
+    \qtab \quad \text{then let} \ v' = \getbx{bx_1}{s}\\
+    \qtab \quad \text{else let} \ v' = \getbx{bx_2}{s}\\
+    \qtab \quad \text{fi} \ {cond_{sv_1} \ s \ v'}; \text{ return } v') \\
+    \tab = \text{if} \ {cond_{sv_1} \ s \ v} \ and \ {cond_{s_1} \ s} \\
+        \qtab \text{then let} \ (s', v') = \pg{bx_1}{s}{v}\\
+        \qtab \text{else let} \ (s', v') = \pg{bx_2}{s}{v}\\
+        \qtab \text{fi} \ cond_{v_1} \ s' \ and \ cond_{sv_1} \ s \ v'; \text{ return } (s',v')$\\
 
 A restriction for $\pginline{case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\bxputinline{case}$ as well as $\bxgetinline{case}$.
+We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$.
 If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\bxputinline{case}$ and $\bxgetinline{case}$ are not always simultaneously satified.
-The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv1} \ s \ v) \ and \ (not (cond_{s1} \ s)))$ and $((not (cond_{sv1} \ s \ v) \ and \ (cond_{s1} \ s))$, which are restricted in this paper.
+This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified.
+The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper.
 This does not happen for the others, which the construction of $\pginline{case}$ uses above.
@@
 $\pg{bx_1 \circ bx_2}{s}{v} \\
-    \hspace*{5mm} = (\bxput{bx_1}{s}{(\bxput{bx_2}{(\bxget{bx_1}{s})}{v})}, \bxget{bx_2}{(\bxget{bx_1}{s})}) \\
-    \hspace*{5mm} = \text{let } a = \bxget{bx_1}{s} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
-        \hspace*{5mm} \ \quad (d, c) \\
-    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in} \\
-        \hspace*{5mm} \ \quad (d, c) \\
-    \hspace*{5mm} = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (b, c) = \pg{bx_2}{a}{v} \text{ in} \\
-        \hspace*{5mm} \ \quad \text{let } (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \text{ in} \\
-        \hspace*{5mm} \ \quad (d, c)$\\
+    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
+    \tab = \text{let} \ a = \getbx{bx_1}{s} \ \text{in} \\
+        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
+        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
+        \qtab (d, c) \\
+    \tab = \text{let} \ (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
+        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
+        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
+        \qtab (d, c) \\
+    \tab = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
+        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
+        \qtab \text{let} \ (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \ \text{in} \\
+        \qtab (d, c)$\\
 
@@
 The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $(\text{let } a = \bxget{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$.
-It is possible if we consider $(\bxget{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid.
-The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\bxput{bx_1}{s}{dummy}$.
+The third one rewrites $(\text{let } a = \getbx{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$.
+It is possible if we consider $(\getbx{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid.
+The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
 This transformation is feasible due to the PUTPUT law.
@@
 
-$fst = \lambda (Con (x1, x2)). x1$\\
-$snd = \lambda (Con (x1, x2)). x2$\\
-$con = \lambda ks1. \lambda ks2. \lambda x. Con(ks1 \ x, ks2 \ x)$\\
+$fst = \lambda (Con (x_1, x_2)). x_1$\\
+$snd = \lambda (Con (x_1, x_2)). x_2$\\
+$con = \lambda ks_1. \lambda ks_2. \lambda x. Con(ks_1 \ x, ks_2 \ x)$\\
 
 $\cpg{bx}{ks}{kv}{s}{v}$ \\
-    \tab $ks, kv$ evaluate actual values\\
+    \tab $ks, kv$ keep modification information\\
     \tab $s, v$ keep evaluated values\\
@@
 $\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks1, kv1, s1, v1) \Leftarrow \cpg{bx1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab ((con \ ks1 \ ks2, con \ kv1 \ kv2, Con \ s1 \ s2, Con \ v1 \ v2)$\\
+    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
+    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
+    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, Con \ s_1 \ s_2, Con \ v_1 \ v_2)$\\
 
-$\cpg{RearrS \ f1 \ f2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f1 \circ ks}{kv}{f1 \ s}{v};\\
-    \qtab (f2 \circ ks, kv, s, v)$\\
+$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
+    \qtab (f_2 \circ ks, kv, s, v)$\\
 
-$\cpg{RearrV \ g1 \ g2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g1 \circ kv}{s}{g1 \ v};\\
-    \qtab (ks, g2 \circ kv, ks, g2 \ v)$\\
+$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
+    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
 
-$\cpg{Case \ c \ bx1 \ bx2}{ks}{kv}{s}{v} =\\
+$\cpg{Case \ c \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
     \tab \text{if} \ c \ s \ v \ \text{then}\\
-        \qtab \cpg{bx1}{ks}{kv}{s}{v}\\
+        \qtab \cpg{bx_1}{ks}{kv}{s}{v}\\
     \tab \text{else} \\ 
-        \qtab \cpg{bx2}{ks}{kv}{s}{v}$\\
+        \qtab \cpg{bx_2}{ks}{kv}{s}{v}$\\
 
-$\cpg{bx1 \circ bx2}{ks}{kv}{s}{v} =\\
-    \tab (ks1, kv1, \underline{s1}, v1) \Leftarrow \cpg{bx1}{ks}{id}{s}{const$\_$dummy \ v};\\
-    \tab (ks2, kv2, s2, v2) \Leftarrow \cpg{bx2}{kv1}{kv}{v1}{v};\\
-        \qtab (ks1 \circ ks2, kv2,  ks1 \ s2, v2)$\\
+$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
+    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
+        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$\\
 
@@
 \item (Slower for right assoc programs)
-\item $s1'$ in composition is redundant.  
+\item $s_1'$ in composition is redundant.  
 \item Next, we will extend this to avoid this redundant computation.
@@
 
-$\kpg{bx}{ks'}{kv'}{ks}{kv}{s}{v}$ \\
-    \tab $ks', kv'$ keep modification information\\
-    \tab $ks, kv$ evaluate actual values\\
+$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$ \\
+    \tab $ks, kv$ keep modification information\\
+    \tab $ks', kv'$ evaluate actual values\\
     \tab $s, v$ keep evaluated values\\
 
-$\kpg{Skip \ h}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \tab s \Leftarrow ks \ s;\\
-    \tab v \Leftarrow kv \ v;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks', kv', id, id, s, v) \ \text{else fail}$\\
+$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
+    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
+    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
 
-$\kpg{Replace}{ks'}{kv'}{ks}{kv}{s}{v} = (kv', ks', kv, ks, v, s)$\\
+$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
 
-$\kpg{bx1 \ prod \ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \textcolor{red}{s \Leftarrow ks \ s; v \Leftarrow kv \ v; ks \Leftarrow id; kv \Leftarrow id;}\\
-    (ks1', kv1', ks1, kv1, s1, v1) \Leftarrow \kpg{bx1}{fst \circ ks'}{fst \circ kv'}{fst \circ ks}{fst \circ kv}{s}{v};\\
-    (ks2', kv2', ks2, kv2, s2, v2) \Leftarrow \kpg{bx2}{snd \circ ks'}{snd \circ kv'}{snd \circ ks}{snd \circ kv}{s}{v};\\
-    \tab ( con \ ks1' \ ks2', con \ kv1' \ kv2', \\
-    \qtab con \ (ks1 \circ fst) \ (ks2 \circ snd), con \ (kv1 \circ fst) \ (kv2 \circ snd), \\
-    \qtab (Con \ s1 \ s2), (Con \ v1 \ v2))$\\
+$\kpg{bx_1 \ prod \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \textcolor{red}{s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;}\\
+    \textcolor{red}{ks' \Leftarrow id; \quad kv' \Leftarrow id;}\\
+    (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
+    (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
+    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
+    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
+    \tab (Con \ s_1 \ s_2), (Con \ v_1 \ v_2))$\\
 
-$\kpg{RearrS \ f1 \ f2 \ bx}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \tab (ks', kv', ks, kv, s, v) \Leftarrow \kpg{bx}{f1 \circ ks'}{kv'}{f1 \circ ks}{kv}{s}{v};\\
-    \qtab (f2 \circ ks', kv', f2 \circ ks, kv, s, v)$\\
+$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
+    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
 
-$\kpg{RearrV \ g1 \ g2 \ bx}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \tab (ks', kv', ks, kv, s, v) \Leftarrow \kpg{bx}{ks'}{g1 \circ kv'}{ks}{g1 \circ kv}{s}{v};\\
-    \qtab (ks', g2 \circ kv', ks, g2 \circ kv, s, v)$\\
+$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
+    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
 
-$\kpg{Case \ c \ bx1 \ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \tab s \Leftarrow ks \ s;\\
-    \tab v \Leftarrow kv \ v;\\
+$\kpg{Case \ c \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
+    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
     \tab \text{if} \ c \ s \ v \ \text{then}\\
-        \qtab \kpg{bx1}{ks'}{kv'}{id}{id}{s}{v}\\
+        \qtab \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
     \tab \text{else}\\\
-        \qtab \kpg{bx2}{ks'}{kv'}{id}{id}{s}{v}$\\
+        \qtab \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}$\\
 
-$\kpg{bx1 \circ bx2}{ks'}{kv'}{ks}{kv}{s}{v} =\\
-    \tab (ks1', kv1', \underline{ks1}, kv1, \underline{s1}, v1) \Leftarrow \kpg{bx1}{ks'}{id}{ks}{id}{s}{const$\_$dummy \ v};\\
-    \tab (ks2', kv2', ks2, kv2, s2, v2) \Leftarrow \kpg{bx2}{kv1'}{kv'}{kv1}{kv}{v1}{v};\\
-    \qtab (ks1' \circ ks2', kv2', ks1' \circ ks2, kv2, s2, v2)$
+$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
+    \qtab (ks_1 \circ ks_2, kv_2', ks_1 \circ ks_2', kv_2', s_2, v_2)$
 
@@
 \item In this definition of kpg, s' and v' will have actual value only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
-\item In composition $ks1'$ and $s1'$ are still redundant. But, they are not evaluated.
+\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
 \item In prod, Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s' and v'.
@@
 $\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks1',kv1', ks1, kv1, s1, v1) \Leftarrow \kpg{bx_1}{id}{id}{\lambda \_.s}{id}{s}{v};\\
-    \tab (s2, v2) \Leftarrow \xpg{bx_2}{kv1 \ v1}{v};\\
-    \qtab (ks1' \ s2, v2)$
+    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
+    \qtab (ks_1 \ s_2, v_2)$
 
@@
 
-\bibitem{ref_lncs1}
+\bibitem{ref_lncs_1}
 Author, F., Author, S.: Title of a proceedings paper. In: Editor,
@@
 schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+Fig.~\ref{fig_1}).
 
 \begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
+\includegraphics[width=\textwidth]{fig_1.eps}
 \caption{A figure caption is always placed below the illustration.
 Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
+justified by the macro package automatically.} \label{fig_1}
 \end{figure}
@@
 a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
+articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
 book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
 and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
+\cite{ref_article1,ref_lncs_1,ref_book1},
 \cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

### 2019-08-19 paper/main.tex
update csv

@@
 \usepackage{amssymb}
+\usepackage{multirow}
+\usepackage{tabularx}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+\usepackage{subfig}
+\usepackage{lmodern}
 % Used for displaying a sample figure. If possible, figure files should
@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\pgfplotsset{compat=1.12}
 
@@
 \newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
+
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
 
@@
 
-$\getrev{(bx_1 \circ bx_2)}{s}{v} =\\
-    \tab let (v_1, g_1) = \getrev{bx_1}{s} in\\
-    \tab let (v_2, g_2) = \getrev{bx_1}{s} in\\
+$\getrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
+    \tab \text{let} \ (v_2, g_2) = \getrev{bx_1}{s} \ \text{in}\\
         \qtab (v_2, (g_1, g_2))$
@@
 
-\textbf{Definition:} \\
-$\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$ \\
+\begin{definition}
+    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
+\end{definition}
 
@@
 
-Proof:\\
+\begin{proof}
 $pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
@@
     \tab = (s, v) \quad [\text{GETPUT}]$
+\end{proof}
 
@@
     \tab $s, v$ keep evaluated values\\
+
 \vspace{3mm}
@@
 
+$s \Leftarrow s_0; v \Leftarrow v_0;\\
+(ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
+    \tab (s; v)$
+
 \begin{itemize}
@@
     \tab $s, v$ keep evaluated values\\
+
+\vspace{3mm}
 
@@
     \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2', ks_1 \circ ks_2', kv_2', s_2, v_2)$
+    \qtab (ks_1 \circ ks_2, kv_2', ks_1 \circ ks_2', kv_2', s_2, v_2)$\\
+
+$s \Leftarrow s_0; v \Leftarrow v_0;\\
+(ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+    \tab (ks' \ s; kv' \ v)$
 
@@
 
-\section{Experimental result}
+\section{Experiment}
 
-\begin{itemize}
-\item Evaluation time for miniBiGul, pg, cpg, kpg, xpg.
-\item Memory allocation of miniBiGul, pg, cpg, kpg, xpg.
-\item []
-\item xpg looks the best one (maybe)
-\item 
-\end{itemize}
+This section describes the experiments involving $minBiGUl$, $pg$, $cpg$, $kpg$ and $xpg$. The evaluation time, the memory allocation and the number of function calls are considered in each test.
 
-\subsection{Theoretical Computational Complexity}
+\subsection{Experiment environment}
+We implemented 5 approaches with Ocaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
+The Ocaml runtime system options and garbage collection parameters are set as default.
+
+\subsection{Test cases}
+We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
+
+\begin{table}[]
+    \centering
+    \caption{Test cases}
+    \label{tab:test-cases}
+    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
+        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
+        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
+        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
+        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
+        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
+        5 & breverse & left & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
+    \end{tabular*}
+\end{table}
+
+In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
+This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
+Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
+
+The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
+For example, \textit{lassoc-comp-replace} looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace} is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
+\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
+While $Replace$ is a primitive BX program, $phead$ is a combination of such programs.
+The following is the definition of $phead$ in BiGUL [?]:
+
+    $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
+    \tab phead = \$(rearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
+        \qtab \qtab \quad \$(rearrV \ [\![\lambda(v) \to (v,())]\!])\$\\
+            \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
+
+We also provide the same $phead$ program in our Ocaml system.
+It rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
+$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
+$\getbx{phead}{s_0}$ returns the head of $s_0$.
+For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.
+If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
+For example:
+
+    \tab $\putbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]}{100} = [[100,2,3],[\ ],[4,5]]$\\
+    \tab $\getbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]} = 1$
+
+Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
+For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ direction on the composition valid.\\
+
+The last test case, $bReverse$, is a left-associative and recursive composition.
+It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minBiGUL.
+What we know about $foldr$, an important higher-order function on lists, is as follows:\\
+    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
+    \tab foldr \ f \ e \ [\ ] = e\\
+    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
+
+We can define function $reverse$ on lists based on $foldr$:\\
+    $\tab reverse \ [\ ] = [\ ]\\
+    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
+    \tab reverse = foldr \ snoc \ [\ ]$
+
+Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
+
+$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
+\tab bFoldr \ bx =\\
+    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
+        \qtab \qtab \Longrightarrow \$(rearrV [| \lambda v \to ((),v) |])\$\\
+            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
+    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
+        \qtab \qtab \Longrightarrow \$(rearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
+            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
+    \qtab \qtab]\\$
+
+The detailed explanation of $bFoldr$ are skipped.
+Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our Ocaml system.
+If we put our eyes on the second branch of $bFoldr$, we can understand why it is recursive and left-associative composition.
+Note that the compose operator has a higher priority than the product operator.
+This leads to seemingly impossible to transform this function from left-associative style to right-associative style.
+
+Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
+
+    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
+    \tab bReverse =\\
+        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
+        \qtab \qtab \quad \lambda s \ v \to v\\
+        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
+        \qtab \qtab \quad \$(rearrS \ [\![ \lambda s \to (s,[\ ]) ]\!])\$\\
+            \qtab \qtab \qquad bFoldr \ bSnoc\\
+       \qtab \qtab ]$
+
+The above definition contains an adaptive branch and it only occurs when the length of the source and the view are different.
+If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
+For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
+Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
+Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
+In $bReverse$, the number of composition equals to the length of the source and the view.
+
+\subsection{Results and discussions}
+We conducted experiments on 5 test cases while changing the number of composition from 10 to 100000.
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north east
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north east
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Evaluation time (s)}
+    \label{fig:evaluation-time}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north west
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north east
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of composition},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize},
+                legend pos = north east
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Memory allocation (MBytes)}
+    \label{fig:memory-allocation}
+\end{figure}
 

### 2019-08-23 paper/main.tex
Update main.tex

@@
 
-This section describes the experiments involving $minBiGUl$, $pg$, $cpg$, $kpg$ and $xpg$. The evaluation time, the memory allocation and the number of function calls are considered in each test.
+This section describes the experiments involving $minBiGUl$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
 
@@
 
-The above definition contains an adaptive branch and it only occurs when the length of the source and the view are different.
+The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
 If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
@@
 Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of composition equals to the length of the source and the view.
+In $bReverse$, the number of compositions equals to the length of the source and the view.
 
 \subsection{Results and discussions}
-We conducted experiments on 5 test cases while changing the number of composition from 10 to 100000.
 
@@
                 ymode = normal,
-                xlabel = {number of composition},
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Number of calls}
+    \label{fig:count-calls}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
                 ylabel = {evaluation time (s)},
                 width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north west
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north west
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
+                xlabel = {number of compositions},
                 ylabel = {evaluation time (s)},
                 width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north west
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north east
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
+                xlabel = {number of compositions},
                 ylabel = {evaluation time (s)},
                 width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north east
+                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
+                xlabel = {number of compositions},
                 ylabel = {memory allocation (MBytes)},
                 width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north west
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
             ]
@@
                 ymode = normal,
-                xlabel = {number of composition},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north east
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
             ]
@@
                 \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
+                \addlegendentry{minbigul};x
                 \addplot table[x=nComp,y=pg]{\data};
@@
                 ymode = normal,
-                xlabel = {number of composition},
+                xlabel = {number of compositions},
                 ylabel = {memory allocation (MBytes)},
                 width = 0.55\linewidth,
-                legend style = {font=\scriptsize},
-                legend pos = north east
+                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
             ]
@@
 \end{figure}
+
+\begin{table}[]
+    \centering
+    \caption{Comparison}
+    \label{tab:comparison}
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{|l|l|l|l|}
+            \hline
+            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
+            lassoc-comp-replace & l, NR & kpg $\sim$ xpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ kpg $\sim$ xpg (95299) $<$ pg (ND) \\ \hline
+            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
+            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ kpg $\sim$ xpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
+            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
+            breverse & l, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
+            \multicolumn{4}{r}{} \\
+            \multicolumn{4}{r}{l/r: left/right associative} \\
+            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
+            \multicolumn{4}{r}{ND: not determined} \\
+            \multicolumn{4}{r}{best $< \ldots <$ worst} \\
+        \end{tabular}%
+    }
+    \end{table}
 

### 2019-08-23 paper/main.tex
update main.tex 190823
+ memo for results & discussions

@@
 
-\begin{itemize}
-\item Information loss causes re-evaluation of get.
-\item To prevent this, we introduced the idea of reversible computation.
-\item In reversible computation we need to keep all information: we can not lose any information.
-\item In BiGul we can do this with a pair of put and get.
-\item pg bx (s, v) = (put bx (s, v), get bx (s))
-\item pg: s * v $\to$ s * v
-\item No garbage
-\item This function is self-inverse. (involution, (f (f input)) = input)
-\item \textcolor{red}{Constraint: In case expression, put and get must always choose the same branch.
-This constraint is not satisfied in general: (e.g, SNOC, the different length of two inputs.) To remove this constraint we need to use garbage.}
-\end{itemize}
-
 For the approaches so far, when evaluating a composition of BX programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
@@
 
+\textcolor{red}{
+    $s \Leftarrow s_0; v \Leftarrow v_0;\\
+    (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
+        \tab (s; v)$
+}
+
 \vspace{3mm}
@@
 
-$s \Leftarrow s_0; v \Leftarrow v_0;\\
-(ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
-    \tab (s; v)$
-
-\begin{itemize}
-\item This is faster than pg for left assoc BX programs.
-\item (Slower for right assoc programs)
-\item $s_1'$ in composition is redundant.  
+\begin{itemize}
+\item $s_1$ in composition is redundant.  
 \item Next, we will extend this to avoid this redundant computation.
@@
     \tab $s, v$ keep evaluated values\\
+
+\textcolor{red}{
+    $s \Leftarrow s_0; v \Leftarrow v_0;\\
+    (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+        \tab (ks' \ s; kv' \ v)$
+}
 
@@
 
-$s \Leftarrow s_0; v \Leftarrow v_0;\\
-(ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-    \tab (ks' \ s; kv' \ v)$
-
-\begin{itemize}
-\item In this definition of kpg, s' and v' will have actual value only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
-\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
-\item In prod, Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s' and v'.
-\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
+\begin{itemize}
+    \item In this definition of kpg, s and v keep actual values only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
+    \item In composition $ks_1'$ and $s_1$ are still redundant. But, they are not evaluated.
+    \item In prod, evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s and v.
+    \item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg). \textcolor{red}{But it makes stack\_overflow comes sooner}.
 \end{itemize}
@@
 
-This section describes the experiments involving $minBiGUl$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
+This section describes the experiments involving $minBIGUL$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
 
 \subsection{Experiment environment}
-We implemented 5 approaches with Ocaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The Ocaml runtime system options and garbage collection parameters are set as default.
+We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
+The OCaml runtime system options and garbage collection parameters are set as default.
 
@@
 
-We also provide the same $phead$ program in our Ocaml system.
+We also provide the same $phead$ program in our OCaml system.
 It rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
@@
 The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our Ocaml system.
+Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
 If we put our eyes on the second branch of $bFoldr$, we can understand why it is recursive and left-associative composition.
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
 \begin{table}[]
@@
             \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & kpg $\sim$ xpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ kpg $\sim$ xpg (95299) $<$ pg (ND) \\ \hline
+            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
             rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ kpg $\sim$ xpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
+            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
             rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
@@
             \multicolumn{4}{r}{} \\
+            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
+            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
             \multicolumn{4}{r}{l/r: left/right associative} \\
             \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined} \\
-            \multicolumn{4}{r}{best $< \ldots <$ worst} \\
+            \multicolumn{4}{r}{ND: not determined}
         \end{tabular}%
@@
 
+\begin{itemize}
+    \item Number of calls
+        \begin{itemize}
+            \item No(cpg) = No(kpg) = No(xpg) = No(put) (\textcolor{blue}{come from function definitions})
+            \item For right-associative \& non-recursive composition (rassoc-comp-replace, rassoc-comp-phead): No(get), No(put), No(pg) are linear where No(get) $<$ No(put) $<$ No(pg)
+            \item For left-associative \& non-recursive composition (lassoc-comp-replace, lassoc-comp-phead): No(put) is still linear, No(get) is quadratic, No(pg) is under exponential distribution.
+No(put) $<$ No(get) $<$ No(pg)
+            \item For left-associative \& recursive composition (breverse): we still have No(put) $<$ No(get) $<$ No(pg) and all three are non-linear.
+In case of breverse, No(put) is quadratic, No(get) is cubic and No(pg) is exponential.
+        \end{itemize}
+    \item Evalutation times (s)
+        \begin{itemize}
+            \item For left-associative compositions:\\
+            - pg time is exponential (\textcolor{blue}{since the number of pg calls in these cases are exponential}) \\
+            \item For left-associative \& non-recursive compositions:\\
+            - kpg time are are approximate to xpg time, and they are the most efficient.\\
+            - In case of lassoc-comp-replace, kpg time and xpg time are linear, cpg time and minbigul time are non-linear.
+(\textcolor{blue}{because the input size is constant, No(get) is quadratic for minbigul while the number of redundant evaluations in cpg is \underline{linear} and ...
+?})\\
+            - In case of lassoc-comp-phead, because of the data size is linear and the numbers of function calls are either linear or quadractic or exponential, then all times are non-linear.
+
+            \item For right-associative \& non-recursive compositions:\\
+            - minbigul time, pg time and xpg time are linear \\
+            - In case of rassoc-comp-replace, all times are small enough to consider all them as linear \\
+            - In case of rassoc-comp-phead, cpg time and kpg time are quadratic (\textcolor{blue}{because data size is linear, the number of redundant evaluations in cpg and kpg are linear})
+
+            \item For left-associative \& recursive composition (breverse):\\
+            - all times should be non-linear (\textcolor{blue}{due to there is no linear number of function calls})
+        \end{itemize}
+    \item Memory allocation (MBytes)
+        \begin{itemize}
+            \item Memory usage depends on the input size and the number of composition
+            \item In general, minbigul uses less memory than the others for non-recursive compositions (all tests except breverse)
+            \item pg only uses memory efficient when handling right-associative composition 
+            \item cpg \& kpg are worst in memory allocation for right-associative
+            \item For left-associative \& recursive composition (breverse), minbigul uses significantly more memory than kpg as well as xpg.
+        \end{itemize}
+\end{itemize}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Number of calls}
+    \label{fig:count-calls}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Evaluation time (s)}
+    \label{fig:evaluation-time}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};x
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Memory allocation (MBytes)}
+    \label{fig:memory-allocation}
+\end{figure}
+
+\newpage
 \section{Conclusion and Future Work}

### 2019-08-23 bigul/Alignment.hs
update bigul

@@ -1,114 +1,0 @@
-{-#  LANGUAGE TemplateHaskell, TypeFamilies, ScopedTypeVariables  #-}
-
-module Alignment where
-
-import Generics.BiGUL
-import Generics.BiGUL.TH
-import Generics.BiGUL.Lib
-import Generics.BiGUL.Interpreter
-
-import Data.Tuple
-import Data.Maybe
-import Data.List
-
-type Source = (Id, (Name, Salary))
-
-type Id     = Int
-type Name   = String
-type Salary = Int
-
-employees :: [Source]
-employees =  [ (0, ("Zhenjiang", 1000))
-             , (1, ("Josh"     ,  400))
-             , (2, ("Jeremy"   , 2000)) ]
-
-type View = (Id, Name)
-
-bx :: BiGUL Source View
-bx =  $(rearrV [| \(id, name) -> (id, (name, ())) |])$
-        Replace `Prod` (Replace `Prod` Skip (const ()))
-
-cr :: View -> Source
-cr (i, n) = (i, (n, 0))
-
-posAlign :: (Show s, Show v) => BiGUL s v -> (v -> s) -> BiGUL [s] [v]
-posAlign b c = Case
-  [ $(normalSV [p| [] |] [p| [] |] [p| [] |])
-    ==> $(update [p| [] |] [p| [] |] [d|  |])
-  , $(normalSV [p| _:_ |] [p| _:_ |] [p| _:_ |])
-    ==> $(update [p| x:xs |] [p| x:xs |] [d| x = b; xs = posAlign b c |])
-  , $(adaptiveSV [p| _:_ |] [p| [] |])
-    ==> \_ _ -> []
-  , $(adaptiveSV [p| [] |] [p| _:_ |])
-    ==> \_ (v:_) -> [c v]
-  ]
-
-updatedEmployees0 :: [View]
-updatedEmployees0 =  [(0, "Zhenjiang"), (2, "Jeremy")]
-
-updatedEmployees1 :: [View]
-updatedEmployees1 =  [(2, "Jeremy"), (0, "Zhenjiang"), (1, "Josh")]
-
-keyAlign :: forall s v k . (Show s, Show v, Eq k)
-         => (s -> k) -> (v -> k) -> BiGUL s v -> (v -> s) -> BiGUL [s] [v]
-keyAlign ks kv b c = Case
-  [ $(normalSV [p| [] |] [p| [] |] [p| [] |])
[…]

### 2019-08-23 bigul/Basic.hs

@@ -93,3 +93,4 @@ pEither x0 =
        $(adaptive [| \_ _ -> True |])
          ==> \s v -> Left v
        ]
+

### 2019-08-23 paper/main.tex
add explanations (except pg & experiment)

--- 2019-08-19 paper/main.tex
@@
 %
-\title{Bx to Rev (tentative) \thanks{Supported by organization x.}}
+\title{An optimization of bidirectional update language based on the idea of reversible computing
+ \thanks{Supported by organization x.}}
 %
@@
 %
+
+Points to discuss
+\begin{itemize}
+\item When will be the deadline of PEPM2020? (mid of October?)
+\item Need to change the style file of latex..
+\item []
+
+\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
+\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
+\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
+\item Related work: suggest?
+\item another pg ?? should be compared?
+\end{itemize}
+
+\begin{tabular}{c|c|c}
+  {} & {Not to keep eval info} & {Keep eval info as a function}\\ \hline
+  {eager evaluation} & {pg} & {cpg}\\ \hline
+  {lazy evaluation} & {??} & {kpg}
+\end{tabular}
+
 \section{Introduction}
@@
 \item Current status of BiGUL: Fastest BX language in the world
-\item A problem: Efficiency of compose evaluation. If we do not save the intermediate states, the number of get is quadratic. This is not good.
-\item To solve this problem we use an idea from reversible computation: not to lose any information.
-\item Based on this idea we introduce pg.
-\item We extend pg with several ideas and produce faster implementation.
+\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
+\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
+%  an idea from reversible computation: not to lose any information.
+\item We extend pg with several ideas to produce faster implementation.
+\item 
 \end{itemize}
@@
 \item Improvement of evaluation efficiency
+  \begin{itemize}
+  \item Quadratic to exponential? 
+  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+  \end{itemize}
 \item Reveal a part of relationship between BX and Reversible.
@@
   \end {itemize}
-\item Introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+
 \end{itemize}
@@
 \begin{itemize}
-\item 
+\item BiGUL 
+\item pg (pair of put and get)
+\item cpg (extention of pg. Keep eval information for reusing)
+\item kpg (extention of cpg. Lazy evaluation)
+\item xpg (combination of pg and kpg)
+\item Experiment
+\item Related work
+\item Conclusion and Future work
 \end{itemize}
@@
 \item our target language is miniBiGUL (BiGUL minus adaptive case) : very well-behaved
+\item Basic explanation of bidirectional transformation
+  \begin{itemize}
   \item get = s $\to$ v
@@
   \item We assume putget and getput hold (explanation about putget and getput)
+  \item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
+  \end{itemize}
 \item explanation about put-based BX language: BiGUL
-\item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
 \end{itemize}
@@
 
-\section{Skip:Reversible langauge: Rwhile}
+\section{Remove?:Reversible langauge: Rwhile}
 
@@
 
-\section{Skip:First approach}
+\section{Remove?:First approach}
 
@@
 \item In reversible computation we need to keep all information: we can not lose any information.
-\item In BiGul we can do this with a pair of put and get.
+\item In BiGUL we can acheive this by tupling put and get.
 \item pg bx (s, v) = (put bx (s, v), get bx (s))
@@
 
-\subsection{Skip:Construction of pg}
+\subsection{Construction of pg}
 
@@
 
-\subsection{How to Evaluate Bidirectional Programs using pg}
-
-\begin{itemize}
-\item Because we combine put and get, we need to supply view for evaluation of get.
-\item introduce dummy?
-\item automatic analysis for domain \& range?
-\item 
-\item partial evaluation of put(s(D), dummy(S)).
-\item \textcolor{red}{Not good?
+\subsection{Evaluation by pg}
+
+\begin{itemize}
+\item In pg approach we always need to supply source and view.
+\item If we would like to evaluate get, we do not have any view.
+\item For this, we require the programmer to write a function to produce view from source. (This can be the same with get??)
+\item Example program: \textcolor{red}{Good example?}
+  Evaluate: $\putbx{bx}{s}{v}$ 
+  Evalaute: $\getbx{bx}{s}$
+  $\pg{bx}{s}{v} = $
+\end{itemize}
+
+%\begin{itemize}
+%\item Because we combine put and get, we need to supply view for evaluation of get.
+%\item introduce dummy?
+%\item automatic analysis for domain \& range?
+%\item 
+%\item partial evaluation of put(s(D), dummy(S)).
+%\item \textcolor{red}{Not good?
 Basically put (BX) is often a recursion on the source.
@@
 So partial evaluation might not work well.}
-\item \textcolor{red}{lazy evaluation is also a possibility?}
-\end{itemize}
+%\item \textcolor{red}{lazy evaluation is also a possibility?}
+%\end{itemize}
 
@@
 \begin{itemize}
-\item If the BX programs are left assoc of composition, pg will be exponential. Because it calls pg bx$_1$ twice.
+\item If the given BX programs are left assoc of composition, the recursive call of pg will be exponential. Because it calls $\pg{bx_1}{\_}{\_}$ twice in compose.
 \item To solve this problem, we introduce accumulation of changes.
@@
 
-$s \Leftarrow s_0; v \Leftarrow v_0;\\
-(ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
-    \tab (s; v)$
-
 \begin{itemize}
@@
 \item (Slower for right assoc programs)
-\item $s_1'$ in composition is redundant.  
-\item Next, we will extend this to avoid this redundant computation.
+
 \end{itemize}
@@
 \subsection{kpg}
+
+\begin{itemize}
+\item In cpg, $s_1'$ in composition is not used, but evaluated.
+\item Next, we will extend cpg to avoid this redundant computation.
+\item For this, we evaluate the values lazily.
+\item []
+\item s' and v' will have actual value only in Skip and case. Except them, the functions for computation will be kept as ks' and kv'
+\item Skip and Case need values of the time. They evaluate s and v with ks' and kv', and ks' and kv' will be reseted to an identity function.
+\item A special thing is in prod. In prod, if we omit red parts,
+Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. Because this causes redundant re-evaluation of ks' and kv', we insert evaluation before recursive calls.
+\item Except them, the functions for computation will be kept as ks' and kv'.
+\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
+\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
+\end{itemize}
 
@@
     \tab $ks, kv$ keep modification information\\
-    \tab $ks', kv'$ evaluate actual values\\
+    \tab $ks', kv'$ for lazy evaluation of actual values\\
     \tab $s, v$ keep evaluated values\\
@@
 $\kpg{bx_1 \ prod \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \textcolor{red}{s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;}\\
-    \textcolor{red}{ks' \Leftarrow id; \quad kv' \Leftarrow id;}\\
-    (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
+    \tab \textcolor{red}{s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;}\\
+    \tab \textcolor{red}{ks' \Leftarrow id; \quad kv' \Leftarrow id;}\\
+    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
+    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
     \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
@@
 
-$s \Leftarrow s_0; v \Leftarrow v_0;\\
-(ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-    \tab (ks' \ s; kv' \ v)$
-
-\begin{itemize}
-\item In this definition of kpg, s' and v' will have actual value only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
-\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
-\item In prod, Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s' and v'.
-\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
-\end{itemize}
-
-\section{xpg}
-
-\begin{itemize}
-\item pg is fastest for right assoc programs and kpg is fastest for left assoc programs.
-\item To combine their strength, we can use the following definition.
+%$s \Leftarrow s_0; v \Leftarrow v_0;\\
+%(ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+%    \tab (ks' \ s; kv' \ v)$
+
+\subsection{xpg}
+
+\begin{itemize}
+\item So far, we focus on left assoc programs. xpg is faster than the others.
+\item On the other hand, pg is fastest for right assoc programs.
+\item To combine their strength, we introfuce new pg that combines pg and kpg.
 \item If there is a composition, we use xpg for evaluation of $bx_1$ and pg for evaluation of $bx_2$.
@@
 
+\section{Related Work}
+
 \section{Conclusion and Future Work}

### 2019-08-23 paper/main.tex

--- 2019-08-23 paper/main.tex
@@
 \subsection{Self-inverse function: pg}
+
+%\begin{itemize}
+%\item Information loss causes re-evaluation of get.
+%\item To prevent this, we introduced the idea of reversible computation.
+%\item In reversible computation we need to keep all information: we can not lose any information.
+%\item In BiGUL we can acheive this by tupling put and get.
+%\item pg bx (s, v) = (put bx (s, v), get bx (s))
+%\item pg: s * v $\to$ s * v
+%\item No garbage
+%\item This function is self-inverse. (involution, (f (f input)) = input)
+%\item \textcolor{red}{Constraint: In case expression, put and get must always choose the same branch.
+This constraint is not satisfied in general: (e.g, SNOC, the different length of two inputs.) To remove this constraint we need to use garbage.}
+%\end{itemize}
 
@@
 \begin{itemize}
+\item This is faster than pg for left assoc BX programs.
+\item (Slower for right assoc programs)
 \item $s_1$ in composition is redundant.  
@@
 
+\begin{itemize}
+\item In cpg, $s_1'$ in composition is not used, but evaluated.
+\item Next, we will extend cpg to avoid this redundant computation.
+\item For this, we evaluate the values lazily.
+\item []
+\item s' and v' will have actual value only in Skip and case. Except them, the functions for computation will be kept as ks' and kv'
+\item Skip and Case need values of the time. They evaluate s and v with ks' and kv', and ks' and kv' will be reseted to an identity function.
+\item A special thing is in prod. In prod, if we omit red parts,
+Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. Because this causes redundant re-evaluation of ks' and kv', we insert evaluation before recursive calls.
+\item Except them, the functions for computation will be kept as ks' and kv'.
+\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
+\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
+\end{itemize}
+
 $\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$ \\
     \tab $ks, kv$ keep modification information\\
-    \tab $ks', kv'$ evaluate actual values\\
+    \tab $ks', kv'$ for lazy evaluation of actual values\\
     \tab $s, v$ keep evaluated values\\
@@
 
-\section{xpg}
-
-\begin{itemize}
-\item pg is fastest for right assoc programs and kpg is fastest for left assoc programs.
-\item To combine their strength, we can use the following definition.
+\subsection{xpg}
+
+\begin{itemize}
+\item So far, we focus on left assoc programs. xpg is faster than the others.
+\item On the other hand, pg is fastest for right assoc programs.
+\item To combine their strength, we introfuce new pg that combines pg and kpg.
 \item If there is a composition, we use xpg for evaluation of $bx_1$ and pg for evaluation of $bx_2$.
@@
 
+\section{Related Work}
+
 \newpage
+
 \section{Conclusion and Future Work}

### 2019-08-23 paper/main.tex
Update main.tex

--- 2019-08-23 paper/main.tex
@@
             \item For left-associative \& recursive composition (breverse):\\
-            - all times should be non-linear (\textcolor{blue}{due to there is no linear number of function calls})
+            - all times should be non-linear (\textcolor{blue}{because there is no linear number of function calls})
         \end{itemize}

### 2019-08-29 paper/main.tex
update tex 2908
+ results and discussions

--- 2019-08-23 paper/main.tex
@@
 This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified.
-The evaluated BXs are different in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper.
+The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper.
 This does not happen for the others, which the construction of $\pginline{case}$ uses above.
@@
 
-This section describes the experiments involving $minBIGUL$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
+This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
 
@@
         4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
+        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
     \end{tabular*}
@@
 
-The last test case, $bReverse$, is a left-associative and recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minBiGUL.
+The last test case, $bReverse$, is a recursive composition.
+It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
 What we know about $foldr$, an important higher-order function on lists, is as follows:\\
@@
 Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-If we put our eyes on the second branch of $bFoldr$, we can understand why it is recursive and left-associative composition.
+In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
+Then this composition can be classified as either left or right associative type.
+However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
 Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform this function from left-associative style to right-associative style.
+This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
 
@@
 \subsection{Results and discussions}
+
+We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
+
+Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
+From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
+Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
+
+For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
+
+For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
+In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
+
+For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
+In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {number of calls},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
+                \addplot table[x=nComp,y=count_get]{\data};
+                \addlegendentry{get};
+                \addplot table[x=nComp,y=count_put]{\data};
+                \addlegendentry{put};
+                \addplot table[x=nComp,y=count_pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=count_xpg]{\data};
+                \addlegendentry{cpg/kpg/xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Number of calls}
+    \label{fig:count-calls}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {evaluation time (s)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Evaluation time (s)}
+    \label{fig:evaluation-time}
+\end{figure}
+
+Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
+
+$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
+
+For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
+In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
+They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
+In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
+
+For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
+In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
+In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
+
+For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
+
+\begin{figure}
+    \centering
+    \subfloat[\textit{lassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-replace}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \hspace{0mm}
+    \subfloat[\textit{lassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \subfloat[\textit{rassoc-comp-phead}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};x
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \vspace{2mm}
+    \subfloat[\textit{breverse}]{
+        \begin{tikzpicture}
+            \begin{axis}[
+                xmode = normal,
+                ymode = normal,
+                xlabel = {number of compositions},
+                ylabel = {memory allocation (MBytes)},
+                width = 0.55\linewidth,
+                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
+            ]
+                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
+                \addplot table[x=nComp,y=minbigul]{\data};
+                \addlegendentry{minbigul};
+                \addplot table[x=nComp,y=pg]{\data};
+                \addlegendentry{pg};
+                \addplot table[x=nComp,y=cpg]{\data};
+                \addlegendentry{cpg};
+                \addplot table[x=nComp,y=kpg]{\data};
+                \addlegendentry{kpg};
+                \addplot table[x=nComp,y=xpg]{\data};
+                \addlegendentry{xpg};
+            \end{axis}
+        \end{tikzpicture}
+    }
+    \caption{Memory allocation (MBytes)}
+    \label{fig:memory-allocation}
+\end{figure}
+
+Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
+In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
+$pg$ only uses memory efficient when handling right-associative compositions.
+Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
+For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
 
@@
             rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
+            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
             \multicolumn{4}{r}{} \\
@@
 
-\begin{itemize}
-    \item Number of calls
-        \begin{itemize}
-            \item No(cpg) = No(kpg) = No(xpg) = No(put) (\textcolor{blue}{come from function definitions})
-            \item For right-associative \& non-recursive composition (rassoc-comp-replace, rassoc-comp-phead): No(get), No(put), No(pg) are linear where No(get) $<$ No(put) $<$ No(pg)
-            \item For left-associative \& non-recursive composition (lassoc-comp-replace, lassoc-comp-phead): No(put) is still linear, No(get) is quadratic, No(pg) is under exponential distribution.
-No(put) $<$ No(get) $<$ No(pg)
-            \item For left-associative \& recursive composition (breverse): we still have No(put) $<$ No(get) $<$ No(pg) and all three are non-linear.
-In case of breverse, No(put) is quadratic, No(get) is cubic and No(pg) is exponential.
-        \end{itemize}
-    \item Evalutation times (s)
-        \begin{itemize}
-            \item For left-associative compositions:\\
-            - pg time is exponential (\textcolor{blue}{since the number of pg calls in these cases are exponential}) \\
-            \item For left-associative \& non-recursive compositions:\\
-            - kpg time are are approximate to xpg time, and they are the most efficient.\\
-            - In case of lassoc-comp-replace, kpg time and xpg time are linear, cpg time and minbigul time are non-linear.
-(\textcolor{blue}{because the input size is constant, No(get) is quadratic for minbigul while the number of redundant evaluations in cpg is \underline{linear} and ...
-?})\\
-            - In case of lassoc-comp-phead, because of the data size is linear and the numbers of function calls are either linear or quadractic or exponential, then all times are non-linear.
-
-            \item For right-associative \& non-recursive compositions:\\
-            - minbigul time, pg time and xpg time are linear \\
-            - In case of rassoc-comp-replace, all times are small enough to consider all them as linear \\
-            - In case of rassoc-comp-phead, cpg time and kpg time are quadratic (\textcolor{blue}{because data size is linear, the number of redundant evaluations in cpg and kpg are linear})
-
-            \item For left-associative \& recursive composition (breverse):\\
-            - all times should be non-linear (\textcolor{blue}{due to there is no linear number of function calls})
-        \end{itemize}
-    \item Memory allocation (MBytes)
-        \begin{itemize}
-            \item Memory usage depends on the input size and the number of composition
-            \item In general, minbigul uses less memory than the others for non-recursive compositions (all tests except breverse)
-            \item pg only uses memory efficient when handling right-associative composition 
-            \item cpg \& kpg are worst in memory allocation for right-associative
-            \item For left-associative \& recursive composition (breverse), minbigul uses significantly more memory than kpg as well as xpg.
-        \end{itemize}
-\end{itemize}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
+Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
+This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
+We conducted tests while increasing the number of compositions from 10 to 100000.
+By default, the stack overflow point in the table is larger than 100000.
+In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
+Then the above points are not determined.
 

### 2019-09-05 paper/main.tex
update tex 0509

@@
 \newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrS}[2]{rearrS \ #1 \ #2}
-\newcommand{\rearrV}[2]{rearrV \ #1 \ #2}
-\newcommand{\casebx}[6]{case [#1][#2] \to #3, [#4][#5] \to #6}
+\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
+\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
+\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
 
@@
 
-% Evaluation result: Comparison with OCaml implementation of miniBiGUL
+% Evaluation result: Comparison with OCaml implementation of minBiGUL
 
@@
 \begin{itemize}
-\item our target language is miniBiGUL (BiGUL minus adaptive case) : very well-behaved
+\item our target language is minBiGUL (BiGUL minus adaptive case) : very well-behaved
 \item Basic explanation of bidirectional transformation
@@
 
-\subsection{Semantics of miniBiGUL}
+\subsection{Semantics of minBiGUL}
 
@@
 
-$\putbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
-    \tab \product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}$
-
-$\putbx{\rearrS{f}{bx}}{s}{v} = f^{-1}(\putbx{bx}{(f \ s)}{v})$
-
-$\putbx{\rearrV{g}{bx}}{s}{v} = \putbx{bx}{s}{(g \ v)}$
-
-$\putbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} = \\
-    \tab \text{if } {cond_{sv_1} \ s \ v} \\
-    \tab \text{then let } s' = \putbx{bx_1}{s}{v} \\
-    \tab \text{else let } s' = \putbx{bx_2}{s}{v} \\
-    \tab \text{fi } cond_{s_1} \ s'; \ \text{return} \ s'$
+$\putbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
+    \tab (\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2})$
+
+$\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
+
+$\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
+
+$\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
+    \tab \text{if} \ {cond_{sv} \ s \ v} \\
+    \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
+    \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
+    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
 
@@
 
-$\getbx{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
-    \tab \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}$
-
-$\getbx{\rearrS{f}{bx}}{s} = \getbx{bx}{(f \ s)}$
-
-$\getbx{\rearrV{g}{bx}}{s} = g^{-1} (\getbx{bx}{s})$
-
-$\getbx{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
-    \tab \text{if} \ {cond_{s_1} \ s}\\
-    \tab \text{then let} \ v' = \getbx{bx_1}{s}\\
-    \tab \text{else let} \ v' = \getbx{bx_2}{s}\\
-    \tab \text{fi} \ {cond_{sv_1} \ s \ v'}; \ \text{return} \ v'$
+$\getbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
+    \tab (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})$
+
+$\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
+
+$\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
+
+$\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
+    \tab \text{if} \ {cond_{s} \ s}\\
+    \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
+    \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
+    \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
 
@@
 \item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * g
-\item get$_{REV}$ = s $\to$ v * g
-\item introduce (put$_{REV}^{-1}$: s * g $\to$ s * v) and (get$_{REV}^{-1}$:v * g $\to$ s)
+\item put$_{REV}$ = s * v $\to$ s * gb
+\item get$_{REV}$ = s $\to$ v * gb
+\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
 \item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
@@
 
-$\putrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} =\\
-    \tab \text{let} \ (s_1, g_1) = \putrev{bx_1}{s_1}{v_1} \ \text{in}\\
-    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_2}{v_2} \ \text{in}\\
-        \qtab ((s_1, s_2), (g_1, g_2))$
-
-$\putrev{\rearrS{f}{bx}}{s}{v} =\\
-    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{(f \ s)}{v} \ \text{in}\\
-        \qtab (f^{-1} s_1, g_1)$
-
-$\putrev{\rearrV{g}{bx}}{s}{v} =\\
-    \tab \text{let} \ (s_1, g_1) = \putrev{bx}{s}{(g \ v)} \ \text{in}\\
-        \qtab (s_1, g_1)$
-
-$\putrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv_1} \ s \ v\\
-    \tab \text{then let} \ (s', g) = \putrev{bx_1}{s}{v} \ \text{in}\\
-    \tab \text{else let} \ (s', g) = \putrev{bx_2}{s}{v} \ \text{in}\\
-    \tab \text{fi} \ cond_{s_1} \ s'; \ \text{return} \ (s', g)$
+$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
+        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
+
+$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
+        \qtab (f_2 \ s_1, gb_1)$
+
+$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
+        \qtab (s_1, gb_1)$
+
+$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
+    \tab \text{if} \ cond_{sv} \ s \ v\\
+    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
+    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
+    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
 
 $\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab \text{let} \ (s_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
-    \tab \text{let} \ (s_2, g_2) = \putrev{bx_2}{s_1}{v} \ \text{in}\\
-    \tab \text{let} \ (s_3, g_3) = \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, (g_1, g_2, g_3))$
+    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
+    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
+        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
 
@@
 
-$\getrev{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}} = \\
-    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s_1} \ \text{in}\\
-    \tab \text{let} \ (v_2, g_2) = \getrev{bx_2}{s_2} \ \text{in}\\
-        \qtab ((v_1, v_2), (g_1, g_2))$
-
-$\getrev{\rearrS{f}{bx}}{s}{v} =\\
-    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{(f \ s)} \ \text{in}\\
-        \qtab (v_1, g_1)$
-
-$\getrev{\rearrV{g}{bx}}{s}{v} =\\
-    \tab \text{let} \ (v_1, g_1) = \getrev{bx}{s} \ \text{in}\\
-        \qtab (g^{-1} v_1, g_1)$
-
-$\getrev{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s_1} \ s\\
-    \tab \text{then let} \ (v', g) = \getrev{bx_1}{s} \ \text{in}\\
-    \tab \text{else let} \ (v', g) = \getrev{bx_2}{s} \ \text{in}\\
-    \tab \text{fi} \ cond_{sv_1} \ s \ v'; \ \text{return} \ (v', g)$
+$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
+        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
+
+$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
+        \qtab (v_1, gb_1)$
+
+$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
+        \qtab (g_2 \ v_1, gb_1)$
+
+$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
+    \tab \text{if} \ cond_{s} \ s\\
+    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
+    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
+    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
 
 $\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab \text{let} \ (v_1, g_1) = \getrev{bx_1}{s} \ \text{in}\\
-    \tab \text{let} \ (v_2, g_2) = \getrev{bx_1}{s} \ \text{in}\\
-        \qtab (v_2, (g_1, g_2))$
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
+        \qtab (v_2, gb_1 :: gb_2)$
 
@@
 It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In miniBiGUL, we can do that by tupling $put$ and $get$.
+In minBiGUL, we can do that by tupling $put$ and $get$.
 The such pair is accepted as input of a function named $pg$ to produce a new one that contains the actual result of a BX program.
@@
 
-$\pg{\product{bx_1}{bx_2}}{\tuple{s_1}{s_2}}{\tuple{v_1}{v_2}} \\
-    \tab = (\product{(\putbx{bx_1}{s_1}{v_1})}{(\putbx{bx_2}{s_2}{v_2})}, \product{(\getbx{bx_1}{s_1})}{(\getbx{bx_2}{s_2})}) \\
-    \tab = \text{let} \ (s_1, v_1) = \pg{bx_1}{s_1}{v_1} \ \text{in} \\
-        \qtab \text{let} \ (s_2, v_2) = \pg{bx_2}{s_2}{v_2} \ \text{in} \\
-        \qtab (\product{s_1}{s_2}, \product{v_1}{v_2})$
-
-$\pg{\rearrS{f}{bx}}{s}{v} \\
-    \tab = (f^{-1}(\putbx{bx}{(f \ s)}{v}), \getbx{bx}{(f \ s)}) \\
-    \tab = \text{let} \ (s, v) = \pg{bx}{f \ s}{v} \ \text{in} \\
-        \qtab (f^{-1} s, v)$
-
-$\pg{\rearrV{g}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g v)}, g^{-1} (\getbx{bx}{s})) \\
-    \tab = \text{let} \ (s, v) = \pg{bx}{s}{g \ v} \ \text{in} \\
-        \qtab (s, g^{-1} v)$\\
-
-Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrS{f}{bx}]\!]$ and $[\![\rearrV{g}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv_1}}{cond_{s_1}}{bx_1}{cond_{sv_2}}{cond_{s_2}}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv_1} \ s \ v} \\
-    \qtab \text{then let} \ s' = \putbx{bx_1}{s}{v} \\
-    \qtab \text{else let} \ s' = \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s_1} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s_1} \ s}\\
-    \qtab \quad \text{then let} \ v' = \getbx{bx_1}{s}\\
-    \qtab \quad \text{else let} \ v' = \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv_1} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv_1} \ s \ v} \ and \ {cond_{s_1} \ s} \\
-        \qtab \text{then let} \ (s', v') = \pg{bx_1}{s}{v}\\
-        \qtab \text{else let} \ (s', v') = \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{v_1} \ s' \ and \ cond_{sv_1} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{case}$ as well as $\getbxinline{case}$.
+$\pg{\product{bx_1}{bx_2}}{s_1 :: s_2}{v_1 :: v_2} \\
+    \tab = ((\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})) \\
+    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
+        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
+        \qtab \tab (s_1 :: s_2, v_1 :: v_2)$
+
+$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
+    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
+    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
+        \qtab \tab (f_2 \ s, v)$
+
+$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
+    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
+    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
+        \qtab \tab (s, g_2 \ v)$\\
+
+Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrs{f_1}{f_2}{bx}]\!]$ and $[\![\rearrv{g_1}{g_2}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
+
+$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
+    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
+    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
+    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
+    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
+    \qtab \quad \text{if} \ {cond_{s} \ s}\\
+    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
+    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
+    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
+    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
+        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
+        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
+        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
+
+A restriction for $\pginline{Case}$ needs to be introduced here.
+We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
 If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{case}$ and $\getbxinline{case}$ are not always simultaneously satified.
-The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv_1} \ s \ v) \ and \ (not (cond_{s_1} \ s)))$ and $((not (cond_{sv_1} \ s \ v) \ and \ (cond_{s_1} \ s))$, which are restricted in this paper.
-This does not happen for the others, which the construction of $\pginline{case}$ uses above.
+This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
+The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+This does not happen for the others, which the construction of $\pginline{Case}$ uses above.
 \\
 
+\textcolor{blue}{(should substitute a, b, c, d, e, f by s1, s2, v1, v2...)}\\
 $\pg{bx_1 \circ bx_2}{s}{v} \\
     \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = \text{let} \ a = \getbx{bx_1}{s} \ \text{in} \\
-        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
-        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
-        \qtab (d, c) \\
-    \tab = \text{let} \ (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
-        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
-        \qtab \text{let} \ (d, e) = \pg{bx_1}{s}{b} \ \text{in} \\
-        \qtab (d, c) \\
-    \tab = \text{let } (f, a) = \pg{bx_1}{s}{dummy} \ \text{in} \\
-        \qtab \text{let} \ (b, c) = \pg{bx_2}{a}{v} \ \text{in} \\
-        \qtab \text{let} \ (d, e) = \pg{bx_1}{\textcolor{red}{f}}{b} \ \text{in} \\
-        \qtab (d, c)$\\
+    \tab = a \Leftarrow \getbx{bx_1}{s}; \\
+        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
+        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
+            \qtab \tab (d, c) \\
+    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
+        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
+        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
+            \qtab \tab (d, c) \\
+    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
+        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
+        \qtab (d, e) \Leftarrow \pg{bx_1}{f}{b}; \\
+            \qtab \tab (d, c)$\\
 
@@
 The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $(\text{let } a = \getbx{bx_1}{s} \text{ in})$ into $(\text{let } (f, a) = \pg{bx_1}{s}{dummy} \text{ in})$.
-It is possible if we consider $(\getbx{bx_1}{s})$ as the second element of $(\pg{bx_1}{s}{dummy})$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ direction is valid.
-The last equality substitutes $s$ in $(\text{let } (d, e) = \pg{bx_1}{s}{b} \text{ in})$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
+The third one rewrites $a \Leftarrow \getbx{bx_1}{s}$ into $(f, a) \Leftarrow \pg{bx_1}{s}{dummy}$.
+It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid.
+The last equality substitutes $s$ in $(d, e) \Leftarrow \pg{bx_1}{s}{b}$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
 This transformation is feasible due to the PUTPUT law.
-Then, $e$ in the result pair $(d,e)$ of the new let-expression equals $dummy$.
+Then, $e$ in the result pair $(d,e)$ equals $dummy$.
 So we can realize that there is no loss information when computing a composition.
@@
 
+To better understand about $pg$, let's take a look at some examples. The following is the definition of $phead$ in BiGUL [?]:
+
+    $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
+    \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
+        \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
+            \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
+
+The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
+$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
+$\getbx{phead}{s_0}$ returns the head of $s_0$.
+For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
+
+In minBiGUL, we can express $phead$ as follows:
+
+    $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
+        \qtab f_1 = \lambda (s:ss).(s,ss)\\
+        \qtab f_2 = \lambda (s,ss).(s:ss)\\
+        \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
+            \qtab \qtab g_1 = \lambda v.(v,())\\
+            \qtab \qtab g_2 = \lambda (v,()).v\\
+            \qtab \qtab bx_v = Replace \ prod \ (Skip (const ()))$\\
+
+Now, we evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
+
+$\putbx{phead}{[1,2,3]}{100}\\
+    \tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
+    \tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
+    \tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
+    \tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
+    \tab = f_2 \ (\putbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{(100,())})\\
+    \tab = f_2 \ ((\putbx{Replace}{1}{100}) :: (\putbx{(Skip (const ()))}{[2,3]}{()}))\\
+    \tab = f_2 \ (100 :: [2,3])\\
+    \tab = f_2 \ (100, [2,3])\\
+    \tab = [100,2,3]$\\
+
+$\getbx{phead}{[1,2,3]}\\
+    \tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
+    \tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
+    \tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
+    \tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
+    \tab = g_2 \ (\getbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])})\\
+    \tab = g_2 \ ((\getbx{Replace}{1}) :: (\getbx{(Skip (const ()))}{[2,3]}))\\
+    \tab = g_2 \ (1 :: ())\\
+    \tab = g_2 \ (1,())\\
+    \tab = 1$\\
+
+By the definition of $pg$, we have:
+$\pg{phead}{[1,2,3]}{100}\\
+    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
+    \tab = ([100,2,3], 1)$\\
+
+On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
+$\pg{phead}{[1,2,3]}{100}\\
+    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
+    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \{\\
+        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
+        \qtab \tab (s, g_2 \ v)\\
+        \qtab \};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \{\\
+        \qtab \tab (s, v) \Leftarrow \pg{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{((100,()))};\\
+        \qtab \tab (s, g_2 \ v)\\
+        \qtab \};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \{\\
+        \qtab \tab (s, v) \Leftarrow \{\\
+            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
+            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip (const ()))}{[2,3]}{()};\\
+            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
+        \qtab \tab \};\\
+        \qtab \tab (s, g_2 \ v)\\
+        \qtab \};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \{\\
+        \qtab \tab (s, v) \Leftarrow \{\\
+            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
+            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
+            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
+        \qtab \tab \};\\
+        \qtab \tab (s, g_2 \ v)\\
+        \qtab \};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow \{\\
+        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
+        \qtab \tab (s, g_2 \ v)\\
+        \qtab \};\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
+        \qtab \tab (f_2 \ s, v)\\
+    \tab = ([100,2,3], 1)$\\
+
+If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
+
+    \tab $\putbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]}{100} = [[100,2,3],[\ ],[4,5]]$\\
+    \tab $\getbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]} = 1$\\
+
+We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
+
+$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}\\
+    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}\\
+    \tab = \getbx{phead_2}{[1,2,3]}\\
+    \tab = 1$\\
+
+$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}{100}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}{100})}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{[100,2,3]}\\
+    \tab = [[100,2,3],[\ ],[4,5]]$\\
+
+We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
+
+$\pg{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}{100}\\
+    \tab = (f, a) \Leftarrow \pg{phead_1}{[[1,2,3],[\ ],[4,5]}{dummy};\\
+        \qtab (b, c) \Leftarrow \pg{phead_2}{a}{100};\\
+        \qtab (d, e) \Leftarrow \pg{phead_1}{f}{b};\\
+            \qtab \tab (d, c)\\
+    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
+        \qtab (b, c) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
+        \qtab (d, e) \Leftarrow \pg{phead_1}{f}{b};\\
+            \qtab \tab (d, c)\\
+    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
+        \qtab (b, c) \Leftarrow ([100,2,3], 1);\\
+        \qtab (d, e) \Leftarrow \pg{phead_1}{[dummy,[\ ],[4,5]]}{[100,2,3]};\\
+            \qtab \tab (d, c)\\
+    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
+        \qtab (b, c) \Leftarrow ([100,2,3], 1);\\
+        \qtab (d, e) \Leftarrow ([[100,2,3],[\ ],[4,5]], dummy);\\
+            \qtab \tab (d, c)\\
+    \tab = ([[100,2,3],[\ ],[4,5]], 1)$\\
+
+In the above evaluation, we do not clearly define what $dummy$ is.
+As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
+In this case, if we care about the type of $dummy$, it should be a list of integers.
+
+\textcolor{red}{
     \begin{itemize}
@@
     \item For this, we require the programmer to write a function to produce view from source. (This can be the same with get??)
-\item Example program: \textcolor{red}{Good example?}
-  Evaluate: $\putbx{bx}{s}{v}$ 
-  Evalaute: $\getbx{bx}{s}$
-  $\pg{bx}{s}{v} = $
     \end{itemize}
+}
 
@@
 
-\subsection{cpg}
-
-\begin{itemize}
-\item If the given BX programs are left assoc of composition, the recursive call of pg will be exponential. Because it calls $\pg{bx_1}{\_}{\_}$ twice in compose.
-\item To solve this problem, we introduce accumulation of changes.
-\end{itemize}
-
-$fst = \lambda (Con (x_1, x_2)). x_1$\\
-$snd = \lambda (Con (x_1, x_2)). x_2$\\
-$con = \lambda ks_1. \lambda ks_2. \lambda x. Con(ks_1 \ x, ks_2 \ x)$\\
-
-$\cpg{bx}{ks}{kv}{s}{v}$ \\
-    \tab $ks, kv$ keep modification information\\
-    \tab $s, v$ keep evaluated values\\
-
-\textcolor{red}{
-    $s \Leftarrow s_0; v \Leftarrow v_0;\\
-    (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
-        \tab (s; v)$
-}
-
-\vspace{3mm}
+\section{cpg}
+
+When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
+If a given BX program is a left-associative composition, the number of $pg$ calls will be exponential.
+Therefore, the runtime inefficiency is inevitable.
+
+In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
+$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
+The output of this function is a 4-tuple $(ks, kv, s, v)$.
+
+To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we will use some following utility functions:
+
+$\tab fst = \lambda (x_1 :: x_2). x_1\\
+ \tab snd = \lambda (x_1 :: x_2). x_2\\
+ \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x) :: (ks_2 \ x))$\\
+
+The construction of $cpg$ is as follows:\\
 
@@
     \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, Con \ s_1 \ s_2, Con \ v_1 \ v_2)$\\
+    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, s_1 :: s_2, v_1 :: v_2)$\\
 
@@
 
-$\cpg{Case \ c \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ c \ s \ v \ \text{then}\\
-        \qtab \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \\ 
-        \qtab \cpg{bx_2}{ks}{kv}{s}{v}$\\
+$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
+    \tab \text{if} \ cond_{sv} \ s \ v\\
+    \tab \text{then} \ (ks, kv, s, v) \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
+    \tab \text{else} \ (ks, kv, s, v) \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
+    \tab \text{fi} \ cond_{s} \ s; \ \text{return} \ (ks, kv, s, v)$\\
 
@@
 
-\begin{itemize}
-\item This is faster than pg for left assoc BX programs.
-\item (Slower for right assoc programs)
-\item $s_1$ in composition is redundant.  
-\item Next, we will extend this to avoid this redundant computation.
-\end{itemize}
-
-\subsection{kpg}
-
-\begin{itemize}
-\item In cpg, $s_1'$ in composition is not used, but evaluated.
-\item Next, we will extend cpg to avoid this redundant computation.
-\item For this, we evaluate the values lazily.
-\item []
-\item s' and v' will have actual value only in Skip and case. Except them, the functions for computation will be kept as ks' and kv'
-\item Skip and Case need values of the time. They evaluate s and v with ks' and kv', and ks' and kv' will be reseted to an identity function.
-\item A special thing is in prod. In prod, if we omit red parts,
-Evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. Because this causes redundant re-evaluation of ks' and kv', we insert evaluation before recursive calls.
-\item Except them, the functions for computation will be kept as ks' and kv'.
-\item In composition $ks_1'$ and $s_1'$ are still redundant. But, they are not evaluated.
-\item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg)
-\end{itemize}
-
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$ \\
-    \tab $ks, kv$ keep modification information\\
-    \tab $ks', kv'$ for lazy evaluation of actual values\\
-    \tab $s, v$ keep evaluated values\\
-
-\textcolor{red}{
-    $s \Leftarrow s_0; v \Leftarrow v_0;\\
-    (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \tab (ks' \ s; kv' \ v)$
-}
-
-\vspace{3mm}
+Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates on accumulative functions $ks$ and $kv$.
+
+For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
+The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, \underline{s_1}, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
+Some computations to get $s_1$ are seen as redundant.
+Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
+The second $cpg$ call uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
+After two first $cpg$ calls, we used a function application, $ks_1 \ s_2$, to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
+
+\textcolor{blue}{(how to explain parameters of $cpg$ in the composition?)}
+
+Suppose that we have a source $s_0$ and a view $v_0$.
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
+
+    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
+        \qtab (s; v)$
+
+\textcolor{blue}{(some explanation for parameters $\lambda \_.s, \lambda \_.v$)}
+
+\section{kpg}
+
+In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
+Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
+While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
+Every values are evaluated lazily in a computation of $kpg$.
+The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$, $kv$ and $s$, $v$ are still used to keep the modifcation information and evaluated values respectively, $ks'$ and $kv'$ are used for lazy evaluation of actual values.
+The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
+
+Suppose that we have a source $s_0$ and a view $v_0$.
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
+
+    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
+    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+        \qtab (ks' \ s; kv' \ v)$
+
+\textcolor{blue}{(some explanation for parameters $\lambda \_.s, id, id, id$)}\\
+
+The construction of $kpg$ is as follows:\\
 
@@
 $\kpg{bx_1 \ prod \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab \textcolor{red}{s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;}\\
-    \tab \textcolor{red}{ks' \Leftarrow id; \quad kv' \Leftarrow id;}\\
+    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
+    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
     \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
@@
     \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (Con \ s_1 \ s_2), (Con \ v_1 \ v_2))$\\
+    \tab s_1 :: s_2, v_1 :: v_2)$\\
 
@@
 
-$\kpg{Case \ c \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
     \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
     \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ c \ s \ v \ \text{then}\\
-        \qtab \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else}\\\
-        \qtab \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}$\\
+    \tab \text{if} \ cond_{sv} \ s \ v \\
+    \tab \text{then} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
+    \tab \text{else} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
+    \tab \text{fi} \ cond_s \ (ks' \ s); \ \text{return} \ (ks, kv, ks', kv', s, v)$\\
 
@@
 
-\begin{itemize}
-    \item In this definition of kpg, s and v keep actual values only in skip and case. Except them, the functions for computation will be kept as ks' and kv'.
-    \item In composition $ks_1'$ and $s_1$ are still redundant. But, they are not evaluated.
-    \item In prod, evaluation of ks' and kv' will be done indendently in kpg$[\![bx_1]\!]$ and kpg$[\![bx_2]\!]$. To remove redundant re-computation we evaluate actual s and v.
-    \item This implementation is faster for left assoc BX programs. (Actually slower for right assoc than cpg). \textcolor{red}{But it makes stack\_overflow comes sooner}.
-\end{itemize}
-
-\subsection{xpg}
+In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept as $ks'$ and $kv'$.
+
+When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
+
+In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+It is possible to evaluate actual value in $s$ and $v$ to remove redundant recomputations.
+
+\section{xpg}
 
@@
     $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-    \tab phead = \$(rearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-        \qtab \qtab \quad \$(rearrV \ [\![\lambda(v) \to (v,())]\!])\$\\
+    \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
+        \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
             \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
@@
     \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(rearrV [| \lambda v \to ((),v) |])\$\\
+        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
             \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
     \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(rearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
+        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
             \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
@@
         \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(rearrS \ [\![ \lambda s \to (s,[\ ]) ]\!])\$\\
+        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\ ]) ]\!])\$\\
             \qtab \qtab \qquad bFoldr \ bSnoc\\

### 2019-09-05 paper/main.tex
update tex 0509-1439

@@
 \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-While $Replace$ is a primitive BX program, $phead$ is a combination of such programs.
-The following is the definition of $phead$ in BiGUL [?]:
-
-    $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-    \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-        \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-            \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-We also provide the same $phead$ program in our OCaml system.
-It rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
-For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]}{100} = [[100,2,3],[\ ],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]} = 1$
-
+
+In section 5, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
 Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.

### 2019-09-08 paper/main.tex
only update formulas !

@@
 
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}
-
 \section{pg}
@@
 
-\textcolor{blue}{(should substitute a, b, c, d, e, f by s1, s2, v1, v2...)}\\
 $\pg{bx_1 \circ bx_2}{s}{v} \\
     \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = a \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
-        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
-            \qtab \tab (d, c) \\
-    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
-        \qtab (d, e) \Leftarrow \pg{bx_1}{s}{b}; \\
-            \qtab \tab (d, c) \\
-    \tab = (f, a) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (b, c) \Leftarrow \pg{bx_2}{a}{v}; \\
-        \qtab (d, e) \Leftarrow \pg{bx_1}{f}{b}; \\
-            \qtab \tab (d, c)$\\
+    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
+        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
+        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
+            \qtab \tab (s_3, v_2) \\
+    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
+        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
+        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
+            \qtab \tab (s_3, v_2) \\
+    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
+        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
+        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
+            \qtab \tab (s_3, v_2)$\\
 
@@
 The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $a \Leftarrow \getbx{bx_1}{s}$ into $(f, a) \Leftarrow \pg{bx_1}{s}{dummy}$.
+The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
 It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid.
-The last equality substitutes $s$ in $(d, e) \Leftarrow \pg{bx_1}{s}{b}$ by $f$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
+The last equality substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
 This transformation is feasible due to the PUTPUT law.
-Then, $e$ in the result pair $(d,e)$ equals $dummy$.
+Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
 So we can realize that there is no loss information when computing a composition.
@@
 $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (ks, kv, s, v) \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s, v) \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s; \ \text{return} \ (ks, kv, s, v)$\\
+    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
+    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
+    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
+    \tab \text{fi} \ cond_{s} \ s' \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
 
 $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
     \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
@@
 
-\textcolor{blue}{(how to explain parameters of $cpg$ in the composition?)}
-
 Suppose that we have a source $s_0$ and a view $v_0$.
@@
     $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{\lambda \_.v}{s}{v};\\
+    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
         \qtab (s; v)$
-
-\textcolor{blue}{(some explanation for parameters $\lambda \_.s, \lambda \_.v$)}
 
@@
         \qtab (ks' \ s; kv' \ v)$
-
-\textcolor{blue}{(some explanation for parameters $\lambda \_.s, id, id, id$)}\\
 
@@
     \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \\
-    \tab \text{then} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s); \ \text{return} \ (ks, kv, ks', kv', s, v)$\\
+    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
+    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
+    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
+    \tab \text{fi} \ cond_s \ (ks' \ s') \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
 
 $\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{const$\_$dummy \ v};\\
+    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
     \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2', ks_1 \circ ks_2', kv_2', s_2, v_2)$\\
+    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$\\
 
@@
 $\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
     \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
@@
 
-In section 5, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
+In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
 Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
@@
 %
+
+\section{Remove?:Reversible langauge: Rwhile}
+
+% Maybe we should skip
+
+\begin{itemize}
+\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
+\item Using reversible language, we can obtain inverse programs easily.
+\item Explanation of Rwhile: we need fi-condition.
+\item syntax, how to work.
+\end{itemize}
+
+\section{Remove?:First approach}
+
+% We should skip this
+
+\begin{itemize}
+\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
+\item put$_{REV}$ = s * v $\to$ s * gb
+\item get$_{REV}$ = s $\to$ v * gb
+\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
+\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
+\end{itemize}
+
+$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
+
+$\putrev{Replace}{s}{v} = (v, s)$
+
+$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
+        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
+
+$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
+        \qtab (f_2 \ s_1, gb_1)$
+
+$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
+        \qtab (s_1, gb_1)$
+
+$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
+    \tab \text{if} \ cond_{sv} \ s \ v\\
+    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
+    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
+    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
+
+$\putrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
+    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
+        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
+
+\vspace{5mm}
+
+$\getrev{Skip \ h}{s} = {(h \ s, s)}$
+
+$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
+
+$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
+        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
+
+$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
+        \qtab (v_1, gb_1)$
+
+$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
+        \qtab (g_2 \ v_1, gb_1)$
+
+$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
+    \tab \text{if} \ cond_{s} \ s\\
+    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
+    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
+    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
+
+$\getrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
+        \qtab (v_2, gb_1 :: gb_2)$
+
+\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-10 paper/main.tex
update tex 20190910-7.55
+ explanation for parameters

@@
     put(s, dummy) can be a complement function.
-  \end {itemize}
+  \end{itemize}
 
@@
 
-\section{Bidirectional Programming Language: BiGUL}
+\section{Bidirectional Programming Language: minBiGUL}
 
@@
 
+\subsection{Syntax}
+
+$data := Unit \\ 
+    \qtab | \ Int \ x \\ 
+    \qtab | \ Con(data_1, data_2)$\\
+
+$bx := Skip \ h \\ 
+    \qtab | \ Replace \\ 
+    \qtab | \ bx_1 \ prod \ bx_2 \ \\ 
+    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
+    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
+    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
+    \qtab | \ bx_1 \circ bx_2$
+
 \subsection{Semantics of minBiGUL}
@@
 \begin{itemize}
-\item To implement this in OCaml, we need some way to produce inverse function for $f$ and $g$.
-\item Avoiding this, our implementation requires users to write inverse function f$^{-1}$ and g$^{-1}$.
+\item To implement these function in OCaml, we need some way to produce inverse function for $f$ and $g$.
+\item $f_2 = f_1^{-1}, g_2 = g_1^{-1}$
 \end{itemize}
@@
 \subsection{Self-inverse function: pg}
-
-%\begin{itemize}
-%\item Information loss causes re-evaluation of get.
-%\item To prevent this, we introduced the idea of reversible computation.
-%\item In reversible computation we need to keep all information: we can not lose any information.
-%\item In BiGUL we can acheive this by tupling put and get.
-%\item pg bx (s, v) = (put bx (s, v), get bx (s))
-%\item pg: s * v $\to$ s * v
-%\item No garbage
-%\item This function is self-inverse. (involution, (f (f input)) = input)
-%\item \textcolor{red}{Constraint: In case expression, put and get must always choose the same branch.
-This constraint is not satisfied in general: (e.g, SNOC, the different length of two inputs.) To remove this constraint we need to use garbage.}
-%\end{itemize}
 
@@
 
-The construction of $cpg$ is as follows:\\
+\begin{definition}
+$\cpg{bx}{ks}{kv}{s}{v}$
 
@@
     \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
+    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
 
 $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
+    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
     \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$\\
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates on accumulative functions $ks$ and $kv$.
+        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
+\end{definition}
+
+Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
 
 For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, \underline{s_1}, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
+The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
+Because there is no real view here, we need to construct a dummy from the source.
+Then the continuation updating on this dummy should be initiated as an identity function.
+The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
 Some computations to get $s_1$ are seen as redundant.
 Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-The second $cpg$ call uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-After two first $cpg$ calls, we used a function application, $ks_1 \ s_2$, to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
+In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
+It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
+After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
 
@@
         \qtab (s; v)$
+
+In general, the begining of a continuation should be an identity function.
+However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
 
@@
 
-The construction of $kpg$ is as follows:\\
+The begining of accumulative functions $kv$, $ks'$, $kv'$ are set as identity, while $ks$ is initiated as a constant function over the source for the same reason as in $cpg$.
+
+\begin{definition}
+$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
 
@@
     \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$\\
+    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
+\end{definition}
 
@@
 In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-It is possible to evaluate actual value in $s$ and $v$ to remove redundant recomputations.
+It is possible to evaluate actual values in $s$ and $v$ to remove redundant recomputations.
 
@@
 
-\begin{itemize}
-\item So far, we focus on left assoc programs. xpg is faster than the others.
-\item On the other hand, pg is fastest for right assoc programs.
-\item To combine their strength, we introfuce new pg that combines pg and kpg.
-\item If there is a composition, we use xpg for evaluation of $bx_1$ and pg for evaluation of $bx_2$.
-\item If $bx_1$ is simple program (= Right assoc), this is almost the same with pg.
-\item If $bx_1$ is complicated program (= Left assoc), this is almost the same with kpg.
-\end{itemize}
-
-$\xpg{E}{s}{v} = \text{same with } pg \text{ if } E \text{ is not compose}$
+So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
+To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
+
+\begin{definition}
+$\xpg{bx}{s}{v}$
+
+$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \text{ is not compose}$
 
@@
     \qtab (ks_1 \ s_2, v_2)$
+\end{definition}
+
+Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the updated pair.
+The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
+Here, $xpg$ is called recursively instead of $pg$.
+For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
+The first call and the function application are come from $kpg$ approaches, while the second call is based on $pg$ approach.
 

### 2019-09-13 paper/main.tex
update tex 20190913-7.53
update from minbigul -> experiment

@@
 \newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \ prod \ #2}
+\newcommand{\product}[2]{#1 \times #2}
 \newcommand{\tuple}[2]{(#1 :: #2)}
@@
 
-\begin{itemize}
-\item our target language is minBiGUL (BiGUL minus adaptive case) : very well-behaved
-\item Basic explanation of bidirectional transformation
-  \begin{itemize}
-  \item get = s $\to$ v
-  \item put = s * v $\to$ s
-  \item explanation about get and put.
-  \item We assume putget and getput hold (explanation about putget and getput)
-  \item Because our target language satisfies very well-behaved (No adaptive case) $\to$ putput hold.
-  \end{itemize}
-\item explanation about put-based BX language: BiGUL
-\end{itemize}
+MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
+BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
+Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
+    $\tab \getbxinline{bx} = fun \ s \to v \\
+    \tab \putbxinline{bx} = fun \ s \ v \to v$
+
+BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
+    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
+    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
+
+The GETPUT law means that if there is no change to the view, there should be no change to the source.
+The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
+
+MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
+In addition, we restrict adaptive cases of BiGUL on minBiGUL.
+Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
+    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
+
+The PUTPUT law means that a source update should overwrite the effect of previous source updates.
+
+Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
+
+In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
 
@@
 
-$data := Unit \\ 
-    \qtab | \ Int \ x \\ 
-    \qtab | \ Con(data_1, data_2)$\\
+The syntax of minBiGUL is briefly written as follows:
 
@@
     \qtab | \ Replace \\ 
-    \qtab | \ bx_1 \ prod \ bx_2 \ \\ 
+    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
     \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
@@
     \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ bx_1 \circ bx_2$
-
-\subsection{Semantics of minBiGUL}
+    \qtab | \ Compose \ bx_1 \ bx_2$
+
+A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
+
+For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
+It is a kind of bijection.
+However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
+In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
+
+To help make demonstration more direct, we provide the following alternatives representation:
+
+    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
+    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
+
+In general, $\circ$ has a higher priority than $\times$.
+Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
+The one of $\circ$ may be left or right, however we do not declare the default.
+
+In the next part, we will introduce the semantics of minBiGUL.
+
+\subsection{Semantics}
+
+\begin{definition}
+    $\putbx{bx}{s}{v}$
 
@@
 
-$\putbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2})$
+    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
+        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
 
@@
     $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-
-\vspace{5mm}
+\end{definition}
+
+\begin{definition}
+    $\getbx{bx}{s}$
 
@@
 
-$\getbx{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})$
+    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
+        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
 
@@
     $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-
-\begin{itemize}
-\item To implement these function in OCaml, we need some way to produce inverse function for $f$ and $g$.
-\item $f_2 = f_1^{-1}, g_2 = g_1^{-1}$
-\end{itemize}
+\end{definition}
+
+In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
+Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
+$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
+If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
+We can write the equivalent if-then-else statement as follows:
+
+$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
+\tab \equiv \text{if }E_1 = true \text{ then}\\ 
+    \qtab \tab X_1;\\ 
+    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
+    \qtab \text{else}\\ 
+    \qtab \tab X_2;\\ 
+    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
+
+The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
+
+Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
+$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
+$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
+$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
+$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
+$\putbx{Replace}{1}{100} = 100$\\
+$\getbx{Replace}{1} = 1$\\
+
+Now, let's consider the definition of $phead$ in BiGUL [?]:
+
+$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
+\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
+    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
+        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
+
+The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
+$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
+$\getbx{phead}{s_0}$ returns the head of $s_0$.
+For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
+
+In minBiGUL, we can express $phead$ as follows:
+
+$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
+    \qtab f_1 = \lambda (s::ss).(s,ss)\\
+    \qtab f_2 = \lambda (s,ss).(s::ss)\\
+    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
+        \qtab \qtab g_1 = \lambda v.(v,())\\
+        \qtab \qtab g_2 = \lambda (v,()).v\\
+        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
+
+Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
+
+$\putbx{phead}{[1,2,3]}{100}\\
+\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
+\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
+\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
+\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
+\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
+\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
+\tab = f_2 \ (100, [2,3])\\
+\tab = [100,2,3]$\\
+
+$\getbx{phead}{[1,2,3]}\\
+\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
+\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
+\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
+\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
+\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
+\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
+\tab = g_2 \ (1,())\\
+\tab = 1$\\
+
+If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
+
+    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
+    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
+
+We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
+
+$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
+    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
+    \tab = \getbx{phead_2}{[1,2,3]}\\
+    \tab = 1$\\
+
+$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
+    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
+    \tab = [[100,2,3],[\,],[4,5]]$\\
+
+One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
+$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
+\tab replaceAll = Case \ [\\
+    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
+    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
+    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
+    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
+    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
+    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
+\tab ]$\\
+
+The following expresses $replaceAll$ in minBiGUL:\\
+$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
+    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
+    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
+    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
+    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
+        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
+            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
+
+We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
+$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
+For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
 
@@
 
-For the approaches so far, when evaluating a composition of BX programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
+In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
 This is a kind of information loss.
-One question is whether it is possible to calculate the above programs without losing information.
+One question is whether it is possible to calculate such programs without losing information.
 And the answer is yes.
@@
 In minBiGUL, we can do that by tupling $put$ and $get$.
-The such pair is accepted as input of a function named $pg$ to produce a new one that contains the actual result of a BX program.
+A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
 
@@
 
-$\pg{\product{bx_1}{bx_2}}{s_1 :: s_2}{v_1 :: v_2} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}) :: (\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}) :: (\getbx{bx_2}{s_2})) \\
+$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
+    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
     \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
         \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab (s_1 :: s_2, v_1 :: v_2)$
+        \qtab \tab ((s_1,s_2), (v_1,v_2))$
 
@@
 
-Constructions of $pg$ for $[\![Replace]\!]$, $[\![\product{bx_1}{bx_2}]\!]$, $[\![\rearrs{f_1}{f_2}{bx}]\!]$ and $[\![\rearrv{g_1}{g_2}{bx}]\!]$ are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
+Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
 
@@
 This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated BXs are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others, which the construction of $\pginline{Case}$ uses above.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
 \\
@@
 The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that make the $put \, [\![bx_1]\!]$ valid.
-The last equality substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{dummy}$.
-This transformation is feasible due to the PUTPUT law.
+It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
+Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
+Then it is feasible to construct a dummy from the source.
+The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
+This transformation is possible due to the PUTPUT law.
 Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
@@
 \subsection{Evaluation by pg}
-
-To better understand about $pg$, let's take a look at some examples. The following is the definition of $phead$ in BiGUL [?]:
-
-    $\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-    \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-        \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-            \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-    $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-        \qtab f_1 = \lambda (s:ss).(s,ss)\\
-        \qtab f_2 = \lambda (s,ss).(s:ss)\\
-        \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-            \qtab \qtab g_1 = \lambda v.(v,())\\
-            \qtab \qtab g_2 = \lambda (v,()).v\\
-            \qtab \qtab bx_v = Replace \ prod \ (Skip (const ()))$\\
-
-Now, we evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-    \tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-    \tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-    \tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-    \tab = f_2 \ (\putbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{(100,())})\\
-    \tab = f_2 \ ((\putbx{Replace}{1}{100}) :: (\putbx{(Skip (const ()))}{[2,3]}{()}))\\
-    \tab = f_2 \ (100 :: [2,3])\\
-    \tab = f_2 \ (100, [2,3])\\
-    \tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-    \tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-    \tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-    \tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-    \tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-    \tab = g_2 \ (\getbx{Replace \ prod \ (Skip (const ()))}{(1,[2,3])})\\
-    \tab = g_2 \ ((\getbx{Replace}{1}) :: (\getbx{(Skip (const ()))}{[2,3]}))\\
-    \tab = g_2 \ (1 :: ())\\
-    \tab = g_2 \ (1,())\\
-    \tab = 1$\\
 
@@
     \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{Replace \ prod \ (Skip (const ()))}{(1,[2,3])}{((100,()))};\\
+        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
         \qtab \tab (s, g_2 \ v)\\
@@
             \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip (const ()))}{[2,3]}{()};\\
-            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
+            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
+            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
         \qtab \tab \};\\
@@
             \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab (s_1 :: s_2, v_1 :: v_2)\\
+            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
         \qtab \tab \};\\
@@
 
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]}{100} = [[100,2,3],[\ ],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\ ],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\ ],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\ ],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\ ],[4,5]]$\\
-
 We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
 
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\ ],[4,5]]}{100}\\
-    \tab = (f, a) \Leftarrow \pg{phead_1}{[[1,2,3],[\ ],[4,5]}{dummy};\\
-        \qtab (b, c) \Leftarrow \pg{phead_2}{a}{100};\\
-        \qtab (d, e) \Leftarrow \pg{phead_1}{f}{b};\\
-            \qtab \tab (d, c)\\
-    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
-        \qtab (b, c) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (d, e) \Leftarrow \pg{phead_1}{f}{b};\\
-            \qtab \tab (d, c)\\
-    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
-        \qtab (b, c) \Leftarrow ([100,2,3], 1);\\
-        \qtab (d, e) \Leftarrow \pg{phead_1}{[dummy,[\ ],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (d, c)\\
-    \tab = (f, a) \Leftarrow ([dummy,[\ ],[4,5]], [1,2,3]);\\
-        \qtab (b, c) \Leftarrow ([100,2,3], 1);\\
-        \qtab (d, e) \Leftarrow ([[100,2,3],[\ ],[4,5]], dummy);\\
-            \qtab \tab (d, c)\\
-    \tab = ([[100,2,3],[\ ],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly define what $dummy$ is.
+$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
+    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
+        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
+        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
+            \qtab \tab (s_3, v_2)\\
+    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
+        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
+        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
+            \qtab \tab (s_3, v_2)\\
+    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
+        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
+        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
+            \qtab \tab (s_3, v_2)\\
+    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
+        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
+        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
+            \qtab \tab (s_3, v_2)\\
+    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
+
+In the above evaluation, we do not clearly contruct what $dummy$ is.
 As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
 In this case, if we care about the type of $dummy$, it should be a list of integers.
-
-\textcolor{red}{
-    \begin{itemize}
-    \item In pg approach we always need to supply source and view.
-    \item If we would like to evaluate get, we do not have any view.
-    \item For this, we require the programmer to write a function to produce view from source. (This can be the same with get??)
-    \end{itemize}
-}
-
-%\begin{itemize}
-%\item Because we combine put and get, we need to supply view for evaluation of get.
-%\item introduce dummy?
-%\item automatic analysis for domain \& range?
-%\item 
-%\item partial evaluation of put(s(D), dummy(S)).
-%\item \textcolor{red}{Not good?
-Basically put (BX) is often a recursion on the source.
-If source is static, partial evaluation will work well.
-On the other hand, view is small data.
-So partial evaluation might not work well.}
-%\item \textcolor{red}{lazy evaluation is also a possibility?}
-%\end{itemize}
+Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
 
@@
 When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given BX program is a left-associative composition, the number of $pg$ calls will be exponential.
+If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
 Therefore, the runtime inefficiency is inevitable.
@@
 
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we will use some following utility functions:
-
-$\tab fst = \lambda (x_1 :: x_2). x_1\\
- \tab snd = \lambda (x_1 :: x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x) :: (ks_2 \ x))$\\
+To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
+
+$\tab fst = \lambda (x_1,x_2). x_1\\
+ \tab snd = \lambda (x_1,x_2). x_2\\
+ \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
 
@@
     \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, s_1 :: s_2, v_1 :: v_2)$\\
+    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
 
@@
 
+Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
+After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
+    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
+    \tab s_2 = [100,2,3]$\\
+where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
+Then:\\ 
+$\tab ks_1 \ s_2 \\ 
+\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
+\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
+\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
+\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
+
+If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
+This is an unexpected result when we target it to be the updated source.
+If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
+This time, the result is what we want to see.
+
+Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
+
 \section{kpg}
@@
 Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$, $kv$ and $s$, $v$ are still used to keep the modifcation information and evaluated values respectively, $ks'$ and $kv'$ are used for lazy evaluation of actual values.
+The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
 The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
@@
 
-The begining of accumulative functions $kv$, $ks'$, $kv'$ are set as identity, while $ks$ is initiated as a constant function over the source for the same reason as in $cpg$.
+The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
 
@@
 
-$\kpg{bx_1 \ prod \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
     \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
@@
     \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab s_1 :: s_2, v_1 :: v_2)$\\
+    \tab (s_1, s_2), (v_1,v_2))$\\
 
@@
     \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
+    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
 
@@
 
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept as $ks'$ and $kv'$.
+In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
 
@@
 
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-It is possible to evaluate actual values in $s$ and $v$ to remove redundant recomputations.
+In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
+It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
 
@@
 
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \text{ is not compose}$
+$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
 
@@
 
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the updated pair.
+Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
 The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Here, $xpg$ is called recursively instead of $pg$.
+Note that, $xpg$ is called recursively instead of $pg$.
 For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application are come from $kpg$ approaches, while the second call is based on $pg$ approach.
+The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
 
@@
 The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace} looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace} is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
+For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
 \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
@@
 Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ direction on the composition valid.\\
+For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
 
@@
     $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\ ] = e\\
+    \tab foldr \ f \ e \ [\,] = e\\
     \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
@@
 We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\ ] = [\ ]\\
+    $\tab reverse \ [\,] = [\,]\\
     \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\ ]$
+    \tab reverse = foldr \ snoc \ [\,]$
 
@@
         \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\ ]) ]\!])\$\\
+        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
             \qtab \qtab \qquad bFoldr \ bSnoc\\

### 2019-09-13 paper/main.tex
add first version of introduction

@@
 %
-\title{An optimization of bidirectional update language based on the idea of reversible computing
- \thanks{Supported by organization x.}}
+\title{An efficient composition of bidirectional programs
+  by tupling and lazy updates 
+%  tupling
+%  accumulating updates = lazy updates
+%  based on the idea of reversible computing  
+  % \thanks{Supported by organization x.}
+}
 %
@@
 \begin{itemize}
-\item When will be the deadline of PEPM2020? (mid of October?)
+\item Schedule:
+\item Introduction (story), Contribution, positions, examples, beginning of September?
+\item zero draft : mid of September
+
+\item When will be the deadline of PEPM2020 (October 18th)
+\item ESOP (?)
 \item Need to change the style file of latex..
@@
 
-\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-\item Related work: suggest?
-\item another pg ?? should be compared?
+\item Related work?
+\item
+
+\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
+%\item right assoc \& Recursive example?
+%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
+%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
+%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
+%\item Related work: suggest?
+%\item another pg ?? should be compared?
 \end{itemize}
 
-\begin{tabular}{c|c|c}
-  {} & {Not to keep eval info} & {Keep eval info as a function}\\ \hline
-  {eager evaluation} & {pg} & {cpg}\\ \hline
-  {lazy evaluation} & {??} & {kpg}
-\end{tabular}
-
 \section{Introduction}
+
+%\begin{itemize}
+%\item Importance of BX, BX is a solution of view update problem in database.
+%\item goodness of 
+%\item Explanation of put-based BX: BiGUL.
+%\item Current status of BiGUL: Fastest BX language in the world
+%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
+%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
+%  an idea from reversible computation: not to lose any information.
+%\item We extend pg with several ideas to produce faster implementation.
+%\item 
+%\end{itemize}
+
+%\begin{itemize}
+%\item Importance of BX, BX is a solution of view update problem in database.
+%\item goodness of 
+%\item Explanation of put-based BX: BiGUL.
+%\item Current status of BiGUL: Fastest BX language in the world
+%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
+%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
+%  an idea from reversible computation: not to lose any information.
+%\item We extend pg with several ideas to produce faster implementation.
+%\item 
+%\end{itemize}
+
+In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
+% A Survey to View Update Problem 
+As a solution for this problem, bidirectional transformation is introduced \cite{}.
+It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
+% The first paper of bidirectional transformation
+
+There are two types of bidirectional transformation.
+One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
+A problem of this approach is that it is possible to derive many correct $put$s.
+To choose one put, hulistics and programmer's intentions are needed.
+Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
+From a $put$ we can derive an unique $get$.
+
+BiGUL \cite{} is one of put-based bidirectional transformation language.
+It is fastest bidirectional language in the world now.
+However, it contains a problem when the BiGUL programs including many compositions.
+BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
+In this case the number of get is quadratic.
+
+\vspace{5mm}
+
+Figure that shows many gets re-evaluation
+
+\vspace{5mm}
+
+To solve this problem and make BiGUL evaluation efficient, we use two techniques.
+The first technique is tupling, well-known technique for efficiency \cite{}.
+Its basic idea is to reduce the common computations by combining two evaluations.
+In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
+Thanks to tupling, we do not lose information like intermediate states.
+The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
+
+The rest of this paper is the following.
+We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
+Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
+We apply three optimizations to $pg$ in sequence.
+The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
+The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
+The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
+In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
+In Section 8 we discuss about related work, and conclude in Section 9.
 
@@
 
-\begin{itemize}
-\item Importance of BX, BX is a solution of view update problem in database.
-\item Explanation of put-based BX: BiGUL.
-\item Current status of BiGUL: Fastest BX language in the world
-\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
+%\begin{itemize}
+%\item Importance of BX, BX is a solution of view update problem in database.
+%\item Explanation of put-based BX: BiGUL.
+%\item Current status of BiGUL: Fastest BX language in the world
+%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
+%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
 %  an idea from reversible computation: not to lose any information.
-\item We extend pg with several ideas to produce faster implementation.
-\item 
-\end{itemize}
+%\item We extend pg with several ideas to produce faster implementation.
+%\item 
+%\end{itemize}
 
@@
   \end{itemize}
-\item Reveal a part of relationship between BX and Reversible.
+\item Tupling is also effective in bidirectional programming.
   \begin{itemize}
-  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
   \item A pair of put and get is self-inverse (involution).
-  \item If we can obtain good parameter (dummy) for a BiGUL program,
-    put(s, dummy) can be a complement function.
+  \end{itemize}
+\item lazy update \& lazy computation is also good optimization for BX.
+%\item Reveal a part of relationship between BX and Reversible.
+%  \begin{itemize}
+%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
+%  \item 
+%  \item If we can obtain good parameter (dummy) for a BiGUL program,
+%    put(s, dummy) can be a complement function.
   \end{itemize}

### 2019-09-13 paper/main.tex
fix an error
change on schedule & points to discuss

@@
 
-Points to discuss
+Schedule
 \begin{itemize}
-\item Schedule:
+
 \item Introduction (story), Contribution, positions, examples, beginning of September?
@@
 
-\item When will be the deadline of PEPM2020 (October 18th)
+\item deadline of PEPM2020 (October 18th)
 \item ESOP (?)
@@
 \item []
-
+\end{itemize}
+
+Points to discuss
+\begin{itemize}
+\item Introduction (story)
+\item Contributions
 \item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
+\item One week to revise current version.
+  After that, Prof. Hu and Robert, could you send comments?
+
+%\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
 %\item right assoc \& Recursive example?
@@
 The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
+We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2.
 Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
 We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
+The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4.
+The second one is lazy evaluation to avoid redundant computations, and we obtain $kpg$ in Section 5.
 The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
+In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
 In Section 8 we discuss about related work, and conclude in Section 9.
@@
 
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
+%\subsection{Construction of this paper}
+
+%\begin{itemize}
+%\item BiGUL 
+%\item pg (pair of put and get)
+%\item cpg (extention of pg. Keep eval information for reusing)
+%\item kpg (extention of cpg. Lazy evaluation)
+%\item xpg (combination of pg and kpg)
+%\item Experiment
+%\item Related work
+%\item Conclusion and Future work
+%\end{itemize}
 
@@
 
+\begin{itemize}
+\item BiGUL papers (PEPM, POPL, tutorial?)
+\item the first BX paper
+\item view update problem paper
+\item get-based 
+\end{itemize}
+
 \newpage

### 2019-09-18 paper/main.tex
discussion memo

@@
 
-\pgfplotsset{compat=1.12}
+\pgfplotsset{compat=1.11}
 
@@
 
+% plan to pages
+% abstraction & title & Introduction: 3pages
+% miniBiGUL: 1.5page
+% pg: 3pages
+% cpg: 2page
+% kpg: 1page
+% xpg:0.5page
+% evaluation:2pages
+% related work:1page
+% conclusion:1page (including reference)
+% page limit : 15
+
 Schedule
@@
 \item deadline of PEPM2020 (October 18th)
-\item ESOP (?)
+\item ESOP (October 24th)
+\item FLOPS (the middle of November??)
 \item Need to change the style file of latex..
@@
 
+% general, concrete problem
+% get-based also have the same problem. even in lenses. more general problem in
+% The example should be from lenses.
+% traversal, not exactly the same computation
+
 To solve this problem and make BiGUL evaluation efficient, we use two techniques.
@@
 
+% get & put independently, but for efficient computation
+% we need to tuple them to help composition
+% 2 or 3 examples to show tupling solves the problem.
+% other examples are for showing kpg and cpg is needed.
+
 \subsection{Contribution}
@@
   \end{itemize}
-\item Tupling is also effective in bidirectional programming.
+\item Get and put more tight
+
+\item (Tupling introduces some constructions)
   \begin{itemize}
@@
   \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
+\item careful lazyness is needed in our implementation for efficiency
+%  lazy update \& lazy computation are also good optimization for BX.
 %\item Reveal a part of relationship between BX and Reversible.
@@
 \begin{itemize}
-\item BiGUL papers (PEPM, POPL, tutorial?)
-\item the first BX paper
-\item view update problem paper
-\item get-based 
+%\item BiGUL papers (PEPM, POPL, tutorial?)
+%\item the first BX paper
+%\item get-based BX
+%\item put-based BX?
+\item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
+\item optimization techniques : tupling, lazy computation, fusion transformation?
+\item (view update problem paper)
 \end{itemize}

### 2019-09-27 paper/1.introduction.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-18 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.11}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-% plan to pages
-% abstraction & title & Introduction: 3pages
-% miniBiGUL: 1.5page
-% pg: 3pages
-% cpg: 2page
-% kpg: 1page
-% xpg:0.5page
-% evaluation:2pages
-% related work:1page
-% conclusion:1page (including reference)
-% page limit : 15
-
-Schedule
-\begin{itemize}
-
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item deadline of PEPM2020 (October 18th)
-\item ESOP (October 24th)
-\item FLOPS (the middle of November??)
-\item Need to change the style file of latex..
-\item []
-\end{itemize}
-
-Points to discuss
-\begin{itemize}
-\item Introduction (story)
-\item Contributions
-\item Related work?
-\item One week to revise current version.
-  After that, Prof. Hu and Robert, could you send comments?
-
-%\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
 \section{Introduction}
@@
 %\subsection{Construction of this paper}
-
-%\begin{itemize}
-%\item BiGUL 
-%\item pg (pair of put and get)
-%\item cpg (extention of pg. Keep eval information for reusing)
-%\item kpg (extention of cpg. Lazy evaluation)
-%\item xpg (combination of pg and kpg)
-%\item Experiment
-%\item Related work
-%\item Conclusion and Future work
-%\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\begin{itemize}
-%\item BiGUL papers (PEPM, POPL, tutorial?)
-%\item the first BX paper
-%\item get-based BX
-%\item put-based BX?
-\item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
-\item optimization techniques : tupling, lazy computation, fusion transformation?
-\item (view update problem paper)
-\end{itemize}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/2.minbigul.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
 \section{Bidirectional Programming Language: minBiGUL}
@@
 Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
+    $\tab \getbxinline{bx} : s \to v,
+    \tab \putbxinline{bx} : s * v \to v$
 
 BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
+    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
+    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$
 
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
+The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source.
+The \textsc{PutGet} law means that we can recover the modified view by applying the forward transformation to the updated source.
 
@@
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
+Then $put$ and $get$ satify one more law, \textsc{PutPut}, like the following:\\
+    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
 
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
+The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates.
+Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.
 
@@
 
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
+$\tab bx := Skip \ h \
+        | \ Replace \
+        | \ Prod \ bx_1 \ bx_2 \
+        | \ RearrS \ f_1 \ f_2 \ bx \
+        | \ RearrV \ g_1 \ g_2 \ bx \\
+    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \
+        | \ Compose \ bx_1 \ bx_2$
 
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
+A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
 
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
+For source/view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
 It is a kind of bijection.
@@
 
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
+    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2,
+    \tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
 
@@
 
-In the next part, we will introduce the semantics of minBiGUL.
-
 \subsection{Semantics}
 
+\begin{multicols}{2}
     \begin{definition}
@@
 
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
+        $\putbx{Skip \ h}{s}{v} = \\
+            \tab \sif{h \ s = v}{s}{\text{fail}}$
 
@@
 
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
+        $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
             \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
 
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
+        $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = \\
+            \tab f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
 
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
+        $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \\
+            \tab \putbx{bx}{s}{(g_1 \ v)}$
 
@@
 
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
+        $\putbx{bx_1 \circ bx_2}{s}{v} = \\
+            \tab \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
     \end{definition}
-
+\columnbreak
     \begin{definition}
@@
 
-    $\getbx{Skip \ h}{s} = h \ s$
+        $\getbx{Skip \ h}{s} = \\ 
+            \tab h \ s$
 
@@
 
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
+        $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \\ 
+            \tab \getbx{bx}{(f_1 \ s)}$
 
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
+        $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = \\ 
+            \tab g_2 \ (\getbx{bx}{s})$
 
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
+        $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} = \\
+            \tab \text{if} \ {cond_{s} \ s} \\
+            \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\ 
+            \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s} \\ 
             \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
 
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
+        $\getbx{bx_1 \circ bx_2}{s} = \\ 
+            \tab \getbx{bx_2}{(\getbx{bx_1}{s})}$
     \end{definition}
+\end{multicols}
 
 In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
+This statement is useful to describe many functions related to $Case$ in this paper.
 Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
@@
 $\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
+\tab \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false; \}\\
+    \qtab \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false; \}$
 
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
+Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:
 
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
+$\tab \putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10 \qtab \ \getbx{Skip \ (\lambda x.(x*x))}{10} = 100\\
+    \tab \putbx{Skip \ (\lambda \_.())}{1}{()} = 1 \qtabs{3} \getbx{Skip \ (\lambda \_.())}{1} = ()\\
+    \tab \putbx{Replace}{1}{100} = 100 \qtabs{3} \getbx{Replace}{1} = 1
+$
 
-Now, let's consider the definition of $phead$ in BiGUL [?]:
+Now, let's consider the definition of $phead$ in minBiGUL:
 
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
+$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
+    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where: } g_1 = \lambda v.(v,()), \, g_2 = \lambda (v,()).v,\\
+        \qtabs{2} bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$
 
 The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
+Intuitively, $\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$, and $\getbx{phead}{s_0}$ returns the head of the list $s_0$.
+For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.
+If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
+For example:\\
 \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}
+\tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$

### 2019-09-27 paper/3.pg.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

\section{pg}
\subsection{Self-inverse function: pg}

In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation. This is a kind of information loss. One question is whether it is possible to calculate such programs without losing information. And the answer is yes. It comes from the the idea of reversible computation where all information during the evaluation need to be kept. In minBiGUL, we can do that by tupling $put$ and $get$. A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of the corresponding minBiGUL program.

\begin{definition}
    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
\end{definition}

$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.

\begin{proof}
$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\textsc{PutGet}] \\
    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\textsc{PutPut}] \\
    \tab = (s, v) \quad [\textsc{GetPut}]$
\end{proof}

\subsection{Construction of pg}

$\tab \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
    \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
    \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{fail}}$\\
There is a trick in the construction of $\pginline{Skip \ h}$. The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. And the last one is quite obivious. What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.

$\tab \pg{Replace}{s}{v} = (v, s)$\\
$\tab \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
    \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
    \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
        \qtab \tab \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
        \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
$\tab \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
        \qtabs{2} (f_2 \ s, v)$\\
$\tab \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \qtabs{2} (s, g_2 \ v)$\\
Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.

$\tab \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
\tabs{2} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
\tab \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtabs{2} \, \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
\tab \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtabs{2} \ \, \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
\tab \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \qtabs{2} \tab \ \, \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')\\
\tab \qtab \quad \text{if} \ {cond_{s} \ s}\\
\tab \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
\tab \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
\tab \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')$

A restriction for $\pginline{Case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$. If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions. This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified. The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper. This does not happen for the others which is used in the construction of $\pginline{Case}$.

$\tab \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{2} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{2} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{2} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{2} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{2} (s_3, v_2)$

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid. Due to \textsc{GetPut} law, we can use $get \ s$ to update the source $s$ in the $put$ direction. Then it is feasible to construct a $dummy$ from the source. The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{(construct\_dummy \ s)}$. This transformation is possible because of the \textsc{PutPut} law. Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$. So we can realize that there is no loss information when computing a composition.

### 2019-09-27 paper/4.cpg.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
 \section{cpg}
@@
 Therefore, the runtime inefficiency is inevitable.
-
 In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
@@
 
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
+$\tab fst = \lambda (x_1,x_2). x_1 \qtab snd = \lambda (x_1,x_2). x_2\\
+ \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$
 
@@
 
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
+$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$
 
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
+$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$
 
@@
     \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
+    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$
 
@@
     \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
+    \qtab (f_2 \circ ks, kv, s, v)$
 
@@
     \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
+    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$
 
@@
     \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
+    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$
 
@@
 
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
+Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.\\
 For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
@@
 It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
+After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.\\
 Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:\\
+    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
+        \qtab (s; v)$\\
 In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
+However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.\\
+Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively.
+Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.
+After the first two assignments in the definition of $cpg$ for the composition, we have: $ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))$ and $s_2 = [100,2,3]$ where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$.
+\\
+Then: $ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
 \tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
+\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$\\
 If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
@@
 This time, the result is what we want to see.
-
 Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/5.kpg.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
 \section{kpg}
@@
 The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
 Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:\\
+    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+        \qtab (ks' \ s; kv' \ v)$\\
 The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
@@
 $\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
+    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
+    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$
 
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
+$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$
 
 $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
+    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
     \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
     \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
+    \qtab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, con \ (ks_1' \circ fst) \ (ks_2' \circ snd),\\
+    \qtab con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
+    \qtab (s_1, s_2), (v_1,v_2))$
 
@@
     \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
+    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$
 
@@
     \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
+    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$
 
 $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
+    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id ; \quad kv' \Leftarrow id;\\
     \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
@@
     \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
+    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' \ v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$
 
@@
 
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
+In the construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$.
+Except them, the functions for computation will be kept in $ks'$ and $kv'$.
 When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
 In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
@@
 It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/6.xpg.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
 \section{xpg}
-
 So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
@@
 \end{definition}
-
 Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
@@
 The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/7.experiments.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+\section{Experiments}
 
-\pgfplotsset{compat=1.12}
+This section describes the experiments involving $minBiGUL$, $pg$, $cpg$, $kpg$ and $xpg$. The evaluation time and the memory allocation are considered in each test.
 
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
+\subsection{Experiment environment \& Test cases}
 We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
@@
 
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
+We also conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
 
-\begin{table}[]
+\begin{table}[hbt!]
     \centering
@@
 This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
+Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
+The first four test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
 For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
+The last test case, \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function for \textit{foldr} which is an important higher-order function on lists.
+This case is more natural than the other since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and $bFoldr$.
+It will look like a composition of more BX programs at some point if we slow down the evaluation.
+Note that $\circ$ has a higher priority than $\times$.
+This leads to seemingly impossible to transform \textit{bReverse} as well as \textit{bFoldr} from left-associative style to right-associative style.
 
@@
 
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
+Firstly, from definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
+Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.
+For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear.
+For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
+For recursive composition (\textit{breverse}), all three are nonlinear.
+The appendix will provide more complete insights about those statements.
+Now we take a closer look at the empirical results.
 
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
+\input{result.tex}
 
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
+Fig \ref{fig:result}a illustrates the evaluation times in 5 test cases.
+$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since No($pg$) in each case is under an exponential distribution.
 For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
+In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minBiGUL$ time are nonlinear.
+They are quite reasonable because the data sizes are constant, No($get$) is quadractic for $minBiGUL$ while there is a linear quantity of redundant evaluations in $cpg$.
+In case of \textit{lassoc-comp-phead}, since the data sizes are linear and No($f$) are either linear or quadractic or exponential, all times are nonlinear.
+For right-associative and non-recursive compositions, $minBiGUL$, $pg$ and $xpg$ times are linear.
 In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
+In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.
 For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
+Fig \ref{fig:result}b shows the memory usage which depends on the input size and number of compositions.
+In general, $minBiGUL$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
 $pg$ only uses memory efficient when handling right-associative compositions.
 Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}
+For recursive compositions (\textit{breverse}), $minBiGUL$ uses significantly more memory than $kpg$ as well as $xpg$.

### 2019-09-27 paper/8.related-work.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
-
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
-
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
-\begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
+\section{Related Work}
 
 \begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+    %\item BiGUL papers (PEPM, POPL, tutorial?)
+    %\item the first BX paper
+    %\item get-based BX
+    %\item put-based BX?
+    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
+    \item optimization techniques : tupling, lazy computation, fusion transformation?
+    \item (view update problem paper)
 \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/appendix.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage{tikz}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{subfig}
-\usepackage{lmodern}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+\appendix
 
-\pgfplotsset{compat=1.12}
+\section{Remove?:First approach}
 
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
+% We should skip this
 
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
-
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
-\newcommand{\product}[2]{#1 \times #2}
-\newcommand{\tuple}[2]{(#1 :: #2)}
-\newcommand{\rearrs}[3]{RearrS \ #1 \ #2 \ #3}
-\newcommand{\rearrv}[3]{RearrV \ #1 \ #2 \ #3}
-\newcommand{\casebx}[4]{Case \ #1 \ #2 \ #3 \ #4}
-
-\newcommand{\putbx}[3]{put \, [\![#1]\!] \ #2 \ #3}
-\newcommand{\putbxinline}[1]{put \, [\![#1]\!]}
-\newcommand{\getbx}[2]{get \, [\![#1]\!] \ #2}
-\newcommand{\getbxinline}[1]{get \, [\![#1]\!]}
-
-\newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
-\newcommand{\getrev}[2]{get_{REV} \, [\![#1]\!] \ {#2}}
-
-\newcommand{\pg}[3]{pg \, [\![#1]\!] (#2, #3)}
-\newcommand{\pginline}[1]{pg \, [\![#1]\!]}
-
-\newcommand{\cpg}[5]{cpg [\![#1]\!] (#2, #3, #4, #5)}
-\newcommand{\cpginline}[1]{cpg \, [\![#1]\!]}
-
-\newcommand{\kpg}[7]{kpg [\![#1]\!] (#2, #3, #4, #5, #6, #7)}
-\newcommand{\kpginline}[1]{kpg \, [\![#1]\!]}
-
-\newcommand{\xpg}[3]{xpg \, [\![#1]\!] (#2, #3)}
-\newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
-
-\begin{document}
-%
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
-%  tupling
-%  accumulating updates = lazy updates
-%  based on the idea of reversible computing  
-  % \thanks{Supported by organization x.}
-}
-%
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
-\authorrunning{F. Author et al.}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
-%
-\maketitle              % typeset the header of the contribution
-%
-\begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
-
-\keywords{First keyword  \and Second keyword \and Another keyword.}
-\end{abstract}
-%
-%
-%
-
-Points to discuss
 \begin{itemize}
-\item Schedule:
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item When will be the deadline of PEPM2020 (October 18th)
-\item ESOP (?)
-\item Need to change the style file of latex..
-\item []
-
-\item Related work?
-\item
-
-\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
+\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
+\item put$_{REV}$ = s * v $\to$ s * gb
+\item get$_{REV}$ = s $\to$ v * gb
+\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
+\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
 \end{itemize}
 
-\section{Introduction}
+$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
 
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
+$\putrev{Replace}{s}{v} = (v, s)$
 
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
+$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
+        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
 
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
+$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
+        \qtab (f_2 \ s_1, gb_1)$
 
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
+$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
+        \qtab (s_1, gb_1)$
 
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
+$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
+    \tab \text{if} \ cond_{sv} \ s \ v\\
+    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
+    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
+    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
+
+$\putrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
+    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
+        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
 
@@
 
-Figure that shows many gets re-evaluation
+$\getrev{Skip \ h}{s} = {(h \ s, s)}$
 
-\vspace{5mm}
+$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
 
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
+$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
+        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
 
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computation, and obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal result of $pg$, $cpg$, $kpg$, $xpg$ and the original miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
+$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
+        \qtab (v_1, gb_1)$
 
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
+$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
+        \qtab (g_2 \ v_1, gb_1)$
 
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
+$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
+    \tab \text{if} \ cond_{s} \ s\\
+    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
+    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
+    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
 
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
+$\getrev{bx_1 \circ bx_2}{s}{v} =\\
+    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
+    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
+        \qtab (v_2, gb_1 :: gb_2)$
 
-\subsection{Contribution}
+\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}
 
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-\end{itemize}
-
-\subsection{Construction of this paper}
-
-\begin{itemize}
-\item BiGUL 
-\item pg (pair of put and get)
-\item cpg (extention of pg. Keep eval information for reusing)
-\item kpg (extention of cpg. Lazy evaluation)
-\item xpg (combination of pg and kpg)
-\item Experiment
-\item Related work
-\item Conclusion and Future work
-\end{itemize}
-
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
+\section{minBiGUL}
 Now, let's consider the definition of $phead$ in BiGUL [?]:
@@
 \tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
+    \qtabs{2} \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
+        \qtabs{2} \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
 
@@
     \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
+        \qtabs{2} g_1 = \lambda v.(v,())\\
+        \qtabs{2} g_2 = \lambda (v,()).v\\
+        \qtabs{2} bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
 
@@
     \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
+    \qtabs{2} \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
+    \qtabs{2} Replace \ `prod\text{'} \ replaceAll\\
 \tab ]$\\
@@
         \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
+            \qtabs{2} bx_v = \product{Replace}{replaceAll}$
 
@@
 
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
+\section{Evaluation by pg}
 
@@
         \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
+            \qtabs{2} (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
+            \qtabs{2} (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
+            \qtabs{2} ((s_1,s_2), (v_1,v_2))\\
         \qtab \tab \};\\
@@
         \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
+            \qtabs{2} (s_1, v_1) \Leftarrow (100, 1);\\
+            \qtabs{2} (s_2, v_2) \Leftarrow ([2,3], ());\\
+            \qtabs{2} ((s_1,s_2), (v_1,v_2))\\
         \qtab \tab \};\\
@@
 
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
+\section{Experiments-1}
 In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
@@
     \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
+        \qtabs{2} \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
+            \qtabs{3} (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
+    \qtabs{2}, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
+        \qtabs{2} \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
+            \qtabs{3} (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
+    \qtabs{2}]\\$
 
@@
         \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
+        \qtabs{2} \quad \lambda s \ v \to v\\
+        \qtabs{2} , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
+        \qtabs{2} \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
+            \qtabs{2} \qquad bFoldr \ bSnoc\\
+       \qtabs{2} ]$
 
@@
 
+\section{A}
+\begin{table}[]
+    \centering
+    \caption{Comparison}
+    \label{tab:comparison}
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{|l|l|l|l|}
+            \hline
+            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
+            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
+            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
+            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
+            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
+            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
+            \multicolumn{4}{r}{} \\
+            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
+            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
+            \multicolumn{4}{r}{l/r: left/right associative} \\
+            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
+            \multicolumn{4}{r}{ND: not determined}
+        \end{tabular}%
+    }
+\end{table}
+
+Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
+This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
+We conducted tests while increasing the number of compositions from 10 to 100000.
+By default, the stack overflow point in the table is larger than 100000.
+In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
+Then the above points are not determined.
+
 \subsection{Results and discussions}
@@
 We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
 
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
+Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
+From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
+Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
+
+For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
+
+For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
+In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
+
+For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
+In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
 
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
 Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
@@
 For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\newpage
-
-\section{Conclusion and Future Work}
-
-% From sample code
-% 
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
-\end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-09-27 paper/main.tex
split tex + compress
missing: abstract, intro, related work, conclusion, reference, appendix

--- 2019-09-13 paper/main.tex
@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage{tabularx}
@@
 \usepackage{pgfplotstable}
-\usepackage{subfig}
 \usepackage{lmodern}
+\usepackage{pgffor}
+
 % Used for displaying a sample figure. If possible, figure files should
@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\pgfplotsset{compat=1.12}
+\usetikzlibrary{matrix}
+\usepgfplotslibrary{groupplots}
+\pgfplotsset{compat=newest}
 
@@
 
-\newcommand{\tab}{\hspace*{5mm}}
-\newcommand{\qtab}{\hspace*{5mm} \ \quad}
+\newcommand{\tab}{\hspace*{3mm}}
+\newcommand{\tabs}[1]{\foreach \n in {1,...,#1}{\tab}}
+\newcommand{\qtab}{\hspace*{3mm} \ \quad}
+\newcommand{\qtabs}[1]{\foreach \n in {1,...,#1}{\qtab}}
 
@@
 \end{itemize}
-
-Points to discuss
-\begin{itemize}
-\item Introduction (story)
-\item Contributions
-\item Related work?
-\item One week to revise current version.
-  After that, Prof. Hu and Robert, could you send comments?
-
-%\item TODO: %relationship between result (original vs \{cpg, kpg, xpg\})
-%\item right assoc \& Recursive example?
-%\item current foldr is not efficient? Are there any efficient foldr? Question to Prof. Hu.
-%\item Language to show semantics: fi?, $f^{-1}$, $cond_{sv}$ and $cond_{s}$
-%\item A good simple example of bx: Used in explanation of BiGUL and evaluation of pg.
-%\item Related work: suggest?
-%\item another pg ?? should be compared?
-\end{itemize}
-
-\section{Introduction}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item goodness of 
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
-% A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
-
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
-
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
-
-\vspace{5mm}
-
-Figure that shows many gets re-evaluation
-
-\vspace{5mm}
-
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
-
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computations, and we obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
-
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
-
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
-
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
-
-\subsection{Contribution}
-
-\begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
-  \end{itemize}
-\item Tupling is also effective in bidirectional programming.
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item lazy update \& lazy computation is also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-%\subsection{Construction of this paper}
 
@@
 
-\section{Bidirectional Programming Language: minBiGUL}
-
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
-BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} = fun \ s \to v \\
-    \tab \putbxinline{bx} = fun \ s \ v \to v$
-
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\text{GETPUT}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\text{PUTGET}]$
-
-The GETPUT law means that if there is no change to the view, there should be no change to the source.
-The PUTGET law means that we can recover the modified view by applying the forward transformation to the updated source.
-
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, PUTPUT, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\text{PUTPUT}]$
-
-The PUTPUT law means that a source update should overwrite the effect of previous source updates.
-
-Due to the satisfaction of three laws, GETPUT, PUTGET and PUTPUT, minBiGUL is very well-behaved.
-
-In the remaining of this section, we will present about the syntax and semantics of minBiGUL through formal definitions and examples.
-
-\subsection{Syntax}
-
-The syntax of minBiGUL is briefly written as follows:
-
-$bx := Skip \ h \\ 
-    \qtab | \ Replace \\ 
-    \qtab | \ Prod \ bx_1 \ bx_2 \\ 
-    \qtab | \ RearrS \ f_1 \ f_2 \ bx \\
-    \qtab | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \\ 
-    \qtab | \ Compose \ bx_1 \ bx_2$
-
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source / view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
-
-For source / view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
-It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
-In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
-
-To help make demonstration more direct, we provide the following alternatives representation:
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2$\\
-    $\tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
-In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
-
-In the next part, we will introduce the semantics of minBiGUL.
-
-\subsection{Semantics}
-
-\begin{definition}
-    $\putbx{bx}{s}{v}$
-
-    $\putbx{Skip \ h}{s}{v} = \sif{h \ s = v}{s}{\text{fail}}$
-
-    $\putbx{Replace}{s}{v} = v$
-
-    $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} =\\
-        \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
-
-    $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
-
-    $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \putbx{bx}{s}{(g_1 \ v)}$
-
-    $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
-        \tab \text{if} \ {cond_{sv} \ s \ v} \\
-        \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-        \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-        \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
-
-    $\putbx{bx_1 \circ bx_2}{s}{v} = \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
-\end{definition}
-
-\begin{definition}
-    $\getbx{bx}{s}$
-
-    $\getbx{Skip \ h}{s} = h \ s$
-
-    $\getbx{Replace}{s} = s$
-
-    $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-        \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
-
-    $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \getbx{bx}{(f_1 \ s)}$
-
-    $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = g_2 \ (\getbx{bx}{s})$
-
-    $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-        \tab \text{if} \ {cond_{s} \ s}\\
-        \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-        \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-        \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
-
-    $\getbx{bx_1 \circ bx_2}{s} = \getbx{bx_2}{(\getbx{bx_1}{s})}$
-\end{definition}
-
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
-We can write the equivalent if-then-else statement as follows:
-
-$\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then}\\ 
-    \qtab \tab X_1;\\ 
-    \qtab \tab \text{if } E_2 = true \text{ then } S \text{ else assert } false;\\
-    \qtab \text{else}\\ 
-    \qtab \tab X_2;\\ 
-    \qtab \tab \text{if } E_2 = false \text{ then } S \text{ else assert } false;$
-
-The if-then-else-fi statement is useful to describe many functions related to $Case$ in this paper.
-
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:\\
-$\putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10$\\
-$\putbx{Skip \ (\lambda \_.())}{1}{()} = 1$\\ 
-$\getbx{Skip \ (\lambda x.(x*x))}{10} = 100$\\
-$\getbx{Skip \ (\lambda \_.())}{1} = ()$\\
-$\putbx{Replace}{1}{100} = 100$\\
-$\getbx{Replace}{1} = 1$\\
-
-Now, let's consider the definition of $phead$ in BiGUL [?]:
-
-$\tab phead :: Show \ s \Rightarrow BiGUL \ [s] \ s\\
-\tab phead = \$(RearrS \ [\![\lambda(s:ss) \to (s,ss)]\!])\$\\
-    \qtab \qtab \quad \$(RearrV \ [\![\lambda v \to (v,())]\!])\$\\
-        \qtab \qtab \qquad Replace \ `prod\text{'} \ (Skip \ (const \ ()))$
-
-The above program rearranges the source, a non-empty list, to a pair of its head element $s$ and its tail $ss$, and the view to a pair $(v, ())$, then we can use $v$ to replace $s$ and $()$ to keep $ss$.
-$\putbx{phead}{s_0}{v_0}$ returns a list whose head is $v_0$ and tail is the tail of $s_0$.
-$\getbx{phead}{s_0}$ returns the head of $s_0$.
-For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.\\
-
-In minBiGUL, we can express $phead$ as follows:
-
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where:}\\
-    \qtab f_1 = \lambda (s::ss).(s,ss)\\
-    \qtab f_2 = \lambda (s,ss).(s::ss)\\
-    \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where:}\\
-        \qtab \qtab g_1 = \lambda v.(v,())\\
-        \qtab \qtab g_2 = \lambda (v,()).v\\
-        \qtab \qtab bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$\\
-
-Then, we can evaluate $\putbxinline{phead}$ and $\getbxinline{phead}$ based on their definitions:
-
-$\putbx{phead}{[1,2,3]}{100}\\
-\tab = \putbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-\tab = f_2 \ (\putbx{bx_s}{(f_1 \ [1,2,3])}{100})\\
-\tab = f_2 \ (\putbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100})\\
-\tab = f_2 \ (\putbx{bx_v}{(1,[2,3])}{(g_1 \ 100}))\\
-\tab = f_2 \ (\putbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{(100,())})\\
-\tab = f_2 \ ((\putbx{Replace}{1}{100}), (\putbx{(Skip \ (\lambda \_.()))}{[2,3]}{()}))\\
-\tab = f_2 \ (100, [2,3])\\
-\tab = [100,2,3]$\\
-
-$\getbx{phead}{[1,2,3]}\\
-\tab = \getbx{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}\\
-\tab = \getbx{bx_s}{(f_1 \ [1,2,3])}\\
-\tab = \getbx{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}\\
-\tab = g_2 \ (\getbx{bx_v}{(1,[2,3])})\\
-\tab = g_2 \ (\getbx{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])})\\
-\tab = g_2 \ ((\getbx{Replace}{1}), (\getbx{(Skip \ (\lambda \_.()))}{[2,3]}))\\
-\tab = g_2 \ (1,())\\
-\tab = 1$\\
-
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$. For example:
-
-    \tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-    \tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$\\
-
-We distingwish two $phead$s in the composition by indices. According to definitions of $\putbxinline{bx}$ and $\getbxinline{bx}$, we have:
-
-$\getbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}\\
-    \tab = \getbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}\\
-    \tab = \getbx{phead_2}{[1,2,3]}\\
-    \tab = 1$\\
-
-$\putbx{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{(\getbx{phead_1}{[[1,2,3],[\,],[4,5]]})}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{(\putbx{phead_2}{[1,2,3]}{100})}\\
-    \tab = \putbx{phead_1}{[[1,2,3],[\,],[4,5]]}{[100,2,3]}\\
-    \tab = [[100,2,3],[\,],[4,5]]$\\
-
-One more example we would like to present is $replaceAll$ whose definition in BiGUL as follows:\\
-$\tab replaceAll :: (Eq \ s, Show \ s) \Rightarrow BiGUL \ [s] \ s \\ 
-\tab replaceAll = Case \ [\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s == 1]\!][\![\lambda s \to length \ s == 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda [x] \to x]\!]) Replace,\\
-    \qtab \$(normal \ [\![\lambda s \ v \to length \ s > 1]\!][\![\lambda s \to length \ s > 1]\!]) \\ 
-    \qtab \Longrightarrow \$(rearrS \ [\![\lambda (x:xs) \to (x,xs)]\!])\$\\ 
-    \qtab \qtab \$(rearrV \ [\![\lambda v \to (v,v)]\!])\$\\ 
-    \qtab \qtab Replace \ `prod\text{'} \ replaceAll\\
-\tab ]$\\
-
-The following expresses $replaceAll$ in minBiGUL:\\
-$\tab replaceAll = Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \ \text{where:}\\ 
-    \qtab cond_{sv} = \lambda s.\lambda v.length \ s == 1\\
-    \qtab cond_{s} = \lambda s.length \ s == 1\\ 
-    \qtab bx_1 = RearrS \ (\lambda [x] \to x) \ (\lambda x \to [x]) \ Replace\\ 
-    \qtab bx_2 = RearrS \ (\lambda (x::xs) \to (x,xs)) \ (\lambda (x,xs) \to (x::xs)) \ bx_s \ \text{where:}\\ 
-        \qtab \tab bx_s = RearrV \ (\lambda v \to (v,v)) \ (\lambda (v,v) \to v) \ bx_v \ \text{where:}\\
-            \qtab \qtab bx_v = \product{Replace}{replaceAll}$
-
-We can also evaluate $\putbxinline{replaceAll}$ and $\getbxinline{replaceAll}$ similarly to the previous examples if we pay attention to conditions $cond_sv$ and $cond_s$ as well as the if-then-esle-fi structure.
-$\putbxinline{replaceAll}$ uses the view to update all elements in the source list, while $\getbxinline{replaceAll}$ returns the an element of the source list with same elements.
-For example, $\putbx{replaceAll}{[1,2,3]}{100} = [100,100,100]$ and $\getbx{replaceAll}{[1,1,1]} = 1$.
-
-\section{pg}
-\subsection{Self-inverse function: pg}
-
-In minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are re-evaluated so many times since no intermediate state is stored during the evaluation.
-This is a kind of information loss.
-One question is whether it is possible to calculate such programs without losing information.
-And the answer is yes.
-It comes from the the idea of reversible computation where all information during the evaluation need to be kept.
-In minBiGUL, we can do that by tupling $put$ and $get$.
-A pair of a source and a view is accepted as the input of a function named $pg$ to produce a new pair that contains the actual result of a corresponding minBiGUL program.
-
-\begin{definition}
-    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
-\end{definition}
-
-$pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
-
-\begin{proof}
-$pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
-    \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
-    \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\text{PUTGET}] \\
-    \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\text{PUTPUT}] \\
-    \tab = (s, v) \quad [\text{GETPUT}]$
-\end{proof}
-
-\subsection{Construction of pg}
-
-$\pg{Skip \ h}{s}{v} \\
-    \tab = (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \tab = \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \tab = \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-There is a trick in the construction of $\pginline{Skip \ h}$.
-The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
-The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-And the last one is quite obivious.
-What we call the trick here is in the second equality where in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.\\
-
-$\pg{Replace}{s}{v}\\
-    \tab = (v, s)$
-
-$\pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)} \\
-    \tab = ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtab \tab ((s_1,s_2), (v_1,v_2))$
-
-$\pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \\
-    \tab = (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtab \tab (f_2 \ s, v)$
-
-$\pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \\
-    \tab = (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \tab = (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtab \tab (s, g_2 \ v)$\\
-
-Constructions of $pg$ for the replacement, the product and the source / view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
-
-$\pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-    \tab = (\text{if} \ {cond_{sv} \ s \ v} \\
-    \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-    \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-    \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s', \\
-    \qtab \quad \text{if} \ {cond_{s} \ s}\\
-    \qtab \quad \text{then} \ v' \Leftarrow \getbx{bx_1}{s}\\
-    \qtab \quad \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-    \qtab \quad \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v') \\
-    \tab = \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \\
-        \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-        \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-        \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$\\
-
-A restriction for $\pginline{Case}$ needs to be introduced here.
-We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$.
-If a tupling $put$ and $get$ occurs, there will be 4 combinations of these conditions.
-This means that two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
-This does not happen for the others which is used in the construction of $\pginline{Case}$ above.
-\\
-
-$\pg{bx_1 \circ bx_2}{s}{v} \\
-    \tab = (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tab = v_1 \Leftarrow \getbx{bx_1}{s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \\
-            \qtab \tab (s_3, v_2) \\
-    \tab = (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tab (s_3, v_2)$\\
-
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
-The first two equalities comes from mentioned definitions and some basic transformations.
-The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to GETPUT law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a dummy from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{construct\_dummy \ s}$.
-This transformation is possible due to the PUTPUT law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
-
-\subsection{Evaluation by pg}
-
-By the definition of $pg$, we have:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = (\putbx{phead}{[1,2,3]}{100}), \getbx{phead}{[1,2,3]})\\
-    \tab = ([100,2,3], 1)$\\
-
-On the other hand, we can evaluate $\pginline{phead}$ due to the construction of $pg$:
-$\pg{phead}{[1,2,3]}{100}\\
-    \tab = \pg{\rearrs{f_1}{f_2}{bx_s}}{[1,2,3]}{100}\\
-    \tab = (s, v) \Leftarrow \pg{bx_s}{f_1 \ [1,2,3]}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \pg{\rearrv{g_1}{g_2}{bx_v}}{(1,[2,3])}{100};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{bx_v}{(1,[2,3])}{g_1 \ 100};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \pg{\product{Replace}{(Skip \ (\lambda \_.()))}}{(1,[2,3])}{((100,()))};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow \pg{Replace}{1}{100};\\
-            \qtab \qtab (s_2, v_2) \Leftarrow \pg{(Skip \ (\lambda \_.()))}{[2,3]}{()};\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow \{\\
-            \qtab \qtab (s_1, v_1) \Leftarrow (100, 1);\\
-            \qtab \qtab (s_2, v_2) \Leftarrow ([2,3], ());\\
-            \qtab \qtab ((s_1,s_2), (v_1,v_2))\\
-        \qtab \tab \};\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow \{\\
-        \qtab \tab (s, v) \Leftarrow ((100,[2,3]), (1,()))\\
-        \qtab \tab (s, g_2 \ v)\\
-        \qtab \};\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = (s, v) \Leftarrow ((100, [2,3]), 1);\\
-        \qtab \tab (f_2 \ s, v)\\
-    \tab = ([100,2,3], 1)$\\
-
-We also have the evaluation of $\pginline{phead_1 \circ phead_2}$ as follows:
-
-$\pg{phead_1 \circ phead_2}{[[1,2,3],[\,],[4,5]]}{100}\\
-    \tab = (s_1, v_1) \Leftarrow \pg{phead_1}{[[1,2,3],[\,],[4,5]}{dummy};\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{v_1}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow \pg{phead_2}{[1,2,3]}{100};\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{s_1}{s_2};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow \pg{phead_1}{[dummy,[\,],[4,5]]}{[100,2,3]};\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = (s_1, v_1) \Leftarrow ([dummy,[\,],[4,5]], [1,2,3]);\\
-        \qtab (s_2, v_2) \Leftarrow ([100,2,3], 1);\\
-        \qtab (s_3, v_3) \Leftarrow ([[100,2,3],[\,],[4,5]], dummy);\\
-            \qtab \tab (s_3, v_2)\\
-    \tab = ([[100,2,3],[\,],[4,5]], 1)$\\
-
-In the above evaluation, we do not clearly contruct what $dummy$ is.
-As we mentioned, $dummy$ is a special value which makes the first $\pginline{phead_1}$ valid.
-In this case, if we care about the type of $dummy$, it should be a list of integers.
-Thus, we can choose $dummy$ is an arbitrary list which has that type, for instance, an empty list or the head element of the source.
-
-\section{cpg}
-
-When evaluating $\pginline{bx_1 \circ bx_2}$, we realize that there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
-
-In this section, we introduce a new function, $cpg$, accumulates changes in the source and the view to solve that problem.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
-The output of this function is a 4-tuple $(ks, kv, s, v)$.
-
-To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:
-
-$\tab fst = \lambda (x_1,x_2). x_1\\
- \tab snd = \lambda (x_1,x_2). x_2\\
- \tab con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$\\
-
-\begin{definition}
-$\cpg{bx}{ks}{kv}{s}{v}$
-
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$\\
-
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$\\
-
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
-    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$\\
-
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-    \qtab (f_2 \circ ks, kv, s, v)$\\
-
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$\\
-
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} s v'; \ \text{return} \ (ks, kv, s', v')$\\
-
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
-\end{definition}
-
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
-Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because it is not used in the later evaluation process.
-Some computations to get $s_1$ are seen as redundant.
-Such values and computations have negative impacts on the runtime as well as the memory allocation in the system.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
-It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling $\cpginline{bx_1}$ one more time like in $\pginline{bx_1 \circ bx_2}$.
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$
-
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
-
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.\\ 
-After the first two assignments in the definition of $cpg$ for the composition, we have:\\ 
-    $\tab ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))\\
-    \tab s_2 = [100,2,3]$\\
-where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$\\
-Then:\\ 
-$\tab ks_1 \ s_2 \\ 
-\tab = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2)))\\
-\tab = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
-
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
-This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, $ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
-This time, the result is what we want to see.
-
-Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to keep the unchanged parts in the source.
-
-\section{kpg}
-
-In the previous section, we have known that there are some variables which are evaluated but not used later when evaluating $\cpginline{bx_1 \circ bx_2}$.
-Now we introduce $kpg$, an extension of $cpg$, for keeping away such redundant computations.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
-
-Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
-
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0;\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$
-
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
-
-\begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
-
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$\\
-
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$\\
-
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-    \tab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, \\
-    \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-    \tab (s_1, s_2), (v_1,v_2))$\\
-
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$\\
-
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$\\
-
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab s \Leftarrow ks' \ s; \tab v \Leftarrow kv' \ v;\\
-    \tab ks' \Leftarrow id ; \tab kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
-    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$\\
-
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-    \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
-    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
-\end{definition}
-
-In this construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, the functions for computation will be kept in $ks'$ and $kv'$.
-
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-
-In $\kpginline{bx_1 \ prod \ bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-
-\section{xpg}
-
-So far, when computing a composition, we only recursively call an unique function, either $pg$ or $cpg$ or $kpg$.
-To be more flexible, we make a new extension, $xpg$, allowing to call external functions.
-
-\begin{definition}
-$\xpg{bx}{s}{v}$
-
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
-\end{definition}
-
-Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair.
-The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$.
-Note that, $xpg$ is called recursively instead of $pg$.
-For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result.
-The first call and the function application come from $kpg$ approaches, while the second call is based on $pg$ approach.
-
-\section{Experiment}
-
-This section describes the experiments involving $minbigul$, $pg$, $cpg$, $kpg$ and $xpg$. The number of function calls, the evaluation time and the memory allocation are considered in each test.
-
-\subsection{Experiment environment}
-We implemented 5 approaches with OCaml 4.07.1 in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4.
-The OCaml runtime system options and garbage collection parameters are set as default.
-
-\subsection{Test cases}
-We are conducted tests on 5 cases, including many composition styles such as left or right associative, non-recursive or recursive. Table \ref{tab:test-cases} shows more details on these cases.
-
-\begin{table}[]
-    \centering
-    \caption{Test cases}
-    \label{tab:test-cases}
-    \begin{tabular*}{\textwidth}{|c @{\extracolsep{\fill}}|l|l|l|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Associative}} & \multicolumn{1}{c|}{\multirow{2}{*}{Recursive}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{5-8}
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lassoc-comp-replace & left & no & 1 & 100 & 100 & 1 \\ \hline
-        2 & rassoc-comp-replace & right & no & 1 & 100 & 100 & 1 \\ \hline
-        3 & lassoc-comp-phead & left & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & rassoc-comp-phead & right & no & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        5 & breverse & left/right & yes & [1..n] & [1..n] & [n..1] & [n..1] \\ \hline
-    \end{tabular*}
-\end{table}
-
-In table \ref{tab:test-cases}, $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and original view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 5 cases mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.\\
-
-The first 4 test cases simply use $n$ compose operators to make a composition of $n + 1$ similar $BX$ programs which are non-recursive.
-For example, \textit{lassoc-comp-replace}, left-associative composition of $Replace$s, looks like $(\ldots((Replace \circ Replace) \circ Replace) \circ \ldots \circ Replace) \circ Replace$, while \textit{rassoc-comp-replace}, right-associative composition of $Replace$s, is like $Replace \circ (Replace \circ (Replace \circ \ldots \circ (Replace \circ Replace)\ldots))$, where there are $n + 1 \ Replace$s in each case.
-\textit{lassoc-comp-phead} and \textit{rassoc-comp-phead} are similar to the above two.
-
-In section 3, we defined a composition $phead \circ phead$ for updating the head element of the head element of a list of lists by using the view.
-Similarly, we can use the view to update head element in the nested lists with a composition of a specific number of $phead$s.
-For both \textit{lassoc-comp-phead} and \textit{rassoc-comp-phead}, we choose input view $v_0$ is 100 and input source $s_0$ is one of the smallest nested lists which make $put$ and $get$ directions on the composition valid.\\
-
-The last test case, $bReverse$, is a recursive composition.
-It is defined in terms of $bFoldr$, a putback function for $foldr$ in BiGUL as well as minbigul.
-What we know about $foldr$, an important higher-order function on lists, is as follows:\\
-    $\tab foldr :: (a \to b \to b) \to b \to [a] \to b\\
-    \tab foldr \ f \ e \ [\,] = e\\
-    \tab foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
-
-We can define function $reverse$ on lists based on $foldr$:\\
-    $\tab reverse \ [\,] = [\,]\\
-    \tab reverse \ (x:xs) = snoc \ x \ (reverse \ xs)\\
-    \tab reverse = foldr \ snoc \ [\,]$
-
-Let's see the following definition of $bFoldr$ which is a restricted version of $lensFoldr$ in [?] without adaptive cases:
-
-$\tab bFoldr :: (Show \ a, Show \ v) \Rightarrow BiGUL \ (a, v) \ v \to BiGUL \ ([a], v) \ v\\
-\tab bFoldr \ bx =\\
-    \qtab Case \ [\$(normal \ [\![ \lambda (x,\_) v \to null \ x ]\!] \ [\![ \lambda (x,\_) \to null \ x ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrV [| \lambda v \to ((),v) |])\$\\
-            \qtab \qtab \qtab (Skip \ (const \ ())) \ `Prod \text{'} \ Replace\\
-    \qtab \qtab, \$(normal \ [\![ \lambda (x,\_) v \to not \ (null \ x) ]\!] \ [\![ \lambda (x,\_) \to not \ (null \ x) ]\!])\\
-        \qtab \qtab \Longrightarrow \$(RearrS \ [\![ \lambda ((x:xs), e) \to (x, (xs,e)) |])\$\\
-            \qtab \qtab \qtab (Replace \ `Prod\text{'} \ bFoldr \ bx) \ \circ \ bx\\
-    \qtab \qtab]\\$
-
-The detailed explanation of $bFoldr$ are skipped.
-Like $phead$, the program equipvalent to the above $bFoldr$ is also provided in our OCaml system.
-In the second branch of $bFoldr$, there is a composition of two BX programs where the first requires a recursive call.
-Then this composition can be classified as either left or right associative type.
-However it will look like a composition of more BX programs at some point if we slow down to evaluate it.
-Note that the compose operator has a higher priority than the product operator.
-This leads to seemingly impossible to transform the above composition from left-associative style to right-associative style.
-
-Suppose that we haved $bSnoc$, a bidirectional version of $snoc$, in BiGUL. Then $bReverse$ can defined as follows:
-
-    $\tab bReverse :: Show \ a \Rightarrow BiGUL \ [a] \ [a]\\
-    \tab bReverse =\\
-        \qtab Case \ [ \$(adaptive \ [\![ \lambda s \ v \to length \ s \neq length \ v ]\!])\\
-        \qtab \qtab \quad \lambda s \ v \to v\\
-        \qtab \qtab , \$(normal \ [\![ \lambda s \ v \to True ]\!] \ [\![ \lambda s \to True ]\!])\\
-        \qtab \qtab \quad \$(RearrS \ [\![ \lambda s \to (s,[\,]) ]\!])\$\\
-            \qtab \qtab \qquad bFoldr \ bSnoc\\
-       \qtab \qtab ]$
-
-The above definition contains an adaptive branch and it only occurs when the lengths of the source and the view are different.
-If we restrict them as two lists whose same lengths, the case analysis will only evaluate the normal branch.
-For test case $breverse$ in table \ref{tab:test-cases}, we choose both $s_0$ and $v_0$ as lists from 1 to $n$.
-Then, the results $s_r$ and $v_r$ are the list from $n$ to 1.
-Explicitly, $\putbx{bReverse}{[1..n]}{[1..n]} = [n..1]$ and $\getbx{bReverse}{[1..n]} = [n..1]$.
-In $bReverse$, the number of compositions equals to the length of the source and the view.
-
-\subsection{Results and discussions}
-
-We will in turn consider the empirical results about the number of function calls, the evaluation time and the memory allocation.\\
-
-Fig \ref{fig:count-calls} depicts the number of function calls including $put$, $get$, $pg$, $cpg$, $kpg$ and $xpg$ for each test case.
-From definitions of $put$, $cpg$, $kpg$ and $xpg$, we have No($cpg$) = No($kpg$) = No($xpg$) = No($put$) where No($f$) is the number of function call $f$.
-Here we consider that No($xpg$) in evaluating $\xpginline{bx_1 \circ bx_2}$ includes both the number of $kpg$s which are used in $\kpginline{bx_1}$ and the number of $xpg$s which are used in $\xpginline{bx_2}$.\\
-
-For right-associative and non-recursive compositions (\textit{rassoc-comp-replace} and \textit{rassoc-comp-phead}), No($get$), No($put$) and No($pg$) are linear where No($get$) $<$ No($put$) $<$ No($pg$).\\
-
-For left-associative and non-recursive compositions (\textit{lassoc-comp-replace} and \textit{lassoc-comp-phead}), No($put$) is still linear while No($get$) is quadratic and No($pg$) is under an exponential distribution.
-In this case, No($put$) $<$ No($get$) $<$ No($pg$).\\
-
-For recursive composition (\textit{breverse}), the results still show that No($put$) $<$ No($get$) $<$ No($pg$) but all three are nonlinear.
-In case of \textit{breverse}, No($put$) is quadratic, No($get$) is cubic and No($pg$) is exponential.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.15,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {number of calls},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.47,0.78)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/count-breverse.csv}\data
-                \addplot table[x=nComp,y=count_get]{\data};
-                \addlegendentry{get};
-                \addplot table[x=nComp,y=count_put]{\data};
-                \addlegendentry{put};
-                \addplot table[x=nComp,y=count_pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=count_xpg]{\data};
-                \addlegendentry{cpg/kpg/xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Number of calls}
-    \label{fig:count-calls}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {evaluation time (s)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(1.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/time-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Evaluation time (s)}
-    \label{fig:evaluation-time}
-\end{figure}
-
-Fig \ref{fig:evaluation-time} illustrates the evaluation times in 5 test cases.\\
-
-$pg$ times are always exponential for both left-associative compositions (\textit{lassoc-comp-replace}, \textit{lassoc-comp-phead}) and recursive compositions (\textit{breverse}) since the number of $pg$s in these cases are under exponential distributions.\\
-
-For left-associative and non-recursive compositions, $kpg$ times are approximate to $xpg$ times, and they are the most efficient.
-In case of \textit{lassoc-comp-replace}, $kpg$ time and $xpg$ time are linear, $cpg$ time and $minbigul$ time are nonlinear.
-They are quite reasonable because the data sizes are constant, No($get$) is quadractic for minbigul while there is a linear redundant evaluation in $cpg$.
-In case of \textit{lassoc-comp-phead}, since the data sizes are linear and the numbers of function calls are either linear or quadractic or exponential, all times are nonlinear\\
-
-For right-associative and non-recursive compositions, $minbigul$, $pg$ and $xpg$ times are linear.
-In case of \textit{rassoc-comp-replace}, all times are small enough to consider all them as linear.
-In case of \textit{rassoc-comp-phead}, both $cpg$ and $kpg$ times are quadratic because the data sizes are linear and the number of redundant evaluations are also linear.\\
-
-For recursive compositions (\textit{breverse}), all times should be nonlinear since there is no linear number of function calls.\\
-
-\begin{figure}
-    \centering
-    \subfloat[\textit{lassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-replace}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.03,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-replace.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \hspace{0mm}
-    \subfloat[\textit{lassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-lassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \subfloat[\textit{rassoc-comp-phead}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.2,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-rassoc-comp-phead.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};x
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \vspace{2mm}
-    \subfloat[\textit{breverse}]{
-        \begin{tikzpicture}
-            \begin{axis}[
-                xmode = normal,
-                ymode = normal,
-                xlabel = {number of compositions},
-                ylabel = {memory allocation (MBytes)},
-                width = 0.55\linewidth,
-                legend style = {font=\scriptsize,at={(0.57,0.75)},anchor=west},
-            ]
-                \pgfplotstableread[col sep=comma]{csv/mem-breverse.csv}\data
-                \addplot table[x=nComp,y=minbigul]{\data};
-                \addlegendentry{minbigul};
-                \addplot table[x=nComp,y=pg]{\data};
-                \addlegendentry{pg};
-                \addplot table[x=nComp,y=cpg]{\data};
-                \addlegendentry{cpg};
-                \addplot table[x=nComp,y=kpg]{\data};
-                \addlegendentry{kpg};
-                \addplot table[x=nComp,y=xpg]{\data};
-                \addlegendentry{xpg};
-            \end{axis}
-        \end{tikzpicture}
-    }
-    \caption{Memory allocation (MBytes)}
-    \label{fig:memory-allocation}
-\end{figure}
-
-Fig \ref{fig:memory-allocation} shows the memory usage which depends on the input size and number of compositions.
-In general, $minbigul$ uses less memory than the others for non-recursive compositions (all tests except \textit{breverse}).
-$pg$ only uses memory efficient when handling right-associative compositions.
-Also for evaluating these  compositions, both $cpg$ and $kpg$ are worst in memory allocation.
-For recursive compositions (\textit{breverse}), $minbigul$ uses significantly more memory than $kpg$ as well as $xpg$.
-
-\begin{table}[]
-    \centering
-    \caption{Comparison}
-    \label{tab:comparison}
-    \resizebox{\textwidth}{!}{%
-        \begin{tabular}{|l|l|l|l|}
-            \hline
-            \multicolumn{1}{|c|}{Name} & \multicolumn{1}{c|}{Properties} & \multicolumn{1}{c|}{Evaluation time order} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Memory allocation order\\ (stack\_overflow point / 100000)\end{tabular}} \\ \hline
-            lassoc-comp-replace & l, NR & xpg $\sim$ kpg $<$ cpg $<$ minbigul $<$ pg & minbigul $<$ cpg $<$ xpg $\sim$ kpg (95299) $<$ pg (ND) \\ \hline
-            rassoc-comp-replace & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (95299) $<$ kpg (80638) \\ \hline
-            lassoc-comp-phead & l, NR & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & minbigul $<$ xpg $\sim$ kpg (80638) $<$ cpg (ND) $<$ pg (ND) \\ \hline
-            rassoc-comp-phead & r, NR & minbigul $<$ pg $<$ xpg $<$ cpg $<$ kpg & minbigul $<$ pg $<$ xpg $<$ cpg (43680) $<$ kpg (61665) \\ \hline
-            breverse & l/r, R & xpg $<$ kpg $<$ minbigul $<$ cpg $<$ pg & xpg (ND) $<$ kpg (ND) $<$ cpg (ND) $<$ minbigul (ND) $<$ pg (ND) \\ \hline
-            \multicolumn{4}{r}{} \\
-            \multicolumn{4}{r}{stack\_overflow point (default: $>$ 100000): can not evaluate a program if the number of compositions is larger than this point} \\
-            \multicolumn{4}{r}{best $< \ldots < \ldots <$ worst} \\
-            \multicolumn{4}{r}{l/r: left/right associative} \\
-            \multicolumn{4}{r}{NR/R: non-recursive/recursive} \\
-            \multicolumn{4}{r}{ND: not determined}
-        \end{tabular}%
-    }
-\end{table}
-
-Table \ref{tab:comparison} indicates a general comparison of the above results over the evaluation time and the memory allocation.
-This table provides some additional information about the point where a stack overflow error occurs in our OCaml system.
-We conducted tests while increasing the number of compositions from 10 to 100000.
-By default, the stack overflow point in the table is larger than 100000.
-In some cases, we limited the number of compositions because the corresponding evaluation times are too long.
-Then the above points are not determined.
-
-\section{Related Work}
-
-\begin{itemize}
-\item BiGUL papers (PEPM, POPL, tutorial?)
-\item the first BX paper
-\item view update problem paper
-\item get-based 
-\end{itemize}
+\input{1.introduction.tex}
 
@@
 
-\section{Conclusion and Future Work}
+\input{2.minbigul.tex}
+
+\input{3.pg.tex}
+
+\input{4.cpg.tex}
+
+\input{5.kpg.tex}
+
+\input{6.xpg.tex}
+
+\input{7.experiments.tex}
+
+\input{8.related-work.tex}
+
+\input{9.conclusion}
 
@@
 \end{thebibliography}
+
+% \input{appendix.tex}
+
 \end{document}
@@
 %
-
-\section{Remove?:Reversible langauge: Rwhile}
-
-% Maybe we should skip
-
-\begin{itemize}
-\item Usually programs receives inputs and produces outputs. It is impossible to produce inputs from outputs in general. Obtaining the inverse program of a program is difficult.
-\item Using reversible language, we can obtain inverse programs easily.
-\item Explanation of Rwhile: we need fi-condition.
-\item syntax, how to work.
-\end{itemize}
-
-\section{Remove?:First approach}
-
-% We should skip this
-
-\begin{itemize}
-\item Make put and get to be reversible. We need garbages (g) to keep lost infomation after evaluation.
-\item put$_{REV}$ = s * v $\to$ s * gb
-\item get$_{REV}$ = s $\to$ v * gb
-\item introduce (put$_{REV}^{-1}$: s * gb $\to$ s * v) and (get$_{REV}^{-1}$:v * gb $\to$ s)
-\item \textcolor{red}{(Missing part?)} Relationship between put$_{REV}$, get$_{REV}$, put$_{REV}^{-1}$, get$_{REV}^{-1}$ 
-\end{itemize}
-
-$\putrev{Skip \ h}{s}{v} =  \sif{h \ s = v}{(s, v)}{\text{fail}}$
-
-$\putrev{Replace}{s}{v} = (v, s)$
-
-$\putrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)}{(v_1 :: v_2)} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx_1}{s_1}{v_1};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_2}{v_2};\\
-        \qtab (s_1 :: s_2, gb_1 :: gb_2)$
-
-$\putrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{(f_1 \ s)}{v};\\
-        \qtab (f_2 \ s_1, gb_1)$
-
-$\putrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \putrev{bx}{s}{(g_1 \ v)};\\
-        \qtab (s_1, gb_1)$
-
-$\putrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-    \tab \text{if} \ cond_{sv} \ s \ v\\
-    \tab \text{then} \ (s', gb) \Leftarrow \putrev{bx_1}{s}{v};\\
-    \tab \text{else} \ (s', gb) \Leftarrow \putrev{bx_2}{s}{v};\\
-    \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ (s', gb)$
-
-$\putrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (s_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (s_2, gb_2) \Leftarrow \putrev{bx_2}{s_1}{v};\\
-    \tab (s_3, gb_3) \Leftarrow \putrev{bx_1}{s}{s_2} \ \text{in (* need to copy s *)}\\
-        \qtab (s_3, gb_1 :: gb_2 :: gb_3)$
-
-\vspace{5mm}
-
-$\getrev{Skip \ h}{s} = {(h \ s, s)}$
-
-$\getrev{Replace}{s} = (s, \textcolor{red}{\text{Nil}})$
-
-$\getrev{\product{bx_1}{bx_2}}{(s_1 :: s_2)} = \\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s_1};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_2}{s_2};\\
-        \qtab (v_1 :: v_2, gb_1 :: gb_2)$
-
-$\getrev{\rearrs{f_1}{f_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{(f_1 \ s)};\\
-        \qtab (v_1, gb_1)$
-
-$\getrev{\rearrv{g_1}{g_2}{bx}}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx}{s};\\
-        \qtab (g_2 \ v_1, gb_1)$
-
-$\getrev{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} =\\
-    \tab \text{if} \ cond_{s} \ s\\
-    \tab \text{then} \ (v', gb) \Leftarrow \getrev{bx_1}{s};\\
-    \tab \text{else} \ (v', gb) \Leftarrow \getrev{bx_2}{s};\\
-    \tab \text{fi} \ cond_{sv} \ s \ v'; \ \text{return} \ (v', gb)$
-
-$\getrev{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (v_1, gb_1) \Leftarrow \getrev{bx_1}{s};\\
-    \tab (v_2, gb_2) \Leftarrow \getrev{bx_1}{s};\\
-        \qtab (v_2, gb_1 :: gb_2)$
-
-\textcolor{red}{If we use Rwhile for target language, we do not need to introduce inverse.}

### 2019-10-03 paper/1.introduction.tex
add intro/related/conclusion

@@
 
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
+In software, there are strong demands for synchronizing data. In database community this is known as ``view update problem'' and researched for a long time \cite{viewupdate}.
 % A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
+As a solution for this problem, bidirectional transformation (BX) is introduced \cite{}.
+As an example, let us consider a small BX program of $pHead$\footnote{The actual program is shown in the next section.}.
+BX provides two functions: $get$ and $put$.
+The following is one example of $pHead$'s evaluation.
 
-There are two types of bidirectional transformation.
+\begin{center}
+  \includegraphics[width=5cm]{./temp_fig/bxexample1.eps}
+\end{center}
+%\end{picture}
+
+Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $pHead$ picks the first element of the given original source as a view.
+%\texttt{get pHead [1,2] = 1}.
+After the view \texttt{v} is obtained, the user can modify the view.
+In this case, he or she modified the view to \texttt{8} from \texttt{1}.
+From this updated view, how can we obtain new updated source? This is ``view update problem.'' For this, BX provides $put$, update on the original source.
+In this example, $put$ of $pHead$ constructs a new list \texttt{[8,2]} from the updated view ($v'$) and the original source ($s$).
+%\texttt{put pHead [1,2] 8 = [8,2]}
+
+For more various evaluations, we can use composition of programs.
+Let us consider another BX example program, three composition of $pHead$: $pHead \circ pHead \circ pHead$.
+The following is one example of its evaluation.
+
+\begin{center}
+  \includegraphics[width=10cm]{./temp_fig/bxexample2.eps}
+\end{center}
+
+In this example, the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]}.
+By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}.
+To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
+The first put evaluation is from \texttt{i2} and \texttt{v'} and we can obtain \texttt{i2'}.
+
+For evaluating put-direction, there are two strategies.
+One is ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i1} and \texttt{i2}.
+If we choose this strategy, the number of $get$s will be quadratic.
+This is a problem, because the evaluation will be slow if the BX programs include many compositions.
+In this example, two $get$s are required for obtaining \texttt{i2} and one $get$ is required for obtaining \texttt{i1}.
+In total, three $get$s are required.
+
+The other strategy is ``keeping all intermediate states and using them when they are needed.'' This strategy causes a storage problem.
+If views includes most information, for evaluating $put$ the needed parts of the original source will be small.
+It is redundant to keep all intermediate states.
+
+In this work, we choose another strategy ``keeping complements and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states.
+Therefore this strategy solves the previous two problems.
+Readers might already know that it is a hard problem to obtain complements \cite{}. In our work, thanks to the following finding it is not hard:
+
+\vspace{2mm}
+In very-well behaved BX programs, 
+$put$ can be a complement function for $get$, $get$ can be a complement function for $put$.
+\vspace{2mm}
+
+For using complement, we combine $put$ and $get$, and produce new function $pg$ in Section 3.
+This is a kind of tupling. Because parts of $put$ and $get$ are doing the same computation, we can make efficient by tupling them.
+
+As an example, let us consider the previous example again: $pHead \circ pHead \circ pHead$.
+
+\begin{center}
+  \includegraphics[width=9cm]{./temp_fig/bxexample3.eps}
+\end{center}
+
+In the figure, \texttt{c1} and \texttt{c2} are complements.
+The points in this figure are two.
+After evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c1} and \texttt{i1}.
+This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
+Actually, the simple combined $pg$ is not effective for left associative compositions.
+To achieve efficient evaluation, we introduce $cpg$, lazy update version of $pg$, in Section 4 and $kpg$, lazy evaluation version of $cpg$, in Section 5.
+In Section 6 we combine $pg$ and $kpg$ to achieve efficiency for both associative compositions.
+In Section 7 we compare all approaches including the original BX approach by the evaluation time and memory allocation.
+
+%There are two types of bidirectional transformation.
 One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
@@
 
-BiGUL \cite{} is one of put-based bidirectional transformation language.
+%BiGUL \cite{} is one of put-based bidirectional transformation language.
 It is fastest bidirectional language in the world now.
@@
 
-\vspace{5mm}
+%\vspace{5mm}
 
-Figure that shows many gets re-evaluation
+%Figure that shows many gets re-evaluation
 
-\vspace{5mm}
+%\vspace{5mm}
 
@@
 
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
+%To solve this problem and make BiGUL evaluation efficient, we use two techniques.
 The first technique is tupling, well-known technique for efficiency \cite{}.
@@
 Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
+%The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
 
-The rest of this paper is the following.
+%The rest of this paper is the following.
 We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2.
 Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
+%We apply three optimizations to $pg$ in sequence.
 The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4.
@@
 In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
+%In Section 8 we discuss about related work, and conclude in Section 9.
 
@@
 
-\subsection{Contribution}
+\subsubsection*{Contribution of this paper}
+
+We can summarize contribution of this paper as follows.
 
@@
   \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+  \item This is both in evaluation time and memory allocation
   \end{itemize}
-\item Get and put more tight
+\item Optimization by tupling
+  \begin{itemize}
+  \item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt
+  \end{itemize}
+\end{itemize}
 
-\item (Tupling introduces some constructions)
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item careful lazyness is needed in our implementation for efficiency
+%\begin{itemize}
+%\item Improvement of evaluation efficiency
+%  \begin{itemize}
+%  \item Quadratic to exponential? 
+%  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+%  \end{itemize}
+%\item Get and put more tight
+
+%\item (Tupling introduces some constructions)
+%  \begin{itemize}
+%  \item A pair of put and get is self-inverse (involution).
+%  \end{itemize}
+%\item careful lazyness is needed in our implementation for efficiency
 %  lazy update \& lazy computation are also good optimization for BX.
@@
 %    put(s, dummy) can be a complement function.
-  \end{itemize}
+%  \end{itemize}
 

### 2019-10-03 paper/8.related-work.tex
add intro/related/conclusion

@@
 
+\subsubsection*{BX langauges and implementations}
+
+There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF} (memo: BXtend, eMoflon, EVL+Strace, JTL, NMF.
+SDMLib should be added?) based on several application scenarios and BX researches \cite{} (add papers for lenses).
+They mostly focus on the theory side, such as semantics and correctness, and do not focus on efficiency so much.
+For practical implementation of BX languages Anjorin et.
+al.
+introduces the first benchmark for BX \cite{BXcomp}.
+We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
+In this paper we focus on BiGUL \cite{josh_pepm} \cite{josh_popl}, its implementation uses ``not keeping any intermediate states and obtaining them by evaluation when they are needed'' strategy.
+Other BX languages might have the same problem, and our approach can be applicable.
+
+\subsubsection*{Optimization techniques}
+
+In this paper, we use several optimization techniques: tupling, lazy update, and lazy computation (\textcolor{red}{also fusion?}).
+Tupling \cite{tupling} is an optimization technique by combining several computations.
+Lazy update .. delta lenses?
+
+Lazy computation \cite{} is also an old but effective widly-used technique.
+
 \begin{itemize}
+\item This part looks just short introduction of  techniques .. 
+\end{itemize}
+
+%\begin{itemize}
     %\item BiGUL papers (PEPM, POPL, tutorial?)
@@
     %\item put-based BX?
-    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
-    \item optimization techniques : tupling, lazy computation, fusion transformation?
-    \item (view update problem paper)
-\end{itemize}
+%    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
+%    \item optimization techniques : tupling, lazy computation, fusion transformation?
+%    \item (view update problem paper)
+%\end{itemize}

### 2019-10-03 paper/9.conclusion.tex
add intro/related/conclusion

\section{Conclusion and Future Work}

In this paper we focus on efficiency of composition of BX programs.
The essential finding comes from the idea of tupling: in very-well behaved BX programs we can use $put$ as a compliment function for $get$, and vice versa.
Based on the idea, we introduced $pg$, and improved it by several optimization techniques. From experimental results, our fastest approach $xpg$ is faster than other approaches for non purely right associative compositions. For right associative compositions, the original approach (miniBiGUL) is faster than $xpg$ because $xpg$ has some overhead cost. However this is not a problem, because usually programs are mix of left and right associative. If programmers know that their programs are purely right associative, they can choose miniBiGUL.

We will continue our work on the following points. First, our target language is limited to very-well behaved, because our main idea requires the put-put property. However, for practical programs, we need to extend our work to overcome the current limitations.

\begin{itemize}
\item Extend our approach -- overcome our limitations
  \begin{itemize}
  \item treat well-behaved programs -- How to treat adaptive cases
  \item the programs that use the different paths (put and get)
  \end{itemize}
\item How to obtain dummy?
\end{itemize}

### 2019-10-03 paper/main.tex
add intro/related/conclusion

@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
+%The abstract should briefly summarize the contents of the paper in
+%15--250 words.
+  Bidirectional transformation (BX) is a solution of view update problems and widely used for synchronizing data.
+Its semantics and correctness are researched well, but the efficiency and optimization are not considered well.
+In this work we focus on the evaluation of composition, because it includes some problems of evaluation time efficiency and memory allocation.
+To solve these problems we make $put$ and $get$ more tight, based on the idea of tupling.
+%42
+  Because simple tupled result includes some redundancies for left associative compositions, we apply two optimization techniques: lazy update and lazy evaluation.
+We show some experimental results that our optimized approach is faster than the original approach that is used in an actual BX language.
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{Bidirectional transformation  \and Tupling \and Optimization \and ??.}
 \end{abstract}
@@
 
-Schedule
+%Schedule
 \begin{itemize}
-
-\item Introduction (story), Contribution, positions, examples, beginning of September?
-\item zero draft : mid of September
-
-\item deadline of PEPM2020 (October 18th)
-\item ESOP (October 24th)
-\item FLOPS (the middle of November??)
-\item Need to change the style file of latex..
-\item []
+%\item Introduction (story), Contribution, positions, examples, beginning of September?
+%\item zero draft : mid of September  
+%\item deadline of PEPM2020 (October 18th)
+%\item ESOP (October 24th)
+\item FLOPS 2020, 15 November 2019 (AoE): Abstract submission
+\item 22 November 2019 (AoE): Submission deadline
+\item keywords, ORCID id, introduction, contribution, future work, related work 
 \end{itemize}
@@
 
+\textcolor{red}{TODO: FIX reference, order, contents ..}
+
 % From sample code
@@
 \begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+\bibitem{viewupdate}
+F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.
 
-\bibitem{ref_lncs_1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+\bibitem{BXcomp}
+Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
+``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019
 
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
+\bibitem{tupling}
+  M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.
 
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
+\bibitem{josh_pepm}
+  Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.
 
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
+\bibitem{josh_popl}
+  Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.
+
+\bibitem{BXtend}
+  Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
+of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.
+
+\bibitem{eMoflon}
+  Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr.
+Developing eMoflon with eMoflon.
+In Davide Di Ruscio and Daniel Varr´o, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014.
+Proceedings, volume 8568 of Lecture Notes in
+Computer Science, pages 138–145. Springer, 2014.
+
+\bibitem{EVL+Strace}
+Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
+Technology, 100:47--72, 2018.
+
+\bibitem{JTL}
+Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio.
+JTL: A bidirectional and change propagating transformation language.
+In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages
+183--202, Springer-Verlag.
+
+\bibitem{NMF}
+Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.
+
+\bibitem{SDMLib}
+  SDMLib paper?
+
+\bibitem{lenses}
+  Lense papers?
+
+\bibitem{tutorial}
+  Tutorial paper of BiGUL
+
+%\bibitem{ref_article1}
+%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+
+%\bibitem{ref_lncs_1}
+%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
+%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
+%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+
+%\bibitem{ref_book1}
+%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
+%Location (1999)
+
+%\bibitem{ref_proc1}
+%Author, A.-B.: Contribution title. In: 9th International Proceedings
+%on Proceedings, pp. 1--2. Publisher, Location (2010)
+
+%\bibitem{ref_url1}
+%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
+%Oct 2017
 \end{thebibliography}

### 2019-10-03 paper/main.tex
fix refenrece

@@
 Developing eMoflon with eMoflon.
-In Davide Di Ruscio and Daniel Varr´o, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014.
+In Davide Di Ruscio and Daniel Varro, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014.
 Proceedings, volume 8568 of Lecture Notes in
@@
 JTL: A bidirectional and change propagating transformation language.
-In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages
-183--202, Springer-Verlag.
+In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages 183--202, Springer-Verlag.
 

### 2019-11-11 bigul/Basic.hs
Update Basic.hs

@@ -28,6 +28,11 @@ pHead =  $(rearrS [| \(s:ss) -> (s, ss) |])$
            $(rearrV [| \v -> (v, ()) |])$
              Replace `Prod` skip1
 
+pTail :: Show s => BiGUL [s] [s]
+pTail = $(rearrS [| \(s:ss) -> (s, ss) |])$
+            $(rearrV [| \v -> ((), v) |])$
+                skip1 `Prod` Replace
+
 pNth   :: Show s => Int -> BiGUL [s] s
 pNth i =  if i == 0 then pHead
                     else $(rearrS [| \(x:xs) -> (x,xs) |]) $

### 2019-11-13 paper/main.tex
update src
+ minbigul_m
revise abstract

@@
 %
-\title{An efficient composition of bidirectional programs
-  by tupling and lazy updates 
+\title{
+  % Strategies for efficient composition of bidirectional programs
+  % by memoization and lazy updates
   %  tupling
@@
   % \thanks{Supported by organization x.}  
+  An efficient composition of bidirectional programs
+  by memoization
 }
@@
 Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
-%
+
+%{Zhenjiang Hu\inst{1}\orcidID{0000-0002-9034-205X}}
+%{Robert Gl\"{u}ck\inst{2}\orcideID{0000-0001-6990-3935}}
+%{Kanae Tsushima\inst{3}\orcidID{}}
+%{Bach-Trong Nguyen\inst{3}\orcidID{}}
 \authorrunning{F. Author et al.}
@@
   Bidirectional transformation (BX) is a solution of view update problems and widely used for synchronizing data.
-Its semantics and correctness are researched well, but the efficiency and optimization are not considered well.
-In this work we focus on the evaluation of composition, because it includes some problems of evaluation time efficiency and memory allocation.
-To solve these problems we make $put$ and $get$ more tight, based on the idea of tupling.
-%42
-  Because simple tupled result includes some redundancies for left associative compositions, we apply two optimization techniques: lazy update and lazy evaluation.
-We show some experimental results that our optimized approach is faster than the original approach that is used in an actual BX language.
+Its semantics and correctness are researched well, but the evaluation efficiency is not considered well.
+As a first attempt we consider efficiency of evaluation in this work.
+To improve evaluation efficiency, we focus on the evaluation of BX compositions, because its evaluation often causes redundant evaluations.
+  For evaluating BX compositions efficiently, we introduce two memoization strategies.
+The first strategy, minBigul\_m, uses so simple memoization, but it works well for many BX programs.
+This keeps the evaluated results in a table and reuse them.
+A disadvantage of this is cost for saving to a table and searching.
+When the inputs are large, the evaluation time becomes slow.
+To treat large inputs, we introduce the second approach, xpg, that uses tupling and some optimization techniques.
+In this approach we shrink the size of the evaluated results and keep them in a closure.
+The experiments of these strategies show the following results: the original strategy is faster for right assoc compositions, minBiGUL\_m is faster for not-large inputs and kpg is faster for large inputs.
 
-\keywords{Bidirectional transformation  \and Tupling \and Optimization \and ??.}
+ % We show experimental results for several BX programs with simple inputs and large inputs.
+
+  %To achieve fast evaluation for complicated data, we introduced a novel approach.
+First, we make $put$ and $get$ more tight, based on the idea of tupling.
+Because simple tupled result includes redundancies for left associative compositions, we apply two optimization techniques: lazy update and lazy evaluation.
+We compare the approaches including the strategy used in the actual BX language, and show that our optimized approach is faster than the other approaches (?).
+
+\keywords{Bidirectional transformation \and
+  Implementation technique \and
+  %Evaluation? \and
+  Efficiency \and 
+  Optimization \and Tupling.
+  %\and Lazy update?
+}
 \end{abstract}
@@
 % miniBiGUL: 1.5page
-% pg: 3pages
-% cpg: 2page
-% kpg: 1page
-% xpg:0.5page
+% miniBiGUL_m : 1 page
+% xpg:5.5page
 % evaluation:2pages
@@
 
+%\begin{itemize}
+%\item FLOPS 2020, 15 November 2019 (AoE): Abstract submission
+%\item 22 November 2019 (AoE): Submission deadline
+%\item Today's agenda: introduction.
+% result of 2nd approach, memory alloc result. introduction, contribution
+%\end{itemize}
+
 %Schedule
-\begin{itemize}
+%\begin{itemize}
 %\item Introduction (story), Contribution, positions, examples, beginning of September?
@@
 %\item ESOP (October 24th)
-\item FLOPS 2020, 15 November 2019 (AoE): Abstract submission
-\item 22 November 2019 (AoE): Submission deadline
-\item keywords, ORCID id, introduction, contribution, future work, related work 
-\end{itemize}
+%\end{itemize}
+
+%Recent result/findings:
+%\begin{itemize}
+%\item miniBigul\_m becomes worse when the number of compositions is increased (in the case of Replace, pHead).
+Using some optimization to avoid comparison between BXs, it becomes faster, but it is still slower than original miniBiGUL.
+%\item bmapreplace is a special case of miniBiGUL\_m. Why??
+%\item miniBigul\_m is better for bmapreplace and breverse than other approaches.
+%\item Compose(breverse, breverse): miniBiGUL can evaluate, but kpg can not. 
+%\end{itemize}
+
+%\begin{tabular}{c|c|c|c}
+%  & 1st(miniBiGUL) & 2nd(memo\_miniBiGUL) & 3rd(kpg) \\ \hline
+%  num of $get$s & linear to quadratic & linear & linear \\ \hline 
+%  (storage) & (linear) & (linear to N(num of composition)) * (size(S)) & (max(size(S), size(S'))) \\ \hline
+%  overhead & low & saving (BX, s) to table \& searching & keeping as functions
+% \end{tabular}
+
+%Time:
+%\begin{itemize}
+%\item Lassoc (comp\_replace \& comp\_phead): xpg $<$ miniBiGUL << miniBiGUL\_m (100x)
+%\item Rassoc (comp\_replace \& comp\_phead): miniBiGUL $<$ xpg $<$ miniBiGUL\_m (1000x)
+%\item breverse: miniBiGUL\_m << xpg (10x) << miniBiGUL(100x)
+%\item bmapreplace: (miniBiGUL\_m <<) xpg << miniBiGUL (100x)
+%\end{itemize}
+
+%Memory allocation:
+%\begin{itemize}
+%\item Lassoc comp\_replace : miniBiGUL $<$ xpg $<$ miniBiGUL\_m
+%\item Lassoc\& Rassoc comp\_phead: miniBiGUL\_m << miniBiGUL < xpg
+%\item Rassoc comp\_replace: miniBiGUL < xpg < miniBiGUL\_m
+%\item breverse: miniBiGUL\_m < xpg << miniBiGUL
+%\item bmapreplace xpg << miniBiGUL 
+%\end{itemize}
+
+%\begin{itemize}
+%\item Rassoc $\to$ miniBiGUL
+%\item 
+%\end{itemize}
 

### 2019-11-14 paper/main.tex
update tex 20191114
change kpg -> xpg

@@
 In this approach we shrink the size of the evaluated results and keep them in a closure.
-The experiments of these strategies show the following results: the original strategy is faster for right assoc compositions, minBiGUL\_m is faster for not-large inputs and kpg is faster for large inputs.
+The experiments of these strategies show the following results: the original strategy is faster for right assoc compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.
 
@@
 % plan to pages
-% abstraction & title & Introduction: 3pages
-% miniBiGUL: 1.5page
-% miniBiGUL_m : 1 page
-% xpg:5.5page
-% evaluation:2pages
-% related work:1page
-% conclusion:1page (including reference)
+% 1.abstraction & title & Introduction: 3pages
+% 2.miniBiGUL: 1.5page
+% 3.miniBiGUL_m : 1 page
+% 4.xpg:5.5page
+% 5.evaluation:2pages
+% 6.related work:1page
+% 7.conclusion:1page (including reference)
 % page limit : 15

### 2019-11-14 paper/3.minbigul_m.tex

\section{Memorizing: minBiGUL$_m$}

For minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are reevaluated so many times since no intermediate state is stored during the evaluation. To avoid reevaluating $get$s, we can save intermediate states in a key-value table where the key is a pair of the program $bx$ and the source $s$, and the value is the result of $\getbx{bx}{s}$. Then, searching an evaluated $get$s on the table is used instead of recomputing it. We define a memory version, minBiGUL$_m$, including $get_m$ and $put_m$ as in definition 3 and 4 respectively. Another function introduced in definition 5, $get_{mh}$, is used to retrieve the value of an evaluated $get$ from the table or compute an unevaluated $get$ by using $get_m$ and store the key-value tuple to the table.

\begin{definition} $\putm{bx}{s}{v}$

    $\putm{bx}{s}{v} = \text{same with } put \text{ if } bx \neq bx_1 \circ bx_2$

    $\putm{bx_1 \circ bx_2}{s}{v} = \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})}$
\end{definition}

\begin{multicols}{2}
    \begin{definition} $\getm{bx}{s}{v}$

        $\getm{bx}{s} = \\
            \tab \text{same with } get \text{ if } bx \neq bx_1 \circ bx_2$

        $\getm{bx_1 \circ bx_2}{s} =\\
            \tab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
            \tab \text{with Not\_found } \to\\
                \qtab v \Leftarrow \getmh{bx_2}{(\getmh{bx_1}{s})}\\
                \qtab \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
                \qtab v$
    \end{definition}
\columnbreak
    \begin{definition} $\getmh{bx}{s}{v}$

        $\getmh{bx}{s} =\\
            \tab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
            \tab \text{with Not\_found } \to\\
                \qtab v \Leftarrow \getm{bx}{s}\\
                \qtab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
                \qtab \text{with Not\_found } \to\\
                    \qtabs{2} \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
                    \qtabs{2} v$
    \end{definition}
\end{multicols}

The special point is that not all states are saved when evaluating a program, only composition evaluations are treated. When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, an application of $get_{mh}$, not $get$ or $get_m$, appears to indicate the saving. When evaluating $\getm{bx_1 \circ bx_2}{s}$, we try to search key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, on the table. The matched value can be used for next steps in the evaluation. If there is no such key, the value $v$ will be calculated by using $get_{mh}$. The further key-value insertion is needed because the later computation could require to reevaluate $\getm{bx_1 \circ bx_2}{s}$ one more time.

### 2019-11-14 paper/4.xpg.tex

\section{Tupling and lazy updates: xpg}
\subsection{pg}

Another solution for saving intermediate states is tupling. If $put$ and $get$ are evaluated simultaneously, it is potential to reduce the number of recomputed $get$s. The following function, $pg$, accepts the pair of a source and a view as the input to produce a new pair that contains the actual result of the corresponding minBiGUL program.

\begin{definition}
    $\pg{bx}{s}{v} = (\putbx{bx}{s}{v}, \getbx{bx}{s})$
\end{definition}

% $pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.

% \begin{proof}
% $pg [\![bx]\!] \ (\pg{bx}{s}{v}) \\
%     \tab = \pg{bx}{(\putbx{bx}{s}{v})}{(\getbx{bx}{s})} \quad [pg \text{ definition}] \\
%     \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}) \ (\getbx{bx}{s}), \getbx{bx}{(\putbx{bx}{s}{v})})  \quad [pg \text{ definition}] \\
%     \tab = (put \ [\![bx]\!] \ (\putbx{bx}{s}{v}, \getbx{bx}{s}), v) \quad [\textsc{PutGet}] \\
%     \tab = (put \ [\![bx]\!] \ (s, \getbx{bx}{s}), v) \quad [\textsc{PutPut}] \\
%     \tab = (s, v) \quad [\textsc{GetPut}]$
% \end{proof}

Now, let's see how we construct $pg$ recursively.

$\tab \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
    \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
    \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{fail}}$\\
The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. This is a trick since in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$. The last equality is relatively obivious.

$\tab \pg{Replace}{s}{v} = (v, s)$\\
$\tab \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
    \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
    \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
        \qtab \tab \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
        \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
$\tab \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
        \qtabs{2} (f_2 \ s, v)$\\
$\tab \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \qtabs{2} (s, g_2 \ v)$\\
Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.

$\tab \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
\tabs{2} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \text{if} \ {cond_{s} \ s} \\
\tab \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtab \tab \, \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\
\tab \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtab \tab \ \, \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
\tab \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
\tabs{2} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
\tab \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
\tab \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
\tab \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$

A restriction for $\pginline{Case}$ needs to be introduced here. We know that there is one entering condition and one exit condition when evaluating $\putbxinline{Case}$ as well as $\getbxinline{Case}$. If a tupling occurs, there will be 4 combinations from these conditions. This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified. The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper. This does not happen for the others which is used in the construction of $\pginline{Case}$.

$\tab \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{2} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{2} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{2} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{2} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{2} (s_3, v_2)$

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid. Due to \textsc{GetPut} law, we can use $get \ s$ to update the source $s$ in the $put$ direction. Then it is feasible to construct a $dummy$ from the source. The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{(construct\_dummy \ s)}$. This transformation is possible because of the \textsc{PutPut} law. Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$. So we can realize that there is no loss information when computing a composition.

\subsection{cpg}

When evaluating $\pginline{bx_1 \circ bx_2}$, there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$. If a given program is a left-associative composition, the number of $pg$ calls will be exponential. Therefore, the runtime inefficiency is inevitable. To solve that, we introduce a new function, $cpg$, accumulates changes in the source and the view. $\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values. The output of this function is a 4-tuple $(ks, kv, s, v)$. To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:\\ $fst = \lambda (x_1,x_2). x_1 , \quad snd = \lambda (x_1,x_2). x_2 , \quad con = \lambda ks_1. \lambda ks_2. \lambda x. ((ks_1 \ x),(ks_2 \ x))$

\begin{definition}
$\cpg{bx}{ks}{kv}{s}{v}$

$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$

$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$

$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \ s}{snd \ v};\\
    \qtab (con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, (s_1,s_2), (v_1,v_2))$

$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
    \qtab (f_2 \circ ks, kv, s, v)$

$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
    \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
    \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$

$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
    \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
    \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
    \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
    \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \text{return} \ (ks, kv, s', v')$

$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
    \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
    \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
\end{definition}

Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.\\
For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls. The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source. Because there is no real view here, we need to construct a dummy from the source. Then the continuation updating on this dummy should be initiated as an identity function. The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because this evaluated variable is not used in the later steps. In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input. It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$. After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.\\
Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:\\
    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
        \qtab (s; v)$\\
In general, the begining of a continuation should be an identity function. However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.\\
Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively. Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$. After the first two assignments in the definition of $cpg$ for the composition, we have: $ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))$ and $s_2 = [100,2,3]$ where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$. \\
Then: $ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$\\
If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$. This is an unexpected result when we target it to be the updated source. If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, the result will be what we want to see.$ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$. Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to retain the discarded things in the source.

\subsection{kpg}

The problem for $cpg$ lies in redundant computation during evaluating. For keeping away such computations, we introduce an extention named $kpg$. While $cpg$ evaluate values eagerly, $kpg$ does the opposite. Every values are evaluated lazily in a computation of $kpg$. The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values. The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
Suppose that we have a source $s_0$ and a view $v_0$. The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:\\
    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
        \qtab (ks' \ s; kv' \ v)$\\
The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.

\begin{definition}
$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$

$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
    \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$

$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$

$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
    \qtab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, con \ (ks_1' \circ fst) \ (ks_2' \circ snd),\\
    \qtab con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
    \qtab (s_1, s_2), (v_1,v_2))$

$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
    \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$

$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
    \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$

$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id ; \quad kv' \Leftarrow id;\\
    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
    \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
    \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' \ v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$

$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
    \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
    \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
    \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
\end{definition}

In the construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$. Except them, functions for the computation will be kept in $ks'$ and $kv'$. When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated. In $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$. It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.

\subsection{xpg}
The most imporant thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that is used to produce new source instead of using more recursive call. This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$. The second call $\kpginline{bx_2}$ just play an syntactic role while it requires many parameters. We can completely reduce this number to 2, including one for the source and one for the view as in a $pg$ call. This idea is realized by function $xpg$ as follows:

\begin{definition}
$\xpg{bx}{s}{v}$

$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$

$\xpg{bx_1 \circ bx_2}{s}{v} =\\
    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
    \qtab (ks_1 \ s_2, v_2)$
\end{definition}

Similar to $pg$, $\xpginline{bx}$ accepts a pair of the source and the view $(s,v)$ to produce the new pair. The constructions of $\xpginline{bx}$ when $bx$ is not a composition are the same as the ones of $\pginline{bx}$. Note that, $xpg$ is called recursively instead of $pg$. For $\xpginline{bx_1 \circ bx_2}$, we use two function calls and a function application to calculate the result. The first call and the function application come from $kpg$, while the second call is based on $pg$.

### 2019-11-15 paper/0.abst.tex
extract abstract from main.tex, update abstract

 Bidirectional transformation (BX) is a solution to the view update problem and widely used for synchronizing data. Its semantics and correctness are researched well, but the evaluation efficiency is not considered so far. As a first attempt we consider efficiency of evaluation in this work. To improve evaluation efficiency, we focus on the evaluation of BX compositions, because its evaluation often causes redundant evaluations. 
 For evaluating BX compositions efficiently, we introduce two memoization methods. The first method, minBigul\_m, uses simple memoization, but it works well for many BX programs. This keeps the evaluated results in a table and reuses them.  A disadvantage of this is cost for saving to a table and searching. When the inputs are large, the evaluation time becomes slow. To treat large inputs, we introduce the second method, xpg, that uses tupling, lazy update and lazy evaluation as optimizations. Lazy update keeps updates in closures and enables to use them later.
 The experiments of these methods show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.

### 2019-11-15 paper/main.tex
update result2.tex
extract abstract from main.tex, update abstract

@@
 \usepackage{pgffor}
+\usepackage{schemata}
 
@@
 
+\newcommand{\putm}[3]{put_m \, [\![#1]\!] \ #2 \ #3}
+\newcommand{\putminline}[1]{put_m \, [\![#1]\!]}
+\newcommand{\getm}[2]{get_m \, [\![#1]\!] \ #2}
+\newcommand{\getminline}[1]{get_m \, [\![#1]\!]}
+
+\newcommand{\getmh}[2]{get_{mh} \, [\![#1]\!] \ #2}
+\newcommand{\getmhinline}[1]{get_{mh} \, [\![#1]\!]}
+
 \newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
@@
   An efficient composition of bidirectional programs
-  by memoization
+  by memoization and lazy update
 }
@@
 %{Kanae Tsushima\inst{3}\orcidID{}}
-%{Bach-Trong Nguyen\inst{3}\orcidID{}}
+%{Bach-Trong Nguyen\inst{3}\orcidID{0000-0002-6342-6020}}
 \authorrunning{F. Author et al.}
@@
 %15--250 words.
-  Bidirectional transformation (BX) is a solution of view update problems and widely used for synchronizing data.
-Its semantics and correctness are researched well, but the evaluation efficiency is not considered well.
-As a first attempt we consider efficiency of evaluation in this work.
-To improve evaluation efficiency, we focus on the evaluation of BX compositions, because its evaluation often causes redundant evaluations.
-  For evaluating BX compositions efficiently, we introduce two memoization strategies.
-The first strategy, minBigul\_m, uses so simple memoization, but it works well for many BX programs.
-This keeps the evaluated results in a table and reuse them.
-A disadvantage of this is cost for saving to a table and searching.
-When the inputs are large, the evaluation time becomes slow.
-To treat large inputs, we introduce the second approach, xpg, that uses tupling and some optimization techniques.
-In this approach we shrink the size of the evaluated results and keep them in a closure.
-The experiments of these strategies show the following results: the original strategy is faster for right assoc compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.
-
+ \input{0.abst.tex}
  % We show experimental results for several BX programs with simple inputs and large inputs.
@@
 
-\input{3.pg.tex}
+\input{3.minbigul_m.tex}
 
-\input{4.cpg.tex}
-
-\input{5.kpg.tex}
-
-\input{6.xpg.tex}
+\input{4.xpg.tex}
 
@@
 
-\input{9.conclusion}
+\input{9.conclusion.tex}
 
@@
 
+\input{result2.tex}
+
 \end{document}

### 2019-11-19 paper/5.experiments.tex
update tex 20191119

\section{Experiments}

We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, Ocaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.

\subsection{Test cases}

We have conducted 12 test cases shown in Table \ref{tab:test-cases}. $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 12 functions mentioned in the table. Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

\begin{table}[hbt!]
    \centering
    \caption{Test cases}
    \label{tab:test-cases}
    \begin{tabular*}{\textwidth}{|l @{\extracolsep{\fill}}|l|l|c|c|c|c|}
        \hline
        \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Type}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{4-7} 
        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
        1 & lcomp-phead & manual & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        2 & rcomp-phead & manual & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
        3 & lcomp-ptail & manual & [1,$\ldots$,n+1] & [1,$\ldots$,10] & [1,$\ldots$,n+1,1,$\ldots$,10] & [\ ] \\ \hline
        4 & rcomp-ptail & manual & [1,$\ldots$,n+1] & [1,$\ldots$,10] & [1,$\ldots$,n+1,1,$\ldots$,10] & [\ ] \\ \hline
        5 & lcomp-ptail-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
        6 & rcomp-ptail-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
        7 & lcomp-bsnoc & manual & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
        8 & rcomp-bsnoc & manual & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
        9 & lcomp-bsnoc-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
        10 & rcomp-bsnoc-ldata & maunal & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ \\ \hline
        11 & breverse & self & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
        12 & breverse-ldata & self & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
    \end{tabular*}
\end{table}

The first 10 cases maunally use $n$ compose operators to make a composition of $n + 1$ similar pure programs. We use the term \textit{pure} here to refer to programs that are not self-compose. The prefix \textit{lcomp} and \textit{rcomp} denote the associate precedences of a manual composition. The suffix \textit{ldata} indicates that the input size is large. The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = \underbrace{[T,\ldots,T]}_{\text{n times}}$,\ $T = \underbrace{[A,\ldots,A]}_{\text{10 times}}$,\ $A = [1,\ldots,5]$

If the associative type and input size are omitted, the table, in general, will comprise 4 programs: composition of many \textit{phead}s, composition of many \textit{ptail}s, composition of many \textit{bsnoc}s and \textit{breverse}.

We introduced the composition $phead \circ phead$ earlier. The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view. We now review the behavior of the remaining 3 programs.

\tab $\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$. $\getbx{ptail}{s}$ return the tail of the source list $s$. The compositions of many $ptail$, in the \textit{put} direction, replace a part of the tail of the source list by the view list, in the \textit{get} direction, returns the such tail from the the source.

\tab $\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$ as two same length lists, to produce a new list by moving the last element of $v$ to its first position. $\getbx{ptail}{s}$ return another list by moving the first element of the list $s$ to its end position. The compositions of many \textit{ptail}, in the \textit{put} direction, create a permutation of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a permutation of the source list.

\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function for \textit{foldr} which is an important higher-order function on lists. This case is more natural than the other since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}. It will look like a composition of more BX programs at some point if we slow down the evaluation. Note that $\circ$ has a higher priority than $\times$. This leads to seemingly impossible to transform \textit{bReverse} as well as \textit{bFoldr} from the left associative style to the right one.

\subsection{Results and discussions}

\input{result2.tex}

Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 12 test cases. $put$ works more efficiently in the cases of right associative compositions, rather than the left ones. This makes sense because the number of \textit{get} when evaluating a right associative composition is linear by the number of compositions, while the respective number when computing a left associative composition is significantly larger. For self-composed program, \textit{breverse}, it could be impossible to change the associate precedence from left to right. Then, $put$ works less efficiently than both $put_m$ and $xpg$. In general, for left associative compositions, if the input size is not too large, $put_m$ is the fastest method. However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table. At that time, $xpg$ proved to be the most effective method.

### 2019-11-19 paper/0.abst.tex

@@
- Bidirectional transformation (BX) is a solution to the view update problem and widely used for synchronizing data.
-Its semantics and correctness are researched well, but the evaluation efficiency is not considered so far.
-As a first attempt we consider efficiency of evaluation in this work.
-To improve evaluation efficiency, we focus on the evaluation of BX compositions, because its evaluation often causes redundant evaluations.
- For evaluating BX compositions efficiently, we introduce two memoization methods.
-The first method, minBigul\_m, uses simple memoization, but it works well for many BX programs.
-This keeps the evaluated results in a table and reuses them.
-A disadvantage of this is cost for saving to a table and searching.
-When the inputs are large, the evaluation time becomes slow.
-To treat large inputs, we introduce the second method, xpg, that uses tupling, lazy update and lazy evaluation as optimizations.
-Lazy update keeps updates in closures and enables to use them later.
- The experiments of these methods show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.
+ Bidirectional transformations (BX) are a solution to the view update problem and widely used for synchronizing data.
+The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
+In this paper, as a first step, we study different evaluation methods to optimized their evaluation.
+We focus on the interpretive evaluation of BX compositions because we found that these compositions are an important cause of redundant computations.
+ For evaluating BX compositions efficiently, we investigate two memoization methods.
+The first method, minBigul\_m, uses memoization, which improves the runtime of many BX programs by keeping intermediate results for later reuse.
+A disadvantage is the familiar tradeoff for keeping and searching values in a table.
+  When inputs become large, the overhead increases and the effectiveness decreases.
+To deal with large inputs, we introduce the second method, xpg, that uses tupling, lazy update and lazy evaluation as optimizations.
+Lazy updates delay updates in closures and enables to use them later.
+ Both evaluation methods were fully implemented for MiniBiGul and experimentally evaluated.
+Our experiments show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.

### 2019-11-19 paper/6.related-work.tex
add authors/change the file names of related work & conclusion

\section{Related Work}

\subsubsection*{BX langauges and implementations}

There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF} (memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?) based on several application scenarios and BX lenses works \cite{} (add papers for lenses). The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{BXcomp}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimized their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Section 1) and our methods (in Section 3 and 4). From experimental results, we find that there is no best approach for all BX programs. It is important to choose good method depending on the BX programs and their inputs. Currently we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
% its implementation uses ``not keeping any intermediate states and obtaining them by evaluation when they are needed'' strategy. Other BX languages might have the same problem, and our approach can be applicable.

\subsubsection*{Optimization techniques}

In this paper, we use several optimization techniques: tupling, lazy update, and lazy computation (\textcolor{red}{also fusion?}).

Tupling \cite{tupling} is an optimization technique by combining several computations. Although tupling is usually used for reducing the evaluation cost\cite{} \cite{}, we use tupling to reduce the size of the things need to keep.

We introduce lazy update to remove repeated evaluation of put. This is different from lazy evaluation. In lazy evaluation we have everything that is needed to evaluate. In our lazy update, because view information is missing, the delay is until we can obtain the view.

, because it is delayed to 
Lazy update .. delta lenses?

Lazy evaluation \cite{} is well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

\begin{itemize}
\item This part looks just short introduction of  techniques .. 
\end{itemize}

%\begin{itemize}
    %\item BiGUL papers (PEPM, POPL, tutorial?)
    %\item the first BX paper
    %\item get-based BX
    %\item put-based BX?
%    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
%    \item optimization techniques : tupling, lazy computation, fusion transformation?
%    \item (view update problem paper)
%\end{itemize}

### 2019-11-19 paper/main.tex
update tex 20191119
add authors/change the file names of related work & conclusion

@@
 %
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and
+  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and
+  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
+  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
 
-%{Zhenjiang Hu\inst{1}\orcidID{0000-0002-9034-205X}}
-%{Robert Gl\"{u}ck\inst{2}\orcideID{0000-0001-6990-3935}}
-%{Kanae Tsushima\inst{3}\orcidID{}}
-%{Bach-Trong Nguyen\inst{3}\orcidID{0000-0002-6342-6020}}
-\authorrunning{F. Author et al.}
+\authorrunning{K. Tsushima et al.}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\institute{National Institute of Informatics, Japan
+  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
+  University of Copenhagen, Denmark
+  \email{glueck@di.ku.dk } \and
+  Peking University, China
+  \email{huzj@pku.edu.cn}}
+
 %
@@
 
-\newpage
-
 \input{2.minbigul.tex}
@@
 
-\input{7.experiments.tex}
+\input{5.experiments.tex}
 
-\input{8.related-work.tex}
+\input{6.related-work.tex}
 
-\input{9.conclusion.tex}
+\input{7.conclusion.tex}
 
@@
 %
+
+\tiny
+
 \begin{thebibliography}{8}
@@
 
-\input{result2.tex}
-
 \end{document}

### 2019-11-19 paper/0.abst.tex
update fig + tex 20191119

@@
  For evaluating BX compositions efficiently, we investigate two memoization methods.
-The first method, minBigul\_m, uses memoization, which improves the runtime of many BX programs by keeping intermediate results for later reuse.
+The first method, minBiGUL$_m$, uses memoization, which improves the runtime of many BX programs by keeping intermediate results for later reuse.
 A disadvantage is the familiar tradeoff for keeping and searching values in a table.
   When inputs become large, the overhead increases and the effectiveness decreases.
-To deal with large inputs, we introduce the second method, xpg, that uses tupling, lazy update and lazy evaluation as optimizations.
+To deal with large inputs, we introduce the second method, $xpg$, that uses tupling, lazy update and lazy evaluation as optimizations.
 Lazy updates delay updates in closures and enables to use them later.
- Both evaluation methods were fully implemented for MiniBiGul and experimentally evaluated.
-Our experiments show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL\_m is faster for not-large inputs and xpg is faster for large inputs.
+ Both evaluation methods were fully implemented for minBiGUL and experimentally evaluated.
+Our experiments show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL$_m$ is faster for not-large inputs and $xpg$ is faster for large inputs.

### 2019-11-19 paper/1.introduction.tex
update fig + tex 20191119

@@
 BX provides two functions: $get$ and $put$.
-The following is one example of $pHead$'s evaluation.
+Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.
 
-\begin{center}
-  \includegraphics[width=5cm]{./temp_fig/bxexample1.eps}
-\end{center}
-%\end{picture}
+\begin{figure}[!htb]
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[height=3.5cm]{./fig/fig1.png}
+    \caption{Evaluating $phead$}
+    \label{fig:eval-phead}
+  \end{minipage}\hfill
+  \begin{minipage}{0.7\textwidth}
+    \centering
+    \includegraphics[height=3.5cm]{./fig/fig2.png}
+    \caption{Evaluating $phead \circ phead \circ phead$}
+    \label{fig:eval-comp-phead}
+  \end{minipage}
+\end{figure}
 
@@
 After the view \texttt{v} is obtained, the user can modify the view.
-In this case, he or she modified the view to \texttt{8} from \texttt{1}.
+In this case, he or she modified the view to \texttt{100} from \texttt{1}.
 From this updated view, how can we obtain new updated source? This is ``view update problem.'' For this, BX provides $put$, update on the original source.
-In this example, $put$ of $pHead$ constructs a new list \texttt{[8,2]} from the updated view ($v'$) and the original source ($s$).
+In this example, $put$ of $pHead$ constructs a new list \texttt{[100,2]} from the updated view ($v'$) and the original source ($s$).
 %\texttt{put pHead [1,2] 8 = [8,2]}
 
-For more various evaluations, we can use composition of programs.
-Let us consider another BX example program, three composition of $pHead$: $pHead \circ pHead \circ pHead$.
-The following is one example of its evaluation.
-
-\begin{center}
-  \includegraphics[width=10cm]{./temp_fig/bxexample2.eps}
-\end{center}
-
-In this example, the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]}.
+For more various evaluations, we can use the composition of programs.
+Let us consider another example, a composition of 3 $pHead$s: $pHead \circ pHead \circ pHead$.
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of this program where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \textit{100}.
 By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}.
@@
 One is ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i1} and \texttt{i2}.
-If we choose this strategy, the number of $get$s will be quadratic.
+If we choose this strategy, the number of $get$s will be linear in the number of compositions.
 This is a problem, because the evaluation will be slow if the BX programs include many compositions.
@@
 
-As an example, let us consider the previous example again: $pHead \circ pHead \circ pHead$.
+As an example, let's see the evaluation of $pHead \circ pHead \circ pHead$ in another way as shown in figure \ref{fig:eval-comp-phead-2}
 
-\begin{center}
-  \includegraphics[width=9cm]{./temp_fig/bxexample3.eps}
-\end{center}
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[height=5cm]{./fig/fig3.png}
+  \caption{Evaluating $phead \circ phead \circ phead$ by keeping intermediate states}
+  \label{fig:eval-comp-phead-2}
+\end{figure}
 
-In the figure, \texttt{c1} and \texttt{c2} are complements.
+In figure \ref{fig:eval-comp-phead-2}, \texttt{c1} and \texttt{c2} are complements.
 The points in this figure are two.

### 2019-11-19 paper/2.minbigul.tex
update tex 20191114
update fig + tex 20191119

@@
 To help make demonstration more direct, we provide the following alternatives representation: 
-
-    $\tab Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2,
-    \tab Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
+    $\quad Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2,
+    \quad Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
 
@@
 
-Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:
+% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:
 
-$\tab \putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10 \qtab \ \getbx{Skip \ (\lambda x.(x*x))}{10} = 100\\
-    \tab \putbx{Skip \ (\lambda \_.())}{1}{()} = 1 \qtabs{3} \getbx{Skip \ (\lambda \_.())}{1} = ()\\
-    \tab \putbx{Replace}{1}{100} = 100 \qtabs{3} \getbx{Replace}{1} = 1
-$
+% $\tab \putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10 \qtab \ \getbx{Skip \ (\lambda x.(x*x))}{10} = 100\\
+%     \tab \putbx{Skip \ (\lambda \_.())}{1}{()} = 1 \qtabs{3} \getbx{Skip \ (\lambda \_.())}{1} = ()\\
+%     \tab \putbx{Replace}{1}{100} = 100 \qtabs{3} \getbx{Replace}{1} = 1
+% $
 
-Now, let's consider the definition of $phead$ in minBiGUL:
+We now consider the definition of $phead$ in minBiGUL:
 

### 2019-11-19 paper/5.experiments.tex
update experiment 20191119
update fig + tex 20191119

@@
 
-\input{result2.tex}
+\input{result.tex}
 
@@
 $put$ works more efficiently in the cases of right associative compositions, rather than the left ones.
-This makes sense because the number of \textit{get} when evaluating a right associative composition is linear by the number of compositions, while the respective number when computing a left associative composition is significantly larger.
+This makes sense because the number of \textit{get} when evaluating a right associative composition is linear in the number of compositions, while the respective number when computing a left associative composition is significantly larger.
 For self-composed program, \textit{breverse}, it could be impossible to change the associate precedence from left to right.

### 2019-11-20 paper/6.related-work.tex
revise contents

@@
 
-There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF} (memo: BXtend, eMoflon, EVL+Strace, JTL, NMF.
-SDMLib should be added?) based on several application scenarios and BX lenses works \cite{} (add papers for lenses).
+There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF}
+%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
+based on several application scenarios and BX lenses works \cite{} (add papers for lenses).
 The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
@@
 It is important to choose good method depending on the BX programs and their inputs.
-Currently we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
+In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
 % its implementation uses ``not keeping any intermediate states and obtaining them by evaluation when they are needed'' strategy.
@@
 Tupling \cite{tupling} is an optimization technique by combining several computations.
-Although tupling is usually used for reducing the evaluation cost\cite{} \cite{}, we use tupling to reduce the size of the things need to keep.
+Although tupling is usually used for reducing the evaluation cost\cite{}, we use it to reduce the size of the things need to keep.
 
@@
 In lazy evaluation we have everything that is needed to evaluate.
-In our lazy update, because view information is missing, the delay is until we can obtain the view.
-
-, because it is delayed to 
-Lazy update .. delta lenses?
+In our lazy update, because view information is missing, we postpone the evaluation.
+(\textcolor{red}{Relation between delta lenses?})
 
@@
 If we use Haskell, a lazy evaluated language, we do not need to introduce them.
-
-\begin{itemize}
-\item This part looks just short introduction of  techniques .. 
-\end{itemize}
 

### 2019-11-20 paper/7.conclusion.tex
revise contents

\section{Conclusion and Future Work}

In this paper, we focus on efficiency of composition of BX programs.
To achieve fast evaluation, we introduce two different methods using memoization.
From the experimental results, we know that there is no best approach for all BX programs. However, if programmer can choose one method based on the BX programs and inputs, we can use better evaluation methods.

We will continue our work on the following points. First is extention of our language. Currently, because $pg$ requires very-well-behaved property

\begin{itemize}
\item overcome our limitations
  \begin{itemize}
  \item very-well-behaved $\to$ only xpg?
  \item In case expressions, path's problem $\to$ only xpg?
  \item How to obtain $\to$ only xpg?
  \end{itemize}
  \item tests for big BX programs
\item Type system \& Typing -- for safety
\end{itemize}

%The essential finding comes from the idea of tupling: in very-well behaved BX programs we can use $put$ as a compliment function for $get$, and vice versa.
%Based on the idea, we introduced $pg$, and improved it by several optimization techniques. From experimental results, our fastest approach $xpg$ is faster than other approaches for non purely right associative compositions. For right associative compositions, the original approach (miniBiGUL) is faster than $xpg$ because $xpg$ has some overhead cost. However this is not a problem, because usually programs are mix of left and right associative. If programmers know that their programs are purely right associative, they can choose miniBiGUL.

%We will continue our work on the following points. First, our target language is limited to very-well behaved, because our main idea requires the put-put property. However, for practical programs, we need to extend our work to overcome the current limitations.

%\begin{itemize}
%\item Extend our approach -- overcome our limitations
%  \begin{itemize}
%  \item treat well-behaved programs -- How to treat adaptive cases
%  \item In case expressions, the programs that use the different paths (put and get)
%  \end{itemize}
%\item How to obtain dummy?
%\item Type system \& Typing -- for safety
%\end{itemize}

### 2019-11-20 paper/5.experiments.tex
add one sentence

@@
         9 & lcomp-bsnoc-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
-        10 & rcomp-bsnoc-ldata & maunal & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ \\ \hline
+        10 & rcomp-bsnoc-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ \\ \hline
         11 & breverse & self & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
@@
 The first 10 cases maunally use $n$ compose operators to make a composition of $n + 1$ similar pure programs.
+The aim of this manually produced program is to know which method is better for specific composition types and inputs.
 We use the term \textit{pure} here to refer to programs that are not self-compose.

### 2019-11-22 paper/5.experiments.tex
update experiment: table + charts (only lcomp)

--- 2019-11-19 paper/5.experiments.tex
@@
 
-We have conducted 12 test cases shown in Table \ref{tab:test-cases}.
+We have conducted 11 test cases shown in Table \ref{tab:test-cases}.
 $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 12 functions mentioned in the table.
+This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 11 functions mentioned in the table.
 Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
@@
         & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lcomp-phead & manual & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        2 & rcomp-phead & manual & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        3 & lcomp-ptail & manual & [1,$\ldots$,n+1] & [1,$\ldots$,10] & [1,$\ldots$,n+1,1,$\ldots$,10] & [\ ] \\ \hline
-        4 & rcomp-ptail & manual & [1,$\ldots$,n+1] & [1,$\ldots$,10] & [1,$\ldots$,n+1,1,$\ldots$,10] & [\ ] \\ \hline
-        5 & lcomp-ptail-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
-        6 & rcomp-ptail-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
-        7 & lcomp-bsnoc & manual & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
-        8 & rcomp-bsnoc & manual & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
-        9 & lcomp-bsnoc-ldata & manual & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
-        10 & rcomp-bsnoc-ldata & maunal & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ \\ \hline
-        11 & breverse & self & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
-        12 & breverse-ldata & self & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
+        1 & lcomp-replace & straightline & 1 & 100 & 100 & 1 \\ \hline
+        2 & lcomp-replace-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
+        3 & lcomp-phead-ldata & straightline & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
+        4 & lcomp-ptail & straightline & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
+        5 & lcomp-ptail-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
+        6 & lcomp-bsnoc & straightline & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
+        7 & lcomp-bsnoc-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
+        8 & bmapreplace & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] \\ \hline
+        9 & bmapreplace-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
+        10 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
+        11 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
     \end{tabular*}
@@
 
-The first 10 cases maunally use $n$ compose operators to make a composition of $n + 1$ similar pure programs.
+The first 7 cases maunally use $n$ compose operators to make straightline compositions from $n + 1$ similar pure programs.
 We use the term \textit{pure} here to refer to programs that are not self-compose.

### 2019-11-22 paper/0.abst.tex
update based on the discussion

@@
 In this paper, as a first step, we study different evaluation methods to optimized their evaluation.
-We focus on the interpretive evaluation of BX compositions because we found that these compositions are an important cause of redundant computations.
+We focus on the interpretive evaluation of BX compositions because we found that these compositions are an important cause of redundant computations if the compositions are not right associative.
  For evaluating BX compositions efficiently, we investigate two memoization methods.
@@
 Lazy updates delay updates in closures and enables to use them later.
- Both evaluation methods were fully implemented for minBiGUL and experimentally evaluated.
-Our experiments show the following results: the original method that is used in the language is faster for right associative compositions, minBiGUL$_m$ is faster for not-large inputs and $xpg$ is faster for large inputs.
+  Both evaluation methods were fully implemented for minBiGUL.
+The experimental results show that our methods are faster than the original method of BiGUL for the non-right associative compositions.
+minBiGUL$_m$ is faster for not-large inputs and $xpg$ is faster for large inputs.

### 2019-11-22 paper/1.introduction.tex
update fig.eps
update based on the discussion

--- 2019-09-27 paper/1.introduction.tex
@@
 
-In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
+In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
 % A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation is introduced \cite{}.
-It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
-% The first paper of bidirectional transformation
+As a solution for this problem, bidirectional transformation (BX) is introduced.
+As an example, let us consider a small BX program of $pHead$\footnote{The actual program is shown in the next section.}.
+BX provides two functions: $get$ and $put$.
+Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.
 
-There are two types of bidirectional transformation.
-One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$.
-A problem of this approach is that it is possible to derive many correct $put$s.
-To choose one put, hulistics and programmer's intentions are needed.
-Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$.
-From a $put$ we can derive an unique $get$.
+\begin{figure}[!htb]
+  \begin{minipage}{0.3\textwidth}
+    \centering
+    \includegraphics[height=3.5cm]{./fig/fig1.eps}
+    \caption{Evaluating $phead$}
+    \label{fig:eval-phead}
+  \end{minipage}\hfill
+  \begin{minipage}{0.7\textwidth}
+    \centering
+    \includegraphics[height=3.5cm]{./fig/fig2.eps}
+    \caption{Evaluating $phead \circ phead \circ phead$}
+    \label{fig:eval-comp-phead}
+  \end{minipage}
+\end{figure}
 
-BiGUL \cite{} is one of put-based bidirectional transformation language.
-It is fastest bidirectional language in the world now.
-However, it contains a problem when the BiGUL programs including many compositions.
-BiGUL does not store the intermediate states when $get$s are re-evaluated so many times.
-In this case the number of get is quadratic.
+Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $pHead$ picks the first element of the given original source as a view.
+%\texttt{get pHead [1,2] = 1}.
+After the view \texttt{v} is obtained, the user can modify the view.
+In this case, he or she modified the view to \texttt{100} from \texttt{1}.
+From this updated view, how can we obtain new updated source? This is ``the view update problem.'' For this, BX provides $put$, an update function on the original source.
+In this example, $put$ of $pHead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
+%\texttt{put pHead [1,2] 8 = [8,2]}
 
-\vspace{5mm}
+In BX programs, composition of programs is essential for writing various interesting evaluations. Because recursions are achieved by compositions, large number of programs include compositions.
+In the BX lens papers \cite{} compositions are also known as central part of BX programs.
+Because the programs including recursions can be complicated easily, we will use a tiny BX program that contains compositions as a running example: two versions of three composition of $pHead$: ($pHead \circ pHead) \circ pHead$ (we call lp3) and $pHead \circ (pHead \circ pHead$) (we call rp3).
 
-Figure that shows many gets re-evaluation
+%For more various evaluations, we can use the composition of programs.
+Let us consider another example, a composition of 3 $pHead$s: $pHead \circ pHead \circ pHead$.
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of this program where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
+%By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}.
+To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
+The first put evaluation is from \texttt{i2} and \texttt{v'} and we can obtain \texttt{i2'}.
 
-\vspace{5mm}
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of this program where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
+By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}.
+To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
+The first put evaluation is from \texttt{i$_2$} and \texttt{v'} and we can obtain \texttt{i$_2$'}.
 
-% general, concrete problem
-% get-based also have the same problem. even in lenses. more general problem in
-% The example should be from lenses.
-% traversal, not exactly the same computation
+First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i1} and \texttt{i$_2$}.
+This strategy is used in a BX language, BiGUL \cite{}.
+The merit of this strategy is clear semantics and easy to implement.
+The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative.
+In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}.
+In total, three $get$s are evaluated.
+We explain this method in Section 2.
 
-To solve this problem and make BiGUL evaluation efficient, we use two techniques.
-The first technique is tupling, well-known technique for efficiency \cite{}.
-Its basic idea is to reduce the common computations by combining two evaluations.
-In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$.
-Thanks to tupling, we do not lose information like intermediate states.
-The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.
+In the simplest cases, we can transform left associative compositions to right associative.
+However, this is not always possible.
+Therefore it is improtant to have a fast method for left associative compositions.
+For this, we introduce two methods that use memoization.
+The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this strategy, therefore this strategy is faster for many BX programs.
+However, this strategy needs overhead for saving to a table and finding from the table.
+Especially, when we use large inputs, it takes time.
+We explain this method in Section 3.
 
-The rest of this paper is the following.
-We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2.
-Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
-We apply three optimizations to $pg$ in sequence.
-The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4.
-The second one is lazy evaluation to avoid redundant computations, and we obtain $kpg$ in Section 5.
-The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs.
-In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
-In Section 8 we discuss about related work, and conclude in Section 9.
+To treat large inputs, we introduce the second method that uses memoization: ``keeping complements in a closure and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states.
+%Therefore this strategy solves the previous two problems.
+Readers might already know that it is a hard problem to obtain complements \cite{} in general. In our work, thanks to the following finding, it is not hard:
 
-% Story ->
-% reversible is needed not to lose any information -> pg is introduced.
-% pg is not good for lassoc -> cpg (xpg_2) is introduced.
-% cpg can be improved -> kpg(xpg) is introduced.
-% kpg is slower than pg for rassoc. combined version -> xpg.
+\vspace{2mm}
+In very-well behaved BX programs, 
+$put$ is a complement function for $get$.
+%, $get$ can be a complement function for $put$.
+\vspace{2mm}
 
-% Evaluation result: Comparison with OCaml implementation of minBiGUL
+For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the original data.
+As an example, let us consider the previous example again: $pHead \circ (pHead \circ pHead)$ (rp3).
 
-%\begin{itemize}
-%\item Importance of BX, BX is a solution of view update problem in database.
-%\item Explanation of put-based BX: BiGUL.
-%\item Current status of BiGUL: Fastest BX language in the world
-%\item But there is a problem: Efficiency of compose evaluation. The current implementation of BiGUL does not save the intermediate states, the number of get is quadratic. This is not good.
-%\item To solve this problem we use an idea: introduce pg : combination of put and get. Then, no information will be lost in a specific condition.
-%  an idea from reversible computation: not to lose any information.
-%\item We extend pg with several ideas to produce faster implementation.
-%\item 
-%\end{itemize}
+As an example, let's see the evaluation of $pHead \circ pHead \circ pHead$ in another way as shown in figure \ref{fig:eval-comp-phead-2}.
 
-% get & put independently, but for efficient computation
-% we need to tuple them to help composition
-% 2 or 3 examples to show tupling solves the problem.
-% other examples are for showing kpg and cpg is needed.
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[height=5cm]{./fig/fig3.eps}
+  \caption{Evaluating $phead \circ phead \circ phead$ by keeping intermediate states}
+  \label{fig:eval-comp-phead-2}
+\end{figure}
 
-\subsection{Contribution}
+\texttt{c1} and \texttt{c2} are complements.
+The points in this figure are two.
+After evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c1} and \texttt{i1}.
+%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
+Actually, the simple combined $pg$ is not effective for left associative compositions.
+To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation.
+We explain the second method in Section 4.
+
+In Section 5, we show experiemntal results of the approaches. We discuss about related work in Section 6, and conclude in Section 7.
+
+The contributions of this paper are the following.
 
 \begin{itemize}
-\item Improvement of evaluation efficiency
-  \begin{itemize}
-  \item Quadratic to exponential? 
-  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
+\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions. Our methods are similar or better than the original method in BiGUL for all test cases.
+\item We show that memoization is effective for achieving evaluation efficiency in BX languages.
+\item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt. 
+\item Although we focus on a BX language BiGUL in this paper, these techniques can be potentially used in other BX languages. 
+  % \begin{itemize}
+    % Although the first strategy is implemented in BiGUL \cite{}, the second and third strategies are introduced by this paper.
+%  \item Thanks to introduction of the strategies, it is possible to compare the evaluation strategies.
+%  \item Improvement of evaluation efficiency
+%  \end{itemize}
+%\item Approaches used in strategy 3
+%  \begin{itemize}
+%  \item 
+% \item Optimization tequniques by tupling, and lazy update
+%  \item These tequniques can be potentially used in other BX languages
+%  \end{itemize}
 \end{itemize}
-\item Get and put more tight
-
-\item (Tupling introduces some constructions)
-  \begin{itemize}
-  \item A pair of put and get is self-inverse (involution).
-  \end{itemize}
-\item careful lazyness is needed in our implementation for efficiency
-%  lazy update \& lazy computation are also good optimization for BX.
-%\item Reveal a part of relationship between BX and Reversible.
-%  \begin{itemize}
-%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
-%  \item 
-%  \item If we can obtain good parameter (dummy) for a BiGUL program,
-%    put(s, dummy) can be a complement function.
-  \end{itemize}
-
-%\subsection{Construction of this paper}

### 2019-11-22 paper/2.minbigul.tex
fix typos

@@
 
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satify the round-trip laws as follows:\\
+BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
     $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
@@
 
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satify two above laws.
+MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws.
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
@@
 
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinatator without adaptive cases or a composition of some minBiGUL programs.
+A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinator without adaptive cases or a composition of some minBiGUL programs.
 
-For source/view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
+For source/view rearrangement, BiGUL uses just one lambda expression to express how to deconstruct as well as reconstruct data.
 It is a kind of bijection.
@@
 
-In defintions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
+In definitions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
 This statement is useful to describe many functions related to $Case$ in this paper.

### 2019-11-23 paper/0.abst.tex
fix typo

@@
  Bidirectional transformations (BX) are a solution to the view update problem and widely used for synchronizing data.
-The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
-In this paper, as a first step, we study different evaluation methods to optimized their evaluation.
+The semantics and correctness of bidirectional programs have been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
+In this paper, as a first step, we study different evaluation methods to optimize their evaluation.
 We focus on the interpretive evaluation of BX compositions because we found that these compositions are an important cause of redundant computations if the compositions are not right associative.
@@
 Lazy updates delay updates in closures and enables to use them later.
-  Both evaluation methods were fully implemented for minBiGUL.
-The experimental results show that our methods are faster than the original method of BiGUL for the non-right associative compositions.
-minBiGUL$_m$ is faster for not-large inputs and $xpg$ is faster for large inputs.
+  Both evaluation methods were fully implemented for minBiGUL. The experimental results show that our methods are faster than the original method of BiGUL for the non-right associative compositions.
+%minBiGUL$_m$ is faster for not-large inputs and $xpg$ is faster for large inputs.

### 2019-11-25 paper/1.introduction.tex
revise

@@
 As a solution for this problem, bidirectional transformation (BX) is introduced.
-As an example, let us consider a small BX program of $pHead$\footnote{The actual program is shown in the next section.}.
+As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}.
 BX provides two functions: $get$ and $put$.
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
   \begin{minipage}{0.3\textwidth}
@@
 
-Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $pHead$ picks the first element of the given original source as a view.
+Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
 %\texttt{get pHead [1,2] = 1}.
@@
 In this case, he or she modified the view to \texttt{100} from \texttt{1}.
-From this updated view, how can we obtain new updated source? This is ``the view update problem.'' For this, BX provides $put$, an update function on the original source.
-In this example, $put$ of $pHead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
+From this updated view, how can we obtain new updated source? This is ``the view update problem.'' For doing this, BX provides $put$, an update function on the original source.
+In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
 %\texttt{put pHead [1,2] 8 = [8,2]}
@@
 In the BX lens papers \cite{} compositions are also known as central part of BX programs.
-Because the programs including recursions can be complicated easily, we will use a tiny BX program that contains compositions as a running example: two versions of three composition of $pHead$: ($pHead \circ pHead) \circ pHead$ (we call lp3) and $pHead \circ (pHead \circ pHead$) (we call rp3).
+In this section we use a tiny BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
+The running examples are two versions of three composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
 
@@
 
-Figure \ref{fig:eval-comp-phead} illustrates the evaluation of this program where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
-By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}.
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
+By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, \texttt{1}.
 To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
@@
 
-First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i1} and \texttt{i$_2$}.
-This strategy is used in a BX language, BiGUL \cite{}.
-The merit of this strategy is clear semantics and easy to implement.
+First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}.
+This method is used in a BX language, BiGUL \cite{}.
+The merit of this method is clear semantics and easy to implement.
 The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative.
@@
 
-In the simplest cases, we can transform left associative compositions to right associative.
+If the programs are simple like these running examples, we can transform left associative compositions to right associative.
 However, this is not always possible.
-Therefore it is improtant to have a fast method for left associative compositions.
-For this, we introduce two methods that use memoization.
-The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this strategy, therefore this strategy is faster for many BX programs.
-However, this strategy needs overhead for saving to a table and finding from the table.
-Especially, when we use large inputs, it takes time.
+Therefore it is important to have a fast evaluation method for left associative compositions.
+For achieving this, we introduce two methods that use memoization.
+The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs.
+However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table.
+Especially, when we use large inputs, the evaluation time will be long.
 We explain this method in Section 3.
@@
 %Therefore this strategy solves the previous two problems.
-Readers might already know that it is a hard problem to obtain complements \cite{} in general. In our work, thanks to the following finding, it is not hard:
+Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
 
@@
 
-For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the original data.
-As an example, let us consider the previous example again: $pHead \circ (pHead \circ pHead)$ (rp3).
-
-As an example, let's see the evaluation of $pHead \circ pHead \circ pHead$ in another way as shown in figure \ref{fig:eval-comp-phead-2}.
-
-\begin{figure}[!htb]
+\begin{figure}[!t]
   \centering
@@
 
-\texttt{c1} and \texttt{c2} are complements.
-The points in this figure are two.
-After evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c1} and \texttt{i1}.
+For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
+As an example, let us consider the previous example again: $phead \circ (phead \circ phead)$ (rp3) in figure \ref{fig:eval-comp-phead-2}.
+\texttt{c$_1$} and \texttt{c$_2$} are complements.
+The points in this figure is two.
+First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}.
+Second, complements are smaller than the intermediate states in the previous figure.
 %This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
-Actually, the simple combined $pg$ is not effective for left associative compositions.
+Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure.
 To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation.
-We explain the second method in Section 4.
+We explain the second method and optimization in Section 4.
 
-In Section 5, we show experiemntal results of the approaches. We discuss about related work in Section 6, and conclude in Section 7.
+In Section 5, we show experimental results of the approaches. We discuss about related work in Section 6, and conclude in Section 7.
 
@@
 \begin{itemize}
-\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions. Our methods are similar or better than the original method in BiGUL for all test cases.
+\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions.
+Our methods are better than or similar to the results by the original method in BiGUL for all test cases.
 \item We show that memoization is effective for achieving evaluation efficiency in BX languages.

### 2019-11-25 paper/2.minbigul.tex
Update 2.minbigul.tex

@@
 
-MinBiGUL, our target language in this paper, is a subset of BiGUL [?] which is a simple yet powerful putback-based bidirectional language.
+MinBiGUL, our target language in this paper, is a subset of BiGUL which is a simple yet powerful putback-based bidirectional language.
 BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
@@
 
-BiGUL is well-behaved since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
+BiGUL is well-behaved [?] since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
     $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
@@
 
-The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source.
-The \textsc{PutGet} law means that we can recover the modified view by applying the forward transformation to the updated source.
+The \textsc{GetPut} [?] law means that if there is no change to the view, there should be no change to the source.
+The \textsc{PutGet} [?] law means that we can recover the modified view by applying the forward transformation to the updated source.
 
@@
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satify one more law, \textsc{PutPut}, like the following:\\
+Then $put$ and $get$ satisfy one more law, \textsc{PutPut} [?], like the following:\\
     $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
@@
 The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates.
-Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.
+Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved [?].
 
@@
 
-For source/view rearrangement, BiGUL uses just one lambda expression to express how to deconstruct as well as reconstruct data.
+For source/view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
 It is a kind of bijection.
@@
 
-To help make demonstration more direct, we provide the following alternatives representation: 
-    $\quad Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2,
-    \quad Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$
-
+To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$.
 In general, $\circ$ has a higher priority than $\times$.
-Since only consider a product of two minBiGUL programs is considered, it is unnecessary to define the associativity precedence of $\times$.
-The one of $\circ$ may be left or right, however we do not declare the default.
+Their associativity precedence can be either left or right or mixture, but are not set by default.
+We need to explicitly write programs that use these operators.
 
@@
 
-% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obivious things as follows:
+% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obvious things as follows:
 

### 2019-11-25 paper/2.minbigul.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos

@@
 
-For source/view rearrangement, BiGUL use just one lambda expression to express how to deconstruct as well as reconstruct data.
+For source/view rearrangement, BiGUL uses just one lambda expression to express how to deconstruct as well as reconstruct data.
 It is a kind of bijection.
@@
 
-In definitions 1 and 2, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
+We define the semantics of $put$ and $get$ as in definitions 1 and 2 respectively.
+Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.
+The later definitions also use this convention.
+
+In the above definitions, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
 This statement is useful to describe many functions related to $Case$ in this paper.

### 2019-11-25 paper/3.minbigul_m.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos

@@
 We define a memory version, minBiGUL$_m$, including $get_m$ and $put_m$ as in definition 3 and 4 respectively.
-Another function introduced in definition 5, $get_{mh}$, is used to retrieve the value of an evaluated $get$ from the table or compute an unevaluated $get$ by using $get_m$ and store the key-value tuple to the table.
+Another function introduced in definition 5, $get_{mh}$, is used to retrieve the value of an evaluated $get$ from the table or compute an unevaluated $get$ by using $get_m$ then store the key-value tuple to the table if necessary.
 
-\begin{definition} $\putm{bx}{s}{v}$
+\begin{definition} $\text{Memo version of } put$
 
-    $\putm{bx}{s}{v} = \text{same with } put \text{ if } bx \neq bx_1 \circ bx_2$
-
-    $\putm{bx_1 \circ bx_2}{s}{v} = \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})}$
+    $\putm{bx}{s}{v} = \text{match } bx \text{ with }\\
+        \tab \vert \ bx_1 \circ bx_2 \to \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})} \\
+        \tab \vert \ \_ \to \text{similar to } put$
 \end{definition}
@@
 \begin{multicols}{2}
-    \begin{definition} $\getm{bx}{s}{v}$
+    \begin{definition} $\text{Memo version of } get$
 
-        $\getm{bx}{s} = \\
-            \tab \text{same with } get \text{ if } bx \neq bx_1 \circ bx_2$
-
-        $\getm{bx_1 \circ bx_2}{s} =\\
-            \tab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
-            \tab \text{with Not\_found } \to\\
-                \qtab v \Leftarrow \getmh{bx_2}{(\getmh{bx_1}{s})}\\
-                \qtab \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
-                \qtab v$
+        $\getm{bx}{s} = \text{match } bx \text{ with }\\
+            \tab \ \vert \ bx_1 \circ bx_2 \to\\
+            \tabs{2} \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
+            \tabs{2} \text{with Not\_found } \to\\
+                \tabs{3} v \Leftarrow \getmh{bx_2}{(\getmh{bx_1}{s})}\\
+                \tabs{3} \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
+                \tabs{3} v \\
+            \tab \ \vert \ \_ \to \text{similar to } get$
     \end{definition}
 \columnbreak
-    \begin{definition} $\getmh{bx}{s}{v}$
+    \begin{definition} $???$
 
@@
             \tab \text{with Not\_found } \to\\
-                \qtab v \Leftarrow \getm{bx}{s}\\
-                \qtab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
-                \qtab \text{with Not\_found } \to\\
-                    \qtabs{2} \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
-                    \qtabs{2} v$
+                \tabs{2} v \Leftarrow \getm{bx}{s}\\
+                \tabs{2} \text{match } bx \text{ with } \\
+                \tabs{2} \ \vert \ bx_1 \circ bx_2 \to v \\
+                \tabs{2} \ \vert \ \_ \to \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
+                \qtabs{2} \ v$
     \end{definition}

### 2019-11-25 paper/4.xpg.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos

@@
 
-% $pg$ is an involution that is self-inverse. An involution is a function $f$ that satifies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
+% $pg$ is an involution that is self-inverse. An involution is a function $f$ that satisfies $f(f(x)) = x \text{ for all } x \text{ in the domain of } f$.
 
@@
 This is a trick since in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.
-The last equality is relatively obivious.
+The last equality is relatively oblivious.
 
@@
 If a tupling occurs, there will be 4 combinations from these conditions.
-This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satified.
+This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
 The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
@@
 It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Due to \textsc{GetPut} law, we can use $get \ s$ to update the source $s$ in the $put$ direction.
-Then it is feasible to construct a $dummy$ from the source.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which is the evaluated result of $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
+We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
+The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
 This transformation is possible because of the \textsc{PutPut} law.
@@
 To be more convenient for presenting the definition of $cpg$ as well as the other functions later, we provide some following utility functions:\\ $fst = \lambda (x_1,x_2).
-x_1 , \quad snd = \lambda (x_1,x_2).
-x_2 , \quad con = \lambda ks_1.
+x_1 , \ snd = \lambda (x_1,x_2).
+x_2 , \ con = \lambda ks_1.
 \lambda ks_2.
@@
 The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because this evaluated variable is not used in the later steps.
-In the next assigment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
+In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
 It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
@@
         \qtab (s; v)$\\
-In general, the begining of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.\\
-Suppose the begining of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively.
+In general, the beginning of a continuation should be an identity function.
+However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
+
+Suppose the beginning of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively.
 Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.
@@
 The problem for $cpg$ lies in redundant computation during evaluating.
-For keeping away such computations, we introduce an extention named $kpg$.
+For keeping away such computations, we introduce an extension named $kpg$.
 While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
 Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modifcation information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
+The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modification information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
 The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
@@
         \qtab (ks' \ s; kv' \ v)$\\
-The begining of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
+The beginning of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
 
@@
 
-In the construction of $kpg$, $s$ and $v$ hold actual values only in case of $Skip$ and $Case$.
-Except them, functions for the computation will be kept in $ks'$ and $kv'$.
+Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
 When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-In $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done indendently in two assingments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be some recomputations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
+In $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+There may be same computations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
 It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
+One more thing, in $\kpginline{Case}$, we need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the fi condition before returning the 6-tuple.
+Such evaluations can be done lazily to make programs run faster.
+We use the above small optimizations in our implementations.
 
 \subsection{xpg}
-The most imporant thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that is used to produce new source instead of using more recursive call.
+The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that is used to produce new source instead of using more recursive call.
 This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$.
@@
 
-$\xpg{bx}{s}{v} = \text{same with } pg \text{ if } bx \neq bx_1 \circ bx_2$
-
-$\xpg{bx_1 \circ bx_2}{s}{v} =\\
-    \tab (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
-    \tab (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
-    \qtab (ks_1 \ s_2, v_2)$
+$\xpg{bx}{s}{v} = \text{match } bx \text{ with }\\
+    \tab \vert \ bx_1 \circ bx_2 \to \\
+        \tabs{2} (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
+        \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
+        \tabs{3} (ks_1 \ s_2, v_2)\\
+    \tab \vert \ \_ \to \text{similar to } put$
 \end{definition}

### 2019-11-25 paper/5.experiments.tex
Create 5.experiments.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos

@@
 
-We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, Ocaml 4.07.1.
+We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, OCaml 4.07.1.
 The OCaml runtime system options and garbage collection parameters are set as default.
@@
 
-We have conducted 11 test cases shown in Table \ref{tab:test-cases}.
-$s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is one of the 11 functions mentioned in the table.
-Note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
+We have conducted 7 test cases shown in Table \ref{tab:test-cases}.
+Although the number of programs is not so much, they are not trivial and can be completely used in practice.
+In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$.
+This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table.
+These $bx$ are left associative composition since we we are only interested in this kind of program in the paper.
+Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
 
@@
         & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
-        1 & lcomp-replace & straightline & 1 & 100 & 100 & 1 \\ \hline
-        2 & lcomp-replace-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
-        3 & lcomp-phead-ldata & straightline & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
-        4 & lcomp-ptail & straightline & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
-        5 & lcomp-ptail-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
-        6 & lcomp-bsnoc & straightline & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
-        7 & lcomp-bsnoc-ldata & straightline & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
-        8 & bmapreplace & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] & [1,$\ldots$,n] \\ \hline
-        9 & bmapreplace-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
-        10 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
-        11 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
+        1 & lcomp-phead-ldata & straight line & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline
+        2 & lcomp-ptail & straight line & [1,$\ldots$,n+1] & [1,$\ldots$,10] & $s_0 \ @ \ v_0$ & [\ ] \\ \hline
+        3 & lcomp-ptail-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n+1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{10 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n+11 times}}$ & [\ ] \\ \hline
+        4 & lcomp-bsnoc & straight line & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] & [1,$\ldots$,n-1] \\ \hline
+        5 & lcomp-bsnoc-ldata & straight line & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n-1 times}}$ & $\underbrace{[L,\ldots,L]]}_{\text{n-1 times}}$ \\ \hline
+        6 & breverse & recursion & [1,$\ldots$,n] & [1,$\ldots$,n] & [n,$\ldots$,1] & [n,$\ldots$,1] \\ \hline
+        7 & breverse-ldata & recursion & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ & $\underbrace{[L,\ldots,L]}_{\text{n times}}$ \\ \hline
     \end{tabular*}
@@
 
-The first 7 cases maunally use $n$ compose operators to make straightline compositions from $n + 1$ similar pure programs.
-We use the term \textit{pure} here to refer to programs that are not self-compose.
-The prefix \textit{lcomp} and \textit{rcomp} denote the associate precedences of a manual composition.
+The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from $n + 1$ similar \textit{pure} programs.
+We use the term \textit{pure} here to refer non-recursion compositions.
+The prefix \textit{lcomp} means that the associate precedence of $\circ$ in a straight-line composition is left.
 The suffix \textit{ldata} indicates that the input size is large.
-The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = \underbrace{[T,\ldots,T]}_{\text{n times}}$,\ $T = \underbrace{[A,\ldots,A]}_{\text{10 times}}$,\ $A = [1,\ldots,5]$
-
-If the associative type and input size are omitted, the table, in general, will comprise 4 programs: composition of many \textit{phead}s, composition of many \textit{ptail}s, composition of many \textit{bsnoc}s and \textit{breverse}.
+The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$.
+These structures are only intended to generate data enough large for observing experimental results.
 
@@
 The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view.
-We now review the behavior of the remaining 3 programs.
+Because of the type of the source, this program is seen as a \textit{ldata} case.
+Next, we briefly explain the behavior of the remaining compositions in table \ref{tab:test-cases}.
 
@@
 
-\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function for \textit{foldr} which is an important higher-order function on lists.
-This case is more natural than the other since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}.
-It will look like a composition of more BX programs at some point if we slow down the evaluation.
+\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function of \textit{foldr} which is an important higher-order function on lists.
+This program, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
+\textit{breverse} is more natural than the others since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}.
+It will look like a composition of more programs at some points in the evaluation if we slow down this process.
 Note that $\circ$ has a higher priority than $\times$.
@@
 
-\subsection{Results and discussions}
+\subsection{Results}
+
+Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
+We simply ignore the experimental results of $pg$, $cpg$ and $kpg$ since they are shown the slowness compared to $xpg$.
+For left associative compositions, $put$ works poorly because of the number of reevaluated $get$s.
+The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.
+The left of figure contains tests using not very big input, and the right includes of the others.
+We see that $put_m$ is the fastest method if the input size is not too large.
+However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table.
+At that time, $xpg$ proved to be the most effective method.
 
 \input{result.tex}
-
-Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 12 test cases.
-$put$ works more efficiently in the cases of right associative compositions, rather than the left ones.
-This makes sense because the number of \textit{get} when evaluating a right associative composition is linear in the number of compositions, while the respective number when computing a left associative composition is significantly larger.
-For self-composed program, \textit{breverse}, it could be impossible to change the associate precedence from left to right.
-Then, $put$ works less efficiently than both $put_m$ and $xpg$.
-In general, for left associative compositions, if the input size is not too large, $put_m$ is the fastest method.
-However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table.
-At that time, $xpg$ proved to be the most effective method.

### 2019-11-25 paper/2.minbigul.tex
Update 2.minbigul.tex
+ explain data structure

@@
 
-MinBiGUL, our target language in this paper, is a subset of BiGUL which is a simple yet powerful putback-based bidirectional language.
+minBiGUL, our target language in this paper, is a subset of BiGUL which is a simple yet powerful putback-based bidirectional language.
 BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
@@
 
-MinBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws.
+minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws.
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
@@
 A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinator without adaptive cases or a composition of some minBiGUL programs.
+We use numbers, pairs and lists to construct the program input including the source and/or the view.
 

### 2019-11-27 paper/4.xpg.tex
small changes

@@
 The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
+It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
             The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
-This transformation is possible because of the \textsc{PutPut} law.
+            Because there is no real view here, we need to construct a dummy from the source. 
+            The replacement of $s$ and $s_1$ is possible because of the \textsc{PutPut} law.
 Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.

### 2019-11-27 paper/4.xpg.tex
add subsection names

@@
 \section{Tupling and lazy updates: xpg}
-\subsection{pg}
-
+\subsection{Tupling: pg}
+s
 Another solution for saving intermediate states is tupling.
@@
 
-\subsection{cpg}
+\subsection{Lazy update: cpg}
 
@@
 
-\subsection{kpg}
+\subsection{Lazy computation: kpg}
 
@@
 
-\subsection{xpg}
+\subsection{Combination of pg and kpg: xpg}
 The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that is used to produce new source instead of using more recursive call.

### 2019-11-27 paper/6.related-work.tex
update tex 20191125
- def of get_mh
- v / v'
- experiments: xpg is faster than pgs
- dummy
- kpg: lazy eval on fi-cond
- figure size
- typos
major update

@@
 
-\subsubsection*{BX langauges and implementations}
+\subsubsection*{BX languages and implementations}
 
@@
 We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
-In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Section 1) and our methods (in Section 3 and 4).
-From experimental results, we find that there is no best approach for all BX programs.
-It is important to choose good method depending on the BX programs and their inputs.
+In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4).
+From experimental results of left associative programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
 In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
+
+%find that there is no best approach for all BX programs. It is important to choose good method depending on the BX programs and their inputs. 
 % its implementation uses ``not keeping any intermediate states and obtaining them by evaluation when they are needed'' strategy.
@@
 
-In this paper, we use several optimization techniques: tupling, lazy update, and lazy computation (\textcolor{red}{also fusion?}).
+Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
+Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we do not use this way.
+  Instead of that, we introduce another way: using tupling \cite{}, lazy update and lazy computation \cite{}.
+Tupling \cite{tupling} is often used for reducint common evaluations, but we use it for reducing the size of complements.
+  Lazy evaluation \cite{} is also a well-known and widly-used optimization technique to reduce the redundant evaluation.
+Because we use OCaml for implementation, we need to introduce this explicitly.
+If we use Haskell, a lazy evaluated language, we do not need to introduce them.
 
-Tupling \cite{tupling} is an optimization technique by combining several computations.
+%In this paper, we use several well-known techniques: tupling, lazy update, and lazy computation (\textcolor{red}{also fusion?}).
+
+%Tupling \cite{tupling} is an optimization technique by combining several computations.
 Although tupling is usually used for reducing the evaluation cost\cite{}, we use it to reduce the size of the things need to keep.
 
-We introduce lazy update to remove repeated evaluation of put.
+%We introduce lazy update to remove repeated evaluation of put.
 This is different from lazy evaluation.
@@
 
-Lazy evaluation \cite{} is well-known and widly-used optimization technique to reduce the redundant evaluation.
+%Lazy evaluation \cite{} is well-known and widly-used optimization technique to reduce the redundant evaluation.
 Because we use OCaml for implementation, we need to introduce this explicitly.

### 2019-11-27 paper/7.conclusion.tex
major update

@@
 To achieve fast evaluation, we introduce two different methods using memoization.
-From the experimental results, we know that there is no best approach for all BX programs.
-However, if programmer can choose one method based on the BX programs and inputs, we can use better evaluation methods.
+From the experimental results of left associative composition of BX programs, we know that $put_m$ is fastest method if input size is not very big and $xpg$ is fastest for other left associative programs.
+This shows that if programmer choose one method based on the their BX programs and inputs, they can use efficient evaluation.
 
-We will continue our work on the following points. First is extention of our language. Currently, because $pg$ requires very-well-behaved property
+%there is no best approach for all BX programs. However, if programmer can choose one method based on the BX programs and inputs, we can use better evaluation methods.
+
+We will continue our work on the following points.
+First is overcomming our limitation in $xpg$.
+In $xpg$, there are several limitations: only very-well-behaved programs, path problem (shown in Section ..).
+To be a practical evlauation method, extention is needed.
+After extention of the language, we can evaluate more interesting programs for experimental results.
+
+Second is analysis about BX programs and inputs to choose the best evaluation method automatically.
+Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If automatic choice is possible, we can reduce programmer's burden.
 
 \begin{itemize}
-\item overcome our limitations
-  \begin{itemize}
-  \item very-well-behaved $\to$ only xpg?
-  \item In case expressions, path's problem $\to$ only xpg?
-  \item How to obtain $\to$ only xpg?
+\item construct\_dummy?
 \end{itemize}
-  \item tests for big BX programs
-\item Type system \& Typing -- for safety
-\end{itemize}
+
+%First is extention of our language. Currently, because $pg$ requires very-well-behaved property
+
+%\begin{itemize}
+%\item overcome our limitations
+%  \begin{itemize}
+%  \item very-well-behaved $\to$ only xpg?
+%  \item In case expressions, path's problem $\to$ only xpg?
+%  \item How to obtain $\to$ only xpg?
+%  \end{itemize}
+%  \item tests for big BX programs
+%  \item Type system \& Typing -- for safety
+%  \item analysis about programs/inputs to choose the best method automatically
+%\end{itemize}
 

### 2019-11-27 paper/main.tex
hide author info

@@
 %
-\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and
-  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and
-  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
-  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
 
-\authorrunning{K. Tsushima et al.}
+%\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and
+%  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and
+%  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
+%  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
+
+\author{anonymous authors}
+
+%\authorrunning{K. Tsushima et al.}
+
+\authorrunning{}
+
 % First names are abbreviated in the running head.
@@
 %
-\institute{National Institute of Informatics, Japan
-  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
-  University of Copenhagen, Denmark
-  \email{glueck@di.ku.dk } \and
-  Peking University, China
-  \email{huzj@pku.edu.cn}}
+
+%\institute{National Institute of Informatics, Japan
+%  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
+%  University of Copenhagen, Denmark
+%  \email{glueck@di.ku.dk } \and
+%  Peking University, China
+%  \email{huzj@pku.edu.cn}}
+
+\institute{anonymous}
 

### 2019-11-27 paper/6.related-work.tex
fix small things

@@
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
-based on several application scenarios and BX lenses works \cite{} (add papers for lenses).
+based on several application scenarios and BX lenses works \cite{} \textcolor{red}{(add papers for lenses)}.
 The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
@@
 We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
-In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4).
-From experimental results of left associative programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
+In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Sect.
+2) and our methods (in Sect.
+3 and 4).
+From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
 In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
@@
   Instead of that, we introduce another way: using tupling \cite{}, lazy update and lazy computation \cite{}.
-Tupling \cite{tupling} is often used for reducint common evaluations, but we use it for reducing the size of complements.
-  Lazy evaluation \cite{} is also a well-known and widly-used optimization technique to reduce the redundant evaluation.
+Tupling \cite{tupling} is often used for reducing common evaluations, but we use it for reducing the size of complements.
+  Lazy evaluation \cite{} is also a well-known and widely-used optimization technique to reduce the redundant evaluation.
 Because we use OCaml for implementation, we need to introduce this explicitly.

### 2019-11-27 paper/7.conclusion.tex
fix small things

@@
 To achieve fast evaluation, we introduce two different methods using memoization.
-From the experimental results of left associative composition of BX programs, we know that $put_m$ is fastest method if input size is not very big and $xpg$ is fastest for other left associative programs.
+From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not very big and $xpg$ is fastest for other left associative programs.
 This shows that if programmer choose one method based on the their BX programs and inputs, they can use efficient evaluation.
@@
 
-We will continue our work on the following points.
+We will continue our work on the following two points.
 First is overcomming our limitation in $xpg$.
-In $xpg$, there are several limitations: only very-well-behaved programs, path problem (shown in Section ..).
-To be a practical evlauation method, extention is needed.
+In $xpg$, there are two limitations: only for very-well-behaved programs, restriction for $\pginline{Case}$ (shown in Sect.
+4.1).
+To be a practical evlauation method, extention of the target language is needed.
 After extention of the language, we can evaluate more interesting programs for experimental results.
 
-Second is analysis about BX programs and inputs to choose the best evaluation method automatically.
+Second is introducing analysis about BX programs and inputs to choose the best evaluation method automatically.
 Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If automatic choice is possible, we can reduce programmer's burden.
@@
 \begin{itemize}
-\item construct\_dummy?
+\item $construct\_dummy$??
 \end{itemize}

### 2019-11-27 paper/1.introduction.tex
small fixes

@@
 
-In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
+%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
 % A Survey to View Update Problem 
-As a solution for this problem, bidirectional transformation (BX) is introduced.
+%As a solution for this problem, bidirectional transformation (BX) is introduced.
+The synchronization of data is a common problem.
+In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{viewupdate}.
+Bidirectional transformation (BX) is an approach to solve this problem.
 As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}.
@@
 
-Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
+Here, let \texttt{[1,2]} be the original source \texttt{s}.
+$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
 %\texttt{get pHead [1,2] = 1}.
 After the view \texttt{v} is obtained, the user can modify the view.
-In this case, he or she modified the view to \texttt{100} from \texttt{1}.
-From this updated view, how can we obtain new updated source? This is ``the view update problem.'' For doing this, BX provides $put$, an update function on the original source.
+In this case, the user
+modified the view
+from \texttt{1} to \texttt{100}.
+From this updated view, how can the source be updated?
+This is ``the view update problem.''
+To solve this problem, BX provides $put$, an update function on the original source.
 In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
+Note that the update to s' uses s and v' as input.
 %\texttt{put pHead [1,2] 8 = [8,2]}
 
-In BX programs, composition of programs is essential for writing various interesting evaluations. Because recursions are achieved by compositions, large number of programs include compositions.
+The composition of BX programs is a fundamental construct to build more complex BX programs. 
+%In BX programs, composition of programs is essential for writing various interesting evaluations.
+Because recursions are achieved by compositions, large number of programs include compositions.
 In the BX lens papers \cite{} compositions are also known as central part of BX programs.
-In this section we use a tiny BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
+In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
 The running examples are two versions of three composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
@@
 First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}.
-This method is used in a BX language, BiGUL \cite{}.
+This method is used in a BX language, BiGUL \cite{josh_pepm} \cite{josh_popl}.
 The merit of this method is clear semantics and easy to implement.
@@
 \vspace{2mm}
-In very-well behaved BX programs, 
+In very-well behaved \cite{} BX programs, 
 $put$ is a complement function for $get$.
@@
   \includegraphics[height=5cm]{./fig/fig3.eps}
-  \caption{Evaluating $phead \circ phead \circ phead$ by keeping intermediate states}
+  \caption{Evaluating $phead \ \circ \ phead \ \circ \ phead$ by keeping intermediate states}
   \label{fig:eval-comp-phead-2}

### 2019-11-27 paper/2.minbigul.tex
small changes
small fixes

@@
 
-The \textsc{GetPut} [?] law means that if there is no change to the view, there should be no change to the source.
-The \textsc{PutGet} [?] law means that we can recover the modified view by applying the forward transformation to the updated source.
+The \textsc{GetPut} \cite{putget} law means that if there is no change to the view, there should be no change to the source.
+The \textsc{PutGet} \cite{putget} law means that we can recover the modified view by applying the forward transformation to the updated source.
 
@@
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satisfy one more law, \textsc{PutPut} [?], like the following:\\
+Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{putput}, like the following:\\
     $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
@@
 The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates.
-Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved [?].
+Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.
 
@@
 The later definitions also use this convention.
+\textcolor{red}{$Leftarrow$ means assignments.}
 

### 2019-11-27 paper/6.related-work.tex
small fixes

@@
 Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
-Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we do not use this way.
-  Instead of that, we introduce another way: using tupling \cite{}, lazy update and lazy computation \cite{}.
-Tupling \cite{tupling} is often used for reducing common evaluations, but we use it for reducing the size of complements.
-  Lazy evaluation \cite{} is also a well-known and widely-used optimization technique to reduce the redundant evaluation.
+Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
+  Instead of that, we introduce another way that uses tupling \cite{tupling}, lazy update and lazy computation \cite{}.
+Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
+  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation.
 Because we use OCaml for implementation, we need to introduce this explicitly.

### 2019-11-27 paper/7.conclusion.tex
change Sect. to Section
add third point

@@
 
-We will continue our work on the following two points.
+We will continue our work on the following three points.
 First is overcomming our limitation in $xpg$.
-In $xpg$, there are two limitations: only for very-well-behaved programs, restriction for $\pginline{Case}$ (shown in Sect.
-4.1).
+In $xpg$, there are two limitations: only for very-well-behaved programs, restriction for $\pginline{Case}$ (shown in Section 4.1).
 To be a practical evlauation method, extention of the target language is needed.
@@
 Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If automatic choice is possible, we can reduce programmer's burden.
+
+\textcolor{red}{Add here}
+Third is introducing type systems to our target language. If we introduce this, we need to investigate more about $construct\_dummy$, because current definition causes type error.
 

### 2019-11-27 paper/5.experiments.tex
add . for Fig
update experiments

@@
 The OCaml runtime system options and garbage collection parameters are set as default.
+There is no type system in our implementation, so instead of constructing some dummies which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
 
@@
 
-Fig \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
+Figure \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
 We simply ignore the experimental results of $pg$, $cpg$ and $kpg$ since they are shown the slowness compared to $xpg$.

### 2019-11-27 paper/1.introduction.tex
add bib

@@
 
-%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
+%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time.
 % A Survey to View Update Problem 
@@
 The synchronization of data is a common problem.
-In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{viewupdate}.
+In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}.
 Bidirectional transformation (BX) is an approach to solve this problem.
@@
 First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}.
-This method is used in a BX language, BiGUL \cite{josh_pepm} \cite{josh_popl}.
+This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}.
 The merit of this method is clear semantics and easy to implement.

### 2019-11-27 paper/6.related-work.tex
add bib

@@
 
-There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF}
+There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
@@
 al.
-introduces the first benchmark for BX languages and compared them \cite{BXcomp}.
+introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}.
 However, the improvement for practical implementation of BX languages is still missing.
@@
 We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
-In this paper we focus on a BX language BiGUL \cite{josh_pepm} \cite{josh_popl}, we compare BiGUL's method (in Sect.
-2) and our methods (in Sect.
-3 and 4).
+In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4).
 From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
@@
 Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
-  Instead of that, we introduce another way that uses tupling \cite{tupling}, lazy update and lazy computation \cite{}.
+  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{}.
 Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.

### 2019-11-27 paper/main.tex
small fixes
change on TODO part
add bib

@@
 \usepackage{schemata}
+\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
+
+\usepackage[backend=bibtex,doi=false,url=false,style=nature,sorting=none]{biblatex}
+\addbibresource{bibliography}
 
@@
 % to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
+\renewcommand\UrlFont{\color{blue}\rmfamily}
+
 \usetikzlibrary{matrix}
@@
 
-\setlength{\parindent}{0pt}
-\renewcommand{\indent}{\hspace*{0pt}}
+% \setlength{\parindent}{0pt}
+% \renewcommand{\indent}{\hspace*{0pt}}
 
@@
 
-\authorrunning{}
+\authorrunning{anonymous authors}
 
@@
 
-\institute{anonymous}
+%\institute{anonymous}
 
@@
 
-\textcolor{red}{TODO: FIX reference, order, contents ..}
+\textcolor{red}{TODO: fix references (order, contents), add missing references}
 
@@
 %
+% \bibliography{bibliography}
 % \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
 
-\tiny
+\printbibliography
 
-\begin{thebibliography}{8}
-\bibitem{viewupdate}
-F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.
+% \tiny
 
-\bibitem{BXcomp}
-Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
-``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019
+% \begin{thebibliography}{8}
+% \bibitem{viewupdate}
+% F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.
 
-\bibitem{tupling}
-  M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.
+% \bibitem{BXcomp}
+% Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
+% ``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019
 
-\bibitem{josh_pepm}
-  Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.
+% \bibitem{tupling}
+%   M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.
 
-\bibitem{josh_popl}
-  Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.
+% \bibitem{josh_pepm}
+%   Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.
 
-\bibitem{BXtend}
-  Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
-of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.
+% \bibitem{josh_popl}
+%   Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.
 
-\bibitem{eMoflon}
-  Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr.
+% \bibitem{BXtend}
+%   Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
+% of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.
+
+% \bibitem{eMoflon}
+%   Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr.
 Developing eMoflon with eMoflon.
@@
 Proceedings, volume 8568 of Lecture Notes in
-Computer Science, pages 138–145. Springer, 2014.
+% Computer Science, pages 138–145. Springer, 2014.
 
-\bibitem{EVL+Strace}
-Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
-Technology, 100:47--72, 2018.
+% \bibitem{EVL+Strace}
+% Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
+% Technology, 100:47--72, 2018.
 
-\bibitem{JTL}
-Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio.
+% \bibitem{JTL}
+% Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio.
 JTL: A bidirectional and change propagating transformation language.
@@
 
-\bibitem{NMF}
-Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.
+% \bibitem{NMF}
+% Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.
 
-\bibitem{SDMLib}
-  SDMLib paper?
+% \bibitem{SDMLib}
+%   SDMLib paper?
 
-\bibitem{lenses}
-  Lense papers?
+% \bibitem{lenses}
+%   Lense papers?
 
-\bibitem{tutorial}
-  Tutorial paper of BiGUL
+% \bibitem{tutorial}
+%   Tutorial paper of BiGUL
+
+% \bibitem{putget}
+%   H. Pacheco, Z. Hu, and S. Fischer. Monadic combinators for “putback” style bidirectional programming. In Partial Evaluation and Program Manipulation, pages 39–50. ACM, 2014.
+
+% \bibitem{putput}
 
@@
 %Oct 2017
-\end{thebibliography}
+% \end{thebibliography}
 

### 2019-11-27 paper/2.minbigul.tex
small thing
fix indent + scaled tick

@@
 BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:\\
-    $\tab \getbxinline{bx} : s \to v,
-    \tab \putbxinline{bx} : s * v \to v$
+Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:
 
-BiGUL is well-behaved [?] since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:\\
-    $\tab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]\\
-    \tab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$
+    $\qtab \getbxinline{bx} : s \to v, \quad \putbxinline{bx} : s * v \to v$
 
-The \textsc{GetPut} \cite{putget} law means that if there is no change to the view, there should be no change to the source.
-The \textsc{PutGet} \cite{putget} law means that we can recover the modified view by applying the forward transformation to the updated source.
+BiGUL is well-behaved [?] since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:
 
-minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws.
+    $\qtab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]$
+
+    $\qtab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$
+
+The \textsc{GetPut} \cite{Pacheco:2014:MCP:2543728.2543737} law means that if there is no change to the view, there should be no change to the source.
+The \textsc{PutGet} \cite{Pacheco:2014:MCP:2543728.2543737} law means that we can recover the modified view by applying the forward transformation to the updated source.
+
+minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which satisfy two above laws.
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{putput}, like the following:\\
-    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
+Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{putput}, like the following:
+
+    $\qtab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
 
@@
 
-$\tab bx := Skip \ h \
+    $bx := Skip \ h \
         | \ Replace \
@@
         | \ RearrV \ g_1 \ g_2 \ bx \\
-    \qtab | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \
+    \qtabs{2} | \ Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2 \
         | \ Compose \ bx_1 \ bx_2$
@@
 
-        $\putbx{Skip \ h}{s}{v} = \\
+        \noindent $\putbx{Skip \ h}{s}{v} = \\
             \tab \sif{h \ s = v}{s}{\text{fail}}$
 
-        $\putbx{Replace}{s}{v} = v$
+        \noindent $\putbx{Replace}{s}{v} = v$
 
-        $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
+        \noindent $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
             \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
 
-        $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = \\
+        \noindent $\putbx{\rearrs{f_1}{f_2}{bx}}{s}{v} = \\
             \tab f_2 \ (\putbx{bx}{(f_1 \ s)}{v})$
 
-        $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \\
+        \noindent $\putbx{\rearrv{g_1}{g_2}{bx}}{s}{v} = \\
             \tab \putbx{bx}{s}{(g_1 \ v)}$
 
-        $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} = \\
+        \noindent $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
             \tab \text{if} \ {cond_{sv} \ s \ v} \\
@@
 
-        $\putbx{bx_1 \circ bx_2}{s}{v} = \\
+        \noindent $\putbx{bx_1 \circ bx_2}{s}{v} = \\
             \tab \putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}$
@@
 
-        $\getbx{Skip \ h}{s} = \\ 
+        \noindent $\getbx{Skip \ h}{s} = \\ 
             \tab h \ s$
 
-        $\getbx{Replace}{s} = s$
+        \noindent $\getbx{Replace}{s} = s$
 
-        $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
+        \noindent $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
             \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
 
-        $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \\ 
+        \noindent $\getbx{\rearrs{f_1}{f_2}{bx}}{s} = \\ 
             \tab \getbx{bx}{(f_1 \ s)}$
 
-        $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = \\ 
+        \noindent $\getbx{\rearrv{g_1}{g_2}{bx}}{s} = \\ 
             \tab g_2 \ (\getbx{bx}{s})$
 
-        $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} = \\
+        \noindent $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} = \\
             \tab \text{if} \ {cond_{s} \ s} \\
@@
 
-        $\getbx{bx_1 \circ bx_2}{s} = \\ 
+        \noindent $\getbx{bx_1 \circ bx_2}{s} = \\ 
             \tab \getbx{bx_2}{(\getbx{bx_1}{s})}$
@@
 Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.
-The later definitions also use this convention.
-\textcolor{red}{$Leftarrow$ means assignments.}
+The later definitions also follow this convention.
 
-In the above definitions, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$.
+In the above definitions, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$ where $\Leftarrow$ means assignments.
 This statement is useful to describe many functions related to $Case$ in this paper.
@@
 $\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tab \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false; \}\\
-    \qtab \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false; \}$
+\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false; \}\\
+    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false; \}$
 
-% Next, let's take a look at some examples to better understand about minBiGUL. We start with quite obvious things as follows:
-
-% $\tab \putbx{Skip \ (\lambda x.(x*x))}{10}{100} = 10 \qtab \ \getbx{Skip \ (\lambda x.(x*x))}{10} = 100\\
-%     \tab \putbx{Skip \ (\lambda \_.())}{1}{()} = 1 \qtabs{3} \getbx{Skip \ (\lambda \_.())}{1} = ()\\
-%     \tab \putbx{Replace}{1}{100} = 100 \qtabs{3} \getbx{Replace}{1} = 1
-% $
-
-We now consider the definition of $phead$ in minBiGUL:
-
+\noindent We now consider the definition of $phead$ in minBiGUL:\\
 $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
@@
 If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
-For example:\\
-\tab $\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$\\
-\tab $\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$
+For example:
+
+$\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$
+
+$\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$

### 2019-11-27 paper/4.xpg.tex
fix indent + scaled tick

@@
 \subsection{Tupling: pg}
-s
+
 Another solution for saving intermediate states is tupling.
@@
 
-Now, let's see how we construct $pg$ recursively.
+\noindent Now, let's see how we construct $pg$ recursively.
 
-$\tab \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-    \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-    \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{fail}}$\\
+    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
+        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
+        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{fail}}$
+
 The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$.
@@
 
-$\tab \pg{Replace}{s}{v} = (v, s)$\\
-$\tab \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
-    \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
-    \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-        \qtab \tab \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-        \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
-$\tab \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-        \qtabs{2} (f_2 \ s, v)$\\
-$\tab \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-    \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \qtabs{2} (s, g_2 \ v)$\\
+    \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
+    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
+        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
+        \tab \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
+            \qtab \tabs{2} \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
+            \tab \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
+    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
+        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
+            \tab \qtabs{2} (f_2 \ s, v)$\\
+    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
+        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
+            \tab \qtabs{2} (s, g_2 \ v)$
+
 Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.
 
-$\tab \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
-\tabs{2} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \text{if} \ {cond_{s} \ s} \\
-\tab \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtab \tab \, \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\
-\tab \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtab \tab \ \, \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
-\tab \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
-\tabs{2} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
-\tab \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
-\tab \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
-\tab \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$
+    \noindent $\tabs{2} \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
+    \tabs{3} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \text{if} \ {cond_{s} \ s} \\
+    \tabs{2} \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtab \tab \, \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\
+    \tabs{2} \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtab \tab \ \, \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
+    \tabs{2} \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
+    \tabs{3} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
+    \tabs{2} \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
+    \tabs{2} \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
+    \tabs{2} \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$
 
@@
 
-$\tab \pg{bx_1 \circ bx_2}{s}{v} \\
-    \tabs{2} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
-    \tabs{2} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
+    \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
+    \tabs{3} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
+    \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
         \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
@@
             \qtab \tabs{2} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
-    \tabs{2} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
+    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
         \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
@@
 The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$.
-It is possible if we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
+This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
+We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
+Since there is no real view, these things are necessary.
 The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
-            Because there is no real view here, we need to construct a dummy from the source. 
 The replacement of $s$ and $s_1$ is possible because of the \textsc{PutPut} law.
@@
 
-$\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$
+    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$
 
-$\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$
+    \noindent $\cpg{Replace}{ks}{kv}{s}{v} = (kv, ks, v, s)$
 
-$\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
+    \noindent $\cpg{\product{bx_1}{bx_2}}{ks}{kv}{s}{v} =\\
         \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \ s}{fst \ v};\\
@@
 
-$\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
+    \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
         \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
@@
 
-$\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
+    \noindent $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
         \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
@@
 
-$\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
+    \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
         \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
@@
 
-$\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
+    \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
         \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
@@
 
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.\\
+Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
+
 For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
 The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Because there is no real view here, we need to construct a dummy from the source.
+Since there is no real view here, we need to construct a dummy from the source.
 Then the continuation updating on this dummy should be initiated as an identity function.
@@
 It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$.
-After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.\\
+After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.
+
 Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:\\
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $cpg$ as follows:
+
     $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
-        \qtab (s; v)$\\
+        \tab \qtab (s; v)$
+
 In general, the beginning of a continuation should be an identity function.
@@
 After the first two assignments in the definition of $cpg$ for the composition, we have: $ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))$ and $s_2 = [100,2,3]$ where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$.
-\\
-Then: $ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$\\
+Then:
+
+$ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
+\tab \quad = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
+\tab \quad = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
+
 If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
@@
 Suppose that we have a source $s_0$ and a view $v_0$.
-The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:\\
+The pair of the updated source and view $(s, v)$ where $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$ can be obtained using $kpg$ as follows:
+
     $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \qtab (ks' \ s; kv' \ v)$\\
+        \tab \qtab (ks' \ s; kv' \ v)$
+
 The beginning of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
@@
 
-$\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
@@
 
-$\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$
+    \noindent $\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$
 
-$\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
@@
 
-$\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
@@
 
-$\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
@@
 
-$\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id ; \quad kv' \Leftarrow id;\\
-    \tab \text{if} \ cond_{sv} \ s \ v \&\& \ cond_{s} \ s\\
+        \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
         \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
@@
 
-$\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
+    \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
         \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
@@
 Such evaluations can be done lazily to make programs run faster.
-We use the above small optimizations in our implementations.
+We use the above small optimizations in our implementation.
 
@@
 This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$.
-The second call $\kpginline{bx_2}$ just play an syntactic role while it requires many parameters.
+The second call $\kpginline{bx_2}$ just play a syntactic role while it requires many parameters.
 We can completely reduce this number to 2, including one for the source and one for the view as in a $pg$ call.
@@
 
-$\xpg{bx}{s}{v} = \text{match } bx \text{ with }\\
+    \noindent $\xpg{bx}{s}{v} = \text{match } bx \text{ with }\\
     \tab \vert \ bx_1 \circ bx_2 \to \\

### 2019-11-27 paper/5.experiments.tex
fix indent + scaled tick

@@
 
-\tab $\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$.
+$\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$.
 $\getbx{ptail}{s}$ return the tail of the source list $s$.
@@
 
-\tab $\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$ as two same length lists, to produce a new list by moving the last element of $v$ to its first position.
+$\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$ as two same length lists, to produce a new list by moving the last element of $v$ to its first position.
 $\getbx{ptail}{s}$ return another list by moving the first element of the list $s$ to its end position.
@@
 
-\tab \textit{breverse}, is defined in terms of \textit{bFoldr}, a putback function of \textit{foldr} which is an important higher-order function on lists.
+\textit{breverse} is defined in terms of \textit{bfoldr}, a putback function of \textit{foldr} which is an important higher-order function on lists.
 This program, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
-\textit{breverse} is more natural than the others since \text{bFoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bFoldr}.
+\textit{breverse} is more natural than the others since \textit{bfoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bfoldr}.
 It will look like a composition of more programs at some points in the evaluation if we slow down this process.
 Note that $\circ$ has a higher priority than $\times$.
-This leads to seemingly impossible to transform \textit{bReverse} as well as \textit{bFoldr} from the left associative style to the right one.
+This leads to seemingly impossible to transform \textit{breverse} as well as \textit{bfoldr} from the left associative style to the right one.
 
 \subsection{Results}
+
+\input{result.tex}
 
@@
 At that time, $xpg$ proved to be the most effective method.
-
-\input{result.tex}

### 2019-11-29 paper/1.introduction.tex
add explanation for breverse and titlerunning

@@
 
-Here, let \texttt{[1,2]} be the original source \texttt{s}.
+Here, let $[1,2]$ be the original source \texttt{s}.
 $get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
@@
 modified the view
-from \texttt{1} to \texttt{100}.
+from $1$ to $100$.
 From this updated view, how can the source be updated?
@@
 To solve this problem, BX provides $put$, an update function on the original source.
-In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
+In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
 Note that the update to s' uses s and v' as input.
@@
 
-Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
-By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, \texttt{1}.
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
+By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$.
 To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
@@
 However, this is not always possible.
+\textcolor{red}{For example, let us consider the following 'breverse' program.}
+
+%$\tab
+%bsnoc = Rearr\\
+%bfoldr bx = Case \ f_1 \ f_2 \ bx_s \text{where: } f_1 = , f_2% = \\
+
+%breverse = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \l%ambda s.(s.()), \, f_2 = \lambda (s.()).s,\\
+%\qtab bx_s = bfoldr bsnoc$
+
+\textcolor{red}{In this program, recursion occurs on the left-hand side of composition.
+Because of priority, it is impossible to tranform this program to right associative compositions.
+At the same time, we can not apply fusion \cite{} to this program because this program is dynamically produced depending on the length of the input list.}
 Therefore it is important to have a fast evaluation method for left associative compositions.
+
 For achieving this, we introduce two methods that use memoization.
@@
 For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
-As an example, let us consider the previous example again: $phead \circ (phead \circ phead)$ (rp3) in figure \ref{fig:eval-comp-phead-2}.
+As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in figure \ref{fig:eval-comp-phead-2}.
 \texttt{c$_1$} and \texttt{c$_2$} are complements.
+\texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}.
 The points in this figure is two.

### 2019-11-29 paper/1.introduction.tex
small fixes

@@
 In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
-The running examples are two versions of three composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
+The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
 
@@
 
-\textcolor{red}{In this program, recursion occurs on the left-hand side of composition.
+\input{breverse.tex}
+
+\noindent \textcolor{red}{In this program, recursion occurs on the left-hand side of composition.
 Because of priority, it is impossible to tranform this program to right associative compositions.

### 2019-11-29 paper/7.conclusion.tex
remove not-needed part
change on TODO part
update third part

@@
 To be a practical evlauation method, extention of the target language is needed.
-After extention of the language, we can evaluate more interesting programs for experimental results.
+After extention of the language, we can evaluate more various programs for experimental results.
 
@@
 
-\textcolor{red}{Add here}
-Third is introducing type systems to our target language. If we introduce this, we need to investigate more about $construct\_dummy$, because current definition causes type error.
+Third is introducing type systems to datatype of sources and view.
+If we introduce this, we can avoid runtime errors like statically-typed functional languages.
+We need to investigate more about $construct\_dummy$, because current definition that is used in Section 6 will cause type errors.
 
-\begin{itemize}
-\item $construct\_dummy$??
-\end{itemize}
+%\begin{itemize}
+%\item $construct\_dummy$??
+%\end{itemize}
 

### 2019-11-29 paper/6.related-work.tex
add references

@@
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
-based on several application scenarios and BX lenses works \cite{} \textcolor{red}{(add papers for lenses)}.
+based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
 The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
@@
 Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
-Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
+Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}.
+However, fusion of BX programs are not considered well \textcolor{red}{(really?)}.
+Therefore we did not use this way.
   Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{}.

### 2019-11-29 paper/2.minbigul.tex
update tex 20191129

@@
 
-$\putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$
+$\tab \putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$
 
-$\getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$
+$\tab \getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$
+
+In the end of this section, let's see the definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
+
+$\tab bfoldr \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+$\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
+$\tabs{3} \ bx_1 = RearrV \ g_1 \ g_2 \ bx_v$ where: \\
+    $\tabs{4} g_1 = g_2^{-1} = \lambda [v].(v,[\ ]), bx_v = (Skip \ (\lambda \_.())) \times Replace$\\
+$\tabs{3} \ bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
+    $\tabs{4} f_1 = f_2^{-1} = \lambda ((x:xs),e).(x,(xs,e))$, 
+    $bx_s = ((Replace \times bfoldr \ bx) \circ bx)$
+
+$bfoldr$ will look like a composition of many programs at some points in the evaluation if we slow down this process.
+Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.

### 2019-11-29 paper/5.experiments.tex
update tex 20191129

@@
 The OCaml runtime system options and garbage collection parameters are set as default.
-There is no type system in our implementation, so instead of constructing some dummies which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
+There is no type system in our implementation\footnote{\url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}, so instead of constructing some dummies which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
 
@@
 
-\textit{breverse} is defined in terms of \textit{bfoldr}, a putback function of \textit{foldr} which is an important higher-order function on lists.
-This program, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
-\textit{breverse} is more natural than the others since \textit{bfoldr} is a Case analysis whose a branch contains a composition $bx_1 \circ bx_2$ where $bx_1$ is a product of a simple function and \textit{bfoldr}.
-It will look like a composition of more programs at some points in the evaluation if we slow down this process.
-Note that $\circ$ has a higher priority than $\times$.
-This leads to seemingly impossible to transform \textit{breverse} as well as \textit{bfoldr} from the left associative style to the right one.
+\textit{breverse} is defined in terms of \textit{bfoldr}, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
 

### 2019-11-29 paper/1.introduction.tex

@@
 Because recursions are achieved by compositions, large number of programs include compositions.
-In the BX lens papers \cite{} compositions are also known as central part of BX programs.
+In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
 In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
@@
 
-If the programs are simple like these running examples, we can transform left associative compositions to right associative.
+\textcolor{red}{If the programs are simple like the running examples, we can transform left associative compositions to right associative.
 However, this is not always possible.
-\textcolor{red}{For example, let us consider the following 'breverse' program.}
-
-%$\tab
-%bsnoc = Rearr\\
-%bfoldr bx = Case \ f_1 \ f_2 \ bx_s \text{where: } f_1 = , f_2% = \\
-
-%breverse = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \l%ambda s.(s.()), \, f_2 = \lambda (s.()).s,\\
-%\qtab bx_s = bfoldr bsnoc$
-
-\input{breverse.tex}
-
-\noindent \textcolor{red}{In this program, recursion occurs on the left-hand side of composition.
-Because of priority, it is impossible to tranform this program to right associative compositions.
-At the same time, we can not apply fusion \cite{} to this program because this program is dynamically produced depending on the length of the input list.}
-Therefore it is important to have a fast evaluation method for left associative compositions.
+For example, let us consider a BX 'reverse' program, $breverse$.
+$breverse$ is constructed by application of $bfoldr$ (the actual program is shown in Section 2) and $bsnoc$.
+A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$.
+In this program, recursion occurs on the left-hand side of composition.
+Because of priority, it is impossible to transform this program to right associative compositions.
+At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because this BX program is dynamically produced depending on the length of the input list.
+Therefore, it is important to have a fast evaluation method for left associative compositions.}
 

### 2019-11-29 paper/4.xpg.tex
update tex 20191129
remove detail explaination of (fun _ -> s)

@@
 In general, the beginning of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as a constant function $\lambda \_.s$.
+However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as $\lambda \_.s$.
+This constant function helps to retain the discarded things in the source.
 
-Suppose the beginning of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively.
-Let's consider $\cpg{phead_1 \circ phead_2}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.
+% Suppose the beginning of continuations $ks$ and $kv$ are $ks_0$ and $id$ respectively.
+Let's consider $\cpg{phead \circ phead}{ks_0}{id}{s}{v}$ where $s = [[1,2,3], [\,], [4,5]]$ and $v = 100$.
 After the first two assignments in the definition of $cpg$ for the composition, we have: $ks_1 = f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))$ and $s_2 = [100,2,3]$ where $f_1 = \lambda (s::ss).(s,ss)$, $f_2 = \lambda (s,ss).(s::ss)$, $g_1 = \lambda v.(v,())$.
@@
 
-$ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
-\tab \quad = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
-\tab \quad = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
+% $ks_1 \ s_2 = (f_2 \circ (con \ (fst \circ g_1 \circ id) \ (snd \circ f_1 \circ ks_0))) \ s_2 \\
+% \tab \quad = f_2 \ (\ (fst(g_1(id(s_2))) \ , \ snd(f_1(ks_0(s_2))))\ )\\
+% \tab \quad = fst(g_1(id(s_2))) :: snd(f_1(ks_0(s_2))) = [100,2,3] :: snd(f_1(ks_0([100,2,3])))$
 
-If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
+% If $ks_0$ is an identity function, $ks_1 \ s_2 = [100,2,3] :: [2,3]$.
 This is an unexpected result when we target it to be the updated source.

### 2019-11-29 paper/1.introduction.tex
small change on the first paragraph of the introduction before boarding, and will revise the paper on the plane.

@@
 In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}.
-Bidirectional transformation (BX) is an approach to solve this problem.
-As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}.
-BX provides two functions: $get$ and $put$.
-Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.
+Bidirectional transformation (BX) provides a systematic approach to solving this problem.
+As an example, consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}, which consists of two functions: $get$ (for getting the head of an input list) and $put$ (for relfecting the output to the head of the input).
+Figure \ref{fig:eval-phead} shows the behaviors of these functions when evaluating $phead$.
 

### 2019-11-29 paper/3.minbigul_m.tex
small changes
fix indent + scaled tick
update tex 20191129
based on Robert's polishment

@@
-\section{Memorizing: minBiGUL$_m$}
+\section{Adding Memoization: minBiGUL$_m$}
 
-For minBiGUL or even BiGUL, when evaluating a composition of many programs, $get$s are reevaluated so many times since no intermediate state is stored during the evaluation.
-To avoid reevaluating $get$s, we can save intermediate states in a key-value table where the key is a pair of the program $bx$ and the source $s$, and the value is the result of $\getbx{bx}{s}$.
-Then, searching an evaluated $get$s on the table is used instead of recomputing it.
-We define a memory version, minBiGUL$_m$, including $get_m$ and $put_m$ as in definition 3 and 4 respectively.
-Another function introduced in definition 5, $get_{mh}$, is used to retrieve the value of an evaluated $get$ from the table or compute an unevaluated $get$ by using $get_m$ then store the key-value tuple to the table if necessary.
+\begin{definition} \label{def:putm} $\text{Memo version of } put$
 
-\begin{definition} $\text{Memo version of } put$
-
-    $\putm{bx}{s}{v} = \text{match } bx \text{ with }\\
+    \noindent $\putm{bx}{s}{v} = \text{match } bx \text{ with }\\
         \tab \vert \ bx_1 \circ bx_2 \to \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})} \\
@@
 \begin{multicols}{2}
-    \begin{definition} $\text{Memo version of } get$
+    \begin{definition} \label{def:getm} $\text{Memo version of } get$
 
-        $\getm{bx}{s} = \text{match } bx \text{ with }\\
+        \noindent $\getm{bx}{s} = \text{match } bx \text{ with }\\
             \tab \ \vert \ bx_1 \circ bx_2 \to\\
@@
 \columnbreak
-    \begin{definition} $???$
+    \begin{definition} \label{def:getmh} $\text{Interface for } get_m$
 
-        $\getmh{bx}{s} =\\
+        \noindent $\getmh{bx}{s} =\\
             \tab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
@@
 
-The special point is that not all states are saved when evaluating a program, only composition evaluations are treated.
+When evaluating a composition of several programs in minBiGUL (and the same holds for full BiGUL), the same $get$s have to be evaluated repeatedly because no intermediate states are kept during the evaluation.
+This problem was illustrated in Fig.~\ref{fig:eval-comp-phead}.
+To avoid reevaluating $get$s, and as our first approach, we introduce memoization in the minBiGUL interpreter.
+% can save 
+To keep it simple, the intermediate states of a composition are saved in a key-value table where the key is a pair of program $bx$ and source $s$, and the value is the result of evaluating $\getbx{bx}{s}$ the first time.
+Then, instead of recomputing the value, the value in the table is used.
+The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Defs.~\ref{def:putm} and~\ref{def:getm}).
+Another function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Def.~\ref{def:getmh}).
+
+Note as a special point that not all states are saved when evaluating a program, only when a composition is evaluated.
 When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, an application of $get_{mh}$, not $get$ or $get_m$, appears to indicate the saving.
-When evaluating $\getm{bx_1 \circ bx_2}{s}$, we try to search key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, on the table.
-The matched value can be used for next steps in the evaluation.
+When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation.
 If there is no such key, the value $v$ will be calculated by using $get_{mh}$.

### 2019-11-29 paper/6.related-work.tex
capitalize Section titles
add references

@@
 
-\subsubsection*{BX languages and implementations}
+\subsubsection*{BX Languages and Implementations}
 
@@
 
-\subsubsection*{Optimization techniques}
+\subsubsection*{Optimization Techniques}
 
-Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
+Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
 Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}.
@@
 Therefore we did not use this way.
-  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{}.
+  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
 Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
@@
 %Tupling \cite{tupling} is an optimization technique by combining several computations.
-Although tupling is usually used for reducing the evaluation cost\cite{}, we use it to reduce the size of the things need to keep.
+Although tupling is usually used for reducing the evaluation cost, we use it to reduce the size of the things need to keep.
 
@@
 
-%Lazy evaluation \cite{} is well-known and widly-used optimization technique to reduce the redundant evaluation.
+%Lazy evaluation is well-known and widly-used optimization technique to reduce the redundant evaluation.
 Because we use OCaml for implementation, we need to introduce this explicitly.

### 2019-11-29 paper/2.minbigul.tex
update references
add no indent

@@
 In addition, we restrict adaptive cases of BiGUL on minBiGUL.
-Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{putput}, like the following:
+Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{Foster:2007:CBT:1232420.1232424}, like the following:
 
@@
 
-$bfoldr$ will look like a composition of many programs at some points in the evaluation if we slow down this process.
+\noindent $bfoldr$ will look like a composition of many programs at some points in the evaluation if we slow down this process.
 Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.

### 2019-11-29 paper/3.minbigul_m.tex
change Fig. -> Figure

@@
 When evaluating a composition of several programs in minBiGUL (and the same holds for full BiGUL), the same $get$s have to be evaluated repeatedly because no intermediate states are kept during the evaluation.
-This problem was illustrated in Fig.~\ref{fig:eval-comp-phead}.
+This problem was illustrated in Figure~\ref{fig:eval-comp-phead}.
 To avoid reevaluating $get$s, and as our first approach, we introduce memoization in the minBiGUL interpreter.
@@
 Then, instead of recomputing the value, the value in the table is used.
-The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Defs.~\ref{def:putm} and~\ref{def:getm}).
-Another function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Def.~\ref{def:getmh}).
+The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}).
+Another function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).
 

### 2019-11-29 paper/1.introduction.tex
add references
shrinking..
add labels to sections

@@
 In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
-Note that the update to s' uses s and v' as input.
+Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
 %\texttt{put pHead [1,2] 8 = [8,2]}
@@
 In total, three $get$s are evaluated.
-We explain this method in Section 2.
+We explain this method in Section~\ref{sect:minbigul}.
 
-\textcolor{red}{If the programs are simple like the running examples, we can transform left associative compositions to right associative.
+\textcolor{red}{If the BX programs are simple like the running examples, we can transform left associative compositions to right associative.
 However, this is not always possible.
 For example, let us consider a BX 'reverse' program, $breverse$.
-$breverse$ is constructed by application of $bfoldr$ (the actual program is shown in Section 2) and $bsnoc$.
+$breverse$ is constructed by application of $bfoldr$ (the actual program is shown in the last part of Section~\ref{sect:minbigul}) and $bsnoc$.
 A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$.
@@
 Especially, when we use large inputs, the evaluation time will be long.
-We explain this method in Section 3.
+We explain this method in Section~\cite{sect:minbigl_m}.
 
@@
 \vspace{2mm}
-In very-well behaved \cite{} BX programs,
+In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
 $put$ is a complement function for $get$.
@@
 To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation.
-We explain the second method and optimization in Section 4.
+We explain the second method and optimization in Section~\ref{sect:xpg}.
 
-In Section 5, we show experimental results of the approaches. We discuss about related work in Section 6, and conclude in Section 7.
+In Section~\ref{sect:experiments}, we show experimental results of the approaches. We discuss about related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.
 
@@
   % \begin{itemize}
-    % Although the first strategy is implemented in BiGUL \cite{}, the second and third strategies are introduced by this paper.
+    % Although the first strategy is implemented in BiGUL, the second and third strategies are introduced by this paper.
 %  \item Thanks to introduction of the strategies, it is possible to compare the evaluation strategies.

### 2019-11-29 paper/1.introduction.tex
small fixes

@@
 Because of priority, it is impossible to transform this program to right associative compositions.
-At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because this BX program is dynamically produced depending on the length of the input list.
+At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because actual program is dynamically produced by recursions depending on the length of the input list.
 Therefore, it is important to have a fast evaluation method for left associative compositions.}
 
-For achieving this, we introduce two methods that use memoization.
+For fast evaluation for left associative compositions, we introduce two methods that use memoization.
 The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs.
@@
 Especially, when we use large inputs, the evaluation time will be long.
-We explain this method in Section~\cite{sect:minbigl_m}.
+We explain this method in Section~\ref{sect:minbigulm}.
 

### 2019-11-29 paper/2.minbigul.tex
add labels to sections
small fixes

@@
-\section{Bidirectional Programming Language: minBiGUL}
+\section{Bidirectional Programming Language: minBiGUL} \label{sect:minbigul}
 
@@
 
-BiGUL is well-behaved [?] since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:
+BiGUL is well-behaved \cite{Pacheco:2014:MCP:2543728.2543737} since two functions $\putbxinline{bx}$ and $\getbxinline{bx}$ satisfy the round-trip laws as follows:
 
@@
 
-The \textsc{GetPut} \cite{Pacheco:2014:MCP:2543728.2543737} law means that if there is no change to the view, there should be no change to the source.
-The \textsc{PutGet} \cite{Pacheco:2014:MCP:2543728.2543737} law means that we can recover the modified view by applying the forward transformation to the updated source.
+The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source.
+The \textsc{PutGet} law means that we can recover the modified view by applying the forward transformation to the updated source.
 
@@
 \begin{multicols}{2}
-    \begin{definition}
+    \begin{definition} \label{def:minbigulput}
         $\putbx{bx}{s}{v}$
@@
 \columnbreak
-    \begin{definition}
+    \begin{definition} \label{def:minbigulget}
         $\getbx{bx}{s}$
@@
 
-We define the semantics of $put$ and $get$ as in definitions 1 and 2 respectively.
+We define the semantics of $put$ and $get$ as in Definitions~\ref{def:minbigulput} and \ref{def:minbigulget} respectively.
 Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.

### 2019-11-29 paper/4.xpg.tex
capitalize Section titles
add labels to sections
small fixes

@@
-\section{Tupling and lazy updates: xpg}
+\section{Tupling and Lazy Updates: xpg} \label{sect:xpg}
 \subsection{Tupling: pg}
@@
 
-\subsection{Lazy update: cpg}
+\subsection{Lazy Update: cpg}
 
@@
 This is an unexpected result when we target it to be the updated source.
-If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, the result will be what we want to see.$ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
+If $ks_0 = \lambda \_.s$ where $s = [[1,2,3], [\,], [4,5]]$, the result will be what we want to see.
+$ks_1 \ s_2 = [100,2,3] :: [[\,], [4,5]] = [[100,2,3], [\,], [4,5]]$.
 Through the above example, using the continuation $ks$ as a constant function at the beginning contributes to retain the discarded things in the source.
 
-\subsection{Lazy computation: kpg}
+\subsection{Lazy Computation: kpg}
 
@@
 It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-One more thing, in $\kpginline{Case}$, we need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the fi condition before returning the 6-tuple.
+One more thing, in $\kpginline{Case}$, we need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the $fi$ condition before returning the 6-tuple.
 Such evaluations can be done lazily to make programs run faster.

### 2019-11-29 paper/2.minbigul.tex
small fixes

@@
 We define the semantics of $put$ and $get$ as in Definitions~\ref{def:minbigulput} and \ref{def:minbigulget} respectively.
-Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.
+Instead of using $v'$ in the $put$ direction like Figures~\ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.
 The later definitions also follow this convention.

### 2019-11-29 paper/7.conclusion.tex
add labels to sections
small fixes

@@
-\section{Conclusion and Future Work}
+\section{Conclusion and Future Work} \label{sect:conclusion}
 
@@
 To achieve fast evaluation, we introduce two different methods using memoization.
-From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not very big and $xpg$ is fastest for other left associative programs.
+From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not large and $xpg$ is fastest for other left associative programs.
 This shows that if programmer choose one method based on the their BX programs and inputs, they can use efficient evaluation.
@@
 First is overcomming our limitation in $xpg$.
-In $xpg$, there are two limitations: only for very-well-behaved programs, restriction for $\pginline{Case}$ (shown in Section 4.1).
+In $xpg$, there are two limitations: only for very-well-behaved programs and restriction for $\pginline{Case}$ in Section~\ref{sect:pg}.
 To be a practical evlauation method, extention of the target language is needed.
@@
 
-Second is introducing analysis about BX programs and inputs to choose the best evaluation method automatically.
-Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If automatic choice is possible, we can reduce programmer's burden.
+Second is introducing an automatic analysis about BX programs and inputs to choose the best evaluation method.
+Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If this is achieved, we can reduce programmer's burden.
 
@@
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-We need to investigate more about $construct\_dummy$, because current definition that is used in Section 6 will cause type errors.
+We need to investigate more about $construct\_dummy$, because current definition that is used in Section~\ref{sect:experiments} will cause type errors.
 

### 2019-11-29 paper/1.introduction.tex
small fixes
no change

@@
 
-For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
-As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in figure \ref{fig:eval-comp-phead-2}.
+\noindent For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
+As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}.
 \texttt{c$_1$} and \texttt{c$_2$} are complements.

### 2019-11-29 paper/1.introduction.tex
revising the introduction (on the way)

--- 2019-11-29 paper/1.introduction.tex
@@
 Bidirectional transformation (BX) provides a systematic approach to solving this problem.
-As an example, consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}, which consists of two functions: $get$ (for getting the head of an input list) and $put$ (for relfecting the output to the head of the input).
-Figure \ref{fig:eval-phead} shows the behaviors of these functions when evaluating $phead$.
+Consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}, which consists of two functions: $get$ (for getting the head of an input list) and $put$ (for reflecting the output to the head of the input).
+Figure \ref{fig:eval-phead} shows an example of the bidirectional behavior of $phead$.
+Let $[1,2]$ be the original source \texttt{s}.
+$get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
+%\texttt{get pHead [1,2] = 1}.
+%After the view \texttt{v} is obtained, the user can modify the view.
+%In this case, the user
+%modified the view
+%from $1$ to $100$.
+%From this updated view, how can the source be updated?
+%This is ``the view update problem.''
+%To solve this problem, BX provides $put$, an update function on the original source.
+%In this example,
+Supposing that the view is updated to $100$,
+$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view (\texttt{v'}) of $2$ and the original source (\texttt{s}) of $[1,2]$.
+%Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
+%\texttt{put pHead [1,2] 8 = [8,2]}
 
@@
 
-Here, let $[1,2]$ be the original source \texttt{s}.
-$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
-%\texttt{get pHead [1,2] = 1}.
-After the view \texttt{v} is obtained, the user can modify the view.
-In this case, the user
-modified the view
-from $1$ to $100$.
-From this updated view, how can the source be updated?
-This is ``the view update problem.''
-To solve this problem, BX provides $put$, an update function on the original source.
-In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
-Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
-%\texttt{put pHead [1,2] 8 = [8,2]}
+The composition of BX programs is a fundamental construct to build more complex BX programs \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
+Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_2 \circ bx_1$ is defined by
+\begin{align}
+get_{bx_2 \circ bx_1}~s &= get_{bx_2} (get_{bx_1}~s)\\
+put_{bx_2 \circ bx_1}~(s,v') &= put_{bx_1} (s,put_{bx_2} (get_{bx_1}~s,v'))
+\end{align}
 
-The composition of BX programs is a fundamental construct to build more complex BX programs.
 %In BX programs, composition of programs is essential for writing various interesting evaluations.
-Because recursions are achieved by compositions, large number of programs include compositions.
-In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
-In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
+%Because recursions are achieved by compositions, large number of programs include compositions.
+%In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
+%In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
 The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
@@
 
-Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
-By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$.
+One interesting thing is that $put_{bx_2 \circ bx_1}$ needs to call $get_{bx_1}$ to compute the intermediate result, which would introduce an efficient problem when we compose many bidirectional programs.
+Generally, if we compose $O(n)$ bidirectional programs, we would call $get$ for $O(n^2)$ times.
+Figure \ref{fig:eval-comp-phead} illustrates the evaluation of our running example:
+\[
+lp3 = (phead \circ phead) \circ phead
+\]
+where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
+%By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$.
 To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times.
@@
 
-First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}.
+%First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}.
 This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}.
@@
 We explain this method in Section~\ref{sect:minbigul}.
+
+\textcolor{red}{I am going to revise the rest of the introduction from here.}
 

### 2019-11-29 paper/5.experiments.tex
update references
add labels to sections

@@
-\section{Experiments}
+\section{Experiments} \label{sect:experiments}
 
@@
 
-\subsection{Test cases}
+\subsection{Test Cases}
 
-We have conducted 7 test cases shown in Table \ref{tab:test-cases}.
+We have conducted 7 test cases shown in Table~\ref{tab:test-cases}.
 Although the number of programs is not so much, they are not trivial and can be completely used in practice.
@@
 Because of the type of the source, this program is seen as a \textit{ldata} case.
-Next, we briefly explain the behavior of the remaining compositions in table \ref{tab:test-cases}.
+Next, we briefly explain the behavior of the remaining compositions in Table~\ref{tab:test-cases}.
 

### 2019-11-29 paper/6.related-work.tex
add labels to sections

@@
-\section{Related Work}
+\section{Related Work} \label{sect:related}
 
@@
 based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
-The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
+The semantics and correctness of BX programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
 Anjorin et.
@@
 However, the improvement for practical implementation of BX languages is still missing.
-As a first step, we study different evaluation methods to optimized their evaluation.
+As a first step, we study different evaluation methods to optimize their evaluation.
 We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
-In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4).
+In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section~\ref{sect:minbigul}) and our methods (in Sections~\ref{sect:minbigulm} and \ref{sect:xpg}).
 From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
-In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
+In this paper we focus on a BX language, BiGUL, however our methods can be applicable to other BX languages which have similar problems.
 
@@
 Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
-Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}.
-However, fusion of BX programs are not considered well \textcolor{red}{(really?)}.
-Therefore we did not use this way.
-  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
+Another way to reduce extra evaluation is program fusion \cite{Wadler:1988:DTP:80099.80104}.
+However, we can not apply fusion to dynamically produced programs by recursion.
+Therefore we did not use fusion in this work.
+  Instead of that, we introduced another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
 Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
@@
 Because we use OCaml for implementation, we need to introduce this explicitly.
-If we use Haskell, a lazy evaluated language, we do not need to introduce them.
+If we use Haskell, a lazy evaluated language, we do not need to introduce it.
 

### 2019-11-30 paper/1.introduction.tex
small
finish revising the introduction

@@
 get_{bx_2 \circ bx_1}~s &= get_{bx_2} (get_{bx_1}~s)\\
-put_{bx_2 \circ bx_1}~(s,v') &= put_{bx_1} (s,put_{bx_2} (get_{bx_1}~s,v'))
+put_{bx_2 \circ bx_1}~s~v' &= put_{bx_1} s~(put_{bx_2}~ get_{bx_1}~s)~v')
 \end{align}
+One feature of this composition is that $put_{bx_2 \circ bx_1}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficient problem if we compute $put$ for composition of many bidirectional programs.
+Generally, for a composition of $O(n)$ bidirectional programs, we eed to call $get$ for $O(n^2)$ times.
+To be concrete, consider the evaluation of the following composition (which will be used as our running example in this paper):
+\[
+lp3 = (phead \circ phead) \circ phead
+\]
+which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source \texttt{s} being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
+To obtain the final updated source \texttt{s'}, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
+The first is from \texttt{i2} and \texttt{v'} to obtain \texttt{i2'}, which needs to call $get$ twice to compute $i2$; the second is from \texttt{i1} and \texttt{i2'} to obtain \texttt{i1'}, which needs to call $get$ once, and the last is from \texttt{s} and \texttt{i1'} to obtain \texttt{s'}, which is just a direct $put$ computation.
 
@@
 
-One interesting thing is that $put_{bx_2 \circ bx_1}$ needs to call $get_{bx_1}$ to compute the intermediate result, which would introduce an efficient problem when we compose many bidirectional programs.
-Generally, if we compose $O(n)$ bidirectional programs, we would call $get$ for $O(n^2)$ times.
-Figure \ref{fig:eval-comp-phead} illustrates the evaluation of our running example:
-\[
-lp3 = (phead \circ phead) \circ phead
-\]
-where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
 %By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$.
@@
 
-\textcolor{red}{I am going to revise the rest of the introduction from here.}
+One direct solution to avoid this repeated computation of $get$ is to compute the composition in a right associative manner. For instance, if we transform  $lp3$ to $rp3$:
+\[
+rp3 = phead \circ (phead \circ phead)
+\]
+then the $put$ for $rp3$ only needs to compute $get$ of $phead$ twice, one time less than that for $lp3$.
+ However, this transformation is not always easy to do.
+For instance, let us consider $breverse$, a bidirectional version of the traditional 'reverse' program for reversing a list.
+It is defined using $\mbox{\it bfoldr}$, a bidrecrtional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}).
+Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
+ \[
+ \mbox{\it bfoldr}~bx~\cdots = \cdots  (\mbox{\it bfoldr}~\cdots) \circ bx \cdots
+ \]
+where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
+This make it hard to do the above transformation statically.
 
-\textcolor{red}{If the BX programs are simple like the running examples, we can transform left associative compositions to right associative.
-However, this is not always possible.
-For example, let us consider a BX 'reverse' program, $breverse$.
-$breverse$ is constructed by application of $bfoldr$ (the actual program is shown in the last part of Section~\ref{sect:minbigul}) and $bsnoc$.
-A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$.
+%A part of the definition of $\mbox{\it bfoldr}$, we have $(\product{Replace}{(\mbox{\it bfoldr} \ bx)}) \circ bx$.
 In this program, recursion occurs on the left-hand side of composition.
@@
 At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because actual program is dynamically produced by recursions depending on the length of the input list.
-Therefore, it is important to have a fast evaluation method for left associative compositions.}
-
-For fast evaluation for left associative compositions, we introduce two methods that use memoization.
-The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' There is no quadratic $get$s in this method, therefore this method improves the runtime of many BX programs.
-However, a disadvantage of this is the familiar tradeoff for keeping and searching values in a table.
-Especially, when we use large inputs, the evaluation time will be long.
-We explain this method in Section~\ref{sect:minbigulm}.
-
-To treat large inputs, we introduce the second method that uses memoization: ``keeping complements in a closure and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states.
-%Therefore this strategy solves the previous two problems.
-Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
-
-\vspace{2mm}
-In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
-$put$ is a complement function for $get$.
-%, $get$ can be a complement function for $put$.
-\vspace{2mm}
+Therefore, it is important to have a fast evaluation method for left associative compositions.
 
@@
 
-\noindent For obtaining complements, we use tupling of $put$ and $get$, and produce new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
+In this paper, we made the first attempt for seriously considering efficiency of evaluation of BX compositions, and
+solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.
+The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' This can avoid repeated computation of  $get$s, therefore it improves the runtime of many BX programs.
+However, this simple memoization needs to keep and search values in a table, which may introduce big cost for large inputs.
+We will explain this method in Section~\ref{sect:minbigulm}.
+
+To treat large inputs, we propose the second method based on memoization: ``keeping complements in a closure and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states.
+%Therefore this strategy solves the previous two problems.
+%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
+%
+%\vspace{2mm}
+%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
+%$put$ is a complement function for $get$.
+%, $get$ can be a complement function for $put$.
+%\vspace{2mm}
+%
+For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
 As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}.
@@
 \texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}.
-The points in this figure is two.
+Here, two points are worth noting.
 First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}.
@@
 
-In Section~\ref{sect:experiments}, we show experimental results of the approaches. We discuss about related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.
+Both methods were fully implemented for minBiGUL, a core bidirectional language. The experimental results show that our methods are much faster than the original evaluation strategy.
+Iwe give detailed experimental results in Section~\ref{sect:experiments}, discuss related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.
 
-The contributions of this paper are the following.
+%The contributions of this paper are the following.
 
-\begin{itemize}
-\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions.
+%\begin{itemize}
+%\item This is the first attempt for seriously considering efficiency of evaluation of BX compositions.
 Our methods are better than or similar to the results by the original method in BiGUL for all test cases.
-\item We show that memoization is effective for achieving evaluation efficiency in BX languages.
-\item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt.
-\item Although we focus on a BX language BiGUL in this paper, these techniques can be potentially used in other BX languages.
+%\item We show that memoization is effective for achieving evaluation efficiency in BX languages.
+%\item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt.
+%\item Although we focus on a BX language BiGUL in this paper, these techniques can be potentially used in other BX languages.
   % \begin{itemize}
@@
 %  \end{itemize}
-\end{itemize}
+%\end{itemize}

### 2019-11-30 paper/1.introduction.tex
minor

@@
 
-Both methods were fully implemented for minBiGUL, a core bidirectional language. The experimental results show that our methods are much faster than the original evaluation strategy.
-Iwe give detailed experimental results in Section~\ref{sect:experiments}, discuss related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.
+Both methods have been fully implemented for minBiGUL, a core bidirectional language. The experimental results show that our methods are much faster than the original evaluation strategy.
+We give detailed experimental results in Section~\ref{sect:experiments}, discuss related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.
 

### 2019-11-30 paper/6.related-work.tex
revised related work.

--- 2019-11-29 paper/6.related-work.tex
@@
-\section{Related Work}
+\section{Related Work} \label{sect:related}
 
-\subsubsection*{BX Languages and Implementations}
+%\subsubsection*{BX Languages and Implementations}
 
-There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
+Since the pioneer work of Lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
-based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
-The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood.
+%based on several application scenarios and BX lenses works \cite{}.
+Although much progress has been made on the semantics and correctness of BX programs for the past years, as far as we are aware, little work has been done on optimization of BX programs \cite{Horn:2018:IRL:3243631.3236769}.
 Anjorin et.
 al.
-introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}.
-However, the improvement for practical implementation of BX languages is still missing.
-As a first step, we study different evaluation methods to optimized their evaluation.
-We can say this paper will be first attempt to improve efficiency of BX composition evaluation.
-In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4).
-From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
-In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.
+introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
+This paper shows the first attempt of improving efficiency of BX composition evaluation.
+
+The baseline of this work is the BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, and we compare BiGUL's method (in Section~\ref{sect:minbigul}) with our methods (in Sections~\ref{sect:minbigulm} and \ref{sect:xpg}).
+From experimental results of left associative BX composition programs, we can see that our memoization methods are faster than the original BiGUL's evaluation method.
+While we focus on BiGUL, our methods are general and should be applicable to other BX languages.
 
@@
 
-\subsubsection*{Optimization Techniques}
+%\subsubsection*{Optimization Techniques}
 
-Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
-Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}.
-However, fusion of BX programs are not considered well \textcolor{red}{(really?)}.
-Therefore we did not use this way.
-  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
-Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
-  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation.
-Because we use OCaml for implementation, we need to introduce this explicitly.
-If we use Haskell, a lazy evaluated language, we do not need to introduce them.
+Our work is related to many known optimization methods for unidirectional programs.
+Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is a technique to avoid repeated redundant computation.
+In our case, we show that two specific memoization methods can be used for bidirectional programs.
+To deal with inefficiency due to compositions, many fusion methods have been studied \cite{Wadler:1988:DTP:80099.80104} to merge a composition of two (recursive) programs into one.
+However, under the context of bidirectional programs, we need to consider not only compositions of recursive programs but also compositions inside a recursive program (as we have seen in \mbox{\it bfoldr}).
+This paper focuses on the composition inside a recursion, where compositions are produced dynamically at runtime.
+We tackled the problem by using tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
 

### 2019-11-30 paper/1.introduction.tex
revise typing
+ bx2 `compose` bx1 => bx1 `compose` bx2
+ \text{s} => $s$

@@
 Figure \ref{fig:eval-phead} shows an example of the bidirectional behavior of $phead$.
-Let $[1,2]$ be the original source \texttt{s}.
+Let $[1,2]$ be the original source $s$.
 $get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
@@
 Supposing that the view is updated to $100$,
-$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view (\texttt{v'}) of $2$ and the original source (\texttt{s}) of $[1,2]$.
+$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view ($v'$) of $2$ and the original source ($s$) of $[1,2]$.
 %Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
@@
 The composition of BX programs is a fundamental construct to build more complex BX programs \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
-Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_2 \circ bx_1$ is defined by
+Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_1 \circ bx_2$ is defined by
 \begin{align}
-get_{bx_2 \circ bx_1}~s &= get_{bx_2} (get_{bx_1}~s)\\
-put_{bx_2 \circ bx_1}~s~v' &= put_{bx_1} s~(put_{bx_2}~ get_{bx_1}~s)~v')
+get_{bx_1 \circ bx_2}~s &= get_{bx_2} (get_{bx_1}~s)\\
+put_{bx_1 \circ bx_2}~s~v' &= put_{bx_1} s~(put_{bx_2}~ (get_{bx_1}~s)~v')
 \end{align}
-One feature of this composition is that $put_{bx_2 \circ bx_1}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficient problem if we compute $put$ for composition of many bidirectional programs.
-Generally, for a composition of $O(n)$ bidirectional programs, we eed to call $get$ for $O(n^2)$ times.
+One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficient problem if we compute $put$ for composition of many bidirectional programs.
+Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times.
 To be concrete, consider the evaluation of the following composition (which will be used as our running example in this paper):
@@
 \]
-which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source \texttt{s} being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
-To obtain the final updated source \texttt{s'}, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
-The first is from \texttt{i2} and \texttt{v'} to obtain \texttt{i2'}, which needs to call $get$ twice to compute $i2$; the second is from \texttt{i1} and \texttt{i2'} to obtain \texttt{i1'}, which needs to call $get$ once, and the last is from \texttt{s} and \texttt{i1'} to obtain \texttt{s'}, which is just a direct $put$ computation.
+which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source $s$ being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
+To obtain the final updated source $s'$, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
+The first is from $i_2$ and $v'$ to obtain $i'_2$, which needs to call $get$ twice to compute $i_2$; the second is from $i_1$ and $i'_2$ to obtain $i'_1$, which needs to call $get$ once, and the last is from $s$ and $i'_1$ to obtain $s'$, which is just a direct $put$ computation.
 
@@
 For instance, let us consider $breverse$, a bidirectional version of the traditional 'reverse' program for reversing a list.
-It is defined using $\mbox{\it bfoldr}$, a bidrecrtional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}).
+It is defined using $\mbox{\it bfoldr}$, a bidrecrtional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}.
 Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
@@
 solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.
-The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed.'' This can avoid repeated computation of  $get$s, therefore it improves the runtime of many BX programs.
+The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed''.
+This can avoid repeated computation of  $get$s, therefore it improves the runtime of many BX programs.
 However, this simple memoization needs to keep and search values in a table, which may introduce big cost for large inputs.
@@
 
-To treat large inputs, we propose the second method based on memoization: ``keeping complements in a closure and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states.
+To treat large inputs, we propose the second method based on memoization: ``keeping complements in a closure and using them when they are needed''.
+Complements are smaller program fragments than the original intermediate states.
 %Therefore this strategy solves the previous two problems.
@@
 As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}.
-\texttt{c$_1$} and \texttt{c$_2$} are complements.
-\texttt{d$_1$} and \texttt{d$_2$} are valid views for \texttt{s} and \texttt{i$_1$}.
+$c_1$ and $c_2$ are complements.
+$d_1$ and $d_2$ are valid views for $s$ and $i_1$.
 Here, two points are worth noting.
-First, after evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c$_1$} and \texttt{i$_1$}.
+First, after evaluation of the first $pg$, we do not need to keep the original source $s$, because all its information is in $c_1$ and $i_1$.
 Second, complements are smaller than the intermediate states in the previous figure.

### 2019-11-30 paper/main.tex
add explanation for breverse and titlerunning
reduce footnote size
capitalize one

@@
 \renewcommand\UrlFont{\color{blue}\rmfamily}
+\renewcommand{\footnotesize}{\scriptsize}
 
@@
   % \thanks{Supported by organization x.}  
-  An efficient composition of bidirectional programs
-  by memoization and lazy update
+  An Efficient Composition of Bidirectional Programs
+  by Memoization and Lazy Update
 }
 %
+\titlerunning{Strategies for efficient evaluation of BX programs}
 %\titlerunning{Abbreviated paper title}
@@
 \input{7.conclusion.tex}
-
-\textcolor{red}{TODO: fix references (order, contents), add missing references}
 

### 2019-11-30 paper/1.introduction.tex
Update 1.introduction.tex
Polished tiny LaTeX details; polished description of Fig.3; corr'ed typo (l.44: 2 --> 100).

@@
 The synchronization of data is a common problem.
-In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}.
+In the database community this problem is known as ``the view update problem'' and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}.
 Bidirectional transformation (BX) provides a systematic approach to solving this problem.
@@
 Figure \ref{fig:eval-phead} shows an example of the bidirectional behavior of $phead$.
-Let $[1,2]$ be the original source $s$.
+Let $[1,2]$ be the original source~$s$.
 $get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
@@
 Supposing that the view is updated to $100$,
-$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view ($v'$) of $2$ and the original source ($s$) of $[1,2]$.
+$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view $v'$ of $100$ and the original source $s$ of $[1,2]$.
 %Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
@@
 \end{align}
-One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficient problem if we compute $put$ for composition of many bidirectional programs.
+One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
 Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times.
@@
 To obtain the final updated source $s'$, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
-The first is from $i_2$ and $v'$ to obtain $i'_2$, which needs to call $get$ twice to compute $i_2$; the second is from $i_1$ and $i'_2$ to obtain $i'_1$, which needs to call $get$ once, and the last is from $s$ and $i'_1$ to obtain $s'$, which is just a direct $put$ computation.
+The first is from $i_2$ and $v'$ to obtain $i'_2$, which needs to call $get$ twice to compute $i_2$; the second is from $i_1$ and $i'_2$ to obtain $i'_1$, which needs to call $get$ once, and the last is from $s$ and $i'_1$ to obtain~$s'$, which is just a direct $put$ computation.
 
@@
  However, this transformation is not always easy to do.
-For instance, let us consider $breverse$, a bidirectional version of the traditional 'reverse' program for reversing a list.
-It is defined using $\mbox{\it bfoldr}$, a bidrecrtional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}.
+For instance, let us consider $breverse$, a bidirectional version of the traditional `reverse' program for reversing a list.
+It is defined using $\mbox{\it bfoldr}$, a bidirectional version of the traditional $\mbox{\it foldr}$, whose definition  is shown in the last part of Section~\ref{sect:minbigul}.
 Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
@@
 
-In this paper, we made the first attempt for seriously considering efficiency of evaluation of BX compositions, and
+In this paper, we made the first attempt for seriously considering the efficiency of evaluating BX compositions, and
 solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.
-The first method uses memoization simply: ``keeping all intermediate states and using them when they are needed''.
-This can avoid repeated computation of  $get$s, therefore it improves the runtime of many BX programs.
+The first method uses straightforward memoization: ``keeping \emph{intermediate states in a table} and using them when needed''.
+This avoids repeated $get$ computations and improves the runtime in many cases.
 However, this simple memoization needs to keep and search values in a table, which may introduce big cost for large inputs.
-We will explain this method in Section~\ref{sect:minbigulm}.
+We explain this method in Section~\ref{sect:minbigulm}.
 
-To treat large inputs, we propose the second method based on memoization: ``keeping complements in a closure and using them when they are needed''.
-Complements are smaller program fragments than the original intermediate states.
+To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
+The key idea is straightforward: Complements are smaller
+% program fragments 
+than
+% the original 
+intermediate states.
 %Therefore this strategy solves the previous two problems.
@@
 For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
-As an example, let us consider the previous example again: $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}.
-$c_1$ and $c_2$ are complements.
-$d_1$ and $d_2$ are valid views for $s$ and $i_1$.
+Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$.
 Here, two points are worth noting.
-First, after evaluation of the first $pg$, we do not need to keep the original source $s$, because all its information is in $c_1$ and $i_1$.
-Second, complements are smaller than the intermediate states in the previous figure.
+First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents is in $c_1$ and~$i_1$.
+Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
 %This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
-Actually, the simple combined $pg$ is not effective for left associative compositions, because this requires other two $put$ in the right part of the figure.
-To achieve efficient evaluation, we use two techniques, lazy update and lazy evaluation.
-We explain the second method and optimization in Section~\ref{sect:xpg}.
+Actually, this simple 
+% combined 
+$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen one the right side of Figure \ref{fig:eval-comp-phead-2}.
+To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation.
+We explain this second method and all optimizations in Section~\ref{sect:xpg}.
 
-Both methods have been fully implemented for minBiGUL, a core bidirectional language. The experimental results show that our methods are much faster than the original evaluation strategy.
+Both methods have been fully implemented for \emph{minBiGUL}, a core bidirectional language, which is a subset of the full bidirectional language \emph{BiGUL}.
+The experimental results show that our methods are much faster than the original evaluation strategy.
 We give detailed experimental results in Section~\ref{sect:experiments}, discuss related work in Section~\ref{sect:related}, and conclude in Section~\ref{sect:conclusion}.

### 2019-12-01 paper/1.introduction.tex

@@
 where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
-This make it hard to do the above transformation statically.
+This makes it hard to do the above transformation statically.
 
@@
 
-In this paper, we made the first attempt for seriously considering the efficiency of evaluating BX compositions, and
+In this paper, we make the first attempt for seriously considering the efficiency of evaluating BX compositions, and
 solve the problem by introducing two methods based on memoization to gain fast evaluation for (left associative) BX compositions.

### 2019-12-01 paper/5.experiments.tex

@@
 
-We have implemented all methods described above in the same environment as follows: macOS 10.14.6, processor Intel Core i7 (2.6 GHz), RAM 16 GB 2400 MHz DDR4, OCaml 4.07.1.
+We have implemented and tested all methods% 
+%
+\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1.
 The OCaml runtime system options and garbage collection parameters are set as default.
-There is no type system in our implementation\footnote{\url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}, so instead of constructing some dummies which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
+No type system in our implementation.}$^,$%HACK!
+%
+\footnote{Anonymous \url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}
+%
+described in the previous sections.
+Instead of constructing some dummies, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
 
@@
 
-We have conducted 7 test cases shown in Table~\ref{tab:test-cases}.
-Although the number of programs is not so much, they are not trivial and can be completely used in practice.
+We have conducted 7 test cases shown in Table~\ref{tab:test-cases}. Although the number of cases is small, 
+% they are not trivial and can be completely used in practice. 
+they represent non-trivial cases of practical significance.
 In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$.
 This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table.
-These $bx$ are left associative composition since we we are only interested in this kind of program in the paper.
+These $bx$ are left associative compositions since we we are only interested in this kind of program in the paper.
 Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
@@
 
-The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from $n + 1$ similar \textit{pure} programs.
-We use the term \textit{pure} here to refer non-recursion compositions.
-The prefix \textit{lcomp} means that the associate precedence of $\circ$ in a straight-line composition is left.
+% The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from the same $n + 1$ \textit{pure} programs. 
+The first five test cases
+% (no.~1--5) 
+are straight-line compositions of the same $n + 1$ non-recursive programs. We refer to them as `pure' compositions.
+% We use the term \textit{pure} here to refer non-recursion compositions. 
+The prefix \textit{lcomp} in a test name indicates that the compositions are left-associative.
+% that the associate precedence of $\circ$ in a straight-line composition is left.
 The suffix \textit{ldata} indicates that the input size is large.
-The symbol $L$ used in \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$.
-These structures are only intended to generate data enough large for observing experimental results.
+The symbol $L$ in the input column of the \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$.
+These structures are only intended to generate test data that is large enough for
+% observing 
+measuring
+% experimental 
+results.
 

### 2019-12-02 paper/1.introduction.tex
small fixes, add one sentence from Robert's comments

@@
 This makes it hard to do the above transformation statically.
+The same efficiency problem occurs in all BX languages.
 
@@
 Here, two points are worth noting.
-First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents is in $c_1$ and~$i_1$.
+First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$.
 Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
@@
 % combined
-$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen one the right side of Figure \ref{fig:eval-comp-phead-2}.
+$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}.
 To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation.

### 2019-12-02 paper/2.minbigul.tex
small fixes, change on explanation of bfoldr

@@
 minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which satisfy two above laws.
-In addition, we restrict adaptive cases of BiGUL on minBiGUL.
+In addition, we restrict 'adaptive case's of BiGUL on minBiGUL.
 Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{Foster:2007:CBT:1232420.1232424}, like the following:
@@
 The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates.
-Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.
+Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved \cite{Foster:2007:CBT:1232420.1232424}.
 
@@
 A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinator without adaptive cases or a composition of some minBiGUL programs.
-We use numbers, pairs and lists to construct the program input including the source and/or the view.
+We use numbers, pairs and lists to construct the program inputs including the source and/or the view.
 
@@
 This statement is useful to describe many functions related to $Case$ in this paper.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, i.e.
-$E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.
-If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
+Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means ``if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, if $E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.'' If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
 We can write the equivalent if-then-else statement as follows:
@@
 $\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false; \}\\
-    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false; \}$
+\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false \}\\
+    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false \}$
 
-\noindent We now consider the definition of $phead$ in minBiGUL:\\
+\noindent As an example of minBiGUL program, consider the definition of $phead$ :\\
 $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
@@
 For instance, $\putbx{phead}{[1,2,3]}{100} = [100,2,3]$ and $\getbx{phead}{[1,2,3]} = 1$.
-If we wanna update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
+If we want to update the head element of the head element of a list of lists by using the view, we can define a composition like $phead \circ phead$.
 For example:
@@
 
-In the end of this section, let's see the definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
+In the end of this section, let's see the minBiGUL definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
 
@@
 
-\noindent $bfoldr$ will look like a composition of many programs at some points in the evaluation if we slow down this process.
+\noindent In $bfoldr$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
 Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.
+%will look like a composition of many programs at some points in the evaluation if we slow down this process.
+Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.

### 2019-12-02 paper/4.xpg.tex
small fixes, add some explanation of construct_dummy

@@
 \section{Tupling and Lazy Updates: xpg} \label{sect:xpg}
-\subsection{Tupling: pg}
+\subsection{Tupling: pg} \label{sect:pg}
 
@@
 This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
 This does not happen for the others which is used in the construction of $\pginline{Case}$.
@@
 This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require a valid view from programmers.
-Since there is no real view, these things are necessary.
-The last equality changes $dummy$ by an application $construct\_dummy \ s$, and also substitutes $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
+            %We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require programmers to give a valid view.
+            Since there is no real view, this $dummy$ is necessary.
+The last equality changes two.
+First one is a change from $dummy$ to an application $construct\_dummy \ s$.
+This shows that we need a way to construct $dummy$ from the current source, $s$.
+We can require programmers to give this $construct\_dummy$ function.
+Second is a substitution on $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
 The replacement of $s$ and $s_1$ is possible because of the \textsc{PutPut} law.

### 2019-12-02 paper/4.xpg.tex
add some explanation of accumulation, move explanation of redundancy to the last of 4.2

@@
 Therefore, the runtime inefficiency is inevitable.
-To solve that, we introduce a new function, $cpg$, accumulates changes in the source and the view.
-$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values.
+To solve that, we introduce a new function, $cpg$, accumulates updates in the source and the view.
+$\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values same as $pg$.
 The output of this function is a 4-tuple $(ks, kv, s, v)$.
@@
 
-Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
+In the places where thrid and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
 
-For $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
+%Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
+
+Thanks to this accumulation, in $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
 The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
@@
 Then the continuation updating on this dummy should be initiated as an identity function.
-The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$ where $s_1$ is redundant because this evaluated variable is not used in the later steps.
+The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$.
 In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input.
@@
 This constant function helps to retain the discarded things in the source.
+
+        Note that, in $\cpginline{bx_1 \circ bx_2}$, $s_1$ is redundant because this evaluated variable is not used in the later steps.
+        In the next session, we will optimize this redundancy.
 

### 2019-12-02 paper/4.xpg.tex
add some explanation of kpg

@@
 Every values are evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modification information, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
+The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
 The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
@@
 
-The beginning of accumulative functions $ks'$ and $kv'$ are set as identity, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
+The beginning of accumulative functions $ks'$ and $kv'$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
 
@@
 
-Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
+In $kpg$, basically, functions for the updates will be kept in $ks'$ and $kv'$.
+$\kpginline{Skip}$ and $\kpginline{Case}$ are special cases, because they require the actual values for evaluation.
+Therefore we evaluate the values by application of $ks' s$ and $kv' v$ and update $ks'$ and $ks'$ with the identity fuctions.
+%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
 When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
-In $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+
+In $kpg$, we did two optimizations.
+First is in $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
 There may be same computations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
-It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that.
-One more thing, in $\kpginline{Case}$, we need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the $fi$ condition before returning the 6-tuple.
+To remove duplicate evaluations, we evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$.
+%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
+Second is in $\kpginline{Case}$.
+We need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the $fi$ condition before returning the 6-tuple.
 Such evaluations can be done lazily to make programs run faster.

### 2019-12-02 paper/5.experiments.tex
small fixes

@@
 The OCaml runtime system options and garbage collection parameters are set as default.
-No type system in our implementation.}$^,$%HACK!
+In our language, there is no type system.}$^,$%HACK!
 %
@@
 described in the previous sections.
-Instead of constructing some dummies, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the updated view that helps a program in the $put$ direction valid.
+Instead of constructing some dummies by $construct\_dummy$s, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the current updated views.
+This helps a program in the $put$ direction valid.
 
@@
 This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table.
-These $bx$ are left associative compositions since we we are only interested in this kind of program in the paper.
+These $bx$s are left associative compositions since we are only interested in this kind of program in the paper.
 Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
@@
 % (no.~1--5) 
-are straight-line compositions of the same $n + 1$ non-recursive programs. We refer to them as `pure' compositions.
+are straight-line compositions of the same $n + 1$ non-recursive programs.
+% We refer to them as `pure' compositions.
 % We use the term \textit{pure} here to refer non-recursion compositions. 
@@
 $\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$.
-$\getbx{ptail}{s}$ return the tail of the source list $s$.
+$\getbx{ptail}{s}$ returns the tail of the source list $s$.
 The compositions of many $ptail$, in the \textit{put} direction, replace a part of the tail of the source list by the view list, in the \textit{get} direction, returns the such tail from the the source.
@@
 $\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$ as two same length lists, to produce a new list by moving the last element of $v$ to its first position.
-$\getbx{ptail}{s}$ return another list by moving the first element of the list $s$ to its end position.
+$\getbx{ptail}{s}$ returns another list by moving the first element of the list $s$ to its end position.
 The compositions of many \textit{ptail}, in the \textit{put} direction, create a permutation of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a permutation of the source list.
@@
 \textit{breverse} is defined in terms of \textit{bfoldr}, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
+Note that compositions is by the recursions of $breverse$ and
+the number of compositions are dynamically determined by the length of the source list.
 
@@
 For left associative compositions, $put$ works poorly because of the number of reevaluated $get$s.
-The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.
-The left of figure contains tests using not very big input, and the right includes of the others.
-We see that $put_m$ is the fastest method if the input size is not too large.
-However, if the input size is large enough, $put_m$ will be slower quickly due to times for manipulating data in the table.
+%The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.
+The left part of the figure contains tests using not-large inputs, and we see that $put_m$ is the fastest method for not-large inputs.
+However if the input size is large enough, in the cases of the right part of the figure, $put_m$ will be slower quickly due to times for manipulating data in the table.
 At that time, $xpg$ proved to be the most effective method.

### 2019-12-02 paper/6.related-work.tex
small fixes

@@
 
-Since the pioneer work of Lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
+Since the pioneer work of lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)

### 2019-12-02 paper/7.conclusion.tex
small fixes

@@
 From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not large and $xpg$ is fastest for other left associative programs.
-This shows that if programmer choose one method based on the their BX programs and inputs, they can use efficient evaluation.
+This shows that if programmers choose one method based on the their BX programs and inputs, they can use an efficient evaluation method.
 
@@
 In $xpg$, there are two limitations: only for very-well-behaved programs and restriction for $\pginline{Case}$ in Section~\ref{sect:pg}.
-To be a practical evlauation method, extention of the target language is needed.
+To be a practical evlauation method, an extention of the target language is needed.
 After extention of the language, we can evaluate more various programs for experimental results.
@@
 Third is introducing type systems to datatype of sources and view.
-If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-We need to investigate more about $construct\_dummy$, because current definition that is used in Section~\ref{sect:experiments} will cause type errors.
+If we introduce this, we can avoid runtime errors like statically-typed functional languages.For this, we need to investigate more about $construct\_dummy$, because the current definition that is used in our experiments %Section~\ref{sect:experiments}
+will cause type errors.
 

### 2019-12-02 paper/1.introduction.tex
reduce some vspaces and add some vspaces

@@
 Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_1 \circ bx_2$ is defined by
+
+\minusvspace
 \begin{align}
@@
 \end{align}
-One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
+\minusvspace
+
+\noindent One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
 Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times.
 To be concrete, consider the evaluation of the following composition (which will be used as our running example in this paper):
+
+\minusvspacetwo
 \[
@@
 \]
-which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source $s$ being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
+\minusvspacetwo
+
+\noindent which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source $s$ being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
 To obtain the final updated source $s'$, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
@@
 One direct solution to avoid this repeated computation of $get$ is to compute the composition in a right associative manner. For instance, if we transform  $lp3$ to $rp3$:
+
+\minusvspacetwo
+\vspace{-4mm}
 \[
@@
 \]
-then the $put$ for $rp3$ only needs to compute $get$ of $phead$ twice, one time less than that for $lp3$.
+\minusvspacetwo
+
+\noindent then the $put$ for $rp3$ only needs to compute $get$ of $phead$ twice, one time less than that for $lp3$.
 However, this transformation is not always easy to do.
@@
 Informally, $\mbox{\it bfoldr}$ is a recursive bidirectional program defined in a way like
+
+\minusvspacetwo
  \[
@@
  \]
-where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
+ \minusvspacetwo
+
+\noindent where the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
 This makes it hard to do the above transformation statically.

### 2019-12-02 paper/2.minbigul.tex
reduce some vspaces and add some vspaces

@@
 
+ \smallvspace
     $\qtab \getbxinline{bx} : s \to v, \quad \putbxinline{bx} : s * v \to v$
+ \smallvspace
 
@@
 
+\smallvspace
     $\qtab \putbx{bx}{s}{(\getbx{bx}{s})} = s \qquad [\textsc{GetPut}]$
@@
     $\qtab \getbx{bx}{(\putbx{bx}{s}{v})} = v \qquad [\textsc{PutGet}]$
+\smallvspace
 
@@
 
+\smallvspace
 $\qtab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$
+\smallvspace
 
@@
 
+\smallvspace
     $bx := Skip \ h \
@@
         | \ Compose \ bx_1 \ bx_2$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else assert } false \}\\
-    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else assert } false \}$
+\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else } undefined \}\\
+    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else } undefined \}$
+\smallvspace
 
 \noindent As an example of minBiGUL program, consider the definition of $phead$ :\\
+\smallvspace
 $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
@@
         \qtabs{2} bx_v = \product{Replace}{(Skip \ (\lambda \_.())}$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$
@@
 $\tab \getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab bfoldr \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
@@
 $bx_s = ((Replace \times bfoldr \ bx) \circ bx)$
+\smallvspace
 

### 2019-12-02 paper/1.introduction.tex
add/reduce spaces

@@
 \end{figure}
+\vspace*{-5.5mm}
 
@@
 
-One direct solution to avoid this repeated computation of $get$ is to compute the composition in a right associative manner. For instance, if we transform  $lp3$ to $rp3$:
+One direct solution to avoid this repeated $get$ computation is to compute compositions in a right associative manner. For instance, if we transform $lp3$ to $rp3$:\break
 
 \minusvspacetwo
-\vspace{-4mm}
+\vspace{-3mm}
 \[
@@
 \minusvspacetwo
+\vspace{-0mm}
 
@@
 
+\vspace{1mm}
 \minusvspacetwo

### 2019-12-02 paper/2.minbigul.tex
small fixes

@@
 minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which satisfy two above laws.
-In addition, we restrict 'adaptive case's of BiGUL on minBiGUL.
+In addition, we restrict `adaptive case's of BiGUL on minBiGUL.
 Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{Foster:2007:CBT:1232420.1232424}, like the following:
@@
 It is a kind of bijection.
-However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need give two functions which one is the inverse of the other.
+However, to be able to implement it in OCaml, the environment used for developing minBiGUL and solutions in the paper, we need to require two functions which one is the inverse of the other.
 In the above syntax, $f_2 = f_1^{-1}$ and $g_2 = g_1^{-1}$.
@@
 
-In the end of this section, let's see the minBiGUL definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
+In the end of this section, let us see the minBiGUL definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
 

### 2019-12-02 paper/4.xpg.tex
add/reduce spaces
small fixes

@@
 
-\noindent Now, let's see how we construct $pg$ recursively.
+\noindent Now, let us see how we construct $pg$ recursively.
 
@@
 The last equality changes two.
-First one is a change from $dummy$ to an application $construct\_dummy \ s$.
+First is a change from $dummy$ to an application $construct\_dummy \ s$.
 This shows that we need a way to construct $dummy$ from the current source, $s$.
@@
 
-In the places where thrid and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
-
+In the places where third and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
 %Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-
 Thanks to this accumulation, in $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
 The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Since there is no real view here, we need to construct a dummy from the source.
+Since there is no real view here, we need to construct a dummy from the source same as $pg$.
 Then the continuation updating on this dummy should be initiated as an identity function.
@@
 
+\smallvspace
     $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
         \tab \qtab (s; v)$
+\smallvspace
 
@@
 For keeping away such computations, we introduce an extension named $kpg$.
-While $cpg$ evaluate values eagerly, $kpg$ does the opposite.
+While $cpg$ evaluates values eagerly, $kpg$ does the opposite.
 Every values are evaluated lazily in a computation of $kpg$.
@@
 
+\smallvspace
     $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
         \tab \qtab (ks' \ s; kv' \ v)$
+\smallvspace
 
@@
 $\kpginline{Skip}$ and $\kpginline{Case}$ are special cases, because they require the actual values for evaluation.
-Therefore we evaluate the values by application of $ks' s$ and $kv' v$ and update $ks'$ and $ks'$ with the identity fuctions.
+Therefore we evaluate the values by application of $ks' s$ and $kv' v$ and update $ks'$ and $kv'$ with the identity functions.
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
@@
 
-In $kpg$, we did two optimizations.
-First is in $\kpginline{bx_1 \times bx_2}$, the evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+Additionally we did two optimizations in $kpg$.
+First is in $\kpginline{bx_1 \times bx_2}$.
+The evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
 There may be same computations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
@@
         \tabs{3} (ks_1 \ s_2, v_2)\\
-    \tab \vert \ \_ \to \text{similar to } put$
+    \tab \vert \ \_ \to \text{similar to } pg$
 \end{definition}

### 2019-12-02 paper/5.experiments.tex
small fixes

@@
 The OCaml runtime system options and garbage collection parameters are set as default.
-In our language, there is no type system.}$^,$%HACK!
+In our target language, there is no type system.}$^,$%HACK!
 %
@@
 % The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from the same $n + 1$ \textit{pure} programs. 
-The first five test cases
+The first 5 test cases
 % (no.~1--5) 
@@
 $\getbx{ptail}{s}$ returns the tail of the source list $s$.
-The compositions of many $ptail$, in the \textit{put} direction, replace a part of the tail of the source list by the view list, in the \textit{get} direction, returns the such tail from the the source.
+The compositions of many $ptail$, in the \textit{put} direction, replace a part of the tail of the source list by the view list, in the \textit{get} direction, returns the such tail from the source.
 
@@
 \textit{breverse} is defined in terms of \textit{bfoldr}, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
-Note that compositions is by the recursions of $breverse$ and
+Note that compositions are by the recursions of $breverse$ and
 the number of compositions are dynamically determined by the length of the source list.
@@
 
-Figure \ref{fig:result} illustrates the evaluation times of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
+Figure \ref{fig:result} illustrates the evaluation time of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
 We simply ignore the experimental results of $pg$, $cpg$ and $kpg$ since they are shown the slowness compared to $xpg$.
@@
 %The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.
-The left part of the figure contains tests using not-large inputs, and we see that $put_m$ is the fastest method for not-large inputs.
-However if the input size is large enough, in the cases of the right part of the figure, $put_m$ will be slower quickly due to times for manipulating data in the table.
-At that time, $xpg$ proved to be the most effective method.
+The left part of the figure contains tests using not-large inputs, and we see that $put_m$ is the fastest method for them.
+However if the input size is large enough, in the cases of the right part of the figure, $put_m$ will be slower quickly due to time for manipulating data in the table.
+At that time, $xpg$ is proved to be the most effective method.

### 2019-12-02 paper/7.conclusion.tex
small fixes

--- 2019-11-29 paper/7.conclusion.tex
@@
 From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not large and $xpg$ is fastest for other left associative programs.
-This shows that if programmer choose one method based on the their BX programs and inputs, they can use efficient evaluation.
+This shows that if programmers choose one method based on their BX programs and inputs, they can use an efficient evaluation method.
 
@@
 We will continue our work on the following three points.
-First is overcomming our limitation in $xpg$.
+First is overcoming our limitation in $xpg$.
 In $xpg$, there are two limitations: only for very-well-behaved programs and restriction for $\pginline{Case}$ in Section~\ref{sect:pg}.
-To be a practical evlauation method, extention of the target language is needed.
-After extention of the language, we can evaluate more various programs for experimental results.
-
+To be a practical evaluation method, an extension of the target language is needed.
+After extension of the language, we can evaluate more various programs for experimental results.
 Second is introducing an automatic analysis about BX programs and inputs to choose the best evaluation method.
-Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If this is achieved, we can reduce programmer's burden.
-
-Third is introducing type systems to datatype of sources and view.
+Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If this analysis is achieved, we can reduce programmers' burden.
+Third is introducing a type system to our target language, especially the datatype of sources and view.
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-We need to investigate more about $construct\_dummy$, because current definition that is used in Section~\ref{sect:experiments} will cause type errors.
+For this, we need to investigate more about $construct\_dummy$, because the current definition that is used in our experiments %Section~\ref{sect:experiments}
+will cause type errors.
 

### 2019-12-02 paper/4.xpg.tex
change fail to undefined

@@
 
-    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{fail}}, h \ s) \\
-        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{fail}} \\
-        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{fail}}$
+    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{undefined}}, h \ s) \\
+        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{undefined}} \\
+        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{undefined}}$
 
@@
 The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-This is a trick since in some cases, the result of $pg$ may be fail although there is no fail when evaluating $\getbxinline{Skip \ h}$.
+This is a trick since in some cases, the result of $pg$ may be undefined although there is no undefined when evaluating $\getbxinline{Skip \ h}$.
 The last equality is relatively oblivious.
@@
 
-    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else fail}$
+    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else undefined}$
 
@@
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else fail}$
+        \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else undefined}$
 

### 2019-12-02 paper/5.experiments.tex
Polished text

@@
 
-We have implemented and tested all methods% 
+We have fully implemented and tested all methods% 
 %
@@
 The OCaml runtime system options and garbage collection parameters are set as default.
-In our target language, there is no type system.}$^,$%HACK!
+Our target language is untyped.}$^,$%HACK!
 %
@@
 
-We have conducted 7 test cases shown in Table~\ref{tab:test-cases}. Although the number of cases is small, 
-% they are not trivial and can be completely used in practice. 
-they represent non-trivial cases of practical significance.
-In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$.
-This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table.
-These $bx$s are left associative compositions since we are only interested in this kind of program in the paper.
-Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.
+We have 
+% conducted 
+selected seven test cases (Table~\ref{tab:test-cases}) to 
+represent
+% assess 
+non-trivial cases of practical significance.
+The test cases use
+% These $bx$s are 
+left associative compositions because we focus on this kind of inefficiency in this paper.
+% Although the number of cases is small, 
+%% they are not trivial and can be completely used in practice. 
+% they represent non-trivial cases of practical significance.
+In the last two columns, $s_r$ and $v_r$ are the updated source and view, respectively.
+They are produced by applying $put$ and $get$ to the original source $s_0$ and view $v_0$.
+That is, $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is the program indicated in the second column of the table.
+% are only interested in this kind of programs in the paper. 
+% Please note that the results
+Results $s_r$ and $v_r$ are independent of the associativity 
+% style 
+of the composition.
 
@@
     \centering
-    \caption{Test cases}
+    \caption{Composition test cases}
     \label{tab:test-cases}
@@
 % The first 5 cases merely use $n$ $\circ$ operators to make straight-line compositions from the same $n + 1$ \textit{pure} programs. 
-The first 5 test cases
-% (no.~1--5) 
-are straight-line compositions of the same $n + 1$ non-recursive programs.
+The first five test cases
+(1--5) 
+are $n$ straight-line (non-recursive) compositions of the same $n + 1$ 
+% non-recursive 
+programs.
 % We refer to them as `pure' compositions.
 % We use the term \textit{pure} here to refer non-recursion compositions. 
-The prefix \textit{lcomp} in a test name indicates that the compositions are left-associative.
+The prefix \textit{lcomp} in the name of a test case indicates that the textual compositions are left associative.
 % that the associate precedence of $\circ$ in a straight-line composition is left.
-The suffix \textit{ldata} indicates that the input size is large.
-The symbol $L$ in the input column of the \textit{ldata} test cases can be expressed as a complex list as follows: $L = [T,\ldots,T] \ (n \ T\text{s}), \ T = [A,\ldots,A] \ (10 \ A\text{s}), \ A = [1,\ldots,5]$.
-These structures are only intended to generate test data that is large enough for
+The suffix \textit{ldata} indicates that the input size is considered large. The symbol $L$ in the input column 
+% \textit{ldata} test cases 
+% can be expressed as a complex list as follows: 
+stands for a list $L = [T,\ldots,T]$ with
+% $(n \ T\text{s}), 
+$T = [A,\ldots,A]$ of length 10
+% \ (10 \ A\text{s}),
+and
+$A = [1,\ldots,5]$. 
+% These structures are 
+They are only intended to generate test data that is large enough for 
 % observing 
@@
 
-We introduced the composition $phead \circ phead$ earlier.
-The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view.
-Because of the type of the source, this program is seen as a \textit{ldata} case.
-Next, we briefly explain the behavior of the remaining compositions in Table~\ref{tab:test-cases}.
+We introduced the composition $phead \circ phead$ earlier in Section~\ref{sec:intro}. The composition of many \textit{phead} 
+% do things
+works similarly. The head of a head element inside a 
+%super 
+deeply nested list, which 
+% expresses 
+is the source, 
+% should be 
+is updated by the changed view. Because of the type of the source, this program is 
+% seen 
+categorized as a \textit{ldata} case.
+
+Next, we briefly explain the behavior of the remaining compositions in 
+\mbox{Table\,\ref{tab:test-cases}.}%HACK!
 
@@
 $\getbx{ptail}{s}$ returns the tail of the source list $s$.
-The compositions of many $ptail$, in the \textit{put} direction, replace a part of the tail of the source list by the view list, in the \textit{get} direction, returns the such tail from the source.
+The composition of many $ptail$, in the \textit{put} direction, replaces a part of the tail of the source list by the view list and, in the \textit{get} direction, returns such a tail from the source.
 
-$\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$ as two same length lists, to produce a new list by moving the last element of $v$ to its first position.
+$\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$, 
+%as two same length lists, 
+which are two lists of the same length,
+and produces a new list by moving the last element of $v$ to its first position.
 $\getbx{ptail}{s}$ returns another list by moving the first element of the list $s$ to its end position.
-The compositions of many \textit{ptail}, in the \textit{put} direction, create a permutation of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a permutation of the source list.
+The composition of many \textit{ptail}, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
 
-\textit{breverse} is defined in terms of \textit{bfoldr}, in the \textit{put} direction, produces a reverse of the view list if the source list has a same length with the view, in the \textit{get} direction, produce a reverse of the source list.
+\textit{breverse} is defined in terms of \textit{bfoldr}.
+In the \textit{put} direction, it produces a reverse of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a reverse of the source list.
 Note that compositions are by the recursions of $breverse$ and
@@
 
-Figure \ref{fig:result} illustrates the evaluation time of $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$ for 7 test cases.
-We simply ignore the experimental results of $pg$, $cpg$ and $kpg$ since they are shown the slowness compared to $xpg$.
+Figure \ref{fig:result} 
+% illustrates 
+shows the evaluation times for each of the seven test cases using the three methods: $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$.
+% We simply ignore 
+For simplicity, we omit the experimental results of $pg$, $cpg$ and $kpg$ because they are shown the slowness compared to $xpg$.
 For left associative compositions, $put$ works poorly because of the number of reevaluated $get$s.
@@
 However if the input size is large enough, in the cases of the right part of the figure, $put_m$ will be slower quickly due to time for manipulating data in the table.
-At that time, $xpg$ is proved to be the most effective method.
+At that time, $xpg$ is
+% proved to be
+the most effective method.

### 2019-12-02 paper/3.minbigul_m.tex
add labels to sections
small fixes
Polished text

@@
-\section{Adding Memoization: minBiGUL$_m$}
+\section{Adding Memoization: minBiGUL$_m$} \label{sect:minbigulm}
+
+When evaluating the composition of several BX programs, 
+% (and the same holds for 
+% full BiGUL), 
+% all BX languages), 
+the same $get$s are evaluated repeatedly.
+% because no intermediate states are kept during the evaluation. 
+This problem was illustrated in Figure~\ref{fig:eval-comp-phead}.
+To avoid reevaluating $get$s, and as our first approach to avoid this inefficiency, we introduce memoization in the minBiGUL interpreter.
+% can save 
+To keep it simple, the intermediate state of a composition is saved in a key-value table where the key is a pair of program $bx$ and source $s$, and the value is the result of evaluating $\getbx{bx}{s}$.
+% the first time. 
+% Then, instead of recomputing the value, the value in the table is used.
+Later the value in the table is used instead of recomputing it.
+
+The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}).
+The auxiliary function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).
 
@@
 
-When evaluating a composition of several programs in minBiGUL (and the same holds for full BiGUL), the same $get$s have to be evaluated repeatedly because no intermediate states are kept during the evaluation.
-This problem was illustrated in Figure~\ref{fig:eval-comp-phead}.
-To avoid reevaluating $get$s, and as our first approach, we introduce memoization in the minBiGUL interpreter.
-% can save 
-To keep it simple, the intermediate states of a composition are saved in a key-value table where the key is a pair of program $bx$ and source $s$, and the value is the result of evaluating $\getbx{bx}{s}$ the first time.
-Then, instead of recomputing the value, the value in the table is used.
-The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}).
-Another function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).
-
-Note as a special point that not all states are saved when evaluating a program, only when a composition is evaluated.
-When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, an application of $get_{mh}$, not $get$ or $get_m$, appears to indicate the saving.
+% Note as a special point 
+% It should be noted 
+Note in particular 
+that the interpreter does not save all states when evaluating a program, only the intermediate states of a composition.
+When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, function $get_{mh}$ is used, not $get$ or $get_m$.
+% appears to indicate the saving. 
 When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation.
 If there is no such key, the value $v$ will be calculated by using $get_{mh}$.
-The further key-value insertion is needed because the later computation could require to reevaluate $\getm{bx_1 \circ bx_2}{s}$ one more time.
+This key-value insertion is necessary because the interpreter may later need to reevaluate $\getm{bx_1 \circ bx_2}{s}$.
+% reevaluate <--> evaluate one more time.

### 2019-12-02 paper/2.minbigul.tex
change fail to undefined
Polished text

@@
 
-minBiGUL, our target language in this paper, is a subset of BiGUL which is a simple yet powerful putback-based bidirectional language.
+The target language in this paper, minBiGUL, is a very-well behaved subset of BiGUL, which is a simple, yet powerful putback-based bidirectional language.
+
 BiGUL supports two transformations: a forward transformation $get$ producing a view from a source and a backward transformation $put$ taking a source and a modified view to produce an updated source.
-Intuitively, if we have a BiGUL program $bx$, these two transformations are following functions:
+Intuitively, if we have a BiGUL program $bx$, these two transformations are the following functions:
 
@@
 
-The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source.
+\noindent The \textsc{GetPut} law means that if there is no change to the view, there should be no change to the source.
 The \textsc{PutGet} law means that we can recover the modified view by applying the forward transformation to the updated source.
 
-minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which satisfy two above laws.
-In addition, we restrict `adaptive case's of BiGUL on minBiGUL.
-Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{Foster:2007:CBT:1232420.1232424}, like the following:
+minBiGUL inherits from BiGUL both 
+% supports 
+transformations, $put$ and $get$, which satisfy the two laws above. 
+% In addition, 
+Because we restrict the `adaptive case' of BiGUL in minBiGUL,
+% In this case, 
+$put$ and $get$ satisfy one more law, namely the \textsc{PutPut} law~\cite{Foster:2007:CBT:1232420.1232424}:
+% like the following:
 
@@
 
-The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates.
-Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved \cite{Foster:2007:CBT:1232420.1232424}.
+\noindent The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates. 
+Because minBiGUL satisfies all three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, it is very well-behaved~\cite{Foster:2007:CBT:1232420.1232424}.
+% Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved \cite{Foster:2007:CBT:1232420.1232424}.
 
@@
 \smallvspace
-    $bx := Skip \ h \
+    $bx ::= Skip \ h \
         | \ Replace \
@@
 
-A minBiGUL program may be either a skip of a function or a replacement or a product of two minBiGUL programs or a source/view rearrangement or a case combinator without adaptive cases or a composition of some minBiGUL programs.
-We use numbers, pairs and lists to construct the program inputs including the source and/or the view.
+A minBiGUL program is either a skip of a function, a replacement, a product of two 
+%minBiGUL 
+programs, a source/view rearrangement, a case combinator (without adaptive cases), or a composition of 
+% some minBiGUL 
+two programs. We use numbers, pairs and lists to construct the program inputs including the source and/or the view.
 
-For source/view rearrangement, BiGUL uses just one lambda expression to express how to deconstruct as well as reconstruct data.
+For source/view rearrangement, BiGUL uses 
+% just one 
+a lambda expression to express how to deconstruct as well as reconstruct data.
 It is a kind of bijection.
@@
 
+The semantics of $put$ and $get$ is shown in Definitions~\ref{def:minbigulput} and \ref{def:minbigulget}, respectively.
+Instead of using the name $v'$ for the updated view in the $put$ direction, like Figures~\ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we
+%just
+simply 
+%use notation 
+use~$v$ below. The later definitions also follow this convention.
+
 \begin{multicols}{2}
@@
         \noindent $\putbx{Skip \ h}{s}{v} = \\
-            \tab \sif{h \ s = v}{s}{\text{fail}}$
+            \tab \sif{h \ s = v}{s}{\text{undefined}}$
 
@@
 
-We define the semantics of $put$ and $get$ as in Definitions~\ref{def:minbigulput} and \ref{def:minbigulget} respectively.
-Instead of using $v'$ in the $put$ direction like Figures~\ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view.
-The later definitions also follow this convention.
-
-In the above definitions, we use if-then-else-fi statements to express semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$ where $\Leftarrow$ means assignments.
+The two definitions use if-then-else-fi statements 
+%express 
+to define the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, where $\Leftarrow$ means assignments.
 This statement is useful to describe many functions related to $Case$ in this paper.
@@
 
-In the end of this section, let us see the minBiGUL definition of $bfoldr$ which is a putback function of an important higher-order function on lists, $foldr$:
+To close this section,
+% In the end of this section, 
+let us see the minBiGUL definition of $\textit{bfoldr}$ which is a putback function of an important higher-order function on lists, $\textit{foldr}$:
 
 \smallvspace
-$\tab bfoldr \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+$\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
 $\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
@@
 
-\noindent In $bfoldr$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
-Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.
+\noindent In $\textit{bfoldr}$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
+Because~$\circ$ has a higher priority than~$\times$, it is seemingly impossible to transform $\mathit{bfoldr}$ from the left associative composition style to the right one.
 %will look like a composition of many programs at some points in the evaluation if we slow down this process.

### 2019-12-03 paper/5.experiments.tex
add reason for ignoring results of pg, cpg, kpg: sec 5.2

@@
 shows the evaluation times for each of the seven test cases using the three methods: $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$.
-% We simply ignore 
-For simplicity, we omit the experimental results of $pg$, $cpg$ and $kpg$ because they are shown the slowness compared to $xpg$.
-For left associative compositions, $put$ works poorly because of the number of reevaluated $get$s.
+We also tried doing similar experiments with $pg$, $cpg$ and $kpg$, but their results are slower than the corresponding ones of $xpg$.
+Then we simply ignore these tedious results.
+As we all know, $put$ works poorly for left associative compositions because of the number of reevaluated $get$s.
 %The figure shows $put_m$ and $xpg$ are better than $put$ in some cases.

### 2019-12-03 paper/4.xpg.tex
undefined -> \it{undefined} in Section 4, this follows Section 2
update sec.4: tiny

@@
 
-    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\text{undefined}}, h \ s) \\
-        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\text{undefined}} \\
-        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\text{undefined}}$
+    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{undefined}, h \ s) \\
+        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{undefined} \\
+        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{undefined}$
 
@@
 
-Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.
+Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
 
@@
 First is a change from $dummy$ to an application $construct\_dummy \ s$.
-This shows that we need a way to construct $dummy$ from the current source, $s$.
+This means we need a way to construct $dummy$ from the current source, $s$.
 We can require programmers to give this $construct\_dummy$ function.
-Second is a substitution on $s$ in $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
-The replacement of $s$ and $s_1$ is possible because of the \textsc{PutPut} law.
-Then, $v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we can realize that there is no loss information when computing a composition.
+Second is a substitution under the \textsc{PutPut} law on $s$ of $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
+$v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
+So we realize that there is no loss information when computing a composition.
 
@@
 When evaluating $\pginline{bx_1 \circ bx_2}$, there are three $pg$ calls, of which twice for $\pginline{bx_1}$ and once for $\pginline{bx_2}$.
-If a given program is a left-associative composition, the number of $pg$ calls will be exponential.
+If a given program is a left associative composition, the number of $pg$ calls will be exponential.
 Therefore, the runtime inefficiency is inevitable.
-To solve that, we introduce a new function, $cpg$, accumulates updates in the source and the view.
+To solve that, we introduce a new function, $cpg$, accumulates updates on the source and the view.
 $\cpg{bx}{ks}{kv}{s}{v}$ is an extension of $\pg{bx}{s}{v}$ where $ks$ and $kv$ are continuations used to hold the modification information, and $s$ and $v$ are used to keep evaluated values same as $pg$.
@@
 
-    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else undefined}$
+    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else } undefined$
 
@@
 %Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
-Thanks to this accumulation, in $\cpginline{bx_1 \circ bx_2}$, there are only two $cpg$ calls.
+Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$.
 The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
@@
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else undefined}$
+        \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else } undefined$
 
@@
 $\kpginline{Skip}$ and $\kpginline{Case}$ are special cases, because they require the actual values for evaluation.
-Therefore we evaluate the values by application of $ks' s$ and $kv' v$ and update $ks'$ and $kv'$ with the identity functions.
+Therefore we evaluate the values by application of $ks' \ s$ and $kv' \ v$ and update $ks'$ and $kv'$ with the identity functions.
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.

### 2019-12-03 paper/2.minbigul.tex
tiny changes in formulas: keywords --> normal text, redundant vars --> underline

@@
         \noindent $\putbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} =\\
-            \tab \text{if} \ {cond_{sv} \ s \ v} \\
-            \tab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
-            \tab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
-            \tab \text{fi} \ cond_{s} \ s'; \ \text{return} \ s'$
+            \tab \textnormal{if} \ {cond_{sv} \ s \ v} \\
+            \tab \textnormal{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \\
+            \tab \textnormal{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \\
+            \tab \textnormal{fi} \ cond_{s} \ s'; \ \textnormal{return} \ s'$
 
@@
         \noindent $\getbx{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s} = \\
-            \tab \text{if} \ {cond_{s} \ s} \\
-            \tab \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\ 
-            \tab \text{else} \ v' \Leftarrow \getbx{bx_2}{s} \\ 
-            \tab \text{fi} \ {cond_{sv} \ s \ v'}; \ \text{return} \ v'$
+            \tab \textnormal{if} \ {cond_{s} \ s} \\
+            \tab \textnormal{then} \ v' \Leftarrow \getbx{bx_1}{s} \\ 
+            \tab \textnormal{else} \ v' \Leftarrow \getbx{bx_2}{s} \\ 
+            \tab \textnormal{fi} \ {cond_{sv} \ s \ v'}; \ \textnormal{return} \ v'$
 
@@
 $\tab \text{if } E_1 \text{ then } X_1 \text{ else } X_2 \text{ fi } E_2; S\\
-\tabs{2} \ \equiv \text{if }E_1 = true \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else } undefined \}\\
-    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = false \text{ then } S \text{ else } undefined \}$
+\tabs{2} \ \equiv \text{if }E_1 = \textit{true} \text{ then } \{ X_1; \text{ if } E_2 = true \text{ then } S \text{ else } \textit{undefined}\}\\
+    \tabs{4} \text{else } \{ X_2; \text{ if } E_2 = \textit{false} \text{ then } S \text{ else } \textit{undefined} \}$
 \smallvspace

### 2019-12-03 paper/4.xpg.tex
tiny changes in formulas: keywords --> normal text, redundant vars --> underline

@@
 
-    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{undefined}, h \ s) \\
-        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{undefined} \\
-        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{undefined}$
+    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\textit{undefined}}, h \ s) \\
+        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\textit{undefined}} \\
+        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\textit{undefined}}$
 
@@
     \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
-        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
-        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
-            \qtab \tabs{2} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
+        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
+        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
+            \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
     \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
-        \tab \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \tab \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
-            \qtab \tabs{2} (s_3, v_2)$
+        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
+        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
+            \qtab \tabs{3} (s_3, v_2)$
 
@@
 
-    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \text{ if } h \ s = v \text{ then } (ks, kv, s, v) \text{ else } undefined$
+    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \textnormal{ if } h \ s = v \text{ then } (ks, kv, s, v) \textnormal{ else } \textit{undefined}$
 
@@
     \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-        \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-        \tab \text{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-        \tab \text{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-        \tab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \text{return} \ (ks, kv, s', v')$
+        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
+        \tab \textnormal{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
+        \tab \textnormal{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
+        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks, kv, s', v')$
 
     \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-        \tab (ks_1, kv_1, s_1, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
+        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
         \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
@@
 
-        Note that, in $\cpginline{bx_1 \circ bx_2}$, $s_1$ is redundant because this evaluated variable is not used in the later steps.
+Note that, in $\cpginline{bx_1 \circ bx_2}$, $\underline{s_1}$ is redundant because this evaluated variable is not used in the later steps.
 In the next session, we will optimize this redundancy.
@@
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab \text{if} \ h \ s = v \ \text{then} \ (ks, kv, ks', kv', s, v) \ \text{else } undefined$
+        \tab \textnormal{if} \ h \ s = v \ \textnormal{then} \ (ks, kv, ks', kv', s, v) \ \textnormal{else } \text{undefined}$
 
@@
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id ; \quad kv' \Leftarrow id;\\
-        \tab \text{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-        \tab \text{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-        \tab \text{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-        \tab \text{fi} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' \ v'); \ \text{return} \ (ks, kv, ks', kv', s', v')$
+        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
+        \tab \textnormal{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
+        \tab \textnormal{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
+        \tab \underline{\textnormal{fi}} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' \ v'); \ \textnormal{return} \ (ks, kv, ks', kv', s', v')$
 
     \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
+        \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
         \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
@@
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $ks_1'$ and $s_1$ in the result of the first assignment are still redundant but they are not evaluated.
+When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{ks_1'}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but they are not evaluated.
 
@@
 Second is in $\kpginline{Case}$.
-We need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the $fi$ condition before returning the 6-tuple.
+We need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple.
 Such evaluations can be done lazily to make programs run faster.
@@
 
-    \noindent $\xpg{bx}{s}{v} = \text{match } bx \text{ with }\\
+    \noindent $\xpg{bx}{s}{v} = \textnormal{match } bx \textnormal{ with }\\
     \tab \vert \ bx_1 \circ bx_2 \to \\

### 2019-12-03 paper/2.minbigul.tex
revise reason for slowness

--- 2019-12-02 paper/2.minbigul.tex
@@
 % In the end of this section, 
-let us see the minBiGUL definition of $\textit{bfoldr}$ which is a putback function of an important higher-order function on lists, $\textit{foldr}$:
+let us see the minBiGUL definition of $\mathit{bfoldr}$ which is a putback function of an important higher-order function on lists, $\mathit{foldr}$:
 
 \smallvspace
-$\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+$\tab \mathit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
 $\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
@@
 
-\noindent In $\textit{bfoldr}$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
-Because~$\circ$ has a higher priority than~$\times$, it is seemingly impossible to transform $\mathit{bfoldr}$ from the left associative composition style to the right one.
+\noindent In $\mathit{bfoldr}$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
+Because~$\circ$ has a higher priority than~$\times$, it is in general not possible to transform $\mathit{bfoldr}$ from the left associative composition style to the right one.
 %will look like a composition of many programs at some points in the evaluation if we slow down this process.

### 2019-12-03 paper/5.experiments.tex
revise reason for slowness

@@
 shows the evaluation times for each of the seven test cases using the three methods: $put$ in minBiGUL, $put_m$ in minBiGUL$_m$ and $xpg$.
-We also tried doing similar experiments with $pg$, $cpg$ and $kpg$, but their results are slower than the corresponding ones of $xpg$.
-Then we simply ignore these tedious results.
+We also did similar experiments with $pg$, $cpg$ and $kpg$, but their results are slower than the corresponding ones of $xpg$.
+The slowness was caused by the exponential number of $pg$ calls in the case of $pg$, redundant evaluations ($cpg$), and redundant overhead for constructing closures ($kpg$).
+Then we simply omit these results in Figure \ref{fig:result}.
 As we all know, $put$ works poorly for left associative compositions because of the number of reevaluated $get$s.

### 2019-12-03 paper/2.minbigul.tex
fix Sec.2: if-then-else-fi

@@
 %express 
-to define the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, where $\Leftarrow$ means assignments.
+to define the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, where $\Leftarrow$ denotes an assignment.
 This statement is useful to describe many functions related to $Case$ in this paper.
-Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means ``if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, if $E_2$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.'' If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
+Statement (if $E_1$ then $X_1$ else $X_2$ fi $E_2$) means ``if the test $E_1$ is true, the statement $X_1$ is executed and the assertion $E_2$ must be true, otherwise, if $E_1$ is false, the statement $X_2$ is executed and the assertion $E_2$ must be false.'' If the values of $E_1$ and $E_2$ are distinct, the if-then-else-fi structure is undefined.
 We can write the equivalent if-then-else statement as follows:

### 2019-12-03 paper/4.xpg.tex
tiny

@@
 This is a trick since in some cases, the result of $pg$ may be undefined although there is no undefined when evaluating $\getbxinline{Skip \ h}$.
-The last equality is relatively oblivious.
+The last equality is relatively obvious.
 
@@
 
-The problem for $cpg$ lies in redundant computation during evaluating.
+The problem for $cpg$ lies in redundant computations during evaluating.
 For keeping away such computations, we introduce an extension named $kpg$.
@@
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{ks_1'}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but they are not evaluated.
+When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{ks_1'}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but not evaluated.
 

### 2019-12-03 paper/5.experiments.tex
edit caption table 1
fix ptail/bsnoc

@@
     \centering
-    \caption{Composition test cases}
+    \caption{Composition test cases (number of compositions = $n$)}
     \label{tab:test-cases}
@@
 and produces a new list by moving the last element of $v$ to its first position.
-$\getbx{ptail}{s}$ returns another list by moving the first element of the list $s$ to its end position.
-The composition of many \textit{ptail}, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
+$\getbx{bsnoc}{s}$ returns another list by moving the first element of the list $s$ to its end position.
+The composition of many \textit{bsnoc}, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
 

### 2019-12-03 paper/5.experiments.tex
tiny

@@
 
-We introduced the composition $phead \circ phead$ earlier in Section~\ref{sec:intro}. The composition of many \textit{phead} 
+We introduced the composition $phead \circ phead$ earlier in Section~\ref{sec:intro}. The composition of many \textit{phead}s 
 % do things
@@
 $\getbx{ptail}{s}$ returns the tail of the source list $s$.
-The composition of many $ptail$, in the \textit{put} direction, replaces a part of the tail of the source list by the view list and, in the \textit{get} direction, returns such a tail from the source.
+The composition of many $ptail$s, in the \textit{put} direction, replaces a part of the tail of the source list by the view list and, in the \textit{get} direction, returns such a tail from the source.
 
@@
 $\getbx{bsnoc}{s}$ returns another list by moving the first element of the list $s$ to its end position.
-The composition of many \textit{bsnoc}, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
+The composition of many \textit{bsnoc}s, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
 

### 2019-12-03 paper/7.conclusion.tex

@@
 To achieve fast evaluation, we introduce two different methods using memoization.
-From the experimental results of left associative BX composition programs, we know that $put_m$ is fastest method if input size is not large and $xpg$ is fastest for other left associative programs.
+From the experimental results of left associative BX composition programs, we know that $xpg$ is fastest method if input size is large and
+$put_m$ is fastest for other left associative programs.
 This shows that if programmers choose one method based on their BX programs and inputs, they can use an efficient evaluation method.
@@
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-For this, we need to investigate more about $construct\_dummy$, because the current definition that is used in our experiments %Section~\ref{sect:experiments}
+For this, we need to investigate more about $construct\_dummy$, because the current definition used in our experiments %Section~\ref{sect:experiments}
 will cause type errors.

### 2019-12-03 paper/4.xpg.tex
small change

@@
 
-\noindent Now, let us see how we construct $pg$ recursively.
+\noindent Now, let us see how we construct $pg$ respectively.
 
@@
 The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-This is a trick since in some cases, the result of $pg$ may be undefined although there is no undefined when evaluating $\getbxinline{Skip \ h}$.
-The last equality is relatively obvious.
+This is a trick since in some cases, the result of $pg$ may be undefined although the result is not
+\end
+undefined when evaluating $\getbxinline{Skip \ h}$. The last equality is relatively obvious.
 

### 2020-02-17 paper/1.introduction.tex
Added section label
revise caption of fig.3
add sentences based on Robert's comment
small changes
reduce spaces: only 3 pages for sec.1
small updates

@@
-\section{Introduction}
+\section{Introduction} \label{sec:intro}
 
@@
 Let $[1,2]$ be the original source~$s$.
-$get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
+The function $get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
 %\texttt{get pHead [1,2] = 1}.
@@
 
-\noindent which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source $s$ being ${[[[1,2],[3]],[4]]}$ and the updated view ${100}$.
+\noindent which is illustrated by Figure \ref{fig:eval-comp-phead} with the original source $s$ being ${[[[1,2],[3]],[[4]]]}$ and the updated view ${100}$.
 To obtain the final updated source $s'$, $put$ for $lp3$ needs to evaluate $put$ of $phead$ three times.
@@
   \includegraphics[height=5cm]{./fig/fig3.eps}
-  \caption{Evaluating $phead \ \circ \ phead \ \circ \ phead$ by keeping intermediate states}
+  \vspace{-0.25cm}
+  \caption{Evaluating $phead \ \circ \ phead \ \circ \ phead$ by keeping complements}
+  \vspace{-0.5cm}
   \label{fig:eval-comp-phead-2}
@@
 
+Although we will introduce the basics of bidirectional transformation in the next session, it is not complete due to space limitations.
+Please refer BiGUL papers \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129} for the details if needed.
+
+%Due to space limitations, we assume the reader is familiar with the basic concepts of bidirectional transformation.
+
 %The contributions of this paper are the following.

### 2020-02-17 paper/3.minbigul_m.tex
tiny changes in formulas: keywords --> normal text, redundant vars --> underline
small updates

@@
 
-    \noindent $\putm{bx}{s}{v} = \text{match } bx \text{ with }\\
+    \noindent $\putm{bx}{s}{v} = \match{bx}\\
         \tab \vert \ bx_1 \circ bx_2 \to \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})} \\
@@
 
-        \noindent $\getm{bx}{s} = \text{match } bx \text{ with }\\
+        \noindent $\getm{bx}{s} = \match{bx}\\
             \tab \ \vert \ bx_1 \circ bx_2 \to\\
-            \tabs{2} \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
-            \tabs{2} \text{with Not\_found } \to\\
+            \tabs{2} \textnormal{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
+            \tabs{2} \textnormal{with } \text{Not\_found } \to\\
                 \tabs{3} v \Leftarrow \getmh{bx_2}{(\getmh{bx_1}{s})}\\
@@
         \noindent $\getmh{bx}{s} =\\
-            \tab \text{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
-            \tab \text{with Not\_found } \to\\
+            \tab \textnormal{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
+            \tab \textnormal{with } \text{Not\_found } \to\\
                 \tabs{2} v \Leftarrow \getm{bx}{s}\\
-                \tabs{2} \text{match } bx \text{ with } \\
+                \tabs{2} \match{bx} \\
                 \tabs{2} \ \vert \ bx_1 \circ bx_2 \to v \\
@@
 that the interpreter does not save all states when evaluating a program, only the intermediate states of a composition.
-When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, function $get_{mh}$ is used, not $get$ or $get_m$.
+When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, the function $get_{mh}$ is used, not $get$ or $get_m$.
 % appears to indicate the saving. 
-When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation.
+When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), where $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation.
 If there is no such key, the value $v$ will be calculated by using $get_{mh}$.

### 2020-02-17 paper/4.xpg.tex
remove \end
small updates

@@
 Another solution for saving intermediate states is tupling.
-If $put$ and $get$ are evaluated simultaneously, it is potential to reduce the number of recomputed $get$s.
+If $put$ and $get$ are evaluated simultaneously, there is potential to reduce the number of recomputed $get$s.
 The following function, $pg$, accepts the pair of a source and a view as the input to produce a new pair that contains the actual result of the corresponding minBiGUL program.
@@
 The second one tuples two results of $put$ and $get$ in the body of the if-expression.
-This is a trick since in some cases, the result of $pg$ may be undefined although the result is not
-\end
-undefined when evaluating $\getbxinline{Skip \ h}$. The last equality is relatively obvious.
+This is a trick since in some cases, the result of $pg$ may be undefined although the result is not undefined when evaluating $\getbxinline{Skip \ h}$.
+The last equality is relatively obvious.
 
@@
 
-Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just paring $put$ and $get$ respectively, then doing basic changes.\\
+Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just pairing $put$ and $get$ respectively, then doing basic changes.\\
 
@@
 
-    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \textnormal{ if } h \ s = v \text{ then } (ks, kv, s, v) \textnormal{ else } \textit{undefined}$
+    \noindent $\cpg{Skip \ h}{ks}{kv}{s}{v} = \textnormal{ if } h \ s = v \text{ then } (ks, kv, s, v) \ \textnormal{ else } \textit{undefined}$
 
@@
 While $cpg$ evaluates values eagerly, $kpg$ does the opposite.
-Every values are evaluated lazily in a computation of $kpg$.
+Every value is evaluated lazily in a computation of $kpg$.
 The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
@@
 Additionally we did two optimizations in $kpg$.
-First is in $\kpginline{bx_1 \times bx_2}$.
+The first is in $\kpginline{bx_1 \times bx_2}$.
 The evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
@@
 \subsection{Combination of pg and kpg: xpg}
-The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that is used to produce new source instead of using more recursive call.
+The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.
 This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$.
@@
 
-    \noindent $\xpg{bx}{s}{v} = \textnormal{match } bx \textnormal{ with }\\
+    \noindent $\xpg{bx}{s}{v} = \match{bx}\\
     \tab \vert \ bx_1 \circ bx_2 \to \\

### 2020-02-17 paper/5.experiments.tex
change anonymous link
small updates

@@
 %
-\footnote{Anonymous \url{https://mega.nz/\#!yUsgyCxA!yrJfINiwtCwhmd4McYMz2xsFa-ljkgGKmMi839ut5gA}}
+\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
 %
 described in the previous sections.
-Instead of constructing some dummies by $construct\_dummy$s, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the current updated views.
+Instead of constructing some dummies by $construct\_dummy$, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the current updated views.
 This helps a program in the $put$ direction valid.

### 2020-02-17 paper/6.related-work.tex
small updates

@@
 
-Since the pioneer work of lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
+Since the pioneering work of lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
 %(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)

### 2020-02-18 paper/2.minbigul.tex
small updates
done: change def of put_m & get_m

@@
  \smallvspace
-    $\qtab \getbxinline{bx} : s \to v, \quad \putbxinline{bx} : s * v \to v$
+    $\qtab \getbxinline{bx} : s \to v, \quad \putbxinline{bx} : s * v \to s$
  \smallvspace
@@
 
+Also in the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, the return statements are used to express clearly the value of functions.
+Variables $s'$ and $v'$ wrapped in these returns are neccessary for checking the fi conditions.
+
+\subsection{Examples}
+
 \noindent As an example of minBiGUL program, consider the definition of $phead$ :\\
@@
 
-\smallvspace
 $\tab \putbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]}{100} = [[100,2,3],[\,],[4,5]]$
@@
 $\tab \getbx{phead \circ phead}{[[1,2,3],[\,],[4,5]]} = 1$
-\smallvspace
+
+Next let us look at another more complex example using a case combinator:
+
+    $\tab bsnoc = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+    $\tabs{3} \ cond_{sv} = \lambda s.\lambda v.(\textnormal{length } v = 1), \ cond_s = \lambda (s_1,s_2).(s_2 = [\ ])$\\
+    $\tabs{3} \ bx_1 =  Replace$\\
+    $\tabs{3} \ bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
+        $\tabs{4} f_1 = f_2^{-1} = \lambda (x:y:ys).(y,(x:ys))$, \\
+        $\tabs{4} bx_s = RearrV \ g_1 \ g_2 \ bx_v$ where:\\
+            $\tabs{5} g_1 = g_2^{-1} = \lambda (v:vs).(v,vs)$\\
+            $\tabs{5} bx_v = Replace \times bsnoc$
 

### 2020-02-18 paper/3.minbigul_m.tex
done: change def of put_m & get_m

@@
 The memoizing version, minBiGUL$_m$, needs only two modifications: $get_m$ and $put_m$ (Definitions~\ref{def:putm} and~\ref{def:getm}).
-The auxiliary function, $get_{mh}$, is used to retrieve the value of an already evaluated $get$ from the table or to compute an unevaluated $get$ by $get_m$ and store the key-value tuple in the table (Definition~\ref{def:getmh}).
 
@@
     \noindent $\putm{bx}{s}{v} = \match{bx}\\
-        \tab \vert \ bx_1 \circ bx_2 \to \putm{bx_1}{s}{(\putm{bx_2}{(\getmh{bx_1}{s})}{v})} \\
+        \tab \vert \ bx_1 \circ bx_2 \to \\
+            \tabs{2} \putm{bx_1}{s}{(\putm{bx_2}{(\getm{bx_1}{s})}{v})} \\
         \tab \vert \ \_ \to \text{similar to } put$
@@
 
-\begin{multicols}{2}
 \begin{definition} \label{def:getm} $\text{Memo version of } get$
@@
     \tabs{2} \textnormal{with } \text{Not\_found } \to\\
-                \tabs{3} v \Leftarrow \getmh{bx_2}{(\getmh{bx_1}{s})}\\
+        \tabs{3} i \Leftarrow \getm{bx_1}{s}; \tab \text{Hashtbl.add } \ table_g \ (bx_1, s) \ i;\\
+        \tabs{3} v \Leftarrow \getm{bx_2}{i}; \tab \text{Hashtbl.add } \ table_g \ (bx_2, i) \ v;\\
         \tabs{3} \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
@@
 \end{definition}
-\columnbreak
-    \begin{definition} \label{def:getmh} $\text{Interface for } get_m$
 
-        \noindent $\getmh{bx}{s} =\\
-            \tab \textnormal{try } (\text{Hashtbl.find} \ table_g \ (bx, s))\\
-            \tab \textnormal{with } \text{Not\_found } \to\\
-                \tabs{2} v \Leftarrow \getm{bx}{s}\\
-                \tabs{2} \match{bx} \\
-                \tabs{2} \ \vert \ bx_1 \circ bx_2 \to v \\
-                \tabs{2} \ \vert \ \_ \to \text{Hashtbl.add } \ table_g \ (bx, s) \ v\\
-                \qtabs{2} \ v$
-    \end{definition}
-\end{multicols}
-
+The evaluation of $\putm{bx_1 \circ bx_2}{s}{v}$ includes two recursive calls of $put_m$ and an external call of $get_m$, which is relatively similar to the evaluation of $\putbx{bx_1 \circ bx_2}{s}{v}$.
+Meanwhile, the evaluation of $\getm{bx_1 \circ bx_2}{s}$ does not merely invoke $get_m$ recursively twice.
+In case that $bx$ is a composition, the key ($bx,s$) needs to be looked up in the table and the corresponding value would be used for the next steps in the evaluation.
+If there is no such key, the value of the intermediate state $i$ and the value of $\getbx{bx}{s}$ in $v$ will be calculated.
+These values along with the corresponding keys will also be stored in the table where the interpreter may later
+leverage instead of reevaluating some states.
 % Note as a special point 
@@
 that the interpreter does not save all states when evaluating a program, only the intermediate states of a composition.
-When evaluating $\putm{bx_1 \circ bx_2}{s}{v}$, the function $get_{mh}$ is used, not $get$ or $get_m$.
-% appears to indicate the saving. 
-When evaluating $\getm{bx_1 \circ bx_2}{s}$, the key ($bx,s$), where $bx \equiv bx_1 \circ bx_2$, is looked up in the table and the corresponding value is used for the next steps in the evaluation.
-If there is no such key, the value $v$ will be calculated by using $get_{mh}$.
-This key-value insertion is necessary because the interpreter may later need to reevaluate $\getm{bx_1 \circ bx_2}{s}$.
-% reevaluate <--> evaluate one more time.

### 2020-02-18 paper/4.xpg.tex
done: change def of put_m & get_m

@@
 
-Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just pairing $put$ and $get$ respectively, then doing basic changes.\\
+Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just pairing $put$ and $get$ respectively, then doing basic changes.
+%
+The values of these $pg$ functions are the values of final expressions respectively. We only use the return keyword to express explicitly the evaluated value of a function in the situation of $Case$.
 
@@
     \tabs{2} \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
-    \tabs{3} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
+    %
+    \tabs{3} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \ \text{then}\\
+        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_1}{s});\\
+        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
+        \tabs{2} \qtab \text{else if} \ {cond_{sv} \ s \ v} \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
+        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_2}{s});\\
+        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
+        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ {cond_{s} \ s} \ \text{then}\\
+        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_1}{s});\\
+        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
+        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
+        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_2}{s});\\
+        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
+    %
+    \tabs{3} \overset{3}{=} (* \text{ with restriction } *)\\
+    \tabs{2} \qtab \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
     \tabs{2} \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
@@
         \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
-        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
+        \tabs{2} \qtab (s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
             \qtab \tabs{3} (s_3, v_2)$
@@
 %We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require programmers to give a valid view.
-Since there is no real view, this $dummy$ is necessary.
+Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$.
 The last equality changes two.
-First is a change from $dummy$ to an application $construct\_dummy \ s$.
+The first is a change from $dummy$ to an application $construct\_dummy \ s$.
 This means we need a way to construct $dummy$ from the current source, $s$.
-We can require programmers to give this $construct\_dummy$ function.
-Second is a substitution under the \textsc{PutPut} law on $s$ of $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ by $s_1$ which equals to $\putbx{bx_1}{s}{(construct\_dummy \ s)}$.
-$v_3$ in the result pair $(s_3,v_3)$ equals $dummy$.
-So we realize that there is no loss information when computing a composition.
+Programmers can be required to give this $construct\_dummy$ function, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is the easiest way to meet our expectation.
+We set up that choice in our experiments but still use the $construct\_dummy$ function in the later definitions to provide a more general meanning.
+%
+The second is the change from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively.
+Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$ is possible.
+The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively.
+Because both variables are no longer used later, this substitution does not affect the outcome of the function.
 

### 2020-02-18 paper/2.minbigul.tex
explain examples + revise experiments descr.

@@
 
-\noindent As an example of minBiGUL program, consider the definition of $phead$ :\\
+\noindent As an example of minBiGUL program, consider the definition of $\textit{phead}$ :\\
 \smallvspace
@@
 
-Next let us look at another more complex example using a case combinator:
+In the same way with $\textit{phead}$, we can define $\textit{ptail}$ in minBiGUL.
+$\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$.
+$\getbx{ptail}{s}$ returns the tail of the source list $s$
 
-    $\tab bsnoc = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
-    $\tabs{3} \ cond_{sv} = \lambda s.\lambda v.(\textnormal{length } v = 1), \ cond_s = \lambda (s_1,s_2).(s_2 = [\ ])$\\
-    $\tabs{3} \ bx_1 =  Replace$\\
-    $\tabs{3} \ bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
+Next let us look at another more complex example, $\textit{bsnoc}$:
+
+    $\tab \textit{bsnoc} = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+    $\tabs{3} \ cond_{sv} = \lambda s.\lambda v.(\textnormal{length } v = 1), \ cond_s = \lambda s.(\textnormal{length } s = 1)$\\
+    $\tabs{3} \ bx_1 =  Replace, \tab bx_2 = RearrS \ f_1 \ f_2 \ bx_s$ where:\\
         $\tabs{4} f_1 = f_2^{-1} = \lambda (x:y:ys).(y,(x:ys))$, \\
         $\tabs{4} bx_s = RearrV \ g_1 \ g_2 \ bx_v$ where:\\
-            $\tabs{5} g_1 = g_2^{-1} = \lambda (v:vs).(v,vs)$\\
-            $\tabs{5} bx_v = Replace \times bsnoc$
+            $\tabs{5} g_1 = g_2^{-1} = \lambda (v:vs).(v,vs), \tab bx_v = Replace \times \textit{bsnoc}$
 
-To close this section,
-% In the end of this section, 
-let us see the minBiGUL definition of $\mathit{bfoldr}$ which is a putback function of an important higher-order function on lists, $\mathit{foldr}$:
+$\putbxinline{bsnoc}$ requires the source $s$ and the view $v$ are non-empty lists and the length of $v$ is not larger than the length of $s$.
+If $cond_{sv}$ is true, i.e.
+$v$ is singleton, a replacement will be executed to produce a new list which should be equal to $v$.
+Because the length of the new list is 1, the exit condition $cond_s$ comes true, so we obtain the updated source.
+If $v$ is a list of more than one elements, there will be two rearrangements on the source and the view before conducting a product.
+The program rearranges the source $x:y:ys$ to a pair of its second element $y$ and a list $x:ys$ created from the remaining elements in the original order, and the view to a pair of its head and tail.
+Then we can use $y$ to replace the head of the view and pair $(x:ys)$ with the tail of the view to form the input of a recursive call $\textit{bsnoc}$.
+The obtained source update in this case should be non-singleton since the value of the exit condition $cond_s$ needs to be false.
+We omit the behavior description of $\getbxinline{bsnoc}$ that accepts a source list $s$, checks $cond_s$ to know how to evaluate the view $v$, then does one more checking, $cond_{sv}$, before resulting.
+%
+Intuitively, $\putbx{bsnoc}{s_0}{v_0}$ produces a new list by moving the last element of $v_0$ to its first position if the length of $v_0$ is not larger than the length of $s_0$.
+$\getbx{bsnoc}{s_0}$ returns another list by moving the first element of the list $s_0$ to its end position.
+For instance, $\putbx{bsnoc}{[1,2,3]}{[4,5,6]} = [6,4,5]$ and $\getbx{bsnoc}{[1,2,3]} = [2,3,1]$.
 
-\smallvspace
-$\tab \mathit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
+Now, let us see the minBiGUL definition of $\textit{bfoldr}$ which is a putback function of an important higher-order function on lists, $\textit{foldr}$:
+
+$\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
 $\tabs{3} \ cond_{sv} = \lambda (s_1,s_2).\lambda v.(s_1 = [\ ]), \ cond_s = \lambda (s_1,s_2).(s_1 = [\ ])$\\
@@
 $bx_s = ((Replace \times bfoldr \ bx) \circ bx)$
-\smallvspace
 
-\noindent In $\mathit{bfoldr}$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
-Because~$\circ$ has a higher priority than~$\times$, it is in general not possible to transform $\mathit{bfoldr}$ from the left associative composition style to the right one.
-%will look like a composition of many programs at some points in the evaluation if we slow down this process.
-Since $\circ$ has a higher priority than $\times$, it is seemingly impossible to transform $bfoldr$ from the left associative composition style to the right one.
+If we think that a minBiGUL program $bx$ has a type of $\textit{MinBiGUL} \ s \ v$, the type of $bfoldr$ will be look like $\textit{MinBiGUL} \ (a, b) \ a \to \textit{MinBiGUL} \ ([a], b) \ b$.
+You can easily find the similarity between the above definition of $\textit{bfoldr}$ with the following definition of $\textit{foldr}$:
+
+$\tab foldr :: (a \to b \to b) \to b \to [a] \to b \\
+\tabs{2} \ \ foldr \ f \ e \ [\ ] = e \\ 
+\tabs{2} \ \ foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
+
+Each branch in a case of $\textit{bfoldr}$ corresponds to a pattern of $foldr$.
+In $\textit{bfoldr}$, the composition is inherently left associative, and the number of composition is dynamically determined by the length of the source list.
+Because~$\circ$ has a higher priority than~$\times$, it is in general not possible to transform $\textit{bfoldr}$ from the left associative composition style to the right one.
+
+Using $\textit{foldr}$, we can define other functions, for example:
+
+$\tab reverse = foldr \ snoc \ [\ ]$
+
+With $\textit{bfoldr}$, we can also write the bidirectional version $breverse$ as follows:
+
+$\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where:\\
+    $\tabs{4} f_1 = f_2^{-1} = \lambda s \to (s, [\ ]), \tab bx = bfoldr \ bsnoc$

### 2020-02-18 paper/4.xpg.tex
tiny + update review-todo
explain examples + revise experiments descr.

@@
             \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
-    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ s}; \\
+    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ v}; \\
         \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
@@
 The last equality changes two.
-The first is a change from $dummy$ to an application $construct\_dummy \ s$.
-This means we need a way to construct $dummy$ from the current source, $s$.
+The first is a change from $dummy$ to an application $construct\_dummy \ v$.
+This means we need a way to construct $dummy$ from the original view $v$.
 Programmers can be required to give this $construct\_dummy$ function, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is the easiest way to meet our expectation.
@@
 The second is the change from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively.
-Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$ is possible.
+Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$.
 The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively.
@@
     \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ s};\\
+        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ v};\\
         \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
@@
 Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ s)$ where $s$ and $ks$ are corresponding to the source and the update over source.
+The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ v)$ where $s$ and $ks$ are corresponding to the source and the update over source.
 Since there is no real view here, we need to construct a dummy from the source same as $pg$.
@@
     \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ s};\\
+        \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ v};\\
         \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
@@
     \tab \vert \ bx_1 \circ bx_2 \to \\
-        \tabs{2} (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ s};\\
+        \tabs{2} (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ v};\\
         \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\

### 2020-02-18 paper/5.experiments.tex
explain examples + revise experiments descr.

@@
 
-$\putbx{ptail}{s}{v}$ accepts a source list $s$ and a view list $v$ to produce a new list by replacing the tail of $s$ with $v$.
-$\getbx{ptail}{s}$ returns the tail of the source list $s$.
 The composition of many $ptail$s, in the \textit{put} direction, replaces a part of the tail of the source list by the view list and, in the \textit{get} direction, returns such a tail from the source.
 
-$\putbx{bsnoc}{s}{v}$ accepts a source $s$ and a view $v$, 
-%as two same length lists, 
-which are two lists of the same length,
-and produces a new list by moving the last element of $v$ to its first position.
-$\getbx{bsnoc}{s}$ returns another list by moving the first element of the list $s$ to its end position.
-The composition of many \textit{bsnoc}s, in the \textit{put} direction, creates a permutation of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a permutation of the source list.
+The composition of many \textit{bsnoc}s, in the \textit{put} direction, creates a permutation of the view list if its length is not larger than the length of the source list and, in the \textit{get} direction, produces a permutation of the source list.
 
 \textit{breverse} is defined in terms of \textit{bfoldr}.
-In the \textit{put} direction, it produces a reverse of the view list if the source list has a same length as the view and, in the \textit{get} direction, produces a reverse of the source list.
-Note that compositions are by the recursions of $breverse$ and
+In the \textit{put} direction, it produces a reverse of the view list if its length is not larger than the length of the source list and, in the \textit{get} direction, produces a reverse of the source list.
+Note that compositions are by the recursions of \textit{breverse} and
 the number of compositions are dynamically determined by the length of the source list.

### 2020-02-18 paper/4.xpg.tex
before renaming vars

@@
 
-\noindent Now, let us see how we construct $pg$ respectively.
+\noindent Now, let us see how we construct $pg$ recursively.
 
@@
     $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
-        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}), (\getbx{bx_2}{s_2})) \\
+        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
         \tab \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
@@
 %
-The values of these $pg$ functions are the values of final expressions respectively. We only use the return keyword to express explicitly the evaluated value of a function in the situation of $Case$.
+The values of these $pg$ functions are obtained from the final expression in the corresponding sequences.
+We only use the return keyword to express explicitly the evaluated value of a function in the situation of $Case$.
 
@@
 
-The problem for $cpg$ lies in redundant computations during evaluating.
-For keeping away such computations, we introduce an extension named $kpg$.
+The problem for $cpg$ lies in redundant computations during the evaluation.
+To prevent such redundant computations from occuring, we introduce an extension named $kpg$.
 While $cpg$ evaluates values eagerly, $kpg$ does the opposite.
@@
         \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-        \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
-        \qtab ( con \ ks_1 \ ks_2, con \ kv_1 \ kv_2, con \ (ks_1' \circ fst) \ (ks_2' \circ snd),\\
-        \qtab con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
-        \qtab (s_1, s_2), (v_1,v_2))$
+        \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \\
+            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
+        \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \\ 
+            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
+        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
+        \qtab \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), \ con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
+        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$
 

### 2020-02-18 paper/4.xpg.tex
rename vars
s: source, v: view
k-: kontinuation
l-: lazy
e-: evaluated
^prime ('): new var for single bx
_index (1 | 2): new var for multiple bxs
textcolor{red}: old version

@@
     \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
+%-{
+    \textcolor{red}{
     $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
@@
         \tab \qtabs{2} (s, g_2 \ v)$
+    }\\
+%-}
+    %+{
+    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
+        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
+        \tab \qtab \overset{2}{=} (s_1', v_1') \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
+            \qtab \tabs{2} \ \, (s_2', v_2') \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
+            \tab \qtabs{2} ((s_1',s_2'), (v_1',v_2'))$\\
+    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
+        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{f_1 \ s}{v};\\
+            \tab \qtabs{2} (f_2 \ s', v')$\\
+    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
+        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
+        \tab \qtabs{2} (s', g_2 \ v')$
+    %+}
 
@@
 
-    \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
+%-{
+    \noindent \textcolor{red}{
+    $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
         \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
         \qtab (f_2 \circ ks, kv, s, v)$
+    }
 
-    \noindent $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
+    \noindent \textcolor{red}{
+    $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
         \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
         \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$
+    }
 
-    \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
+    \noindent \textcolor{red}{
+    $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
         \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
@@
         \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks, kv, s', v')$
+    }
+%-}
+
+    %+{
+    \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
+        \tab (ks', kv', s', v') \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
+        \qtab (f_2 \circ ks', kv', s', v')$
+
+    \noindent $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
+        \tab (ks', kv', s', v') \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
+        \qtab (ks', g_2 \circ kv', ks', g_2 \ v')$
+
+    \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
+        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
+        \tab \textnormal{then} \ (ks', kv', s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
+        \tab \textnormal{else} \ (ks', kv', s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
+        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks', kv', s', v')$
+    %+}
 
@@
 Every value is evaluated lazily in a computation of $kpg$.
-The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, ks', kv', s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $ks'$ and $kv'$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, ks', kv', s, v)$.\\
+The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values.
+The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.\\
 Suppose that we have a source $s_0$ and a view $v_0$.
@@
 \smallvspace
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
-        \tab \qtab (ks' \ s; kv' \ v)$
+    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+        \tab \qtab (lks \ s; lkv \ v)$
 \smallvspace
 
-The beginning of accumulative functions $ks'$ and $kv'$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
+The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
 
 \begin{definition}
-$\kpg{bx}{ks}{kv}{ks'}{kv'}{s}{v}$
+$\kpg{bx}{ks}{kv}{lks}{lkv}{s}{v}$
 
-    \noindent $\kpg{Skip \ h}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab \textnormal{if} \ h \ s = v \ \textnormal{then} \ (ks, kv, ks', kv', s, v) \ \textnormal{else } \text{undefined}$
+%-{
+    \noindent 
+    \textcolor{red}{
+        $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
+        \tab \textnormal{if} \ h \ s = v \ \textnormal{then} \ (ks, kv, lks, lkv, s, v) \ \textnormal{else } \text{undefined}$
+    }
 
-    \noindent $\kpg{Replace}{ks}{kv}{ks'}{kv'}{s}{v} = (kv, ks, kv', ks', v, s)$
+    \noindent 
+    \textcolor{red}{
+        $\kpg{Replace}{ks}{kv}{lks}{lkv}{s}{v} = (kv, ks, lkv, lks, v, s)$
+    }
 
-    \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id; \quad kv' \Leftarrow id;\\
-        \tab (ks_1, kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \\
-            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ ks'}{fst \circ kv'}{s}{v};\\
-        \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \\ 
-            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ ks'}{snd \circ kv'}{s}{v};\\
+    \noindent 
+    \textcolor{red}{
+        $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
+        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
+            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ lks}{fst \circ lkv}{s}{v};\\
+        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
+            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ lks}{snd \circ lkv}{s}{v};\\
         \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
-        \qtab \tab con \ (ks_1' \circ fst) \ (ks_2' \circ snd), \ con \ (kv_1' \circ fst) \ (kv_2' \circ snd), \\
+        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
+        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$
+    }
+
+    \noindent 
+    \textcolor{red}{
+        $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ lks}{lkv}{s}{v};\\
+        \qtab (f_2 \circ ks, kv, f_2 \circ lks, lkv, s, v)$
+    }
+
+    \noindent 
+    \textcolor{red}{
+        $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{lks}{g_1 \circ lkv}{s}{v};\\
+        \qtab (ks, g_2 \circ kv, lks, g_2 \circ lkv, s, v)$
+    }
+
+    \noindent 
+    \textcolor{red}{
+        $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
+        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
+        \tab \textnormal{then} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{lks}{lkv}{s}{v}\\
+        \tab \textnormal{else} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{lks}{lkv}{s}{v}\\
+        \tab \underline{\textnormal{fi}} \ cond_s \ (lks \ s') \ \&\& \ cond_{sv} \ s \ (lkv \ v'); \ \textnormal{return} \ (ks, kv, lks, lkv, s', v')$
+    }
+
+    \noindent 
+    \textcolor{red}{
+        $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
+        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
+        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
+    }
+%-}
+
+    %+{
+    \noindent $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
+        \tab \textnormal{if} \ h \ es = ev \ \textnormal{then} \ (ks, kv, elks, elkv, es, ev) \ \textnormal{else } \text{undefined}$
+
+    \noindent $\kpg{Replace}{ks}{kv}{lks}{lkv}{s}{v} = (kv, ks, lkv, lks, v, s)$
+
+    \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
+        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
+            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ elks}{fst \circ elkv}{es}{ev};\\
+        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
+            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ elks}{snd \circ elkv}{es}{ev};\\
+        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
+        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
         \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$
 
-    \noindent $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ ks'}{kv'}{s}{v};\\
-        \qtab (f_2 \circ ks, kv, f_2 \circ ks', kv', s, v)$
+    \noindent $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ lks}{lkv}{s}{v};\\
+        \qtab (f_2 \circ ks', kv', f_2 \circ lks', lkv', s', v')$
 
-    \noindent $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab (ks, kv, ks', kv', s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{ks'}{g_1 \circ kv'}{s}{v};\\
-        \qtab (ks, g_2 \circ kv, ks', g_2 \circ kv', s, v)$
+    \noindent $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{lks}{g_1 \circ lkv}{s}{v};\\
+        \qtab (ks', g_2 \circ kv', lks', g_2 \circ lkv', s', v')$
 
-    \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab s \Leftarrow ks' \ s; \quad v \Leftarrow kv' \ v; \quad ks' \Leftarrow id ; \quad kv' \Leftarrow id;\\
-        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-        \tab \textnormal{then} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{ks'}{kv'}{s}{v}\\
-        \tab \textnormal{else} \ (ks, kv, ks', kv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{ks'}{kv'}{s}{v}\\
-        \tab \underline{\textnormal{fi}} \ cond_s \ (ks' \ s') \ \&\& \ cond_{sv} \ s \ (kv' \ v'); \ \textnormal{return} \ (ks, kv, ks', kv', s', v')$
+    \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
+        \tab \textnormal{if} \ cond_{sv} \ es \ ev \ \&\& \ cond_{s} \ es\\
+        \tab \textnormal{then} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{elks}{elkv}{es}{ev}\\
+        \tab \textnormal{else} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{elks}{elkv}{es}{ev}\\
+        \tab \underline{\textnormal{fi}} \ cond_s \ (lks' \ s') \ \&\& \ cond_{sv} \ \textbf{es} \ (lkv' \ v'); \ \textnormal{return} \ (ks', kv', lks', lkv', s', v')$
 
-    \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{ks'}{kv'}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{ks_1'}, kv_1', \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{ks'}{id}{s}{construct\_dummy \ v};\\
-        \tab (ks_2, kv_2, ks_2', kv_2', s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{kv_1'}{kv'}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ ks_2', kv_2', s_2, v_2)$
+    \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
+        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
+        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
+        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
+    %+}
 \end{definition}
 
-In $kpg$, basically, functions for the updates will be kept in $ks'$ and $kv'$.
+\textcolor{red}{
+In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
 $\kpginline{Skip}$ and $\kpginline{Case}$ are special cases, because they require the actual values for evaluation.
-Therefore we evaluate the values by application of $ks' \ s$ and $kv' \ v$ and update $ks'$ and $kv'$ with the identity functions.
+Therefore we evaluate the values by application of $lks \ s$ and $lkv \ v$ and update $lks$ and $lkv$ with the identity functions.
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{ks_1'}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but not evaluated.
-
+When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{lks_1}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but not evaluated.
+\\
 Additionally we did two optimizations in $kpg$.
 The first is in $\kpginline{bx_1 \times bx_2}$.
-The evaluation of $ks'$ and $kv'$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be same computations in $fst \circ ks'$ and $snd \circ ks'$ as well as $fst \circ kv'$ and $snd \circ kv'$.
+The evaluation of $lks$ and $lkv$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+There may be same computations in $fst \circ lks$ and $snd \circ lks$ as well as $fst \circ lkv$ and $snd \circ lkv$.
 To remove duplicate evaluations, we evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$.
@@
 Second is in $\kpginline{Case}$.
-We need to evaluate $ks' \ s'$ and $kv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple.
+We need to evaluate $lks \ s'$ and $lkv \ v'$ to check the \underline{fi} condition before returning the 6-tuple.
 Such evaluations can be done lazily to make programs run faster.
 We use the above small optimizations in our implementation.
+}
 
@@
     \tab \vert \ bx_1 \circ bx_2 \to \\
-        \tabs{2} (ks_1,kv_1, ks_1', kv_1', s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ v};\\
-        \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{kv_1' \ v_1}{v};\\
+        \tabs{2} (ks_1,kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ v};\\
+        \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{lkv_1 \ v_1}{v};\\
         \tabs{3} (ks_1 \ s_2, v_2)\\

### 2020-02-18 paper/4.xpg.tex
update initialization because of renaming

@@
 \smallvspace
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s}{id}{s}{v};\\
+    $\tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s_0}{id}{s_0}{v_0};\\
         \tab \qtab (s; v)$
@@
 In general, the beginning of a continuation should be an identity function.
-However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source $s$ needs to be initiated as $\lambda \_.s$.
+However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source needs to be initiated as the constant function from that source.
 This constant function helps to retain the discarded things in the source.
@@
 \smallvspace
-    $\tab s \Leftarrow s_0; v \Leftarrow v_0; (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s}{id}{id}{id}{s}{v};\\
+    $\tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s_0}{id}{id}{id}{s_0}{v_0};\\
         \tab \qtab (lks \ s; lkv \ v)$

### 2020-02-18 paper/5.experiments.tex
update initialization because of renaming

@@
 % they represent non-trivial cases of practical significance.
-In the last two columns, $s_r$ and $v_r$ are the updated source and view, respectively.
+In the last two columns, $s$ and $v$ are the updated source and view, respectively.
 They are produced by applying $put$ and $get$ to the original source $s_0$ and view $v_0$.
-That is, $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is the program indicated in the second column of the table.
+That is, $s = \putbx{bx}{s_0}{v_0}$ and $v = \getbx{bx}{s_0}$, where $bx$ is the program indicated in the second column of the table.
 % are only interested in this kind of programs in the paper. 
 % Please note that the results
-Results $s_r$ and $v_r$ are independent of the associativity 
+Results $s$ and $v$ are independent of the associativity 
 % style 
@@
         \multirow{2}{*}{No} & \multicolumn{1}{c|}{\multirow{2}{*}{Name}} & \multicolumn{1}{c|}{\multirow{2}{*}{Type}} & \multicolumn{2}{c|}{Input} & \multicolumn{2}{c|}{Output} \\ \cline{4-7} 
-        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s_r$} & \multicolumn{1}{c|}{$v_r$} \\ \hline
+        & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{} & \multicolumn{1}{c|}{$s_0$} & \multicolumn{1}{c|}{$v_0$} & \multicolumn{1}{c|}{$s$} & \multicolumn{1}{c|}{$v$} \\ \hline
         1 & lcomp-phead-ldata & straight line & $\underbrace{[[\ldots[1]\ldots]]}_{\text{n+1 times}}$ & 100 & $\underbrace{[[\ldots[100]\ldots]]}_{\text{n+1 times}}$ & 1 \\ \hline

### 2020-02-19 paper/1.introduction.tex
add todo list

@@
 \end{figure}
-\vspace*{-5.5mm}
+%\vspace*{-5.5mm}
 
@@
 \begin{align}
-get_{bx_1 \circ bx_2}~s &= get_{bx_2} (get_{bx_1}~s)\\
+get_{bx_1 \circ bx_2}~s &= get_{bx_2} (get_{bx_1}~s) \AlignFootnote{Note that the order of composition is the reverse order compared to function composition.
+We use this order, because it is standard in BX community.}\\
 put_{bx_1 \circ bx_2}~s~v' &= put_{bx_1} s~(put_{bx_2}~ (get_{bx_1}~s)~v')
@@
 
+\textcolor{red}{TODO-B3,C1: add explanation of complements.}
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.

### 2020-02-19 paper/4.xpg.tex
tiny
small updates
add todo list

@@
 This is a trick since in some cases, the result of $pg$ may be undefined although the result is not undefined when evaluating $\getbxinline{Skip \ h}$.
-The last equality is relatively obvious.
+The last equality is a function application.
 
@@
 This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{If the BX programs have efficient inverse way of evaluation, they are not affected by this restriction.
+Of course, some programs does not have such efficient way.
+An example restricted by this is evaluating $snoc$ with different length of a source and a view.}}.
 This does not happen for the others which is used in the construction of $\pginline{Case}$.
@@
     $\tab (ks, kv, s, v) \Leftarrow \cpg{bx}{\lambda \_.s_0}{id}{s_0}{v_0};\\
-        \tab \qtab (s; v)$
+        \tab \qtab (s, v)$
 \smallvspace
@@
     $\tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s_0}{id}{id}{id}{s_0}{v_0};\\
-        \tab \qtab (lks \ s; lkv \ v)$
+        \tab \qtab (lks \ s, lkv \ v)$
 \smallvspace
@@
 
-\textcolor{red}{
+\textcolor{red}{TODO-C6: RearrS, RearrV just keeping, Really evaluated is skip, product}
 In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
@@
 We use the above small optimizations in our implementation.
-}
 
 \subsection{Combination of pg and kpg: xpg}
+\textcolor{red}{TODO-D3: change explanation clearer.}
 The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.

### 2020-02-19 paper/5.experiments.tex
add todo list

@@
 
-\textit{breverse} is defined in terms of \textit{bfoldr}.
+\textit{breverse} is defined in terms of \textit{bfoldr}, that appeared in Section \ref{sect:minbigul}.
 In the \textit{put} direction, it produces a reverse of the view list if its length is not larger than the length of the source list and, in the \textit{get} direction, produces a reverse of the source list.
@@
 
+\textcolor{red}{TODO-D4: emphasize overhead of memoization. it is not for free.}
 Figure \ref{fig:result} 
@@
 The slowness was caused by the exponential number of $pg$ calls in the case of $pg$, redundant evaluations ($cpg$), and redundant overhead for constructing closures ($kpg$).
-Then we simply omit these results in Figure \ref{fig:result}.
+Therefore we simply omit these results in Figure \ref{fig:result}.
 As we all know, $put$ works poorly for left associative compositions because of the number of reevaluated $get$s.
@@
 the most effective method.
+\textcolor{red}{TODO-B2: add short evaluation result of memory.}

### 2020-02-19 paper/main.tex
Tiny
Capitalized \title, \titlerunning.
reduce some vspaces and add some vspaces
tiny changes in formulas: keywords --> normal text, redundant vars --> underline
small updates
rm \newcommand{getmh}
add todo list

@@
 
-\newcommand{\sif}[3]{\text{if } #1 \text{ then } #2 \text{ else } #3}
+\newcommand{\sif}[3]{\textnormal{if} \ #1 \ \textnormal{then} \ #2 \ \textnormal{else} \ #3}
 \newcommand{\product}[2]{#1 \times #2}
@@
 
-\newcommand{\getmh}[2]{get_{mh} \, [\![#1]\!] \ #2}
-\newcommand{\getmhinline}[1]{get_{mh} \, [\![#1]\!]}
-
 \newcommand{\putrev}[3]{put_{REV} \, [\![#1]\!] \ {#2} \ {#3}}
@@
 \newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
+\newcommand{\smallvspace}{\vspace{0.3mm}}
+\newcommand{\minusvspace}{\vspace*{-5.5mm}}
+\newcommand{\minusvspacetwo}{\vspace*{-3.5mm}}
+
+\newcommand{\match}[1]{\textnormal{match } #1 \ \textnormal{ with }}
+
+\makeatletter
+\newcommand{\AlignFootnote}[1]{%
+    \ifmeasuring@
+    \else
+        \footnote{#1}%
+    \fi
+}
+\makeatother
 
@@
 %
-\titlerunning{Strategies for efficient evaluation of BX programs}
+\titlerunning{An Efficient Composition of BX Programs}
 %\titlerunning{Abbreviated paper title}
@@
 
-%\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and
-%  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and
-%  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
-%  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
+\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and \break
+  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
+  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
+  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
 
-\author{anonymous authors}
+%\author{anonymous authors}
 
-%\authorrunning{K. Tsushima et al.}
+\authorrunning{K. Tsushima et al.}
 
@@
 
-%\institute{National Institute of Informatics, Japan
-%  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
-%  University of Copenhagen, Denmark
-%  \email{glueck@di.ku.dk } \and
-%  Peking University, China
-%  \email{huzj@pku.edu.cn}}
+\institute{National Institute of Informatics, Japan
+  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
+  University of Copenhagen, Denmark
+  \email{glueck@di.ku.dk } \and
+  Peking University, China
+  \email{huzj@pku.edu.cn}}
 

### 2020-02-19 paper/4.xpg.tex
finish renaming + revise dummy

--- 2020-02-18 paper/4.xpg.tex
@@
 This is a trick since in some cases, the result of $pg$ may be undefined although the result is not undefined when evaluating $\getbxinline{Skip \ h}$.
-The last equality is relatively obvious.
+The last equality is a function application.
 
     \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
-%-{
-    \textcolor{red}{
-    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
-        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
-        \tab \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
-            \qtab \tabs{2} \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
-            \tab \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
-    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
-        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
-            \tab \qtabs{2} (f_2 \ s, v)$\\
-    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
-        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
-        \tab \qtabs{2} (s, g_2 \ v)$
-    }\\
-%-}
-    %+{
     $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
@@
         \tab \qtabs{2} (s', g_2 \ v')$
-    %+}
 
@@
             \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
-    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ v}; \\
+    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
         \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
@@
 This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-%We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require programmers to give a valid view.
 Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$.
-The last equality changes two.
-The first is a change from $dummy$ to an application $construct\_dummy \ v$.
-This means we need a way to construct $dummy$ from the original view $v$.
-Programmers can be required to give this $construct\_dummy$ function, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is the easiest way to meet our expectation.
-We set up that choice in our experiments but still use the $construct\_dummy$ function in the later definitions to provide a more general meanning.
-%
-The second is the change from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively.
+In general, $dummy$ depends on the source $s$, the view $v$ and/or the program $bx_1$.
+Programmers can be required to give a way to construct $dummy$, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is one of the easiest ways to meet our expectation.
+That setting is used in our experiments.
+The last equality changes from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively.
 Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$.
@@
 
-%-{
-    \noindent \textcolor{red}{
-    $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
-        \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
-        \qtab (f_2 \circ ks, kv, s, v)$
-    }
-
-    \noindent \textcolor{red}{
-    $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
-        \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
-        \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$
-    }
-
-    \noindent \textcolor{red}{
-    $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
-        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-        \tab \textnormal{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
-        \tab \textnormal{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
-        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks, kv, s', v')$
-    }
-%-}
-
-    %+{
     \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
@@
         \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks', kv', s', v')$
-    %+}
 
     \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ v};\\
+        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{dummy};\\
         \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
@@
 Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$.
-The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ v)$ where $s$ and $ks$ are corresponding to the source and the update over source.
-Since there is no real view here, we need to construct a dummy from the source same as $pg$.
+The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, dummy)$ where $s$ and $ks$ are corresponding to the source and the update over source.
+Since there is no real view here, we need a dummy same as $pg$.
 Then the continuation updating on this dummy should be initiated as an identity function.
@@
 This constant function helps to retain the discarded things in the source.
+
+The result pair $(s, v)$ obtained from $cpg$ as above should be same with the result of $\pg{bx}{s_0}{v_0}$. More generally, we have the following relationship:
+
+$\tab \cpg{bx}{ks}{kv}{s}{v} = \pg{bx}{ks \ s}{kv \ v}$
 
@@
 The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
+The relationship among $kpg$, $cpg$ and $pg$ can be shown as follows:
+
+$\tab \kpg{bx}{ks}{kv}{lks}{lkv}{s}{v} = \cpg{bx}{ks \circ lks}{kv \circ lkv}{s}{v}\\
+    \tabs{5} = \pg{bx}{ks \ (lks \ s)}{kv \ (lkv \ v)}$
 
@@
 
-%-{
-    \noindent 
-    \textcolor{red}{
-        $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
-        \tab \textnormal{if} \ h \ s = v \ \textnormal{then} \ (ks, kv, lks, lkv, s, v) \ \textnormal{else } \text{undefined}$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{Replace}{ks}{kv}{lks}{lkv}{s}{v} = (kv, ks, lkv, lks, v, s)$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
-        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
-            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ lks}{fst \circ lkv}{s}{v};\\
-        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
-            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ lks}{snd \circ lkv}{s}{v};\\
-        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
-        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
-        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ lks}{lkv}{s}{v};\\
-        \qtab (f_2 \circ ks, kv, f_2 \circ lks, lkv, s, v)$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{lks}{g_1 \circ lkv}{s}{v};\\
-        \qtab (ks, g_2 \circ kv, lks, g_2 \circ lkv, s, v)$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
-        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
-        \tab \textnormal{then} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{lks}{lkv}{s}{v}\\
-        \tab \textnormal{else} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{lks}{lkv}{s}{v}\\
-        \tab \underline{\textnormal{fi}} \ cond_s \ (lks \ s') \ \&\& \ cond_{sv} \ s \ (lkv \ v'); \ \textnormal{return} \ (ks, kv, lks, lkv, s', v')$
-    }
-
-    \noindent 
-    \textcolor{red}{
-        $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
-        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
-        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
-    }
-%-}
-
     %+{
     \noindent $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
-        \tab \textnormal{if} \ h \ es = ev \ \textnormal{then} \ (ks, kv, elks, elkv, es, ev) \ \textnormal{else } \text{undefined}$
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
+        \tab \textnormal{if} \ h \ es = ev \ \textnormal{then} \ (ks, kv, id, id, es, ev) \ \textnormal{else } \text{undefined}$
 
@@
     \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
         \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
-            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ elks}{fst \circ elkv}{es}{ev};\\
+            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst}{fst}{es}{ev};\\
         \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
-            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ elks}{snd \circ elkv}{es}{ev};\\
+            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd}{snd}{es}{ev};\\
         \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
@@
     \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
+        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
         \tab \textnormal{if} \ cond_{sv} \ es \ ev \ \&\& \ cond_{s} \ es\\
-        \tab \textnormal{then} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{elks}{elkv}{es}{ev}\\
-        \tab \textnormal{else} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{elks}{elkv}{es}{ev}\\
-        \tab \underline{\textnormal{fi}} \ cond_s \ (lks' \ s') \ \&\& \ cond_{sv} \ \textbf{es} \ (lkv' \ v'); \ \textnormal{return} \ (ks', kv', lks', lkv', s', v')$
+        \tab \textnormal{then} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{id}{id}{es}{ev}\\
+        \tab \textnormal{else} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{id}{id}{es}{ev}\\
+        \tab \underline{\textnormal{fi}} \ cond_s \ (lks' \ s') \ \&\& \ cond_{sv} \ es \ (lkv' \ v'); \ \textnormal{return} \ (ks', kv', lks', lkv', s', v')$
 
     \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
-        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
+        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{dummy};\\
         \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
@@
     \tab \vert \ bx_1 \circ bx_2 \to \\
-        \tabs{2} (ks_1,kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ v};\\
+        \tabs{2} (ks_1,kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{dummy};\\
         \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{lkv_1 \ v_1}{v};\\

### 2020-02-19 paper/5.experiments.tex
finish renaming + revise dummy

--- 2020-02-18 paper/5.experiments.tex
@@
 \footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1.
-The OCaml runtime system options and garbage collection parameters are set as default.
-Our target language is untyped.}$^,$%HACK!
+The OCaml runtime system options and garbage collection parameters are set as default.}$^,$%HACK!
 %
@@
 described in the previous sections.
-Instead of constructing some dummies by $construct\_dummy$, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the current updated views.
+Our target language is untyped.
+Some dummies used for $pg$, $cpg$, $kpg$ and $xpg$ are replaced with the current updated views.
 This helps a program in the $put$ direction valid.

### 2020-02-19 paper/main.tex
rename

@@
 \author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and \break
-  {Bach-Trong Nguyen\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
+  {Bach Nguyen Trong\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
   {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and

### 2020-02-20 paper/1.introduction.tex
fix based on the reviews

--- 2020-02-17 paper/1.introduction.tex
@@
 
-\noindent One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
+\noindent
+\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
+One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
 Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times.
@@
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
+\textcolor{red}{Here, complements are minimal parts that are needed to evaluate $put$. 
+  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead-2},
+  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$.
+However, $[1,2]$ is simpley replaced by $[100,2]$ and not used to coustruct the result.
+In this case we can use $[..., [3]]$ as a complement.}
 The key idea is straightforward: Complements are smaller

### 2020-02-20 paper/4.xpg.tex
fix based on the reviews

@@
 This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{If the BX programs have efficient inverse way of evaluation, they are not affected by this restriction.
-Of course, some programs does not have such efficient way.
-An example restricted by this is evaluating $snoc$ with different length of a source and a view.}}.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction.
+Of course, all BX programs do not always have such efficient way.
+An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}.
 This does not happen for the others which is used in the construction of $\pginline{Case}$.
@@
 
-\textcolor{red}{TODO-C6: RearrS, RearrV just keeping, Really evaluated is skip, product}
+\textcolor{red}{
 In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
-$\kpginline{Skip}$ and $\kpginline{Case}$ are special cases, because they require the actual values for evaluation.
-Therefore we evaluate the values by application of $lks \ s$ and $lkv \ v$ and update $lks$ and $lkv$ with the identity functions.
-%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.
-When evaluating $\kpginline{bx_1 \circ bx_2}$, $\underline{lks_1}$ and $\underline{s_1}$ in the result of the first assignment are still redundant but not evaluated.
-\\
-Additionally we did two optimizations in $kpg$.
-The first is in $\kpginline{bx_1 \times bx_2}$.
-The evaluation of $lks$ and $lkv$ will be done independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
-There may be same computations in $fst \circ lks$ and $snd \circ lks$ as well as $fst \circ lkv$ and $snd \circ lkv$.
-To remove duplicate evaluations, we evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$.
+In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
+The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$. At the same time, $lks$ and $lkv$ are updated with the identity functions.
+This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
+Thanks to this accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as possible.}
+%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}
+
+\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
+  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
+  However, if we accumulate $lks$ and $lkv$, both might be evlauated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+This includes the same computation.
+To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$.
+and $\kpginline{bx_2}$}
 %It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,

### 2020-02-20 paper/5.experiments.tex
fix based on the reviews

@@
 %
-\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
+\footnote{\textcolor{red}{The implementation is available: in GitHub?} \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
 %
@@
 
-\textcolor{red}{TODO-D4: emphasize overhead of memoization. it is not for free.}
 Figure \ref{fig:result} 
@@
 the most effective method.
+
+\textcolor{red}{Note that this result focuses on the BX programs that includes many compositions.
+If the number of compositions are small in the BX programs, $put$ will be fastest, because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
+
 \textcolor{red}{TODO-B2: add short evaluation result of memory.}

### 2020-02-20 paper/6.related-work.tex
fix based on the reviews

@@
 al.
-introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
+introduces the first benchmark\footnote{\textcolor{red}{The BX programs in their benchmark are almost single BX programs.
+Because we focus on the BX programs that includes many compositions, their benchmark is not applicable for our evaluation.}} for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
 This paper shows the first attempt of improving efficiency of BX composition evaluation.

### 2020-02-20 paper/7.conclusion.tex
add todo list
fix based on the reviews

@@
 
-We will continue our work on the following three points.
+We will continue our work on the following four points.
 First is overcoming our limitation in $xpg$.
@@
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-For this, we need to investigate more about $construct\_dummy$, because the current definition used in our experiments %Section~\ref{sect:experiments}
+For this, we need to investigate more about \textcolor{red}{how to construct dummy values}, because the current definition used in our experiments %Section~\ref{sect:experiments}
 will cause type errors.
+\textcolor{red}{Fourth is using a lazy language. Although we used a strict language OCaml in this paper, if we use a lazy language, we might skip one of the optimizations.}
 

### 2020-02-20 paper/4.xpg.tex
change slightly

@@
 \subsection{Combination of pg and kpg: xpg}
-\textcolor{red}{TODO-D3: change explanation clearer.}
+\textcolor{red}{TODO-D3/D22: change explanation clearer.}
 The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.

### 2020-02-20 paper/1.introduction.tex
modified based on the discussion with Bach-san

@@
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
-\textcolor{red}{Here, complements are minimal parts that are needed to evaluate $put$. 
-  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead-2},
+\textcolor{red}{Here, complements are minimal parts of sources that are needed to evaluate $put$. 
+  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
   we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$.
-However, $[1,2]$ is simpley replaced by $[100,2]$ and not used to coustruct the result.
+However, $[1,2]$ is simply replaced by $[100,2]$ and not used to coustruct the result.
 In this case we can use $[..., [3]]$ as a complement.}

### 2020-02-20 paper/4.xpg.tex
modified based on the discussion with Bach-san

@@
 In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
-The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$. At the same time, $lks$ and $lkv$ are updated with the identity functions.
+The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$.
+At the same time, the third argument and fourth argument are updated with the identity functions.
 This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
@@
 %It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
-Second is in $\kpginline{Case}$.
-We need to evaluate $lks \ s'$ and $lkv \ v'$ to check the \underline{fi} condition before returning the 6-tuple.
+Second is in $\kpginline{Case}$ and not shown in the definition.
+We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple.
 Such evaluations can be done lazily to make programs run faster.

### 2020-02-20 paper/2.minbigul.tex
before renaming vars
insert newline to author names, spell checked

@@
         \noindent $\putbx{\product{bx_1}{bx_2}}{(s_1, s_2)}{(v_1, v_2)} = \\
-            \tab ((\putbx{bx_1}{s_1}{v_1}),(\putbx{bx_2}{s_2}{v_2}))$
+            \tab (\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2})$
 
@@
         \noindent $\getbx{\product{bx_1}{bx_2}}{(s_1,s_2)} = \\
-            \tab ((\getbx{bx_1}{s_1}),(\getbx{bx_2}{s_2}))$
+            \tab (\getbx{bx_1}{s_1},\getbx{bx_2}{s_2})$
 
@@
 Also in the semantics of $\putbxinline{Case}$ and $\getbxinline{Case}$, the return statements are used to express clearly the value of functions.
-Variables $s'$ and $v'$ wrapped in these returns are neccessary for checking the fi conditions.
+Variables $s'$ and $v'$ wrapped in these returns are necessary for checking the fi conditions.
 
@@
     $\tabs{4} f_1 = f_2^{-1} = \lambda ((x:xs),e).(x,(xs,e))$, 
-$bx_s = ((Replace \times bfoldr \ bx) \circ bx)$
+$bx_s = ((Replace \times \textit{bfoldr} \ bx) \circ bx)$
 
@@
 
-$\tab foldr :: (a \to b \to b) \to b \to [a] \to b \\
-\tabs{2} \ \ foldr \ f \ e \ [\ ] = e \\ 
-\tabs{2} \ \ foldr \ f \ e \ (x:xs) = f \ x \ (foldr \ f \ e \ xs)$
+$\tab \textit{foldr} :: (a \to b \to b) \to b \to [a] \to b \\
+\tabs{2} \ \ \textit{foldr} \ f \ e \ [\ ] = e \\ 
+\tabs{2} \ \ \textit{foldr} \ f \ e \ (x:xs) = f \ x \ (\textit{foldr} \ f \ e \ xs)$
 
@@
 
-Using $\textit{foldr}$, we can define other functions, for example:
+Using $\textit{foldr}$, we can define other functions like $reverse = \textit{foldr} \ snoc \ [\ ]$. With $\textit{bfoldr}$, we can also write the bidirectional version $breverse$ as follows:
 
-$\tab reverse = foldr \ snoc \ [\ ]$
-
-With $\textit{bfoldr}$, we can also write the bidirectional version $breverse$ as follows:
-
-$\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where:\\
-    $\tabs{4} f_1 = f_2^{-1} = \lambda s \to (s, [\ ]), \tab bx = bfoldr \ bsnoc$
+$\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where: \\
+$\tabs{4} f_1 = f_2^{-1} = \lambda s \to (s, [\ ]), \tab bx = \textit{bfoldr} \ bsnoc$

### 2020-02-20 paper/4.xpg.tex
insert newline to author names, spell checked

@@
 The problem for $cpg$ lies in redundant computations during the evaluation.
-To prevent such redundant computations from occuring, we introduce an extension named $kpg$.
+To prevent such redundant computations from occurring, we introduce an extension named $kpg$.
 While $cpg$ evaluates values eagerly, $kpg$ does the opposite.
@@
   Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
-  However, if we accumulate $lks$ and $lkv$, both might be evlauated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
+  However, if we accumulate $lks$ and $lkv$, both might be evaluated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
 This includes the same computation.

### 2020-02-20 paper/main.tex
insert newline to author names, spell checked

@@
   {Bach Nguyen Trong\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
-  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and
+  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and \break
   {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
@@
 %\item pg (pair of put and get)
-%\item cpg (extention of pg. Keep eval information for reusing)
-%\item kpg (extention of cpg. Lazy evaluation)
+%\item cpg (extension of pg. Keep eval information for reusing)
+%\item kpg (extension of cpg. Lazy evaluation)
 %\item xpg (combination of pg and kpg)

### 2020-02-20 paper/1.introduction.tex
insert newline to author names, spell checked
Update 1.intro
Polished English sentence.

@@
 \noindent
-\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
+\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right. We use this order because it is standard in the BX community.}
+%\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
 One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
@@
   we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$.
-However, $[1,2]$ is simply replaced by $[100,2]$ and not used to coustruct the result.
+However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result.
 In this case we can use $[..., [3]]$ as a complement.}

### 2020-02-21 paper/4.xpg.tex
Tiny update
Footnote 5: English polished.

@@
 This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{BX programs that have an efficient inverse evaluation are not affected by this restriction, but not all BX programs have this property.
+An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}
+% If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction.
 Of course, all BX programs do not always have such efficient way.

### 2020-02-21 paper/1.introduction.tex
fixed based on Prof. Hu's comment

@@
 \noindent
-\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right. We use this order because it is standard in the BX community.}
-%\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
+\textcolor{red}{Note that the order of composition is different from that of the common function composition.
+We use this order, because it is helpful to understand the behavior if we consider data goes from left to right.}
 One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
@@
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
-\textcolor{red}{Here, complements are minimal parts of sources that are needed to evaluate $put$. 
+\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
   In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},

### 2020-02-21 paper/4.xpg.tex
Tiny update
English polished.
fixed based on Prof. Hu's comment

--- 2020-02-20 paper/4.xpg.tex
@@
 This means two entering conditions of $\putbxinline{Case}$ and $\getbxinline{Case}$ are not always simultaneously satisfied.
-The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper\footnote{\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction.
+The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+This does not happen for the others which is used in the construction of $\pginline{Case}$.
+% {\textcolor{red}{If the BX programs have efficient inverse way for evaluation, they are not affected by this restriction.
 Of course, all BX programs do not always have such efficient way.
-An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}.
-This does not happen for the others which is used in the construction of $\pginline{Case}$.
+An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}
 
@@
 \textcolor{red}{
-In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
+In $kpg$, basically, functions for the updates are kept (but not evaluated) in $lks$ and $lkv$.
 In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
-The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$.
-At the same time, the third argument and fourth argument are updated with the identity functions.
+These two functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by applications of $lks \ s$ and $lkv \ v$.
+At the same time, the third arguments and fourth arguments of recursive calls are updated with the identity functions.
 This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
-Thanks to this accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as possible.}
+Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.}
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}
@@
 To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$.
-and $\kpginline{bx_2}$}
+and $\kpginline{bx_2}$}.
 %It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
-Second is in $\kpginline{Case}$ and not shown in the definition.
+The second is in $\kpginline{Case}$ and not shown in the definition.
 We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple.

### 2020-02-21 paper/5.experiments.tex
modified based on the discussion with Bach-san
Very tiny update
Footnote 7: spacing reduced.
fixed based on Prof. Hu's comment

--- 2020-02-19 paper/5.experiments.tex
@@
 %
-\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
+\footnote{\textcolor{red}{The implementation is available: } \url{https://github.com/k-tsushima/pgs}}
 %
@@
 the most effective method.
-\textcolor{red}{TODO-B2: add short evaluation result of memory.}
+
+\textcolor{red}{Note that this result focuses on the BX programs that include many compositions.
+If the number of compositions is small in the BX programs, the original $put$ without any memoization will be fastest, because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
+
+%\textcolor{red}{TODO-B2: add short evaluation result of memory.}

### 2020-02-21 paper/6.related-work.tex
fixed based on Prof. Hu's comment

--- 2020-02-17 paper/6.related-work.tex
@@
 al.
-introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
+introduces the first benchmark\footnote{\textcolor{red}{The BX programs in their benchmark are basically BX programs without composition.
+Because we focus on the BX programs that include many compositions, their benchmark is not applicable for our evaluation.}} for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
 This paper shows the first attempt of improving efficiency of BX composition evaluation.

### 2020-02-21 paper/main.tex
change anonymus -> our name
tiny update main
(RG email updated)
fixed based on Prof. Hu's comment

@@
 
+\newcommand{\newcirc}{\tilde{\circ}}
+
 \makeatletter
@@
 
-\author{{Kanae Tsushima\inst{1}\orcidID{0000-0002-3383-3389}} \and \break
-  {Bach Nguyen Trong\inst{1}\orcidID{0000-0002-6342-6020}} \and \break
-  {Robert Gl\"{u}ck\inst{2}\orcidID{0000-0001-6990-3935}} \and \break
-  {Zhenjiang Hu\inst{3}\orcidID{0000-0002-9034-205X}}}
+\author{{Kanae Tsushima\inst{1}} \and \break% \orcidID{0000-0002-3383-3389}
+  {Bach Nguyen Trong\inst{1}} \and \break %\orcidID{0000-0002-6342-6020}
+  {Robert Gl\"{u}ck\inst{2}} \and \break %\orcidID{0000-0001-6990-3935}
+  {Zhenjiang Hu\inst{3}}} %\orcidID{0000-0002-9034-205X}
 
@@
 
-\authorrunning{anonymous authors}
+%\authorrunning{anonymous authors}
 
@@
 \institute{National Institute of Informatics, Japan 
-  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
+  (\email{\{k\_tsushima,bach\}@nii.ac.jp}) \and
   University of Copenhagen, Denmark 
-  \email{glueck@di.ku.dk } \and
+  (\email{glueck@di.ku.dk}) \and
   Peking University, China
-  \email{huzj@pku.edu.cn}}
+  (\email{huzj@pku.edu.cn})}
 

### 2020-02-21 paper/5.experiments.tex
Tiny update
Polished English of last paragraph.

--- 2020-02-19 paper/5.experiments.tex
@@
 %
-\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
+\footnote{\textcolor{red}{The implementation is available:} \url{https://github.com/k-tsushima/pgs}}
 %
@@
 the most effective method.
-\textcolor{red}{TODO-B2: add short evaluation result of memory.}
+
+\textcolor{red}{Note that this result concerns BX programs that uses many compositions.
+If the number of compositions is small, $put$ will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
+
+%\textcolor{red}{TODO-B2: add short evaluation result of memory.}

### 2020-02-21 paper/1.introduction.tex
small fix

@@
   we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$.
-However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result.
+However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$.
 In this case we can use $[..., [3]]$ as a complement.}
-The key idea is straightforward: Complements are smaller
+The key idea of the second approach is straightforward: Complements are smaller
 % program fragments

### 2020-02-21 paper/2.minbigul.tex
small fix

@@
 
-If we think that a minBiGUL program $bx$ has a type of $\textit{MinBiGUL} \ s \ v$, the type of $bfoldr$ will be look like $\textit{MinBiGUL} \ (a, b) \ a \to \textit{MinBiGUL} \ ([a], b) \ b$.
+If we think that a minBiGUL program $bx$ has a type of $\textit{MinBiGUL} \ s \ v$, the type of $bfoldr$ will be look like $\textit{MinBiGUL} \ (a, b) \ b \to \textit{MinBiGUL} \ ([a], b) \ b$.
 You can easily find the similarity between the above definition of $\textit{bfoldr}$ with the following definition of $\textit{foldr}$:

### 2020-02-21 paper/4.xpg.tex
small fix

@@
 The evaluated branches are distinct in the $put$ and $get$ directions for combinations $((cond_{sv} \ s \ v) \ \&\& \ (not (cond_{s} \ s)))$ and $((not (cond_{sv} \ s \ v)) \ \&\& \ (cond_{s} \ s))$, which are restricted in this paper.
+Because they evaluate different $bx$ for $put$ and $get$, we can not evaluate them efficiently.
 This does not happen for the others which is used in the construction of $\pginline{Case}$.
@@
 
-The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function.
+The construction of $\pginline{bx_1 \circ bx_2}$ is the most important part in the $pg$ function.
 The first two equalities comes from mentioned definitions and some basic transformations.
@@
 This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
-Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$.
+Since there is no real view, this $dummy$ is necessary to pair with the original source $s$ to form the input of $put \, [\![bx_1]\!]$.
 In general, $dummy$ depends on the source $s$, the view $v$ and/or the program $bx_1$.

### 2020-02-21 paper/1.introduction.tex

--- 2020-02-20 paper/1.introduction.tex
@@
 \noindent
-\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right. We use this order because it is standard in the BX community.}
+\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right.
+We use this order because it is helpful to understand the behavior if we consider data goes from left to right.}
 %\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
@@
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
-\textcolor{red}{Here, complements are minimal parts of sources that are needed to evaluate $put$. 
+\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
   In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
   we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$.
-However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result.
+However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$.
 In this case we can use $[..., [3]]$ as a complement.}
-The key idea is straightforward: Complements are smaller
+The key idea of the second approach is straightforward: Complements are smaller
 % program fragments

### 2020-02-21 paper/5.experiments.tex

@@
 \textcolor{red}{Note that this result concerns BX programs that uses many compositions.
-If the number of compositions is small, $put$ will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
+If the number of compositions is small, the original $put$ withoug any memoization will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
 

### 2020-02-21 paper/main.tex
small fix

@@
 \newcommand{\xpginline}[1]{xpg \, [\![#1]\!]}
-\newcommand{\smallvspace}{\vspace{0.3mm}}
+\newcommand{\smallvspace}{\vspace{1.2mm}}
 \newcommand{\minusvspace}{\vspace*{-5.5mm}}
@@
   University of Copenhagen, Denmark
-  (\email{glueck@di.ku.dk}) \and
+  (\email{glueck@acm.org}) \and
   Peking University, China

### 2020-02-21 paper/7.conclusion.tex
fixed based on Prof. Hu's comment
removing comma before because

@@
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-For this, we need to investigate more about \textcolor{red}{how to construct dummy values}, because the current definition used in our experiments %Section~\ref{sect:experiments}
+For this, we need to investigate more about \textcolor{red}{how to construct dummy values} because the current definition used in our experiments %Section~\ref{sect:experiments}
 will cause type errors.
-\textcolor{red}{Fourth is using a lazy language. Although we used a strict language OCaml in this paper, if we use a lazy language, we might skip one of the optimizations.}
+\textcolor{red}{Fourth is using a lazy language. Although we used a strict language OCaml in this paper, if we use a lazy language, we might get laziness for free.}
 

### 2020-02-21 paper/4.xpg.tex
merged

@@
 Of course, all BX programs do not always have such efficient way.
+An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}
+
+    %{\textcolor{red}{BX programs that have an efficient inverse evaluation are not affected by this restriction, but not all BX programs have this property.
 An example restricted by this is evaluating $bsnoc$ with different length of a source and a view.}}
@@
 In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
-These two functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by applications of $lks \ s$ and $lkv \ v$.
-At the same time, the third arguments and fourth arguments of recursive calls are updated with the identity functions.
-This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
+The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by applications of $lks \ s$ and $lkv \ v$.
+At the same time, the third and fourth argument of recursive calls are updated with the identity function.
+This evaluation is needed because these definitions require the actual values, $es$ and $ev$.
 Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.}

### 2020-02-21 paper/1.introduction.tex
small changes: \circ -> \tilde{\circ} in intro. sec., \smallvspace

@@
     \includegraphics[height=3.5cm]{./fig/fig2.eps}
-    \caption{Evaluating $phead \circ phead \circ phead$}
+    \caption{Evaluating $phead~\tilde{\circ}~phead~\tilde{\circ}~phead$}
     \label{fig:eval-comp-phead}
@@
 The composition of BX programs is a fundamental construct to build more complex BX programs \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}.
-Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_1 \circ bx_2$ is defined by
+Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_1 \,\tilde{\circ}\, bx_2$ is defined by
 
@@
 \begin{align}
-get_{bx_1 \circ bx_2}~s &= get_{bx_2} (get_{bx_1}~s)\\
-put_{bx_1 \circ bx_2}~s~v' &= put_{bx_1} s~(put_{bx_2}~ (get_{bx_1}~s)~v')
+get_{bx_1 \tilde{\circ} bx_2}~s &= get_{bx_2} (get_{bx_1}~s)\\
+put_{bx_1 \tilde{\circ} bx_2}~s~v' &= put_{bx_1} s~(put_{bx_2}~ (get_{bx_1}~s)~v')
 \end{align}
@@
 %\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
-One feature of this composition is that $put_{bx_1 \circ bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
+One feature of this composition is that $put_{bx_1 \tilde{\circ} bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs.
 Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times.
@@
 \[
-lp3 = (phead \circ phead) \circ phead
+lp3 = (phead~\tilde{\circ}~phead)~\tilde{\circ}~phead
 \]
@@
 %In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily.
-The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).
+The running examples are two versions of two composition of $phead$: ($phead \tilde{\circ} phead) \tilde{\circ} phead$ (we call lp3) and $phead \tilde{\circ} (phead \tilde{\circ} phead$) (we call rp3).
 
 %For more various evaluations, we can use the composition of programs.
-Let us consider another example, a composition of 3 $pHead$s: $pHead \circ pHead \circ pHead$.
+Let us consider another example, a composition of 3 $pHead$s: $pHead \tilde{\circ} pHead \tilde{\circ} pHead$.
 Figure \ref{fig:eval-comp-phead} illustrates the evaluation of this program where the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]} and the updated view is \texttt{100}.
@@
 \[
-rp3 = phead \circ (phead \circ phead)
+rp3 = phead~\tilde{\circ}~(phead~\tilde{\circ}~phead)
 \]
@@
  \[
- \mbox{\it bfoldr}~bx~\cdots = \cdots  (\mbox{\it bfoldr}~\cdots) \circ bx \cdots
+ \mbox{\it bfoldr}~bx~\cdots = \cdots  (\mbox{\it bfoldr}~\cdots)~\tilde{\circ}~bx \cdots
  \]
@@
 
-%A part of the definition of $\mbox{\it bfoldr}$, we have $(\product{Replace}{(\mbox{\it bfoldr} \ bx)}) \circ bx$.
+%A part of the definition of $\mbox{\it bfoldr}$, we have $(\product{Replace}{(\mbox{\it bfoldr} \ bx)}) \tilde{\circ} bx$.
 In this program, recursion occurs on the left-hand side of composition.
@@
   \vspace{-0.25cm}
-  \caption{Evaluating $phead \ \circ \ phead \ \circ \ phead$ by keeping complements}
+  \caption{Evaluating $phead \ \tilde{\circ} \ phead \ \tilde{\circ} \ phead$ by keeping complements}
   \vspace{-0.5cm}
@@
 For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
-Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$.
+Let us reconsider our example $phead~\tilde{\circ}~phead~\tilde{\circ}~phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$.
 Here, two points are worth noting.

### 2020-02-21 paper/2.minbigul.tex
small changes: \circ -> \tilde{\circ} in intro. sec., \smallvspace

@@
 To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$.
+\textcolor{blue}{The compose symbol $\tilde{\circ}$ used in the previous section will be replaced with the more common one, $\circ$.}
 In general, $\circ$ has a higher priority than $\times$.
@@
 
-\noindent As an example of minBiGUL program, consider the definition of $\textit{phead}$ :\\
+\noindent As an example of minBiGUL program, consider the definition of $\textit{phead}$:
+
 \smallvspace
-$\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
+\noindent $\tab phead = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \lambda (s::ss).(s,ss), \, f_2 = \lambda (s,ss).(s::ss),\\
     \qtab bx_s = RearrV \ g_1 \ g_2 \ bx_v \ \text{where: } g_1 = \lambda v.(v,()), \, g_2 = \lambda (v,()).v,\\
@@
 
+\smallvspace
     $\tab \textit{bsnoc} = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
@@
             $\tabs{5} g_1 = g_2^{-1} = \lambda (v:vs).(v,vs), \tab bx_v = Replace \times \textit{bsnoc}$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \textit{bfoldr} \ bx = Case \ cond_{sv} \ cond_s \ bx_1 \ bx_2$ where:\\
@@
 $bx_s = ((Replace \times \textit{bfoldr} \ bx) \circ bx)$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \textit{foldr} :: (a \to b \to b) \to b \to [a] \to b \\
@@
 \tabs{2} \ \ \textit{foldr} \ f \ e \ (x:xs) = f \ x \ (\textit{foldr} \ f \ e \ xs)$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab breverse = RearrS \ f_1 \ f_2 \ bx$ where: \\

### 2020-02-21 paper/1.introduction.tex
remove red color part & small fix

@@
 \noindent
-\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right.
-We use this order because it is helpful to understand the behavior if we consider data goes from left to right.}
+Unlike function composition, the composition of bidirectional programs is read left-to-right.
+We use this order because it is helpful to understand the behavior if we consider data goes from left to right.
 %\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
@@
 To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
-\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
+Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
   In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
@@
 However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$.
-In this case we can use $[..., [3]]$ as a complement.}
+In this case we can use $[..., [3]]$ as a complement.
 The key idea of the second approach is straightforward: Complements are smaller

### 2020-02-21 paper/2.minbigul.tex
remove red color part & small fix

@@
 To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$.
-\textcolor{blue}{The compose symbol $\tilde{\circ}$ used in the previous section will be replaced with the more common one, $\circ$.}
+The compose symbol $\tilde{\circ}$ used in the previous section will be replaced with the more common one, $\circ$.
 In general, $\circ$ has a higher priority than $\times$.

### 2020-02-21 paper/4.xpg.tex
small changes: \circ -> \tilde{\circ} in intro. sec., \smallvspace
remove red color part & small fix

@@
 
+\smallvspace
     \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\textit{undefined}}, h \ s) \\
@@
         \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\textit{undefined}}$
+\smallvspace
 
@@
 
+\smallvspace
     \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
@@
         \tab \qtabs{2} (s', g_2 \ v')$
+\smallvspace
 
@@
 
+\smallvspace
     \noindent $\tabs{2} \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
@@
     \tabs{2} \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$
+\smallvspace
 
@@
 
+\smallvspace
     \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
@@
             \qtab \tabs{3} (s_3, v_2)$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \cpg{bx}{ks}{kv}{s}{v} = \pg{bx}{ks \ s}{kv \ v}$
+\smallvspace
 
@@
 
+\smallvspace
 $\tab \kpg{bx}{ks}{kv}{lks}{lkv}{s}{v} = \cpg{bx}{ks \circ lks}{kv \circ lkv}{s}{v}\\
     \tabs{5} = \pg{bx}{ks \ (lks \ s)}{kv \ (lkv \ v)}$
+\smallvspace
 
@@
 
-\textcolor{red}{
 In $kpg$, basically, functions for the updates are kept (but not evaluated) in $lks$ and $lkv$.
@@
 This evaluation is needed because these definitions require the actual values, $es$ and $ev$.
-Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.}
+Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.
 %Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}
 
-\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
+Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
   Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
@@
 This includes the same computation.
-To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$.
-and $\kpginline{bx_2}$}.
+To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
 %It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,

### 2020-02-21 paper/5.experiments.tex
remove red color part & small fix

--- 2020-02-21 paper/5.experiments.tex
@@
 %
-\footnote{\textcolor{red}{The implementation is available:} \url{https://github.com/k-tsushima/pgs}}
+\footnote{The implementation is available: \url{https://github.com/k-tsushima/pgs}}
 %
@@
 
-\textcolor{red}{Note that this result concerns BX programs that uses many compositions.
-If the number of compositions is small, $put$ will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}
+Note that this result concerns BX programs that uses many compositions.
+If the number of compositions is small, the original $put$ without any memoization will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).
 

### 2020-02-21 paper/6.related-work.tex
remove red color part & small fix

--- 2020-02-17 paper/6.related-work.tex
@@
 al.
-introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
+introduces the first benchmark\footnote{The BX programs in their benchmark are basically BX programs without composition.
+Because we focus on the BX programs that include many compositions, their benchmark is not applicable for our evaluation.} for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing.
 This paper shows the first attempt of improving efficiency of BX composition evaluation.

### 2020-02-21 paper/7.conclusion.tex
remove red color part & small fix

--- 2019-12-03 paper/7.conclusion.tex
@@
 If we introduce this, we can avoid runtime errors like statically-typed functional languages.
-For this, we need to investigate more about $construct\_dummy$, because the current definition used in our experiments %Section~\ref{sect:experiments}
+For this, we need to investigate more about how to construct dummy values because the current definition used in our experiments %Section~\ref{sect:experiments}
 will cause type errors.
+Fourth is using a lazy language. Although we used a strict language OCaml in this paper, if we use a lazy language, we might get laziness for free.
 

### 2020-02-21 paper/4.xpg.tex
fixed small part of explanation

@@
 
-Constructions of $pg$ for the replacement, the product and the source/view rearrangements are very clear when just pairing $put$ and $get$ respectively, then doing basic changes.
+Constructions of $pg$ for the replacement, the product and the source/view rearrangements are simple. We just pair $put$ and $get$, and change them to $pg$.
 %

### 2020-02-21 paper/main.tex
add acknowledgement about reviewers and KAKENHI
fixed small part of explanation

@@
 
-\author{{Kanae Tsushima\inst{1}} \and \break% \orcidID{0000-0002-3383-3389}
-  {Bach Nguyen Trong\inst{1}} \and \break %\orcidID{0000-0002-6342-6020}
-  {Robert Gl\"{u}ck\inst{2}} \and \break %\orcidID{0000-0001-6990-3935}
+\author{{Kanae Tsushima\inst{1}} \and %\break% \orcidID{0000-0002-3383-3389}
+  {Bach Nguyen Trong\inst{1}} \and %\break %\orcidID{0000-0002-6342-6020}
+  {Robert Gl\"{u}ck\inst{2}} \and %\break %\orcidID{0000-0001-6990-3935}
   {Zhenjiang Hu\inst{3}}} %\orcidID{0000-0002-9034-205X}
@@
 \input{7.conclusion.tex}
+
+\subsubsection*{Acknowledgement}
+We thank the anonymous reviewers for valuable comments and suggestions. This work has been partially supported by JSPS KAKENHI Grant Number JP17H06099.
 

### 2020-02-21 paper/4.xpg.tex
modified the beginning of 4.4

@@
 If a given program is a left associative composition, the number of $pg$ calls will be exponential.
-Therefore, the runtime inefficiency is inevitable.
+Therefore, the runtime inefficiency of $pg$ for left associative BX programs is inevitable.
 To solve that, we introduce a new function, $cpg$, accumulates updates on the source and the view.
@@
 Since there is no real view here, we need a dummy same as $pg$.
-Then the continuation updating on this dummy should be initiated as an identity function.
+Then the continuation updating on this dummy should be initiated as the identity function.
 The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$.
@@
 
-In general, the beginning of a continuation should be an identity function.
+In general, the beginning of a continuation should be the identity function.
 However, to be able to use the function application to get the result of $\cpginline{bx_1 \circ bx_2}$, the accumulative function on the source needs to be initiated as the constant function from that source.
@@
 The input of $\kpginline{bx}$ is expanded to a 6-tuple $(ks, kv, lks, lkv, s, v)$ where $ks$ and $kv$ keep the modification information same as $cpg$, $s$ and $v$ hold evaluated values, and $lks$ and $lkv$ are used for lazy evaluation of actual values.
-The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.\\
+The output of this function is also a 6-tuple $(ks, kv, lks, lkv, s, v)$.
+
 Suppose that we have a source $s_0$ and a view $v_0$.
@@
 
-The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
+The beginning of accumulative functions $lks$ and $lkv$ are set as the identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.
 The relationship among $kpg$, $cpg$ and $pg$ can be shown as follows:
@@
 \subsection{Combination of pg and kpg: xpg}
-\textcolor{red}{TODO-D3/D22: change explanation clearer.}
-The most important thing when using $kpg$ to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.
-This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$.
+
+The purpose we introduced $cpg$ and $kpg$ is to avoid redundant recursive call and keep the dropped parts from the source in a function.
+On the other hand, these accumulations in $cpg$ and $kpg$ will be an overhead if they are not necessary.
+The problem in $\pginline{bx_1 \circ bx_2}$ is that there are two recursive calls of $\pginline{bx_1}$ and there is no problem in the recursive call of $\pginline{bx_2}$.
+Therefore, we combine $pg$ and $kpg$ to take advantage of both approaches.
+
+%The most important thing when using to evaluate a composition $bx_1 \circ bx_2$ is keeping the dropped parts from the source in a function that are used to produce new source instead of using more recursive call.
+
+%This is reflected in the first call $\kpginline{bx_1}$ and the last application in $\kpginline{bx_1 \circ bx_2}$.
 The second call $\kpginline{bx_2}$ just play a syntactic role while it requires many parameters.

### 2020-02-21 paper/main.tex
modified the beginning of 4.4

@@
 
-\institute{National Institute of Informatics, Japan
-  (\email{\{k\_tsushima,bach\}@nii.ac.jp}) \and
-  University of Copenhagen, Denmark
-  (\email{glueck@acm.org}) \and
-  Peking University, China
-  (\email{huzj@pku.edu.cn})}
+\institute{National Institute of Informatics, Japan \break
+  \email{\{k\_tsushima,bach\}@nii.ac.jp} \and
+  University of Copenhagen, Denmark \break
+  \email{glueck@acm.org} \and
+  Peking University, China \break
+  \email{huzj@pku.edu.cn}}
 
 %\institute{anonymous}
+
+\pagestyle{empty}
 

### 2020-02-21 paper/main.tex
removed unneeded part in main.tex

@@
   Implementation technique \and
-  %Evaluation? \and
   Efficiency \and 
   Optimization \and Tupling.
-  %\and Lazy update?
 }
 \end{abstract}
-%
-%
-%
-
-% plan to pages
-% 1.abstraction & title & Introduction: 3pages
-% 2.miniBiGUL: 1.5page
-% 3.miniBiGUL_m : 1 page
-% 4.xpg:5.5page
-% 5.evaluation:2pages
-% 6.related work:1page
-% 7.conclusion:1page (including reference)
-% page limit : 15
-
-%\begin{itemize}
-%\item FLOPS 2020, 15 November 2019 (AoE): Abstract submission
-%\item 22 November 2019 (AoE): Submission deadline
-%\item Today's agenda: introduction.
-% result of 2nd approach, memory alloc result. introduction, contribution
-%\end{itemize}
-
-%Schedule
-%\begin{itemize}
-%\item Introduction (story), Contribution, positions, examples, beginning of September?
-%\item zero draft : mid of September  
-%\item deadline of PEPM2020 (October 18th)
-%\item ESOP (October 24th)
-%\end{itemize}
-
-%Recent result/findings:
-%\begin{itemize}
-%\item miniBigul\_m becomes worse when the number of compositions is increased (in the case of Replace, pHead).
-Using some optimization to avoid comparison between BXs, it becomes faster, but it is still slower than original miniBiGUL.
-%\item bmapreplace is a special case of miniBiGUL\_m. Why??
-%\item miniBigul\_m is better for bmapreplace and breverse than other approaches.
-%\item Compose(breverse, breverse): miniBiGUL can evaluate, but kpg can not. 
-%\end{itemize}
-
-%\begin{tabular}{c|c|c|c}
-%  & 1st(miniBiGUL) & 2nd(memo\_miniBiGUL) & 3rd(kpg) \\ \hline
-%  num of $get$s & linear to quadratic & linear & linear \\ \hline 
-%  (storage) & (linear) & (linear to N(num of composition)) * (size(S)) & (max(size(S), size(S'))) \\ \hline
-%  overhead & low & saving (BX, s) to table \& searching & keeping as functions
-% \end{tabular}
-
-%Time:
-%\begin{itemize}
-%\item Lassoc (comp\_replace \& comp\_phead): xpg $<$ miniBiGUL << miniBiGUL\_m (100x)
-%\item Rassoc (comp\_replace \& comp\_phead): miniBiGUL $<$ xpg $<$ miniBiGUL\_m (1000x)
-%\item breverse: miniBiGUL\_m << xpg (10x) << miniBiGUL(100x)
-%\item bmapreplace: (miniBiGUL\_m <<) xpg << miniBiGUL (100x)
-%\end{itemize}
-
-%Memory allocation:
-%\begin{itemize}
-%\item Lassoc comp\_replace : miniBiGUL $<$ xpg $<$ miniBiGUL\_m
-%\item Lassoc\& Rassoc comp\_phead: miniBiGUL\_m << miniBiGUL < xpg
-%\item Rassoc comp\_replace: miniBiGUL < xpg < miniBiGUL\_m
-%\item breverse: miniBiGUL\_m < xpg << miniBiGUL
-%\item bmapreplace xpg << miniBiGUL 
-%\end{itemize}
-
-%\begin{itemize}
-%\item Rassoc $\to$ miniBiGUL
-%\item 
-%\end{itemize}
-
-%\begin{itemize}
-%\item BiGUL 
-%\item pg (pair of put and get)
-%\item cpg (extension of pg. Keep eval information for reusing)
-%\item kpg (extension of cpg. Lazy evaluation)
-%\item xpg (combination of pg and kpg)
-%\item Experiment
-%\item Related work
-%\item Conclusion and Future work
-%\end{itemize}
 
@@
 
-% \tiny
-
-% \begin{thebibliography}{8}
-% \bibitem{viewupdate}
-% F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.
-
-% \bibitem{BXcomp}
-% Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
-% ``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019
-
-% \bibitem{tupling}
-%   M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.
-
-% \bibitem{josh_pepm}
-%   Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.
-
-% \bibitem{josh_popl}
-%   Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.
-
-% \bibitem{BXtend}
-%   Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
-% of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.
-
-% \bibitem{eMoflon}
-%   Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr.
-Developing eMoflon with eMoflon.
-In Davide Di Ruscio and Daniel Varro, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014.
-Proceedings, volume 8568 of Lecture Notes in
-% Computer Science, pages 138–145. Springer, 2014.
-
-% \bibitem{EVL+Strace}
-% Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
-% Technology, 100:47--72, 2018.
-
-% \bibitem{JTL}
-% Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio.
-JTL: A bidirectional and change propagating transformation language.
-In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages 183--202, Springer-Verlag.
-
-% \bibitem{NMF}
-% Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.
-
-% \bibitem{SDMLib}
-%   SDMLib paper?
-
-% \bibitem{lenses}
-%   Lense papers?
-
-% \bibitem{tutorial}
-%   Tutorial paper of BiGUL
-
-% \bibitem{putget}
-%   H. Pacheco, Z. Hu, and S. Fischer. Monadic combinators for “putback” style bidirectional programming. In Partial Evaluation and Program Manipulation, pages 39–50. ACM, 2014.
-
-% \bibitem{putput}
-
-%\bibitem{ref_article1}
-%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
-
-%\bibitem{ref_lncs_1}
-%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-%Springer, Heidelberg (2016). \doi{10.10007/1234567890}
-
-%\bibitem{ref_book1}
-%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-%Location (1999)
-
-%\bibitem{ref_proc1}
-%Author, A.-B.: Contribution title. In: 9th International Proceedings
-%on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-%\bibitem{ref_url1}
-%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-%Oct 2017
-% \end{thebibliography}
-
-% \input{appendix.tex}
-
 \end{document}
-
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
-
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
-
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
-
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig_1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig_1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig_1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs_1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs_1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%

### 2020-02-21 paper/5.experiments.tex
small fix

@@
 
-Note that this result concerns BX programs that uses many compositions.
-If the number of compositions is small, the original $put$ without any memoization will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).
+Note that this result concerns BX programs that use many compositions.
+If the number of compositions is small, the original $put$ without memoization will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).
 

### 2020-02-21 paper/main.tex
fix typo
small fix

@@
 
-\subsubsection*{Acknowledgement}
-We thank the anonymous reviewers for valuable comments and suggestions. This work has been partially supported by JSPS KAKENHI Grant Number JP17H06099.
+\subsubsection*{Acknowledgment}
+We thank the anonymous reviewers for valuable comments and suggestions. This work has been partially supported by JSPS KAKENHI Grant Number JP17H06099, and ROIS NII Open Collaborative Research 2018.
 
